### 2024-02-15 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
Initial commit

---
title: "Multiple regression: TRELIS-Q WP5 data"
author: "Niklas Karlsen"
date: "2024-02-15"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/bookchapter.TRELIS-Q-WP5.html')) })
---

# Initialize

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) #,results="asis")

Sys.setenv(LANG = "no_NB.UTF-8")            # for norwegian æ,ø og å. Kilde:
Sys.setlocale("LC_ALL","Norwegian")
# https://stackoverflow.com/questions/16347731/how-to-change-the-locale-of-r
```

## Load libraries

```{r load_libraries, echo=FALSE, warning=FALSE, message=FALSE}
library(pander)                             # prettyprint for tables

library(psych)                              # for ICC, SD

library(mice)                               # for md.pattern (missing)
library(viridis)

# cluster-variable
# kilde: https://www.r-bloggers.com/2021/05/clustered-standard-errors-with-r/
library(estimatr)                           # for cluster corrected SE
library(sandwich)
library(lmtest)
```

## Help functions

```{r}
# describe data (mean, sd, skew, curtosis)
describe_data <- function (datalist, QUIET=1) {
  im <- mean(datalist, na.rm=TRUE)
  isd <- sd(datalist, na.rm=TRUE)
  isk <- psych::skew(datalist, na.rm=TRUE) # from psych
  ik <- kurtosi(datalist, na.rm=TRUE) # from psych
  mis <- sum(is.na(datalist)) # missing
  ir <- max(datalist, na.rm=TRUE) -          
    min(datalist, na.rm=TRUE) # range

  # source: https://stackoverflow.com/questions/34351598/getting-name-of-variable-in-r
  if (QUIET != 1) {
    print(sprintf("%7s = %s", "Item", deparse(substitute(datalist))))
    print(sprintf("%7s = %.3f", "mean", im))
    print(sprintf("%7s = %.3f", "std.dev", isd))
    print(sprintf("%7s = %.3f", "skew", isk))
    print(sprintf("%7s = %.3f", "kurt", ik))
    print(sprintf("%7s = %.3f", "range", ir))
    print(sprintf("%7s = %i", "missing", mis))

    if (isk > 2 | ik > 7) {
      print("                            Skew or curtosis TOO LARGE.")
    }
  }

  # outliers
  lowerq = quantile(datalist, na.rm=TRUE)[2]
  upperq = quantile(datalist, na.rm=TRUE)[4]
  iqr = upperq - lowerq #Or use IQR(df)
  # gang med 3 for extreme outliers, 1.5 for mild outliers

  mild.threshold.upper = (iqr * 3) + upperq
  mild.threshold.lower = lowerq - (iqr * 3)

  # print outliers
  # kilde: https://stackoverflow.com/questions/19353116/how-do-i-print-values-in-a-list-that-are-greater-than-a-certain-number-along-wit#19353192
  outliers_upper <- which(datalist>mild.threshold.upper)
  outliersu <- datalist[outliers_upper]
  outliers_lower <- which(datalist<mild.threshold.lower)
  outliersl <- datalist[outliers_lower]

  if (length(outliersu) > 0 | length(outliersl) > 0) {
    print("outliers upper =")
    print(outliersu)
    print("outliers lower =")
    print(outliersl)
  }

  #hist (datalist, main=i) # histogram
  #readline(prompt = "Press enter")

  return(c("mean"=round(im,3), "skew"=round(isk,3), "kurtosi"=round(ik,3)))
           #round(outliersu,3), round(outliersl,3)))
}
```

## Define constructs

```{r load_constructs}
## TPACK-PST for TRELIS-Q
TK_pos_trelisq    <- c("TK1b","TK2","TK34") 
PCK_pos_trelisq   <- c("CK123","CK5","PCK2","PCK5")
TPACK_pos_trelisq <- c("TPACK2","TPACK4","TCK1","TCK4","TPK1","TPK4")

TPACK_PST_list    <- list(TK=TK_pos_trelisq,
                     PCK=PCK_pos_trelisq,
                     TPACK=TPACK_pos_trelisq)
```

# Data management

## Load data

```{r load_data_cfa}
# Rdata-file contains reduced / prepared datasets
load(file="../data/processed_data/tpack-TRELIS-Q_2023.Rdata")
```

## Prepare data

OBS! Anonymiser data i datafil på studiested før det legges ut åpent på nett

```{r}
# identify NA as studiested 5(?) 
data_trelis_q_ren[data_trelis_q_ren$studiested == "Nord universitet","Trinn"] <- "Trinn 5-10"
```

## Check for missing

```{r}
# covariates
md.pattern(data_trelis_q_ren[,c("studiested","Gender", "Trinn", "semester")])

# items / likert data
md.pattern(data_trelis_q_ren[, unlist(TPACK_PST_list)])
```

```{r fix_missing, eval=T}
# remove NMAR + NTNU (19)
data_clean <- data_trelis_q_ren[-c(19,34,44,45,46,95,102),]

# jeg skjuler Annet, siden det bare er en person; alternativt NA?
data_clean$Trinn[data_clean$Trinn == "Annet"] <- NA # "Trinn 5-10"
#trinn_r[is.na(trinn_r)] <- 1

# jeg skulet Annet / ønsker ikke å oppgi, siden det bare er to personer
# skjuler som Kvinne, siden det er flest av de
data_clean$Gender[data_clean$Gender == "Annet/ønsker ikke å oppgi"] <- NA # "Kvinne"
#gender_r[is.na(gender_r)] <- 1
```

# Descriptive statistics

```{r}
scores_v23 <- scoreItems(TPACK_PST_list, data_clean)

describe(scores_v23$scores)

# constructs for multiple regression
TK_v23    <- scores_v23$scores[,"TK"]
PCK_v23   <- scores_v23$scores[,"PCK"]
TPACK_v23 <- scores_v23$scores[,"TPACK"]

# add constructs to data.frame
data_clean$TK    <- TK_v23
data_clean$PCK   <- PCK_v23
data_clean$TPACK <- TPACK_v23
```

## Check for interactions

```{r}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Trinn),
                 trace.factor = factor(data_clean$Gender),
                 response = data_clean$TK,
                 ylab = "TK",
                 xlab = "Target grade level",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Gender")
```

```{r}
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Trinn),
                 trace.factor = factor(data_clean$Gender),
                 response = data_clean$TPACK,
                 ylab = "TPACK",
                 xlab = "Target grade level",
                 col = viridis(2), # c("pink", "blue"),
                 trace.label = "Gender")
```

```{r}
interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TPACK,
                 ylab = "TPACK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

```{r}
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TPACK,
                 ylab = "TPACK",
                 xlab = "Gender",
                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

# Regression analysis

## Check assumptions for regression

```{r normality}
# covariates; 80/20 distribution
# ref http://stackoverflow.com/questions/21639392/ddg#21639445
barplot(prop.table(table(data_clean$Gender)))
barplot(prop.table(table(data_clean$Trinn)))
barplot(prop.table(table(data_clean$semester)))
#hist(oslomet_r)
#hist(innlandet_r)
#hist(HVL_r)
#hist(UIT_r)

# variable
hist(data_clean$TK)
hist(data_clean$PCK)
hist(data_clean$TPACK)
```

```{r}
describe_data(data_clean$TK)
describe_data(data_clean$PCK)
describe_data(data_clean$TPACK)
```

```{r scatterplot_linearity, eval=F}
plot(data_clean$Gender,        data_clean$TPACK)
plot(data_clean$Trinn,         data_clean$TPACK)
plot(data_clean$semester,      data_clean$TPACK)
#plot(data_clean$oslomet_r,         data_clean$TPACK_v23)
#plot(data_clean$innlandet_r,       data_clean$TPACK_v23)
#plot(data_clean$HVL_r,             data_clean$TPACK_v23)
#plot(data_clean$UIT_r,             data_clean$TPACK_v23)

plot(data_clean$TK,            data_clean$TPACK)
#plot(data_clean$TK[oslomet_r],data_clean$TPACK[oslomet_r]) # gr.A
#plot(data_clean$TK[innlandet_r],data_clean$TPACK[innlandet_r]) # gr.B
#plot(data_clean$TK[HVL_r],data_clean$TPACK[HVL_r]) # gr.A
#plot(data_clean$TK[UIT_r],data_clean$TPACK[UIT_r]) # gr.B

#scatterplot(TPACK_regdata, c("TK", "TPACK"), factor(studiested))
#scatterplot(TPACK_regdata, c("TK", "TPACK"), factor(semester))

plot(data_clean$PCK,           data_clean$TPACK)
```

```{r multicollinearity, eval=F}
#cor(TK_v23, studiested_r)
#cor(PCK_v23, studiested_r)
cor(data_clean$PCK,   data_clean$semester)
#cor(TPACK_v23, studiested_r)
cor(data_clean$TPACK, data_clean$Gender)

cor(data_clean$TK,    data_clean$Gender)
cor(data_clean$TK,    data_clean$Trinn)

cor(data_clean$TK,    data_clean$PCK)
cor(data_clean$TK,    data_clean$TPACK)
cor(data_clean$PCK,   data_clean$TPACK)
```

TK korrelererer veldig sterkt (0.84) med TPACK, noe som indikerer kollinearitet. (Jeong and Jung, 2016)

## Hierarchical regression analysis

Level 1: students  
Level 2: institution

We are using cluster corrected standard errors to correct for clustering on institution.

Add student covariates

```{r level1}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(scale(TPACK, center=T, scale=T) ~ factor(Gender) * factor(Trinn) * factor(semester), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r level2}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m2 <- lm_robust(scale(TPACK) ~ factor(Gender) * factor(Trinn) * factor(semester) * factor(studiested), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

Calculate F-stat to see if they are significantly different

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$TPACK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r PCK}
# PCK
m3 <- lm_robust(TPACK ~ factor(Gender) * factor(Trinn) * factor(semester) * factor(studiested) + PCK, 
                clusters = factor(studiested), data=data_clean)
summary(m3)

fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r TK}
# TK
m4 <- lm_robust(TPACK ~ factor(Gender) * factor(Trinn) * factor(semester) * factor(studiested) + PCK + TK, 
                clusters = factor(studiested), data=data_clean)
summary(m4)

fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

### 2024-02-15 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
Fixed analysis.

@@
 #gender_r[is.na(gender_r)] <- 1
+
+# merge 2. and 3. year (Naturfag 1)
+#data_clean$semester[data_clean$semester == "2. år"] <- "3. år"
 ```
@@
 
-```{r}
+```{r TK_gender_trinn}
 # ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
@@
 
-```{r}
+Different signs on slope, a clear indication of interaction effects on TK from gender and target grade level.
+
+### Level 1
+
+```{r TPACK_gender_trinn}
+# https://www.statology.org/interaction-plot-r/
+interaction.plot(x.factor = factor(data_clean$Gender),
+                 trace.factor = factor(data_clean$Trinn),
+                 response = data_clean$TPACK,
+                 type = "b",
+                 ylab = "TPACK",
+                 xlab = "Gender",
+                 col = viridis(2), # c("pink", "blue"),
+                 trace.label = "Target grade level")
+```
+
+Not so clear interaction effect, similar slopes? They do cross eventually.
+
+```{r TPACK_gender_semester}
+# https://www.statology.org/interaction-plot-r/
+interaction.plot(x.factor = factor(data_clean$Gender),
+                 trace.factor = factor(data_clean$semester),
+                 response = data_clean$TPACK,
+                 type = "b",
+                 ylab = "TPACK",
+                 xlab = "Gender",
+                 col = viridis(2), # c("pink", "blue"),
+                 trace.label = "Study year")
+```
+
+```{r TPACK_trinn_semester}
 # https://www.statology.org/interaction-plot-r/
 interaction.plot(x.factor = factor(data_clean$Trinn),
-                 trace.factor = factor(data_clean$Gender),
+                 trace.factor = factor(data_clean$semester),
                  response = data_clean$TPACK,
+                 type = "b",
                  ylab = "TPACK",
@@
                  col = viridis(2), # c("pink", "blue"),
-                 trace.label = "Gender")
+                 trace.label = "Study year")
 ```
 
-```{r}
+### Level 2
+
+```{r TPACK_semester_institute}
 interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
@@
                  response = data_clean$TPACK,
+                 type = "b",
                  ylab = "TPACK",
@@
 
-```{r}
+A clear interaction effect between institute and semester on TPACK.
+
+```{r TPACK_trinn_institute}
+interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
+                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
+                 response = data_clean$TPACK,
+                 type = "b",
+                 ylab = "TPACK",
+                 xlab = "Year of study",
+                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
+                 trace.label = "Institute")
+```
+
+```{r TPACK_gender_institute}
 interaction.plot(x.factor = factor(data_clean$Gender),
@@
 ```
+
+Not so clear interaction on TPACK from gender and institute.
 
@@
 
+### Level 1
+
 ```{r level1}
@@
 # sjekk: er det riktig å skalere?
-m1 <- lm_robust(scale(TPACK, center=T, scale=T) ~ factor(Gender) * factor(Trinn) * factor(semester), 
+m1 <- lm_robust(scale(TPACK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
                 clusters = factor(studiested), data=data_clean)
@@
 
+### Level 2
+
 ```{r level2}
 # studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
-m2 <- lm_robust(scale(TPACK) ~ factor(Gender) * factor(Trinn) * factor(semester) * factor(studiested), 
+m2 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+                               factor(studiested) * (factor(semester) + factor(Gender)),    # level 2
                 clusters = factor(studiested), data=data_clean)
@@
 
+### PCK
+
 ```{r PCK}
 # PCK
-m3 <- lm_robust(TPACK ~ factor(Gender) * factor(Trinn) * factor(semester) * factor(studiested) + PCK, 
+m3 <- lm_robust(TPACK ~ factor(Gender) * (factor(Trinn) + factor(semester)) +         # level 1
+                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
+                        PCK, 
                 clusters = factor(studiested), data=data_clean)
@@
 
+### TK
+
 ```{r TK}
 # TK
-m4 <- lm_robust(TPACK ~ factor(Gender) * factor(Trinn) * factor(semester) * factor(studiested) + PCK + TK, 
+m4 <- lm_robust(TPACK ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+                        factor(studiested) * (factor(semester) + factor(Gender))     # level 2
+                        + PCK + TK, 
                 clusters = factor(studiested), data=data_clean)

### 2024-02-15 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
Checked interaction effects more closely

@@
 
+### TK Level 1
+
 ```{r TK_gender_trinn}
@@
 # https://www.statology.org/interaction-plot-r/
-interaction.plot(x.factor = factor(data_clean$Trinn),
-                 trace.factor = factor(data_clean$Gender),
+interaction.plot(x.factor = factor(data_clean$Gender),
+                 trace.factor = factor(data_clean$Trinn),
                  response = data_clean$TK,
+                 type = "b",
                  ylab = "TK",
-                 xlab = "Target grade level",
+                 xlab = "Gender",
                  col = viridis(2), # c("pink", "blue"), 
-                 trace.label = "Gender")
+                 trace.label = "Target grade level")
 ```
@@
 
-### Level 1
+```{r TK_gender_semester}
+# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
+# https://www.statology.org/interaction-plot-r/
+interaction.plot(x.factor = factor(data_clean$Gender),
+                 trace.factor = factor(data_clean$semester),
+                 response = data_clean$TK,
+                 type = "b",
+                 ylab = "TK",
+                 xlab = "Gender",
+                 col = viridis(2), # c("pink", "blue"), 
+                 trace.label = "Year of study")
+```
+
+```{r TK_semester_trinn}
+# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
+# https://www.statology.org/interaction-plot-r/
+interaction.plot(x.factor = factor(data_clean$Trinn),
+                 trace.factor = factor(data_clean$semester),
+                 response = data_clean$TK,
+                 type = "b",
+                 ylab = "TK",
+                 xlab = "Target grade level",
+                 col = viridis(2), # c("pink", "blue"), 
+                 trace.label = "Year of study")
+```
+
+### TK Level 2
+
+```{r TK_semester_institute}
+interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
+                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
+                 response = data_clean$TK,
+                 type = "b",
+                 ylab = "TK",
+                 xlab = "Year of study",
+                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
+                 trace.label = "Institute")
+```
+
+A clear interaction effect between institute and semester on TPACK.
+
+```{r TK_trinn_institute}
+interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
+                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
+                 response = data_clean$TK,
+                 type = "b",
+                 ylab = "TK",
+                 xlab = "Year of study",
+                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
+                 trace.label = "Institute")
+```
+
+```{r TK_gender_institute}
+interaction.plot(x.factor = factor(data_clean$Gender),
+                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
+                 response = data_clean$TK,
+                 type = "b",
+                 ylab = "TK",
+                 xlab = "Gender",
+                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
+                 trace.label = "Institute")
+```
+
+### PCK Level 1
+
+```{r PCK_gender_trinn}
+# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
+# https://www.statology.org/interaction-plot-r/
+interaction.plot(x.factor = factor(data_clean$Gender),
+                 trace.factor = factor(data_clean$Trinn),
+                 response = data_clean$PCK,
+                 type = "b",
+                 ylab = "PCK",
+                 xlab = "Gender",
+                 col = viridis(2), # c("pink", "blue"), 
+                 trace.label = "Target grade level")
+```
+
+Different signs on slope, a clear indication of interaction effects on PCK from gender and target grade level.
+
+```{r PCK_gender_semester}
+# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
+# https://www.statology.org/interaction-plot-r/
+interaction.plot(x.factor = factor(data_clean$Gender),
+                 trace.factor = factor(data_clean$semester),
+                 response = data_clean$PCK,
+                 type = "b",
+                 ylab = "PCK",
+                 xlab = "Gender",
+                 col = viridis(2), # c("pink", "blue"), 
+                 trace.label = "Year of study")
+```
+
+```{r PCK_semester_trinn}
+# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
+# https://www.statology.org/interaction-plot-r/
+interaction.plot(x.factor = factor(data_clean$Trinn),
+                 trace.factor = factor(data_clean$semester),
+                 response = data_clean$PCK,
+                 type = "b",
+                 ylab = "PCK",
+                 xlab = "Target grade level",
+                 col = viridis(2), # c("pink", "blue"), 
+                 trace.label = "Year of study")
+```
+
+### PCK Level 2
+
+```{r PCK_semester_institute}
+interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
+                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
+                 response = data_clean$PCK,
+                 type = "b",
+                 ylab = "PCK",
+                 xlab = "Year of study",
+                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
+                 trace.label = "Institute")
+```
+
+A clear interaction effect between institute and semester on TPACK.
+
+```{r PCK_trinn_institute}
+interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
+                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
+                 response = data_clean$PCK,
+                 type = "b",
+                 ylab = "PCK",
+                 xlab = "Year of study",
+                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
+                 trace.label = "Institute")
+```
+
+```{r PCK_gender_institute}
+interaction.plot(x.factor = factor(data_clean$Gender),
+                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
+                 response = data_clean$PCK,
+                 type = "b",
+                 ylab = "PCK",
+                 xlab = "Gender",
+                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
+                 trace.label = "Institute")
+```
+
+### TPACK Level 1
 
@@
 
-### Level 2
+### TPACK Level 2
 
@@
 
-## Hierarchical regression analysis
+## Hierarchical RA for TK
 
@@
 
-```{r level1}
+```{r TKlevel1_main}
 # kovariater; we use scale to standardize when using interaction effects
 # sjekk: er det riktig å skalere?
-m1 <- lm_robust(scale(TPACK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
+m1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                 clusters = factor(studiested), data=data_clean)
@@
 
+```{r TKlevel1_interaction}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+m2 <- lm_robust(scale(TK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
+                clusters = factor(studiested), data=data_clean)
+summary(m2)
+```
+
+Calculate F-stat to see if they are significantly different
+
+```{r}
+# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
+# source: p. 257 i Howitt & Cramer, 2014
+fstat <- m2$adj.r.squared / m1$adj.r.squared
+df1 <- length(data_clean$TK) - 1
+df2 <- df1
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
 ### Level 2
 
-```{r level2}
+```{r TKlevel2_main}
 # studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
-m2 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
-                               factor(studiested) * (factor(semester) + factor(Gender)),    # level 2
+m3 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
+                clusters = factor(studiested), data=data_clean)
+summary(m3)
+```
+
+```{r}
+# main
+fstat <- m3$adj.r.squared / m1$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+
+# interaction
+fstat <- m3$adj.r.squared / m2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+```{r TKlevel2_interaction}
+# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
+m4 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
+                clusters = factor(studiested), data=data_clean)
+summary(m4)
+```
+
+```{r}
+# interaction
+fstat <- m4$adj.r.squared / m3$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+## Hierarchical RA for PCK
+
+Level 1: students  
+Level 2: institution
+
+We are using cluster corrected standard errors to correct for clustering on institution.
+
+Add student covariates
+
+### Level 1
+
+```{r PCKlevel1_main}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+m1 <- lm_robust(PCK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
+                clusters = factor(studiested), data=data_clean)
+summary(m1)
+```
+
+```{r PCKlevel1_interaction}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+m2 <- lm_robust(scale(PCK, center=T, scale=T) ~ factor(semester) * (factor(Gender) + factor(Trinn)), 
+                clusters = factor(studiested), data=data_clean)
+summary(m2)
+```
+
+Calculate F-stat to see if they are significantly different
+
+```{r}
+# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
+# source: p. 257 i Howitt & Cramer, 2014
+fstat <- m2$adj.r.squared / m1$adj.r.squared
+df1 <- length(data_clean$PCK) - 1
+df2 <- df1
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+### Level 2
+
+```{r PCKlevel2_main}
+# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
+m3 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
+                clusters = factor(studiested), data=data_clean)
+summary(m3)
+```
+
+```{r}
+# main
+fstat <- m3$adj.r.squared / m1$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+
+# interaction
+fstat <- m3$adj.r.squared / m2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+```{r PCKlevel2_interaction}
+# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
+m4 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
+                clusters = factor(studiested), data=data_clean)
+summary(m4)
+```
+
+```{r}
+fstat <- m4$adj.r.squared / m3$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+## Hierarchical RA for TPACK
+
+Level 1: students  
+Level 2: institution
+
+We are using cluster corrected standard errors to correct for clustering on institution.
+
+Add student covariates
+
+### Level 1
+
+```{r level1_main}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+m1 <- lm_robust(TPACK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
+                clusters = factor(studiested), data=data_clean)
+summary(m1)
+```
+
+```{r level1_interaction}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+m2 <- lm_robust(scale(TPACK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
                 clusters = factor(studiested), data=data_clean)
@@
 
+### Level 2
+
+```{r level2_main}
+# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
+m3 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+                               factor(studiested) + (factor(semester) + factor(Gender)),    # level 2
+                clusters = factor(studiested), data=data_clean)
+summary(m3)
+
+#main
+fstat <- m3$adj.r.squared / m1$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+
+# interaction
+fstat <- m3$adj.r.squared / m2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+```{r level2_interaction}
+# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
+m4 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+                               factor(studiested) * (factor(semester) + factor(Gender)),    # level 2
+                clusters = factor(studiested), data=data_clean)
+summary(m4)
+
+fstat <- m4$adj.r.squared / m3$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
 ### PCK
@@
 # PCK
-m3 <- lm_robust(TPACK ~ factor(Gender) * (factor(Trinn) + factor(semester)) +         # level 1
+m5 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +         # level 1
                         factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
-                        PCK, 
+                        scale(PCK), 
                 clusters = factor(studiested), data=data_clean)
-summary(m3)
+summary(m5)
 
-fstat <- m3$adj.r.squared / m2$adj.r.squared
+# main
+fstat <- m5$adj.r.squared / m3$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+
+# interaction
+fstat <- m5$adj.r.squared / m4$adj.r.squared
 pf(fstat, df1, df2, lower.tail = FALSE)
@@
 # TK
-m4 <- lm_robust(TPACK ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
-                        factor(studiested) * (factor(semester) + factor(Gender))     # level 2
-                        + PCK + TK, 
+m6 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
+                        scale(PCK) + scale(TK), 
                 clusters = factor(studiested), data=data_clean)
-summary(m4)
+summary(m6)
 
-fstat <- m4$adj.r.squared / m3$adj.r.squared
+fstat <- m6$adj.r.squared / m5$adj.r.squared
 pf(fstat, df1, df2, lower.tail = FALSE)

### 2024-02-20 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
Added barplot

@@
 author: "Niklas Karlsen"
-date: "2024-02-15"
+date: "2024-02-20"
 output: 
@@
 ```{r load_libraries, echo=FALSE, warning=FALSE, message=FALSE}
-library(pander)                             # prettyprint for tables
-
-library(psych)                              # for ICC, SD
+#library(pander)                             # prettyprint for tables
+library(plyr)
+library(ggplot2)
+library(viridis)
 
 library(mice)                               # for md.pattern (missing)
-library(viridis)
+library(psych)                              # for ICC, SD
 
@@
 library(estimatr)                           # for cluster corrected SE
-library(sandwich)
-library(lmtest)
+#library(sandwich)
+#library(lmtest)
 ```
@@
 }
+
+###
+# helper functions from http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/
+###
+
+## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
+##   data: a data frame.
+##   measurevar: the name of a column that contains the variable to be summariezed
+##   groupvars: a vector containing names of columns that contain grouping variables
+##   na.rm: a boolean that indicates whether to ignore NA's
+##   conf.interval: the percent range of the confidence interval (default is 95%)
+summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
+                      conf.interval=.95, .drop=TRUE) {
+  #library(plyr)
+
+  # New version of length which can handle NA's: if na.rm==T, don't count them
+  length2 <- function (x, na.rm=FALSE) {
+    if (na.rm) sum(!is.na(x))
+    else       length(x)
+  }
+
+  # This does the summary. For each group's data frame, return a vector with
+  # N, mean, and sd
+  datac <- ddply(data, groupvars, .drop=.drop,
+                 .fun = function(xx, col) {
+                   c(N    = length2(xx[[col]], na.rm=na.rm),
+                     mean = mean   (xx[[col]], na.rm=na.rm),
+                     sd   = sd     (xx[[col]], na.rm=na.rm)
+                   )
+                 },
+                 measurevar
+  )
+
+  # Rename the "mean" column  ; from plyr package
+  datac <- plyr::rename(datac, c("mean" = measurevar))
+  #datac <- rename(datac, measurevar = "mean") # oppdatert til ny versjon av dplyr
+
+  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
+
+  # Confidence interval multiplier for standard error
+  # Calculate t-statistic for confidence interval: 
+  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
+  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
+  datac$ci <- datac$se * ciMult
+
+  return(datac)
+}
+
+## Norms the data within specified grupper in a data frame; it normalizes each
+## subject (identified by idvar) so that they have the same mean, within each group
+## specified by betweenvars.
+##   data: a data frame.
+##   idvar: the name of a column that identifies each subject (or matched subjects)
+##   measurevar: the name of a column that contains the variable to be summariezed
+##   betweenvars: a vector containing names of columns that are between-subjects variables
+##   na.rm: a boolean that indicates whether to ignore NA's
+normDataWithin <- function(data=NULL, idvar, measurevar, betweenvars=NULL,
+                           na.rm=FALSE, .drop=TRUE) {
+  #library(plyr)
+
+  # Measure var on left, idvar + between vars on right of formula.
+  data.subjMean <- ddply(data, c(idvar, betweenvars), .drop=.drop,
+                         .fun = function(xx, col, na.rm) {
+                           c(subjMean = mean(xx[,col], na.rm=na.rm))
+                         },
+                         measurevar,
+                         na.rm
+  )
+
+  # Put the subject means with original data
+  data <- merge(data, data.subjMean)
+
+  # Get the normalized data in a new column
+  measureNormedVar <- paste(measurevar, "_norm", sep="")
+  data[,measureNormedVar] <- data[,measurevar] - data[,"subjMean"] +
+    mean(data[,measurevar], na.rm=na.rm)
+
+  # Remove this subject mean column
+  data$subjMean <- NULL
+
+  return(data)
+}
+
 ```
@@
 data_clean$TPACK <- TPACK_v23
+```
+
+## Bargraphs: comparing institutes
+
+```{r}
+# group institutes into three: low, medium, high
+data_clean$studiested2 <- factor(data_clean$studiested, 
+                                 level=c("HVL","UiT","Nord universitet","Høgskolen i Innlandet","OsloMet"),
+                                 label=c("A", "A", "B", "C", "C"))
+
+dc1 <- summarySE(data_clean, measurevar="TK", groupvars="studiested2")
+dc2 <- summarySE(data_clean, measurevar="PCK", groupvars="studiested2")
+dc3 <- summarySE(data_clean, measurevar="TPACK", groupvars="studiested2")
+
+# rename; kilde: https://www.statology.org/rename-single-column-in-r/
+colnames(dc1)[colnames(dc1) == "TK"] <- "var"
+colnames(dc2)[colnames(dc2) == "PCK"] <- "var"
+colnames(dc3)[colnames(dc3) == "TPACK"] <- "var"
+
+# merge data
+dc <- rbind(dc1,dc2)
+dc <- rbind(dc,dc3)
+
+# add column med navn
+dc["TPACK"] <- c("TK","TK","TK","PCK","PCK","PCK","TPACK","TPACK","TPACK")
+```
+
+```{r}
+ggplot(dc, aes(x=studiested2, y=var, fill=studiested2)) + 
+  geom_bar(stat="identity", position="dodge") +
+  #scale_fill_manual(values=viridis(3)) + 
+  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
+                  linewidth=.3,    # Thinner lines
+                  width=.2,
+                  position=position_dodge(.9)) +
+  ggtitle("PCK, TK og TPACK fordelt på studiested") +
+  guides(fill="none") +
+  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
+  coord_cartesian(ylim = c(1,5)) +
+  ylab("verdi") +
+  xlab("studiested gruppe") +
+  theme_bw() +
+  facet_wrap(~TPACK)
 ```

### 2024-04-02 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
Lagre bilde til fil

@@
 author: "Niklas Karlsen"
-date: "2024-02-20"
+date: "2024-04-02"
 output: 
@@
 # https://stackoverflow.com/questions/16347731/how-to-change-the-locale-of-r
+
+plot_to_file = 1
 ```
@@
 
+Flytt origo
+
+```{r}
+data_trelis_q_ren[,unlist(TPACK_PST_list)] <- data_trelis_q_ren[,unlist(TPACK_PST_list)] - 3 
+```
+
 OBS! Anonymiser data i datafil på studiested før det legges ut åpent på nett
@@
 ```{r}
-ggplot(dc, aes(x=studiested2, y=var, fill=studiested2)) + 
+stolpediagram <- ggplot(dc, aes(x=studiested2, y=var, fill=studiested2)) + 
   geom_bar(stat="identity", position="dodge") +
@@
   # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
-  coord_cartesian(ylim = c(1,5)) +
+#  coord_cartesian(ylim = c(1,5)) +
   ylab("verdi") +
@@
   facet_wrap(~TPACK)
+
+print(stolpediagram) # vis plot
+```
+
+```{r}
+# save plot
+if (plot_to_file) {
+  ggsave("../output/TPACK_stolpediagram.png") #,height=8,width=15)
+  }
+
+```
+
+```{r}
+ggplot(dc, aes(x=TPACK, y=var, fill=studiested2)) + 
+  geom_bar(stat="identity", position="dodge") +
+  #scale_fill_manual(values=viridis(3)) + 
+  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
+                  linewidth=.3,    # Thinner lines
+                  width=.2,
+                  position=position_dodge(.9)) +
+  ggtitle("PCK, TK og TPACK fordelt på studiested") +
+  guides(fill="none") +
+  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
+#  coord_cartesian(ylim = c(1,5)) +
+  ylab("verdi") +
+  xlab("") +
+  theme_bw() +
+  facet_wrap(~studiested2)
 ```

### 2024-05-27 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
Added NFSUN-plot

@@
 ---
+
+Changelog:
+
+* 16.5.2024: added plot for NFSUN2024
 
@@
 
+```
+
+For NFSUN2024:
+
+```{r}
+stolpediagram <- ggplot(dc, aes(x=studiested2, y=var, fill=studiested2)) + 
+  geom_bar(stat="identity", position="dodge") +
+  #scale_fill_manual(values=viridis(3)) + 
+  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
+                  linewidth=.3,    # Thinner lines
+                  width=.2,
+                  position=position_dodge(.9)) +
+  ggtitle("Mean PCK, TK and TPACK for science teacher institutes") +
+  guides(fill="none") +
+  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
+#  coord_cartesian(ylim = c(1,5)) +
+  ylab("Value") +
+  xlab("Institute group") +
+  theme_bw() +
+  facet_wrap(~TPACK)
+
+print(stolpediagram) # vis plot
 ```

### 2024-05-28 Rmd/Paper1.Rmd
Initial commit

---
title: "Instrument development"
author: "Niklas Karlsen, e-post: ..."
date: "2024-05-28"
bibliography: "../data/meta_data/references.bib"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/Paper1.html')) })
---

TODO

* gjøre om referanser til APA7

# Introduction

The main purpose of this Rmd-script is to reproduce the results in paper 1. The whole process from reducing the raw data to doing the EFA and CFA are discussed.

This R-markdown describes the analyses used to support the validity and reliability of the inferences based on the measure developed in the thesis. This includes exploratory factor analysis (EFA) and the confirmatory factor analysis (CFA) that was done to get the results in paper 1.

The development of a measure usually follows three stages (@boateng_best_2018):

1. Item development
2. Scale development
3. Scale evaluation

The first stage is described in Paper 1. This computational essay (@odden_using_2023) describes the analysis behind the results in paper 1, and emphasises the quantitative analyses done in the final two stages. This concerns item reduction and extraction of factors (EFA) in stage two, and test of dimensionality (CFA) and tests of reliability and validity in stage three.

From stage 2 Boateng et al. (2013) describes three steps which will be discussed here: step 4 (gather data), step 5 (item reduction) and step 6 (factor extraction, i.e., EFA). From stage 3 there are three steps which will be discussed: step 7 (tests of dimensionality, i.e., CFA), step 8 (tests of reliability) and step 9 (tests of validity).

# Stage 2

## Step 4: Data gathering

!NB Anonymisere raw data (blant annet studiested) før jeg legger det ut offentlig.

This step concerns gathering data for the EFA and the CFA, using separate samples for each analysis. The recommended sample size is 10 respondents per item and/or 200-300 observations in total (Boateng et al., 2018). The sample size is discussed more closely in the thesis (ref sec page).

Data were collected for two pilots (EFA) and the CFA. The three datasets were reduced individually.

Reducing the raw data ...

## Step 5: Item reduction

The aim of this step is to reduce the number of items to the lowest necessary by removing redundant items based on individual item tests. Since our measure is a self-report questionnaire, the relevant tests are based on correlations (step 5.2 and 5.3) and consideration of missing data (step 5.5). Step 5.1 and 5.4 are omitted as they concern correct or wrong answers to tests, which is not relevant for a self-report measure.

Factor analysis is based on linear analysis (i.e., ordinary least squares-regresion (OLS)). Assumptions underlying OLS-regresion are according to @christophersen_introduksjon_2018 that the residuals have both an average value of 0, constant variance (i.e,. homoscedasticity) and independence (i.e., no autocorrelation). In addition, the dependent variables should depend linearly on the independent variables, which should have no multicollinearity (i.e., meaning that no residuals(?) of independent variable should be strongly correlated with any of the other independent variables).

In addition, according to @flora_old_2012, it is important to decide whether to treat the data as ordinal or interval, as this has consequences for how the items are screened. Treating the data as interval, requires checking of missing data, linearity (checking for outliers and colinearity, i.e., very strong correlation between variables), and normally distributed residuals, checks which are not necessary to do if the data are treated as ordinal.

### Ordinal or interval data?

When we have as few as five answer alternatives (i.e., 5-pt Likert scale), there are weighty arguments for considering the data as ordinal (Flora et al., 2012; @izquierdo_exploratory_2014; @watkins_exploratory_2018). Five answer alternatives do however begin to imitate an interval scale (i.e., a continous scale). Thus, researchers have also argued for considering 5-pt scales as being continous (Christophersen, 2018; @norman_likert_2010). In order to use the full arsenal of parametric methods (i.e., the properties of the sample are decided by a few parameters), such as maximum likelihood (ML), we decided to consider the data as interval. To support this choice further, we also presented the five alternatives as points on a linear scale in our questionnaire to help the respondents visualise the alternatives as being on an interval scale.

### Outliers?

Outliers may impact calculations of standard deviations. Flora et al. (2012) and Watkins (2018) recommend, however, not to remove outliers. Christophersen (2018) recommends to compare calculations with and without outliers to determine their significance.

### Normal distribution of residuals?

Statistical tests assume the residuals are normally distributed. Common criteria are that skewness should be below 2 or 3, and that curtosis (i.e., flatness, which is related to outliers) should be below 7 or 10 (Flora et al., 2012; Christophersen, 2018).

### Missing?

Missing data reduces the size of the dataset used for analysis, which leads to lower power (i.e., larger chances for Type II-error, that is not detecting significant effects, obtaining false negatives). There are different advice for how to handle missing data, depending on whether they are missing at random (MAR), completely at random (MCAR) or not at random (NMAR) (@rubin_inference_1975; @schafer_missing_2002). Data that are missing at random are correlated with the questionnaire (i.e., a question may be missing from a specific version of a questionnaire, which is in a sense random). Data that are missing completely at random, do however not correlate with anything obvious. They are in a sense, seemingly truly random, with no clear rationale for their missing other than circumstance. Data missing not at random are questions that correlate with the question itself (i.e., several respondents may avoid a specific question, i.e., about gender, there seems to be a rationale, that is not a random cause, for data missing for this specific item).

Ways to handle missing data is predictive mean matching when less than 10 % are missing, otherwise ML-methods (@weston_brief_2006; Flora et al., 2012; Watkins, 2018; @mcneish_challenging_2017; @gallagher_introduction_2013).

# Session info
<!-- # ref.: https://intro2r.com/proj_doc.html -->
<!-- #sessionInfo() -->

```{r}
xfun::session_info()
```

# References

<!-- citation(package) -->

### 2024-05-28 Rmd/Paper1.Rmd
Gjorde om til apa7 ref.: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html

@@
 bibliography: "../data/meta_data/references.bib"
+csl: "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl"
 output: 
@@
 
-TODO
-
-* gjøre om referanser til APA7
-
 # Introduction
@@
 
-The development of a measure usually follows three stages (@boateng_best_2018):
+The development of a measure usually follows three stages [@boateng_best_2018]:
 
@@
 The first stage is described in Paper 1.
-This computational essay (@odden_using_2023) describes the analysis behind the results in paper 1, and emphasises the quantitative analyses done in the final two stages.
+This computational essay [@odden_using_2023] describes the analysis behind the results in paper 1, and emphasises the quantitative analyses done in the final two stages.
 This concerns item reduction and extraction of factors (EFA) in stage two, and test of dimensionality (CFA) and tests of reliability and validity in stage three.
@@
 
-When we have as few as five answer alternatives (i.e., 5-pt Likert scale), there are weighty arguments for considering the data as ordinal (Flora et al., 2012; @izquierdo_exploratory_2014; @watkins_exploratory_2018).
+When we have as few as five answer alternatives (i.e., 5-pt Likert scale), there are weighty arguments for considering the data as ordinal [@flora_old_2012; @izquierdo_exploratory_2014; @watkins_exploratory_2018].
 Five answer alternatives do however begin to imitate an interval scale (i.e., a continous scale).
-Thus, researchers have also argued for considering 5-pt scales as being continous (Christophersen, 2018; @norman_likert_2010).
+Thus, researchers have also argued for considering 5-pt scales as being continous [@christophersen_introduksjon_2018; @norman_likert_2010].
 In order to use the full arsenal of parametric methods (i.e., the properties of the sample are decided by a few parameters), such as maximum likelihood (ML), we decided to consider the data as interval.
@@
 Missing data reduces the size of the dataset used for analysis, which leads to lower power (i.e., larger chances for Type II-error, that is not detecting significant effects, obtaining false negatives).
-There are different advice for how to handle missing data, depending on whether they are missing at random (MAR), completely at random (MCAR) or not at random (NMAR) (@rubin_inference_1975; @schafer_missing_2002).
+There are different advice for how to handle missing data, depending on whether they are missing at random (MAR), completely at random (MCAR) or not at random (NMAR) [@rubin_inference_1975; @schafer_missing_2002].
 Data that are missing at random are correlated with the questionnaire (i.e., a question may be missing from a specific version of a questionnaire, which is in a sense random).
@@
 
-Ways to handle missing data is predictive mean matching when less than 10 % are missing, otherwise ML-methods (@weston_brief_2006; Flora et al., 2012; Watkins, 2018; @mcneish_challenging_2017; @gallagher_introduction_2013).
+Ways to handle missing data is predictive mean matching when less than 10 % are missing, otherwise ML-methods [@weston_brief_2006; @flora_old_2012; @watkins_exploratory_2018; @mcneish_challenging_2017; @gallagher_introduction_2013].
 

### 2024-05-28 Rmd/Paper1.Rmd
Improved structure of document

@@
 ---
-title: "Instrument development"
-author: "Niklas Karlsen, e-post: ..."
+title: "Paper 1: Instrument development"
+author: "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
 date: "2024-05-28"
 bibliography: "../data/meta_data/references.bib"
-csl: "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl"
+csl: "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl" # (source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html)
 output: 
@@
     toc_float: true
+    number_sections: true
 
@@
 
-The main purpose of this Rmd-script is to reproduce the results in paper 1. The whole process from reducing the raw data to doing the EFA and CFA are discussed.
+The main purpose of this Rmd-script is to reproduce the results in paper 1.
+The whole process from reducing the raw data to doing the EFA and CFA is discussed.
+The document is written as a computational essay [@odden_using_2023].
 
-This R-markdown describes the analyses used to support the validity and reliability of the inferences based on the measure developed in the thesis.
+This R-markdown describes the analyses used to support the validity and reliability of the inferences about PSTs' TPACK based on the measure developed in the thesis.
 This includes exploratory factor analysis (EFA) and the confirmatory factor analysis (CFA) that was done to get the results in paper 1.
@@
 The first stage is described in Paper 1.
-This computational essay [@odden_using_2023] describes the analysis behind the results in paper 1, and emphasises the quantitative analyses done in the final two stages.
+This computational essay describes the analysis behind the results in paper 1, and emphasises the quantitative analyses done in the final two stages.
 This concerns item reduction and extraction of factors (EFA) in stage two, and test of dimensionality (CFA) and tests of reliability and validity in stage three.
@@
 
+## Imports
+
+```{r}
+# libraries
+```
+
 # Stage 2
 
-## Step 4: Data gathering
-
-!NB Anonymisere raw data (blant annet studiested) før jeg legger det ut offentlig.
+## Step 4: Data gathering / cleaning
 
@@
 
-Reducing the raw data ...
+Reducing the raw data ... (transforming from ods to Rdata)
+
+* Step 4a: convert from ods to Rdata; rename items using TPACK typology; treat missing
 
@@
 
-Factor analysis is based on linear analysis (i.e., ordinary least squares-regresion (OLS)).
-Assumptions underlying OLS-regresion are according to @christophersen_introduksjon_2018 that the residuals have both an average value of 0, constant variance (i.e,.
+Factor analysis is based on linear analysis (i.e., ordinary least squares-regression (OLS)).
+Assumptions underlying OLS-regression are according to @christophersen_introduksjon_2018 that the residuals have both an average value of 0, constant variance (i.e,.
 homoscedasticity) and independence (i.e., no autocorrelation).
 In addition, the dependent variables should depend linearly on the independent variables, which should have no multicollinearity (i.e., meaning that no residuals(?) of independent variable should be strongly correlated with any of the other independent variables).
+In addition, according to @flora_old_2012, it is important to decide whether to treat the data as ordinal or interval, as this has consequences for how the items are screened.
+Treating the data as interval, requires checking of missing data, linearity (checking for outliers and collinearity, i.e., very strong correlation between variables), and normally distributed residuals, checks which are not necessary to do if the data are treated as ordinal.
 
-In addition, according to @flora_old_2012, it is important to decide whether to treat the data as ordinal or interval, as this has consequences for how the items are screened.
-Treating the data as interval, requires checking of missing data, linearity (checking for outliers and colinearity, i.e., very strong correlation between variables), and normally distributed residuals, checks which are not necessary to do if the data are treated as ordinal.
+* Step 5a: check residuals (normality, autocorrelation)
+* Step 5b: check correlations (collinearity, multicollinearity)
+* Step 5c: check outliers (skew, kurtosis, range)
+* Step 5d: check missing (pattern: MCAR, MAR, NMAR)
+* Step 5e: check eigenvalues, item-test correlation, reliability [@izquierdo_exploratory_2014]
 
@@
 Five answer alternatives do however begin to imitate an interval scale (i.e., a continous scale).
-Thus, researchers have also argued for considering 5-pt scales as being continous [@christophersen_introduksjon_2018; @norman_likert_2010].
+Thus, researchers have also argued for considering 5-pt scales as being continuous [@christophersen_introduksjon_2018; @norman_likert_2010].
 In order to use the full arsenal of parametric methods (i.e., the properties of the sample are decided by a few parameters), such as maximum likelihood (ML), we decided to consider the data as interval.
@@
 
+## Step 6: Factor extraction (i.e., EFA)
+
+# Stage 3
+
+## Step 7: Tests of dimensionality (i.e., CFA)
+
+## Step 8: Tests of reliability
+
+## Step 9: Tests of validity
+
+# Discussion and conclusion
+
+# References
+
+<!-- citation(package) -->
+
+<div id="refs"></div>
+
+<!-- source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html
+
+# (APPENDIX) Appendix {-} -->
+
 # Session info
@@
 ```
-
-# References
-
-<!-- citation(package) -->

### 2024-08-20 Rmd/Paper1.Rmd
Added comments

@@
 Data were collected for two pilots (EFA) and the CFA. The three datasets were reduced individually.
+
+The data for the first pilot were collected using almost the same questionnaire, but four of the questions were slightly modified to use the term "coding" instead of "programming".
+"Programming" instead of coding is used in the final version, so this is what we choose also when reducing the dataset.
+
+We used data from respondents who did not explicitly deny this.
+A few had answered blank, which we interpreted as "yes", although strictly speaking this should probably be interpreted as "no" to be on the safe side.
+The data are anyways anonymous.
 

### 2024-08-20 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
Added ICC and updated graph

--- 2024-04-02 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
@@
 author: "Niklas Karlsen"
-date: "2024-04-02"
+date: "2024-08-13"
 output: 
@@
 ---
+
+TODO
+
+* endre navn til paper 2
+
+Changelog:
+
+* 13.8.2024: added check of ICC
+* 16.5.2024: added plot for NFSUN2024
 
@@
 #library(lmtest)
+
+library(nlme)                               # for ICC - variance HLM
 ```
@@
 
+For NFSUN2024:
+
 ```{r}
+# anonymise institutes
+data_clean$studiested3 <- factor(data_clean$studiested, 
+                                 level=c("HVL","UiT","Nord universitet","Høgskolen i Innlandet","OsloMet"),
+                                 label=c("I1", "I2", "I3", "I4", "I5"))
+
+dc1 <- summarySE(data_clean, measurevar="TK", groupvars="studiested3")
+dc2 <- summarySE(data_clean, measurevar="PCK", groupvars="studiested3")
+dc3 <- summarySE(data_clean, measurevar="TPACK", groupvars="studiested3")
+
+# rename; kilde: https://www.statology.org/rename-single-column-in-r/
+colnames(dc1)[colnames(dc1) == "TK"] <- "var"
+colnames(dc2)[colnames(dc2) == "PCK"] <- "var"
+colnames(dc3)[colnames(dc3) == "TPACK"] <- "var"
+
+# merge data
+dc <- rbind(dc1,dc2)
+dc <- rbind(dc,dc3)
+
+# add column med navn
+dc["TPACK"] <- c("TK","TK","TK","TK","TK",
+                 "PCK","PCK","PCK","PCK","PCK",
+                 "TPACK","TPACK","TPACK","TPACK","TPACK")
+```
+
+```{r}
+stolpediagram <- ggplot(dc, aes(x=studiested3, y=var, fill=studiested3)) + 
+  geom_bar(stat="identity", position="dodge") +
+  #scale_fill_manual(values=viridis(3)) + 
+  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
+                  linewidth=.3,    # Thinner lines
+                  width=.2,
+                  position=position_dodge(.9)) +
+  ggtitle("Mean and standard deviation for PCK, TK and TPACK") +
+  guides(fill="none") +
+  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
+#  coord_cartesian(ylim = c(1,5)) +
+  ylab("Value") +
+  xlab("Institute") +
+  theme_bw() +
+  facet_wrap(~TPACK)
+
+print(stolpediagram) # vis plot
+```
+
+```{r}
+# save plot
+if (plot_to_file) {
+  ggsave("../output/NFSUN2024_stolpediagram.png",dpi=300) #height=8,width=15)
+  }
+
+```
+
+```{r eval=F}
 ggplot(dc, aes(x=TPACK, y=var, fill=studiested2)) + 
@@
 
+### ICC
+
+ICC: Christophersen (2018, p. 109-111)
+
+```{r ICC}
+# variance decomposition
+null.mod <- lme(TPACK ~ 1, random =~ 1|studiested, data = data_clean)
+vcn <- VarCorr(null.mod)
+vcn
+```
+
+The variance of indviduals is given by the residual.
+The variance of institutions is givene by the intercept.
+If the variance of the intercept is larger than 0.05 of the total variance (intercept + residual), then the institutional variance contributes to explain the outcome.
+
+### Normality
+
 ```{r normality}
@@
 ```
+
+### Linearity
 

### 2024-08-22 Rmd/Paper4.Rmd
Initial commit

---
title: "Paper 4"
author: "Niklas Karlsen"
date: "2024-08-22"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/paper4.html')) })
---

This is a computational essay?

Changelog

* 22.8.2024: added comparison of two hybrid models

TODO

# Preamble / technical details

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) #,results="asis")

Sys.setenv(LANG = "no_NB.UTF-8")            # for æ,ø og å. Kilde:
Sys.setlocale("LC_ALL","Norwegian")
# https://stackoverflow.com/questions/16347731/how-to-change-the-locale-of-r

# variable
plot_to_file = 0
```

```{r last_bibliotek, echo=FALSE, warning=FALSE, message=FALSE}
## Last bibliotek: rekkefølge - de siste overstyrer de første

## bibliotek for tabell-behandling
# last plyr før dplyr
#library(plyr)                               # for plot of mean and sd
library(dplyr)
library(tidyr)
library(pander)                             # prettyprint for tables
# kilde: https://stackoverflow.com/questions/31626419/in-knitr-no-output-from-pander-in-for-loop
#panderOptions('knitr.auto.asis', FALSE)     # to pander in loop
#library(tidyverse)

## bibliotek for plotting
#library(plotrix)                            # SE(m) (plot mean+CI)
library(ggplot2)
#library(GGally)                             # for plot av ggpairs
# kilde: https://ggplot2-book.org/arranging-plots.html#arranging-plots
#library(patchwork)                          # for inset_element

## bibliotek for statistikk
#library(bruceR,quietly=TRUE)                           # for manova
#library(DescTools)                          # for lambda, categorical data
library(psych)                              # for ICC, SD
library(effsize)                            # for cohen's d
#library(corrplot)
#library(rcompanion)                          # for Cramers V
#library(Rfit)                               # ranked regression # https://journal.r-project.org/archive/2013/RJ-2013-027/RJ-2013-027.pdf

### regression libraries
library(lme4)                                 # for MLM
#library(lmerTest)                             # not recommended? Avoid the lmerTest package (https://utstat.toronto.edu/~brunner/workshops/mixed/NormalWithR.pdf)
library(car)                                   # # for p-values i lme4 (anova, test="F"). durbinWatson
library(nlme)                                 # for lme
#library(rmcorr)
library(plm)                                  # paneldata analysis

# t-tests
library(WRS2)                                 # Yuen-Welch test ref.: Mair & Wilcox, 2020

# SE calculation
library(pcse)                                 # panel-corrected SE
library(sandwich) # cluster corrected error
library(lmtest) # logLikelihood tests, AIC, BIC ,...
#library(MuMIn) # AICc
library(aod) # wald.test

library(texreg)                               # screenreg (print compare results regresion)
library(viridis)

#library(ppcor) # partiell correlation
#library(mice) # missing
```

## Definer noen hjelpefunksjoner.

```{r last_hjelpefunksjoner}
source("../../scripts/help-functions.R")
```

## Definer constructs for lettere access i kolonnene

```{r load_constructs}
source("../../scripts/constructs.R")
```

# Introduction

The purpose of this script is to do the calculations presented in paper 4 for prototype 1 and 2.

* Difference (gain) scores / Hedges' g
* Fixed-effects analysis
* Within-between (hybrid) model

# Methods

## Samples

Properties of samples

Load data

```{r load_data}
# Rdata-file contains reduced / prepared datasets
load(file="../../data/processed_data/tpack-pilot_2022.Rdata")
load(file="../../data/processed_data/tpack-emne2_2023.Rdata")

# fix bug: data_emne2_ren is list not dataframe
data_emne2_ren <- data.frame(data_emne2_ren)
```

Synchronize data structures / naming conventions

```{r create_dataframe_v23}
# [c(1:4,6)] / unlist(TPACK_PST_emne2_v23)
data_emne2_v23 <- data_emne2_ren[,colnames(data_emne2_ren) %in% unlist(TPACK_PST_emne2_v23)]  #- 3

#transfer covariates
data_emne2_v23$Gender   <- c(data_pre1_emne2$Gender, 
                             rep(NA, nrow(data_pre2_emne2)),
                             data_post_emne2$Gender) # data_emne2_ren$Gender
data_emne2_v23$Trinn    <- c(data_pre1_emne2$Trinn, 
                             rep(NA, nrow(data_pre2_emne2)),
                             data_post_emne2$Trinn) # data_emne2_ren$Trinn
data_emne2_v23$praksis  <- c(data_pre1_emne2$'Har du brukt programmering i praksis sammen med skoleelever?', 
                             rep(NA, nrow(data_pre2_emne2)),
                             data_post_emne2$'Har du brukt programmering i praksis sammen med skoleelever?')

data_emne2_v23$full_treatment <- rep(0,nrow(data_emne2_v23))

data_emne2_v23$datasett <- data_emne2_ren$datasett
data_emne2_v23$Subject  <- c(data_pre1_emne2$Subject, 
                             data_pre2_emne2$Subject,
                             data_post_emne2$Subject)
```

sjekk at full_treatment blir registert

```{r}
# denne måtte skilles ut for å bli registrert riktig!?
data_emne2_v23$full_treatment[data_emne2_v23$Subject %in% data_post_emne2$Subject] <- 1

table(data_emne2_v23$full_treatment)
```

```{r create_dataframe_v22}

# make working copy
data_emne2_v22    <-  data_TPACK_pilot  [data_TPACK_pilot$datasett == "pre" | 
                                           data_TPACK_pilot$datasett == "post",
                                         !colnames(data_TPACK_pilot) %in%
                                           c("semester","studiested",
                                             "XK1","XK2","XK3","XK4","XK5","XK6","CK123","CK6")]

```

Sort datasets

```{r sort_students}
# sort students
# sorter på person
data_emne2_v22$datasett <- factor(data_emne2_v22$datasett, level=c("pre","post"))
data_emne2_v22 <-  data_emne2_v22[
                          with(data_emne2_v22, 
                          order(Subject,datasett, # rekkefølge
                          method="radix", # for faktorer, brukes med c nedenfor
                          decreasing = c(0,0))),] # bare datasett (1)

# set correct order of dataset
data_emne2_v23$datasett <- factor(data_emne2_v23$datasett, level=c("pre1","pre2","post"))
data_emne2_v23 <-  data_emne2_v23[
                          with(data_emne2_v23, 
                          order(Subject,datasett, # rekkefølge
                          method="radix", # for faktorer, brukes med c nedenfor
                          decreasing = c(0,0))),] # bare datasett (1) 
```

add group information

```{r add_groups_v22}
# add column with group numbers
data_emne2_v22$group <- c(1:nrow(data_emne2_v22))
data_emne2_v22[data_emne2_v22$Subject == 'Gabriel' |
                 data_emne2_v22$Subject == 'Hulda' |
                 data_emne2_v22$Subject == 'Isak' |
                 data_emne2_v22$Subject == 'Maria' |
                 data_emne2_v22$Subject == 'Nicoline',"group"] <-
                 "armageddon" #1

data_emne2_v22[data_emne2_v22$Subject == 'Conny' |
                 data_emne2_v22$Subject == 'Fride' |
                 data_emne2_v22$Subject == 'Jonas' |
                 data_emne2_v22$Subject == 'Julia' |
                 data_emne2_v22$Subject == 'Lea',"group"] <-
                 "fasetilstander" # 2

data_emne2_v22[data_emne2_v22$Subject == 'Berta' |
                 data_emne2_v22$Subject == 'Dina' |
                 data_emne2_v22$Subject == 'Gloria' |
                 data_emne2_v22$Subject == 'Klaudia',"group"] <- 
                 "atomeksitasjon" # 3

data_emne2_v22[data_emne2_v22$Subject == 'Benjamin' |
                 data_emne2_v22$Subject == 'Ester' |
                 data_emne2_v22$Subject == 'Harald' |
                 data_emne2_v22$Subject == 'Inga' |
                 data_emne2_v22$Subject == 'Olava',"group"] <- 
                 "smittespredning" # 4

data_emne2_v22[data_emne2_v22$Subject == 'David' |
                 data_emne2_v22$Subject == 'Elias' |
                 data_emne2_v22$Subject == 'Filip',"group"] <-
                 "trykkbeholder" # 5

data_emne2_v22 <- data_emne2_v22 %>% relocate(group, .after = Subject)           # bruker dplyr
```

```{r add_groups_v23}
# add column with group numbers
data_emne2_v23$group <- rep(NA,nrow(data_emne2_v23))
data_emne2_v23[data_emne2_v23$Subject == 'Berta' |
                 data_emne2_v23$Subject == 'Fride' |
                 data_emne2_v23$Subject == 'Gloria' |
                 data_emne2_v23$Subject == 'Nora',"group"] <-
                 "kokeplate" #1/A

data_emne2_v23[data_emne2_v23$Subject == 'Conny' |
                 data_emne2_v23$Subject == 'Idun' |
                 data_emne2_v23$Subject == 'Magrit' |
                 data_emne2_v23$Subject == 'Vaarin',"group"] <-
                 "arv øyefarge" # 2/B

data_emne2_v23[data_emne2_v23$Subject == 'Julia' |
                 data_emne2_v23$Subject == 'Dina' |
                 data_emne2_v23$Subject == 'Ruben',"group"] <- 
                 "partikkelmodellen gass" # 3/C

data_emne2_v23[data_emne2_v23$Subject == 'Emmanuel' |
                 data_emne2_v23$Subject == 'Klaudia' |
                 data_emne2_v23$Subject == 'Noah' | 
                 data_emne2_v23$Subject == "Christian","group"] <- 
                 "partikkelmodellen vannglass" # 4/E

data_emne2_v23[data_emne2_v23$Subject == 'Gabriel' |
                 data_emne2_v23$Subject == 'Harald' |
                 data_emne2_v23$Subject == 'Lukas' |
                 data_emne2_v23$Subject == 'Mikael',"group"] <-
                 "friksjon" # 5/F

data_emne2_v23[data_emne2_v23$Subject == 'Olava' |
               data_emne2_v23$Subject == "Yosita" | 
               data_emne2_v23$Subject == "Nicoline" ,"group"] <-
                 "faseoverganger grønn rute" # 6/D

data_emne2_v23 <- data_emne2_v23 %>% relocate(group, .after = Subject)           # bruker dplyr
```

## Instruments (data collection)

Data were collected using the measure developed by Karlsen et al. (2024) / paper 1.

The measure used in paper 4 was an early version of the instrument developed in paper 1. This led to the TK-items being somewhat different from the ones in paper 1. Here I use TK2, TK5 and TK6, while in paper 1 I use TK1, TK2 and TK3.

## Analyses

For easier comparison between the different analysis methods, the same sample is used for the analyses (i.e., n=19 for prototype1 and n=13 for prototype 2).

Select students

```{r select_students_v22}
###
# v22
###

# bare se på studenter som er med både pre og post
stud_alle_v22  <- intersect(data_emne2_v22$Subject[data_emne2_v22$datasett == "pre"],
                         data_emne2_v22$Subject[data_emne2_v22$datasett == "post"])

# hvilke studenter var ikke med på pre, men på post? 
stud_bare_pre_eller_post_v22  <- data_emne2_v22$Subject[!data_emne2_v22$Subject %in%
                                                          stud_alle_v22]

# hvilke var med bare på post?
stud_bare_post_v22 <- data_emne2_v22$Subject[
  data_emne2_v22$datasett == "post" &
    data_emne2_v22$Subject %in% stud_bare_pre_eller_post_v22]

data_emne2_v22$testing <- rep(0,nrow(data_emne2_v22))
data_emne2_v22$testing[data_emne2_v22$Subject == stud_bare_post_v22] <- 1

alle_studenter_v22 <- data_emne2_v22$Subject %in% stud_alle_v22
```

```{r select_students_v23} 
###
# v23
###

# bare se på studenter som er med både pre og post
stud_alle_v23  <- intersect(data_emne2_v23$Subject[data_emne2_v23$datasett == "pre1"],
                         data_emne2_v23$Subject[data_emne2_v23$datasett == "pre2"])

# hvilke var med bare på pre2?
stud_bare_pre2_v23 <- data_emne2_v23$Subject[
  data_emne2_v23$datasett == "pre2" &
    !data_emne2_v23$Subject %in% stud_alle_v23]

stud_alle_v23  <- intersect(stud_alle_v23,
                         data_emne2_v23$Subject[data_emne2_v23$datasett == "post"])

# hvilke var med bare på post?
stud_bare_post_v23 <- data_emne2_v23$Subject[
  data_emne2_v23$datasett == "post" &
    !data_emne2_v23$Subject %in% stud_alle_v23]

data_emne2_v23$testing <- rep(0,nrow(data_emne2_v23))
data_emne2_v23$testing[data_emne2_v23$Subject %in% stud_bare_pre2_v23] <- 1

#data_emne2_v22 <- data_emne2_v22[data_emne2_v22$Subject %in% stud_alle_v22,]
#data_emne2_v23 <- data_emne2_v23[data_emne2_v23$Subject %in% stud_alle_v23,]

alle_studenter_v23 <- data_emne2_v23$Subject %in% stud_alle_v23
```

### Difference scores

To decide which statistical tests we should use to compare pre-post scores, the Levene’s (Mean) test for equal variances and the Kolmogorov-Smirnov test for normality were used as preliminary tests (Pearce and Derrick, 2019). Due to the small sample sizes, the data are probably non-normally distributed. If the variance of the pre-post differences are unequal, this suggests the use of the non-parametric Yuen d-test with trimmed means (Mair & Wilcox, 2020), which show good power for low samples (Wanishsakpong et al., 2023) and can be appropriate when one is not interested in generalizing to a population (Ng  & Cribbie, 2019). If, on the other hand, the pre-post data have equal variance, this suggests the use of the Mann-Whitney (i.e., Wilcox) test with arithmetic mean.

Hedges' g was also calculated, since it is often used in meta-analyses.

#### Prototype 1

```{r descriptive_stat_v22}
scores_v22 <- scoreItems(TPACK_PST_emne2_v22, data_emne2_v22)
data_emne2_v22$TK <- scores_v22$scores[,"TK"]
data_emne2_v22$PCK <- scores_v22$scores[,"PCK"]
data_emne2_v22$TPACK <- scores_v22$scores[,"TPACK"]

#describe(data_emne2_v22[data_emne2_v22$datasett == "pre",])
scores_v22_pre <- scoreItems(TPACK_PST_emne2_v22,
                             data_emne2_v22[data_emne2_v22$datasett == "pre" &
                                            data_emne2_v22$Subject %in%
                                              stud_alle_v22,])
describe(scores_v22_pre$scores)

#describe(data_emne2_v22[data_emne2_v22$datasett == "post",])
scores_v22_post <- scoreItems(TPACK_PST_emne2_v22,
                              data_emne2_v22[data_emne2_v22$datasett == "post" &
                                             data_emne2_v22$Subject %in%
                                               stud_alle_v22,])
describe(scores_v22_post$scores)
```

```{r create_table_v22}
#!NB Sjekk om rekkefølgen her er riktig? Eller alternativt bruk data_emne2_v22-dataframen

# gjør om til loop

# pre
TK_pre <- scores_v22_pre$scores[,"TK"]
TK_alpha_pre <- scores_v22_pre$corrected["TK","TK"]

PCK_pre <- scores_v22_pre$scores[,"PCK"]
PCK_alpha_pre <- scores_v22_pre$corrected["PCK","PCK"]

TPACK_pre <- scores_v22_pre$scores[,"TPACK"]
TPACK_alpha_pre <- scores_v22_pre$corrected["TPACK","TPACK"]

#post
TK_post <- scores_v22_post$scores[,"TK"]
TK_alpha_post <- scores_v22_post$corrected["TK","TK"]

PCK_post <- scores_v22_post$scores[,"PCK"]
PCK_alpha_post <- scores_v22_post$corrected["PCK","PCK"]

TPACK_post <- scores_v22_post$scores[,"TPACK"]
TPACK_alpha_post <- scores_v22_post$corrected["TPACK","TPACK"]

# cohens d
TK_cd <- effsize::cohen.d(TK_post, TK_pre, 
                 hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)
PCK_cd <- effsize::cohen.d(PCK_post, PCK_pre, 
                  hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)
TPACK_cd <- effsize::cohen.d(TPACK_post, TPACK_pre, 
                    hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)

# tabell
TK_rad <- c(mean(TK_pre), mean(TK_post), TK_cd$estimate, TK_cd$conf.int, TK_alpha_pre, TK_alpha_post)
PCK_rad <- c(mean(PCK_pre), mean(PCK_post), PCK_cd$estimate, PCK_cd$conf.int, PCK_alpha_pre, PCK_alpha_post)
TPACK_rad <- c(mean(TPACK_pre), mean(TPACK_post), TPACK_cd$estimate, TPACK_cd$conf.int, TPACK_alpha_pre, TPACK_alpha_post)
tabell_v22 <- t(data.frame(TK_rad, #TCK_rad, TPK_rad, 
                           PCK_rad, TPACK_rad))
rownames(tabell_v22) <- c("TK", "PCK", "TPACK")
colnames(tabell_v22) <- c("pre","post", "cohen's d", "ci lower", "ci upper", "alpha_pre", "alpha_post")

pander(tabell_v22)
```

Er det noen signifikant forskjell mellom de ulike periodene?

sjekk forutsetninger for t-test (normalitet og lik varians)

```{r}
# sjekk av antagelser
# ref. Pearce and Derrick (2019)

TK22_pre <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "pre", "TK"]
TK22_post <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "post", "TK"]
PCK22_pre <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "pre", "PCK"]
PCK22_post <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "post", "PCK"]
TPACK22_pre <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "pre", "TPACK"]
TPACK22_post <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "post", "TPACK"]

df22_prepost <- data.frame(TK = c(TK22_pre, TK22_post),
                    PCK = c(PCK22_pre, PCK22_post),
                    TPACK = c(TPACK22_pre, TPACK22_post),
                    prototype = c(rep("pre", length(TK22_pre)), 
                                  rep("post", length(TK22_post))))

```

```{r}
# Levene's (Mean) test for equal variances
# ref https://www.statology.org/levenes-test-r/
#library(car)
leveneTest(TK ~ prototype, center=mean, data = df22_prepost)
leveneTest(PCK ~ prototype, center=mean, data = df22_prepost)
leveneTest(TPACK ~ prototype, center=mean, data = df22_prepost)
```

Ulik varians.

```{r}
ks.test(TK22_pre, "pnorm")
ks.test(TK22_post, "pnorm")
ks.test(PCK22_pre, "pnorm")
ks.test(PCK22_post, "pnorm")
ks.test(TPACK22_pre, "pnorm")
ks.test(TPACK22_post, "pnorm")
```

Ikke normalfordelt.

```{r}
# Yuen-Welch test
# ref.: Mair & Wilcox, 2020
#library(WRS2)
yuend(TK22_pre, TK22_post)
yuend(PCK22_pre, PCK22_post)
yuend(TPACK22_pre, TPACK22_post)
```

```{r}
t.test(data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="pre", "TK"],
       data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="post", "TK"], 
       paired=T)

t.test(data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="pre", "PCK"],
       data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="post", "PCK"], 
       paired=T)

t.test(data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="pre", "TPACK"],
       data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="post", "TPACK"], 
       paired=T)
```

Se på forandring i TPACK

```{r calc_diff_v22}
calc_diff_TPACK_v22 <- function (TPACK_element) {

df <- data_emne2_v22[data_emne2_v22$datasett == "pre" | 
                     data_emne2_v22$datasett == "post",
                     c("Subject","datasett",TPACK_element)] # "group",

# sorter på datasett
df <-  df[
                          with(df, 
                          order(datasett,Subject, # rekkefølge
                          method="radix"#, # for faktorer, brukes med c nedenfor
                          #decreasing = c(-1,0))
                          ))
                          ,] # bare datasett (1)

# regn ut TPACK pre/post for hver person
ifelse (length(TPACK_element) > 1, {
TPACK_pre <- rowMeans(df[df$datasett=="pre",3:ncol(df)],na.rm=TRUE)
TPACK_post <- rowMeans(df[df$datasett=="post",3:ncol(df)],na.rm=TRUE)
},
{
  TPACK_pre <- df[df$datasett=="pre",TPACK_element]
  TPACK_post <- df[df$datasett=="post",TPACK_element]
})
TPACK_merge <- c(TPACK_pre,TPACK_post)

# legg til resultat
df$TPACK_pre_post <- TPACK_merge

# sorter på person; 
df <-  df[
                          with(df, 
                          order(Subject,datasett, # rekkefølge
                          method="radix"))#, # for faktorer, brukes med c nedenfor
                          #decreasing = c(0,1)))
                          ,] # bare datasett (1)

# behold bare personer som er med i  både pre/post
df_filtered <- df[!df$Subject %in% c("Benjamin","Gabriel","Olava"),]

# beregn differanse; pre inneholder riktig diff-verdi
df_filtered$diff <-
  c(diff(df_filtered[,ncol(df_filtered)]),0)

diff_TPACK <- df_filtered[df_filtered$datasett=="pre",c(1,ncol(df_filtered))]

return (diff_TPACK)
}

TK_diff_v22 <- calc_diff_TPACK_v22(TK_PST22_pos)
PCK_diff_v22 <- calc_diff_TPACK_v22(PCK_PST22_pos)
TPACK_diff_v22 <- calc_diff_TPACK_v22(TPACK_PST22_pos)

# lag dataframe
diff_full_v22 <- data.frame("Subject"=TK_diff_v22[,1], #"group" = TK_diff[,"group"],
  "TK"=TK_diff_v22[,2],
  "PCK"=PCK_diff_v22[,2],
  "TPACK"=TPACK_diff_v22[,2])
```

##### reliable change in individual scores

beregn SID (Estrada et al., 2019).

```{r}
SID <- TK_diff_v22[,2] / SD(TK_diff_v22[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

```{r}
SID <- PCK_diff_v22[,2] / SD(PCK_diff_v22[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

```{r}
SID <- TPACK_diff_v22[,2] / SD(TPACK_diff_v22[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

##### Sjekk for RTM

RTM = regression to the mean

Ref.: Marsden og Torgerson (2012)

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v22[,"TK"], 
                  "TK" = data_emne2_v22[data_emne2_v22$datasett == "pre" & 
                                        data_emne2_v22$Subject %in% stud_alle_v22,"TK"]),
       aes(x=TK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v22[,"PCK"], 
                  "PCK" = data_emne2_v22[data_emne2_v22$datasett == "pre" & 
                                        data_emne2_v22$Subject %in% stud_alle_v22,"PCK"]),
       aes(x=PCK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v22[,"TPACK"], 
                  "TPACK" = data_emne2_v22[data_emne2_v22$datasett == "pre" & 
                                        data_emne2_v22$Subject %in% stud_alle_v22,"TPACK"]),
       aes(x=TPACK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

#### Prototype 2

```{r descriptive_stat_v23}
scores_v23 <- scoreItems(TPACK_PST_emne2_v23, data_emne2_v23)
data_emne2_v23$TK <- scores_v23$scores[,"TK"]
data_emne2_v23$PCK <- scores_v23$scores[,"PCK"]
data_emne2_v23$TPACK <- scores_v23$scores[,"TPACK"]

#describe(data_emne2_v23[data_emne2_v23$datasett == "pre",])
scores_v23_pre <- scoreItems(TPACK_PST_emne2_v23, data_emne2_v23[data_emne2_v23$datasett == "pre1" & data_emne2_v23$Subject %in% stud_alle_v23,])
describe(scores_v23_pre$scores)

#describe(data_emne2_v23[data_emne2_v23$datasett == "post",])
scores_v23_post <- scoreItems(TPACK_PST_emne2_v23,
                              data_emne2_v23[data_emne2_v23$datasett == "post" &
                                             data_emne2_v23$Subject %in%
                                               stud_alle_v23,])
describe(scores_v23_post$scores)
```

```{r create_table_v23}
#!NB Sjekk om rekkefølgen her er riktig? Eller alternativt bruk data_emne2_v23-dataframen

# gjør om til loop

# pre
TK_pre <- scores_v23_pre$scores[,"TK"]
TK_alpha_pre <- scores_v23_pre$corrected["TK","TK"]

PCK_pre <- scores_v23_pre$scores[,"PCK"]
PCK_alpha_pre <- scores_v23_pre$corrected["PCK","PCK"]

TPACK_pre <- scores_v23_pre$scores[,"TPACK"]
TPACK_alpha_pre <- scores_v23_pre$corrected["TPACK","TPACK"]

#post
TK_post <- scores_v23_post$scores[,"TK"]
TK_alpha_post <- scores_v23_post$corrected["TK","TK"]

PCK_post <- scores_v23_post$scores[,"PCK"]
PCK_alpha_post <- scores_v23_post$corrected["PCK","PCK"]

TPACK_post <- scores_v23_post$scores[,"TPACK"]
TPACK_alpha_post <- scores_v23_post$corrected["TPACK","TPACK"]

# cohens d
TK_cd <- effsize::cohen.d(TK_post, TK_pre, 
                 hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)
PCK_cd <- effsize::cohen.d(PCK_post, PCK_pre, 
                  hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)
TPACK_cd <- effsize::cohen.d(TPACK_post, TPACK_pre, 
                    hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)

# tabell
TK_rad <- c(mean(TK_pre), mean(TK_post), TK_cd$estimate, TK_cd$conf.int, TK_alpha_pre, TK_alpha_post)
PCK_rad <- c(mean(PCK_pre), mean(PCK_post), PCK_cd$estimate, PCK_cd$conf.int, PCK_alpha_pre, PCK_alpha_post)
TPACK_rad <- c(mean(TPACK_pre), mean(TPACK_post), TPACK_cd$estimate, TPACK_cd$conf.int, TPACK_alpha_pre, TPACK_alpha_post)
tabell_v23 <- t(data.frame(TK_rad, #TCK_rad, TPK_rad, 
                           PCK_rad, TPACK_rad))
rownames(tabell_v23) <- c("TK", "PCK", "TPACK")
colnames(tabell_v23) <- c("pre","post", "cohen's d", "ci lower", "ci upper", "alpha_pre", "alpha_post")

pander(tabell_v23)
```

Er det noen signifikant forskjell mellom de ulike periodene?

sjekk forutsetninger for t-test (normalitet og lik varians)

```{r}
# sjekk av antagelser
# ref. Pearce and Derrick (2019)

TK23_pre <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "pre1", "TK"]
TK23_post <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "post", "TK"]
PCK23_pre <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "pre1", "PCK"]
PCK23_post <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "post", "PCK"]
TPACK23_pre <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "pre1", "TPACK"]
TPACK23_post <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "post", "TPACK"]

df23_prepost <- data.frame(TK = c(TK23_pre, TK23_post),
                    PCK = c(PCK23_pre, PCK23_post),
                    TPACK = c(TPACK23_pre, TPACK23_post),
                    prototype = c(rep("pre1", length(TK23_pre)), 
                                  rep("post", length(TK23_post))))

```

```{r}
# Levene's (Mean) test for equal variances
# ref https://www.statology.org/levenes-test-r/
#library(car)
leveneTest(TK ~ prototype, center=mean, data = df23_prepost)
leveneTest(PCK ~ prototype, center=mean, data = df23_prepost)
leveneTest(TPACK ~ prototype, center=mean, data = df23_prepost)
```

Ulik varians.

```{r}
ks.test(TK23_pre, "pnorm")
ks.test(TK23_post, "pnorm")
ks.test(PCK23_pre, "pnorm")
ks.test(PCK23_post, "pnorm")
ks.test(TPACK23_pre, "pnorm")
ks.test(TPACK23_post, "pnorm")
```

Ikke normalfordelt.

```{r}
# Wilcoxon signed-rank test (Mann-Whitney)
wilcox.test(TK23_pre, TK23_post, paired=T)
wilcox.test(PCK23_pre, PCK23_post, paired=T)
wilcox.test(TPACK23_pre, TPACK23_post, paired=T)
```

```{r}
# Yuen-Welch test
# ref.: Mair & Wilcox, 2020
#library(WRS2)
yuend(TK23_pre, TK23_post)
yuend(PCK23_pre, PCK23_post)
yuend(TPACK23_pre, TPACK23_post)
```

```{r}
t.test(data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="pre1", "TK"],
       data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="post", "TK"], 
       paired=T)

t.test(data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="pre1", "PCK"],
       data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="post", "PCK"], 
       paired=T)

t.test(data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="pre1", "TPACK"],
       data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="post", "TPACK"], 
       paired=T)
```

Se på forandring i TPACK

```{r calc_diff_v23}
calc_diff_TPACK_v23 <- function (TPACK_element) {

df <- data_emne2_v23[data_emne2_v23$datasett == "pre1" | 
                     data_emne2_v23$datasett == "post",
                     c("Subject","datasett",TPACK_element)] # "group",

# sorter på datasett
df <-  df[
                          with(df, 
                          order(datasett,Subject, # rekkefølge
                          method="radix"#, # for faktorer, brukes med c nedenfor
                          #decreasing = c(-1,0))
                          ))
                          ,] # bare datasett (1)

# regn ut TPACK pre/post for hver person
ifelse (length(TPACK_element) > 1, {
TPACK_pre <- rowMeans(df[df$datasett=="pre1",3:ncol(df)],na.rm=TRUE)
TPACK_post <- rowMeans(df[df$datasett=="post",3:ncol(df)],na.rm=TRUE)
},
{
  TPACK_pre <- df[df$datasett=="pre1",TPACK_element]
  TPACK_post <- df[df$datasett=="post",TPACK_element]
})
TPACK_merge <- c(TPACK_pre,TPACK_post)

# legg til resultat
df$TPACK_pre_post <- TPACK_merge

# sorter på person; 
df <-  df[
                          with(df, 
                          order(Subject,datasett, # rekkefølge
                          method="radix"))#, # for faktorer, brukes med c nedenfor
                          #decreasing = c(0,1)))
                          ,] # bare datasett (1)

# behold bare personer som er med i  både pre/post
df_filtered <- df[df$Subject %in% stud_alle_v23,]

# beregn differanse; pre inneholder riktig diff-verdi
df_filtered$diff <-
  c(diff(df_filtered[,ncol(df_filtered)]),0)

diff_TPACK <- df_filtered[df_filtered$datasett=="pre1",c(1,ncol(df_filtered))]

return (diff_TPACK)
}

TK_diff_v23 <- calc_diff_TPACK_v23(TK_PST23_pos)
PCK_diff_v23 <- calc_diff_TPACK_v23(PCK_PST23_pos)
TPACK_diff_v23 <- calc_diff_TPACK_v23(TPACK_PST23_pos)

# lag dataframe
diff_full_v23 <- data.frame("Subject"=TK_diff_v23[,1], #"group" = TK_diff[,"group"],
  "TK"=TK_diff_v23[,2],
  "PCK"=PCK_diff_v23[,2],
  "TPACK"=TPACK_diff_v23[,2])
```

##### reliable change in individual scores

beregn SID (Estrada et al., 2019).

```{r}
SID <- TK_diff_v23[,2] / SD(TK_diff_v23[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

```{r}
SID <- PCK_diff_v23[,2] / SD(PCK_diff_v23[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

```{r}
SID <- TPACK_diff_v23[,2] / SD(TPACK_diff_v23[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

##### Sjekk for RTM

RTM = regression to the mean

Ref.: Marsden og Torgerson (2012)

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v23[,"TK"], 
                  "TK" = data_emne2_v23[data_emne2_v23$datasett == "pre1" & 
                                        data_emne2_v23$Subject %in% stud_alle_v23,"TK"]),
       aes(x=TK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v23[,"PCK"], 
                  "PCK" = data_emne2_v23[data_emne2_v23$datasett == "pre1" & 
                                        data_emne2_v23$Subject %in% stud_alle_v23,"PCK"]),
       aes(x=PCK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v23[,"TPACK"], 
                  "TPACK" = data_emne2_v23[data_emne2_v23$datasett == "pre1" & 
                                        data_emne2_v23$Subject %in% stud_alle_v23,"TPACK"]),
       aes(x=TPACK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

### Fixed-effects

Before doing the regression analysis, we need to decide how to model the data. Should we model it through pooled OLS, fixed effects or random effects?

... Legg til sjekker av hvilken modell som bør brukes

```{r}
within22 <- plm(TPACK ~ TK + PCK + factor(Subject) + factor(datasett), 
              data=data_emne2_v22,
              index=c("Subject","datasett"), model="within")
firstdifference22 <- plm(TPACK ~ TK + PCK + factor(Subject),# + factor(datasett),
              data=data_emne2_v22,
              index=c("Subject","datasett"), model="fd")
```

```{r within_3pt}
within23 <- plm(TPACK ~ TK + PCK + factor(Subject) + factor(datasett), 
              data=data_emne2_v23,
              index=c("Subject","datasett"), model="within")
firstdifference23 <- plm(TPACK ~ TK + PCK + factor(Subject),# + factor(datasett),
              data=data_emne2_v23,
              index=c("Subject","datasett"), model="fd")
```

```{r within_2pt}
within23_prepost <- plm(TPACK ~ TK + PCK + factor(Subject) + factor(datasett), 
              data=data_emne2_v23[data_emne2_v23$datasett == "pre1" | data_emne2_v23$datasett == "post",],
              index=c("Subject","datasett"), model="within")
firstdifference23_prepost <- plm(TPACK ~ TK + PCK + factor(Subject),# + factor(datasett),
              data=data_emne2_v23[data_emne2_v23$datasett == "pre1" | data_emne2_v23$datasett == "post",],
              index=c("Subject","datasett"), model="fd")
```

```{r}
screenreg(list("within22" = within22,
               "fd22" = firstdifference22,
               "within23" = within23,
               "fd23" = firstdifference23,
               "within23pp" = within23_prepost,
               "fd23pp" = firstdifference23_prepost))
```

### Within-between models

To calculate the within-between model for prototype 2, all three datapoints were needed. This was not done in paper 4.

Prototype 1

prepare centred and demeaned variabels

create a dummy variable for each individual
or
gruppesentrering, _S

```{r prot1_centering}
# TK
# calculate mean for each subject / id; ref.: https://stackoverflow.com/a/34502529
data_emne2_v22$TK_mean <- with(data_emne2_v22, ave(x = TK, Subject, FUN = mean))
data_emne2_v22$TK_S <- data_emne2_v22$TK - data_emne2_v22$TK_mean

# PCK interval
data_emne2_v22$PCK_mean <- with(data_emne2_v22, ave(x = PCK, Subject, FUN = mean))
data_emne2_v22$PCK_S <- data_emne2_v22$PCK - data_emne2_v22$PCK_mean

# TPACK
data_emne2_v22$TPACK_mean <- with(data_emne2_v22, ave(x = TPACK, Subject, FUN = mean))
data_emne2_v22$TPACK_S <- data_emne2_v22$TPACK - data_emne2_v22$TPACK_mean
```

```{r prot2_centering}
# TK
# calculate mean for each subject / id; ref.: https://stackoverflow.com/a/34502529
data_emne2_v23$TK_mean <- with(data_emne2_v23, ave(x = TK, Subject, FUN = mean))
data_emne2_v23$TK_S <- data_emne2_v23$TK - data_emne2_v23$TK_mean

# PCK interval
data_emne2_v23$PCK_mean <- with(data_emne2_v23, ave(x = PCK, Subject, FUN = mean))
data_emne2_v23$PCK_S <- data_emne2_v23$PCK - data_emne2_v23$PCK_mean

# TPACK
data_emne2_v23$TPACK_mean <- with(data_emne2_v23, ave(x = TPACK, Subject, FUN = mean))
data_emne2_v23$TPACK_S <- data_emne2_v23$TPACK - data_emne2_v23$TPACK_mean
```

Hybrid with variable coefficients was chosen since there was only one significant dependent variable in the fixed-effects model (Bell et al., 2019).

Ville det vært bedre å sammenligne med en enkel hybrid model uten variable koeffisienter først, for mer direkte sammenligning med FE-modellen?

```{r eval=T}
# TK_S within effects
# TK_mean between effects
# factor(datasett) fixed time effects
# TK_S|id variable coefficient TK_S per individual

hybrid_TK <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~1|Subject, method="REML", data=data_emne2_v22[alle_studenter_v22,])

summary(hybrid_TK)
```

bare pre-post

```{r eval=T}
# TK_S within effects
# TK_mean between effects
# factor(datasett) fixed time effects
# TK_S|id variable coefficient TK_S per individual

hybrid_TK <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~TK_S|Subject, method="REML", data=data_emne2_v22[alle_studenter_v22,])

summary(hybrid_TK)
```

alle studentene

```{r eval=T}
# TK_S within effects
# TK_mean between effects
# factor(datasett) fixed time effects
# TK_S|id variable coefficient TK_S per individual

hybrid_TK <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~TK_S|Subject, method="REML", data=data_emne2_v22)

summary(hybrid_TK)
```

#### Comparison hybrid models

Use FIML (ML) for model comparison, REML for parameter estimates (Christophersen, 2018).

Prototype 1

```{r eval=T}
# TK_S within effects
# TK_mean between effects
# factor(datasett) fixed time effects
# TK_S|id variable coefficient TK_S per individual

prot1_within_between <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~1|Subject, method="ML", data=data_emne2_v22[alle_studenter_v22,])

prot1_var_coeff <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~TK_S|Subject, method="ML", data=data_emne2_v22[alle_studenter_v22,])
```

Prototype 2

```{r prot2_wb, eval=T}
# TK_S within effects
# TK_mean between effects
# factor(datasett) fixed time effects
# TK_S|id variable coefficient TK_S per individual

# only 13 from both pre-post
prot2_within_between_13 <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~1|Subject, method="REML", 
                   data=data_emne2_v23[alle_studenter_v23 & !data_emne2_v23$datasett == "pre2",])

# all 30 for comparison with var coeff below
prot2_within_between <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~1|Subject, method="ML", data=data_emne2_v23[ !data_emne2_v23$datasett == "pre2",]) # [alle_studenter_v23,]
```

```{r prot2_vc}
# does not converge with only 13 respondents
prot2_var_coeff <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~TK_S|Subject, method="ML", data=data_emne2_v23[ !data_emne2_v23$datasett == "pre2",]) # alle_studenter_v23 &
```

```{r}
# print summary of models
screenreg(list("P1_wb" = prot1_within_between, 
               "P1_vc" = prot1_var_coeff,
               "P2_wb" = prot2_within_between, 
               "P2_vc" = prot2_var_coeff,
               "P2_13" = prot2_within_between_13))
```

chi-square comparison, significant difference between models?

```{r}
anova(prot1_within_between, prot1_var_coeff)

anova(prot2_within_between, prot2_var_coeff)
```

AICc for small samples

```{r}
library(MuMIn) # AICc
AICc(prot1_within_between, prot1_var_coeff, REML=F)
AICc(prot2_within_between, prot2_var_coeff, REML=F)
```

summary of details for best fit models

```{r}
prot1_var_coeff_REML <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~TK_S|Subject, method="REML", data=data_emne2_v22[alle_studenter_v22,])

prot2_within_between_13_REML <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~1|Subject, method="REML", 
                   data=data_emne2_v23[alle_studenter_v23 & !data_emne2_v23$datasett == "pre2",])

summary(prot1_var_coeff_REML)
summary(prot2_within_between_13_REML)
```

# Results

### 2024-08-22 Rmd/thesis-reanalyses.Rmd
Initial commit

---
title: "Reanalyses done for the synopsis"
author: "Niklas Karlsen"
date: "2024-08-21"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/reanalyses-thesis.html')) })
---

```{r}
# load libraries
library(nlme)
library(psych)
library(dplyr)
library(pander)
```

# Introduction

Here I present the reanalyses work that was done as part of writing the synopsis. 
# Methods

## Samples

Load data

```{r load_data_survey}
# Rdata-file contains reduced / prepared datasets
load(file="../../data/processed_data/tpack-TRELIS-Q_2023.Rdata")
```

```{r fix_missing, eval=T}
# remove NMAR + NTNU (19)
data_clean <- data_trelis_q_ren[-c(19,34,44,45,46,95,102),] 
```

## Instruments

### Define constructs

```{r load_constructs}
## TPACK-PST for TRELIS-Q
TK_pos_trelisq    <- c("TK1b","TK2","TK34") 
PCK_pos_trelisq   <- c("CK123","CK5","PCK2","PCK5")
TPACK_pos_trelisq <- c("TPACK2","TPACK4","TCK1","TCK4","TPK1","TPK4")

TPACK_PST_list    <- list(TK=TK_pos_trelisq,
                     PCK=PCK_pos_trelisq,
                     TPACK=TPACK_pos_trelisq)
```

## Calculate mean values

```{r}
scores_v23 <- scoreItems(TPACK_PST_list, data_clean)

describe(scores_v23$scores)

# constructs for multiple regression
TK_v23    <- scores_v23$scores[,"TK"]
PCK_v23   <- scores_v23$scores[,"PCK"]
TPACK_v23 <- scores_v23$scores[,"TPACK"]

# add constructs to data.frame
data_clean$TK    <- TK_v23
data_clean$PCK   <- PCK_v23
data_clean$TPACK <- TPACK_v23
```

# Examine correlation between TK and TPACK

## Survey

level 2-analysis

```{r}
#lme(TPACK ~ TK, )

hlm <- lme(TPACK ~ TK, random=~1|studiested, method="REML", data=data_clean)
summary(hlm)
```

The correlation between the intercept (value of TPACK at each teacher education institute (TEI)) and TK is negative, indicating that the higher TPACK is at a TEI, the lower the correlation with TK (i.e., the slope) becomes (Brown, 2021). This could be a ceiling effect, but since TPACK is relatively low, it probably isn't?

```{r}
# per institute and year of study
mTK <- aggregate(x=data_clean$TK,
          by=list(data_clean$studiested,data_clean$semester),
          FUN=mean)

mTPACK <- aggregate(x=data_clean$TPACK,
          by=list(data_clean$studiested,data_clean$semester),
          FUN=mean)

colnames(mTK) <- c("studiested", "semester", "mTK")
colnames(mTPACK) <- c("studiested", "semester", "mTPACK")

plot(mTK[,3], mTPACK[,3]) # x contains mean value
```

Plot correlation (TK/TPACK) vs TPACK

```{r}
# ref: https://stats.stackexchange.com/questions/4040/r-compute-correlation-by-group [21.8.2024]
#r_TK_TPACK <- by(data_clean[,c("TK", "TPACK")], data_clean$studiested, function(x) {cor(x$TK, x$TPACK)})

# ref: https://www.geeksforgeeks.org/how-to-calculate-correlation-by-group-in-r/ [21.8.2024]
r_TK_TPACK <- data_clean %>%
  group_by(studiested, semester) %>%
  summarise(correlation = cor(TK, TPACK))

# ref: https://www.statology.org/r-merge-multiple-columns/ [21.8.2024]
cor_TPACK_table <- merge(r_TK_TPACK, mTK, by.x=c('studiested', 'semester'), by.y=c('studiested', 'semester'))

cor_TPACK_table <- merge(cor_TPACK_table, mTPACK, by.x=c('studiested', 'semester'), by.y=c('studiested', 'semester'))

pander(cor_TPACK_table)

plot(cor_TPACK_table$correlation, cor_TPACK_table$mTPACK)
```

```{r}
# remove outlier correlation, which is also a indication of TK and TPACK being partially unrelated
linmod <- lm(correlation ~ mTPACK,
             data=cor_TPACK_table[cor_TPACK_table$correlation>0.5,])

summary(linmod)
```

## Learning sequence

Correlation between pre-values compared to correlation between post-values.

### 2024-08-22 Rmd/Paper4.Rmd
Fixed error with mean being calculated from the whole dataset, while the model was estimated only using the pre-post values. (for prototype 2)

@@
 #library(mice) # missing
+library(MuMIn) # AICc
+
 ```
@@
 
+```{r prot2_centering_prepost}
+# do not include pre2-data
+data_emne2_v23_pp <- data_emne2_v23[!data_emne2_v23$datasett == "pre2",]
+
+alle_studenter_v23_pp <-
+  alle_studenter_v23[!data_emne2_v23$datasett == "pre2"]
+
+# TK
+# calculate mean for each subject / id; ref.: https://stackoverflow.com/a/34502529
+data_emne2_v23_pp$TK_mean <- with(data_emne2_v23_pp, ave(x = TK, Subject, FUN = mean))
+data_emne2_v23_pp$TK_S <- data_emne2_v23_pp$TK - data_emne2_v23_pp$TK_mean
+
+# PCK interval
+data_emne2_v23_pp$PCK_mean <- with(data_emne2_v23_pp, ave(x = PCK, Subject, FUN = mean))
+data_emne2_v23_pp$PCK_S <- data_emne2_v23_pp$PCK - data_emne2_v23_pp$PCK_mean
+
+# TPACK
+data_emne2_v23_pp$TPACK_mean <- with(data_emne2_v23_pp, ave(x = TPACK, Subject, FUN = mean))
+data_emne2_v23_pp$TPACK_S <- data_emne2_v23_pp$TPACK - data_emne2_v23_pp$TPACK_mean
+```
+
 Hybrid with variable coefficients was chosen since there was only one significant dependent variable in the fixed-effects model (Bell et al., 2019).
@@
                    random=~1|Subject, method="REML", 
-                   data=data_emne2_v23[alle_studenter_v23 & !data_emne2_v23$datasett == "pre2",])
+                   data=data_emne2_v23_pp[alle_studenter_v23_pp,])
 
@@
 
+using data_emne2_v23[ !data_emne2_v23$datasett == "pre2",] gives an error in the estimate, since TK_mean, etc.
+are calculated using the mean for all three datapoints, not only pre1 and post.
+It does not converge if I use the _pp dataset.
+
 ```{r prot2_vc}
@@
 prot2_var_coeff <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
-                   random=~TK_S|Subject, method="ML", data=data_emne2_v23[ !data_emne2_v23$datasett == "pre2",]) # alle_studenter_v23 &
+                   random=~TK_S|Subject, method="ML", data=data_emne2_v23[ !data_emne2_v23$datasett == "pre2",]) # 
 ```
@@
 ```{r}
-library(MuMIn) # AICc
 AICc(prot1_within_between, prot1_var_coeff, REML=F)

### 2024-08-28 Rmd/thesis-reanalyses.Rmd
Added plot

@@
 author: "Niklas Karlsen"
-date: "2024-08-21"
+date: "2024-08-28"
 output: 
@@
 
+Changelog
+
+* 28.8.2024: added plot with regression line of cor(TK/TPACK) vs mTPACK
+
 ```{r}
@@
 library(pander)
+library(ggplot2)
 ```
@@
 ```{r}
+ggplot(cor_TPACK_table[cor_TPACK_table$correlation>0.5,], aes(x=mTPACK,y=correlation)) +
+  geom_point() +
+  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
+```
+
+```{r}
 # remove outlier correlation, which is also a indication of TK and TPACK being partially unrelated

### 2024-09-02 Rmd/Paper1.Rmd
Added data cleaning code

--- 2024-05-28 Rmd/Paper1.Rmd
@@
 The whole process from reducing the raw data to doing the EFA and CFA is discussed.
-The document is written as a computational essay [@odden_using_2023].
-
-This R-markdown describes the analyses used to support the validity and reliability of the inferences about PSTs' TPACK based on the measure developed in the thesis.
-This includes exploratory factor analysis (EFA) and the confirmatory factor analysis (CFA) that was done to get the results in paper 1.
+The document is inspired by the idea of a computational essay [@odden_using_2023].
 
@@
 
-The first stage is described in Paper 1.
-This computational essay describes the analysis behind the results in paper 1, and emphasises the quantitative analyses done in the final two stages.
+The first stage is described in @karlsen_assessing_2024.
+This computational essay describes the analysis behind the results in this paper, and emphasises the quantitative analyses done in the final two stages.
 This concerns item reduction and extraction of factors (EFA) in stage two, and test of dimensionality (CFA) and tests of reliability and validity in stage three.
-
-From stage 2 Boateng et al.
-(2013) describes three steps which will be discussed here: step 4 (gather data), step 5 (item reduction) and step 6 (factor extraction, i.e., EFA).
-From stage 3 there are three steps which will be discussed: step 7 (tests of dimensionality, i.e., CFA), step 8 (tests of reliability) and step 9 (tests of validity).
 
@@
 # libraries
+
+library(readODS)                            # read .ods-files. Kilde: https://github.com/ropensci/readODS/
+library(readxl)                             # read .xlsx-files. Kilde: https://readxl.tidyverse.org/
+library(dplyr)                              # rename columns
+
 ```
@@
 
+From stage 2 Boateng et al. (2013) describes three steps which will be discussed here: step 4 (gather data), step 5 (item reduction) and step 6 (factor extraction, i.e., EFA).
+
 ## Step 4: Data gathering / cleaning
 
-This step concerns gathering data for the EFA and the CFA, using separate samples for each analysis.
+This step concerns data gathering for the EFA and the CFA using separate samples for each analysis.
 The recommended sample size is 10 respondents per item and/or 200-300 observations in total (Boateng et al., 2018).
-The sample size is discussed more closely in the thesis (ref sec page).
+The sample size is discussed more closely in @karlsen_assessing_2024.
 
@@
 
-Reducing the raw data ... (transforming from ods to Rdata)
+The data for the first pilot were collected using two almost identical questionnaires.
+Four of the items were however slightly modified to use the term "coding" instead of "programming".
+"Programming" is used in the final published version, so this is what we choose also when reducing the dataset.
+
+All respondents were explicitly asked to agree to their answers being treated in the research.
+All answered "yes", except a few that answered blank, perhaps forgetting to check the answer, as it was not obligatory.
+We interpreted blank answers as "yes", although strictly speaking this should probably be interpreted as "no" to be on the safe side.
+For the respondents who answered blank to the qeustion to agree to participate in the research, all the questions in the questionnaire were filled in.
+The data contained no identifying information, and thus were anonymous.
+
+### Reducing the raw data ... (transforming from xlsx/ods to Rdata)
+
+The raw data were downloaded from nettskjema.no in the xlsx-format and transformed to the Rdata-format in Rstudio.
+To get enough respondents for the EFA, the pilot data were combined with data from the learning sequence we developed (codename: emne2).
+
+* pilot: data-257861-2022-06-15-1302-utf.xlsx
+* emne2: predata-257846-2022-03-28-1311-utf.ods
+* trelis-q pilot: Datafil-utproving-trelis-q-institute4-institute1-aug2022.xlsx
+* TRELIS-Q (dec-feb): data-trelis-q-tatt-ut-2feb2023.xlsx
+
+Pilot is the pilot for the questionnaire published in @karlsen_assessing_2024.
+trelis-q pilot is the pilot for the TRELIS-Q questionnaire.
+The pilot was incorporated into TRELIS-Q.
+TRELIS-Q is a larger questionnaire containing questions related to the three working packages of TRELIS related to inquiry, programming and professional development.
+My work is related to the programming work package (WP5).
+The emne2-dataset was saved in the ods-format as the participant names were anonymized.
 
 * Step 4a: convert from ods to Rdata; rename items using TPACK typology; treat missing
+
+#### pilot
+
+The aim is to create the following dataset:
+
+* data_TPACK_pilot, which uses TK, CK, etc for column names
+
+The dataframe is built up with the different datasets as follows:
+
+df[1:21,] =   emne2_pre        (n=21)  
+df[22:41,] =  emne2_post       (n=20)  
+df[42:57,] =  pilot            (n=16)  
+df[58:105,] = trelis_q pilot   (n=48)
+
+#### Load datasets
+
+```{r load_emne2}
+folder_emne2    <- "../data/rawdata/"
+
+pre_emne2       <- "predata-257846-2022-03-28-1311-utf.ods"
+post_emne2      <- "postdata-257846-2022-05-10-1048-utf.ods"
+
+path_pre_emne2  <- paste(folder_emne2, pre_emne2, sep='')
+path_post_emne2 <- paste(folder_emne2, post_emne2, sep='')
+
+data_pre_emne2  <- read_ods(path_pre_emne2)
+data_post_emne2 <- read_ods(path_post_emne2)
+```
+
+```{r load_pilot}
+folder_pilot    <- "../data/rawdata/"
+pilot           <- "data-257861-2022-06-15-1302-utf.xlsx"
+
+path_pilot      <- paste(folder_pilot, pilot, sep='')
+
+data_pilot      <- read_excel(path_pilot)
+```
+
+```{r load_pilot_trelis_q}
+folder_trelis_q <- "../data/rawdata/"
+trelis_q        <- "Datafil-utproving-trelis-q-institute4-institute1-aug2022.xlsx"
+
+path_trelis_q   <- paste(folder_trelis_q, trelis_q, sep='')
+
+data_trelis_q   <- read_excel(path_trelis_q, sheet=1)
+```
+
+##### Data cleaning
+
+Rename items using the term "coding" so they use "programming" instead to enable merging of the datasets.
+
+```{r emne2_cleaning}
+# keep some columns
+data_pre_emne2  <- rename(data_pre_emne2,
+                             "Subject" =
+                             "Hva heter du (fornavn og etternavn)?")
+data_post_emne2 <- rename(data_post_emne2,
+                             "Subject" =
+                             "Hva heter du (fornavn og etternavn)?")
+data_pre_emne2  <- rename(data_pre_emne2,
+                             "Trinn" =
+                             "Går du på grunnskolelærerutdanning for trinn 1-7 eller trinn 5-10?")
+data_post_emne2 <- rename(data_post_emne2,
+                             "Trinn" =
+                             "Går du på grunnskolelærerutdanning for trinn 1-7 eller trinn 5-10?")
+data_pre_emne2  <- rename(data_pre_emne2,
+                             "andre_fag" =
+                             "Hva er ditt andre fag?")
+data_post_emne2 <- rename(data_post_emne2,
+                             "andre_fag" =
+                             "Hva er ditt andre fag?")
+
+# add institute and semester
+data_pre_emne2$studiested      <- rep("Institute4",21)
+data_pre_emne2$semester        <- rep(8,21)      # 8. semester (syklus 2, vår)
+data_post_emne2$studiested     <- rep("Institute4",20)
+data_post_emne2$semester       <- rep(8,20)      # 8. semester
+
+# add full_treatment; 1 = deltok på all undervisning; 0 = ikke all undv.
+data_pre_emne2$full_treatment  <- rep(1,21)
+data_post_emne2$full_treatment <- rep(1,20)
+# de som ikke deltok på alt
+data_pre_emne2[data_pre_emne2$Subject   == "Ester","full_treatment"]  <- 0
+data_pre_emne2[data_pre_emne2$Subject   == "Harald","full_treatment"] <- 0
+data_pre_emne2[data_pre_emne2$Subject   == "Olava","full_treatment"]  <- 0
+data_post_emne2[data_post_emne2$Subject == "Ester","full_treatment"]  <- 0
+data_post_emne2[data_post_emne2$Subject == "Harald","full_treatment"] <- 0
+data_post_emne2[data_post_emne2$Subject == "Olava","full_treatment"]  <- 0
+
+# add gender; 0 = male; 1 = female
+data_pre_emne2$Gender <- c("Kvinne", "Mann", "Kvinne", "Mann", "Kvinne", "Mann", "Kvinne", "Kvinne", "Kvinne", "Kvinne", "Mann", "Mann", "Kvinne", "Mann", "Kvinne", "Mann", "Kvinne", "Kvinne", "Kvinne", "Kvinne", "Kvinne")
+data_post_emne2$Gender <-c("Kvinne", "Mann", "Mann", "Mann", "Mann", "Kvinne", "Kvinne", "Kvinne", "Kvinne", "Kvinne", "Kvinne", "Kvinne", "Kvinne", "Mann", "Kvinne", "Kvinne", "Mann", "Mann", "Kvinne", "Kvinne")
+
+# bytt rekkefølgen på noen items
+data_pre_emne2  <- data_pre_emne2  %>% relocate(Trinn,      .before = "Subject")
+data_post_emne2 <- data_post_emne2 %>% relocate(Trinn,      .before = "Subject")
+data_pre_emne2  <- data_pre_emne2  %>% relocate(Gender,     .after  = "Subject")
+data_post_emne2 <- data_post_emne2 %>% relocate(Gender,     .after  = "Subject")
+data_pre_emne2  <- data_pre_emne2  %>% relocate(andre_fag,  .before = "Subject")
+data_post_emne2 <- data_post_emne2 %>% relocate(andre_fag,  .before = "Subject")
+data_pre_emne2  <- data_pre_emne2  %>% relocate(studiested, .before = "Trinn")
+data_post_emne2 <- data_post_emne2 %>% relocate(studiested, .before = "Trinn")
+data_pre_emne2  <- data_pre_emne2  %>% relocate(semester,   .before = "Trinn")
+data_post_emne2 <- data_post_emne2 %>% relocate(semester,   .before = "Trinn")
+data_pre_emne2  <- data_pre_emne2  %>% relocate(full_treatment, .before = "studiested")
+data_post_emne2 <- data_post_emne2 %>% relocate(full_treatment, .before = "studiested")
+```
+
+Rename items
+
+```{r emne2_rename_items}
+ren_emne2 <- function (data_name) {rename(data_name, 
+"TK2"    = "Jeg kan lage algoritmer ved hjelp av penn og papir og flytdiagrammer.",
+"TK1"    = "Jeg kan bruke algoritmisk tenkning når jeg lager dataprogrammer.",
+"TK3"    = "Jeg kan lage dataprogrammer ved å bruke blokk-basert programmering.",        
+"TK4"    = "Jeg kan lage dataprogrammer ved å bruke tekst-basert programmering.",             
+"TK5"    = "Jeg kan bruke variabler, når jeg programmerer.",
+"TK6"    = "Jeg kan bruke vilkår (if, else), når jeg programmerer.",
+"TK7"    = "Jeg kan bruke løkker (for, while), når jeg programmerer.",
+"TK8"    = "Jeg kan lage dataprogrammer som inneholder funksjoner.",
+"CK1"    = "Jeg forstår det faglige innholdet i naturfag som handler om energi og materie.",                                        
+"CK2"    = "Jeg forstår det faglige innholdet i naturfag som handler om jorda og livet på jorda.",                                         
+"CK3"    = "Jeg forstår det faglige innholdet i naturfag som handler om kropp og helse.",                                          
+"CK4"    = "Jeg forstår hvordan ulike typer teknologi virker, f.eks. hvordan teknologi kan styres med dataprogrammer.",
+"CK5"    = "Jeg har kunnskaper om naturvitenskapelige tenkemåter og praksiser.",                                        
+"TCK1"   = "Jeg kan programmere enkle teknologiske systemer som består av deler som virker sammen.",                                             
+"TCK2"   = "Jeg kan bruke programmering og modellering til å utvikle roboter og mikrokontrollere, f.eks. micro:bit.",
+"TCK3"   = "Jeg kan utvikle modeller av naturfaglige fenomener ved hjelp av programmering og simulering.",                                              
+"TCK4"   = "Jeg kan bruke programmering til å utforske naturfaglige fenomener.",                                                   
+"PK1"    = "Jeg vet hvordan jeg kan støtte elevenes refleksjon over egen læring.",                                              
+"PK2"    = "Jeg vet hvordan jeg kan støtte elevene emosjonelt i deres læring, f.eks deres utholdenhet, evne til å arbeide målrettet og takle motgang.",
+"PK3"    = "Jeg vet hvordan jeg kan støtte elevene sosialt i deres læring, f.eks deres nysgjerrighet, kreativitet og samarbeidsevne.",
+"PK4"    = "Jeg vet hvordan jeg kan tilpasse utfordringer til elevene.",
+"PK5"    = "Jeg vet hvordan jeg kan tilpasse tilbakemeldinger til elevene.",
+"PK6"    = "Jeg vet hvordan jeg kan bruke vurdering til å støtte elevene underveis i deres læring.",
+"TPK1"   = "Jeg vet hvordan jeg kan bruke programmering på måter som beriker undervisningen.",                                          
+"TPK2"   = "Jeg vet hvordan jeg kan bruke programmering på måter som støtter elevenes læring.",                                           
+"TPK3"   = "Jeg vet hvordan jeg kan tilpasse bruken av programmering til forskjellige undervisningsaktiviteter.",
+"TPK4"   = "Jeg tenker kritisk gjennom hvordan jeg kan bruke programmering i undervisningen.",                                          
+"TPK5"   = "Jeg vet hvordan jeg kan legge frem programmering på måter som motiverer for læring.",                                             
+"PCK1"   = "Jeg vet hvordan jeg kan legge til rette for elevenes begrepslæring i naturfag.",                                        
+"PCK2"   = "Jeg vet hvordan jeg kan støtte elevene i å utvikle kritisk tenkning og argumentasjonsferdigheter i naturfag.",
+"PCK3"   = "Jeg vet hvordan jeg kan bruke utforskende arbeidsmåter i naturfag.",                                                           
+"PCK4"   = "Jeg vet hvordan jeg kan bruke modellering som arbeidsmåte i naturfag.",                                    
+"PCK5"   = "Jeg vet hvordan jeg kan vurdere elevenes måloppnåelse på ulike måter i naturfag.",                                            
+"TPACK1" = "Jeg vet hvordan jeg kan motivere elevene for å bruke programmering til å lære naturfag, for eksempel ved å knytte det til elevenes interesser eller vise hvilken relevans det kan ha.",
+"TPACK2" = "Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå mer av naturfaglige begreper og fenomener.",
+"TPACK3" = "Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå mer av naturvitenskapelige tenkemåter og praksiser.",                                     
+"TPACK4" = "Jeg vet hvordan jeg kan bruke programmering i naturfag til å lære elever hvordan teknologiske systemer fungerer.",
+"TPACK5" = "Jeg føler meg klar til å bruke programmering til å undervise naturfag",
+"XK1"    = "Jeg vet hvilke utfordringer elevene kan ha når de programmerer.",
+"XK2"    = "Jeg har kunnskap om elevers ulike forutsetninger for å ta i bruk programmering.",                                          
+"XK3"    = "Jeg vet hvordan programmering i naturfag bidrar til skolens samfunnsoppdrag.",                                               
+"XK4"    = "Jeg kan være en ressursperson for kollegaene i skolen når det gjelder programmering i naturfag",
+"XK5"    = "Jeg vet hvordan bruk av programmeringsteknologi kan påvirke læringssituasjonen i klasserommet.",
+"XK6"    = "Jeg vet hvordan jeg kan lede en undervisningssituasjon i et klasserom der programmering er en sentral aktivitet.")
+ }
+
+data_pre_emne2_ren  <- ren_emne2(data_pre_emne2)
+data_post_emne2_ren <- ren_emne2(data_post_emne2)
+
+# select relevant questions
+data_pre_emne2_ren  <- data_pre_emne2_ren [,5:55]
+data_post_emne2_ren <- data_post_emne2_ren[,5:55]
+```
+
+One of the respondents chose the wrong institution, because their institution lacked amongst the options available.
+This was later corrected in the online questionnaire.
+I therefore have to change the affiliation of row 12 (NR 21571938) from Institute4 to Institute5, and "hovedfag" from Matematikk to Naturfag.
+(This is documented through the e-post correspondence with Institute5, which show the time this happened.)
+
+```{r pilot_correct_entry}
+# row number was found with View(data_pilot)
+# column number was found with colnames(data_pilot)
+data_pilot[12,50] <- "Institute5"
+data_pilot[12,52] <- "Naturfag"
+data_pilot[2,50]  <- "Institute7"  # Institute8 deltok ikke
+```
+
+```{r pilot_cleaning}
+# rename columns
+data_pilot <- rename(data_pilot,
+                             "andre_fag" = 
+                             "Hva er ditt hovedfag?")
+
+data_pilot <- rename(data_pilot,
+                             "Trinn" =
+                             "Går du på grunnskolelærerutdanning for trinn 1-7 eller trinn 5-10 eller tar du videreutdanning?")
+
+data_pilot <- rename(data_pilot,
+                             "studiested" = 
+                             "Hvilket studiested studerer du ved?")
+
+# add semester
+data_pilot$semester                                        <- rep(0,16)
+data_pilot$semester[data_pilot$studiested == "Institute6"] <- 6 
+data_pilot$semester[data_pilot$studiested == "Institute4"] <- 4 
+# Institute7/8, ?. semester 2/3?
+data_pilot$semester[data_pilot$studiested == "Institute5"] <- 8
+
+# reorder some columns
+data_pilot <- data_pilot %>% relocate(andre_fag,  .before = "Jeg kan bruke algoritmisk tenkning når jeg lager kode.")
+data_pilot <- data_pilot %>% relocate(Trinn,      .before = "andre_fag")
+data_pilot <- data_pilot %>% relocate(studiested, .before = "Trinn")
+data_pilot <- data_pilot %>% relocate(semester,   .before = "Trinn")
+```
+
+```{r pilot_rename}
+data_pilot_ren <- data_pilot %>% rename(
+"TK1"    = "Jeg kan bruke algoritmisk tenkning når jeg lager kode.",
+"TK2"    = "Jeg kan lage algoritmer ved hjelp av penn og papir og flytdiagrammer.",
+"TK3"    = "Jeg kan lage kode ved å bruke blokk-basert programmering.",        
+"TK4"    = "Jeg kan lage kode ved å bruke tekst-basert programmering.",             
+"TK5"    = "Jeg kan bruke variabler, når jeg programmerer.",
+"TK6"    = "Jeg kan bruke vilkår (if, else), når jeg programmerer.",
+"TK7"    = "Jeg kan bruke løkker (for, while), når jeg programmerer.",
+"TK8"    = "Jeg kan lage dataprogrammer som inneholder funksjoner.
+(Funksjoner kan for eksempel være noen linjer med kode som har blitt skilt ut som en egen algoritme som kan gjenbrukes, f.eks.
+fordi den beskriver noe som gjøres ofte i programmet.)",
+"CK1"    = "Jeg forstår det faglige innholdet i naturfag som handler om energi og materie.",                                        
+"CK2"    = "Jeg forstår det faglige innholdet i naturfag som handler om jorda og livet på jorda.",                                         
+"CK3"    = "Jeg forstår det faglige innholdet i naturfag som handler om kropp og helse.",                                          
+"CK4"    = "Jeg forstår hvordan ulike typer teknologi virker, f.eks. hvordan teknologi kan styres med dataprogrammer.",
+"CK5"    = "Jeg har kunnskaper om naturvitenskapelige tenkemåter og praksiser.",                                        
+"TCK1"   = "Jeg kan programmere enkle teknologiske systemer som består av deler som virker sammen.",                                             
+"TCK2"   = "Jeg kan bruke programmering og modellering til å utvikle roboter og mikrokontrollere, f.eks. micro:bit.",
+"TCK3"   = "Jeg kan utvikle modeller av naturfaglige fenomener ved hjelp av programmering og simulering.",                                              
+"TCK4"   = "Jeg kan bruke programmering til å utforske naturfaglige fenomener.",                                                   
+"PK1"    = "Jeg vet hvordan jeg kan støtte elevenes refleksjon over egen læring.",                                              
+"PK2"    = "Jeg vet hvordan jeg kan støtte elevene emosjonelt i deres læring, f.eks deres utholdenhet, evne til å arbeide målrettet og takle motgang.",
+"PK3"    = "Jeg vet hvordan jeg kan støtte elevene sosialt i deres læring, f.eks deres nysgjerrighet, kreativitet og samarbeidsevne.",
+"PK4"    = "Jeg vet hvordan jeg kan tilpasse utfordringer til elevene.",
+"PK5"    = "Jeg vet hvordan jeg kan tilpasse tilbakemeldinger til elevene.",
+"PK6"    = "Jeg vet hvordan jeg kan bruke vurdering til å støtte elevene underveis i deres læring.",
+"TPK1"   = "Jeg vet hvordan jeg kan bruke programmering på måter som beriker undervisningen.",                                          
+"TPK2"   = "Jeg vet hvordan jeg kan bruke programmering på måter som støtter elevenes læring.",                                           
+"TPK3"   = "Jeg vet hvordan jeg kan tilpasse bruken av programmering til forskjellige undervisningsaktiviteter.",
+"TPK4"   = "Jeg tenker kritisk gjennom hvordan jeg kan bruke programmering i undervisningen.",                                          
+"TPK5"   = "Jeg vet hvordan jeg kan legge frem programmering på måter som motiverer for læring.",                                             
+"PCK1"   = "Jeg vet hvordan jeg kan legge til rette for elevenes begrepslæring i naturfag.",                                        
+"PCK2"   = "Jeg vet hvordan jeg kan støtte elevene i å utvikle kritisk tenkning og argumentasjonsferdigheter i naturfag.",
+"PCK3"   = "Jeg vet hvordan jeg kan bruke utforskende arbeidsmåter i naturfag.",                                                           
+"PCK4"   = "Jeg vet hvordan jeg kan bruke modellering som arbeidsmåte i naturfag.",                                    
+"PCK5"   = "Jeg vet hvordan jeg kan vurdere elevenes måloppnåelse på ulike måter i naturfag.",                                            
+"TPACK1" = "Jeg vet hvordan jeg kan motivere elevene for å bruke programmering til å lære naturfag, for eksempel ved å knytte det til elevenes interesser eller vise hvilken relevans det kan ha.",
+"TPACK2" = "Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå mer av naturfaglige begreper og fenomener.",
+"TPACK3" = "Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå mer av naturvitenskapelige tenkemåter og praksiser.",                                     
+"TPACK4" = "Jeg vet hvordan jeg kan bruke programmering i naturfag til å lære elever hvordan teknologiske systemer fungerer.",
+"TPACK5" = "Jeg føler meg klar til å bruke programmering til å undervise naturfag",
+"XK1"    = "Jeg vet hvilke utfordringer elevene kan ha når de programmerer.",
+"XK2"    = "Jeg har kunnskap om elevers ulike forutsetninger for å ta i bruk programmering.",                                          
+"XK3"    = "Jeg vet hvordan programmering i naturfag bidrar til skolens samfunnsoppdrag.",                                               
+"XK4"    = "Jeg kan være en ressursperson for kollegaene i skolen når det gjelder programmering i naturfag",
+"XK5"    = "Jeg vet hvordan bruk av programmeringsteknologi kan påvirke læringssituasjonen i klasserommet.",
+"XK6"    = "Jeg vet hvordan jeg kan lede en undervisningssituasjon i et klasserom der programmering er en sentral aktivitet.")
+
+# select TPACK-questions
+data_pilot_ren <- data_pilot_ren[,3:50]
+```
+
+The order of the two first questions was switched between emne2 and the pilot. Correct this. Use pilot as correct order.
+
+* TK1: Jeg kan bruke algoritmisk tenkning når jeg lager kode.
+* TK2: Jeg kan lage algoritmer ved hjelp av penn og papir og flytdiagrammer.
+
+```{r emne2_switch_order_TK1_2}
+data_pre_emne2_ren  <- data_pre_emne2_ren  %>% relocate(TK2, .after = TK1)
+data_post_emne2_ren <- data_post_emne2_ren %>% relocate(TK2, .after = TK1)
+```
+
+```{r pilot_trelis_q_cleaning}
+# keep some columns
+data_trelis_q <- rename(data_trelis_q,
+                            "Trinn" =
+                            "80. Går du på grunnskolelærerutdanning for trinn 1-7 eller trinn 5-10?")
+data_trelis_q <- rename(data_trelis_q,
+                            "Gender" =
+                            "82. Er du:")
+data_trelis_q <- rename(data_trelis_q,
+                            "studiested" =
+                            "79. Hvilket studiested studerer du ved?")
+data_trelis_q <- rename(data_trelis_q,
+                            "semester" =
+                            "81. Hvilket studieår er du i?")
+
+# change from year to semester
+data_trelis_q$semester <- rep(7, 52)  # 4 år =  7. semester (høst)
+
+# reorder some columns
+data_trelis_q <- data_trelis_q %>% relocate(Trinn,      .before = "58. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
+data_trelis_q <- data_trelis_q %>% relocate(Gender,     .before = "58. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
+data_trelis_q <- data_trelis_q %>% relocate(studiested, .before = "Trinn")
+data_trelis_q <- data_trelis_q %>% relocate(semester,   .before = "Trinn")
+```
+
+```{r pilot_trelis_q_rename}
+# rename columns: (bruker dplyr)
+# kilde: https://www.sharpsightlabs.com/blog/rename-columns-in-r/
+data_trelis_q_ren <- rename(data_trelis_q, 
+'CK123'  = '58. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.',
+"CK6"    = "59. Jeg har kunnskap om hvordan forskere i naturvitenskap arbeider",
+'CK5'    = '60. Jeg har kunnskaper om naturvitenskapelige tenkemåter og praksiser',
+'PCK1'   = '61. Jeg vet hvordan jeg kan legge til rette for elevenes begrepslæring i naturfag',
+'PCK2'   = '62. Jeg vet hvordan jeg kan støtte elevene i å utvikle kritisk tenkning og argumentasjonsferdigheter i naturfag',
+'TK2'    = '63. Jeg kan lage algoritmer ved hjelp av penn og papir og flytdiagrammer.',
+'TK5'    = '64. Jeg kan bruke variabler når jeg programmerer',
+'TK6'    = '65. Jeg kan bruke vilkår (if, else), når jeg programmerer',
+'TCK3'   = '66. Jeg kan utvikle modeller av naturfaglige fenomener ved hjelp av programmering og simulering',
+'TCK4'   = '67. Jeg kan bruke programmering til å utforske naturfaglige fenomener',
+'PK3'    = '68. Jeg kan støtte elevene sosialt i deres læring, f.eks deres nysgjerrighet, kreativitet og samarbeidsevne',
+'PK1'    = '69. Jeg vet hvordan jeg kan støtte elevenes refleksjon over egen læring',
+'PK6'    = '70. Jeg kan bruke vurdering underveis til å støtte elevene i deres læring',
+'PCK4'   = '71. Jeg vet hvordan jeg kan bruke modellering som arbeidsmåte i naturfag',
+'TPK1'   = '72. Jeg vet hvordan jeg kan bruke programmering på måter som beriker undervisningen',
+'TPK4'   = '73. Jeg tenker kritisk gjennom hvordan jeg kan bruke programmering i undervisningen',
+'TPK5'   = '74. Jeg vet hvordan jeg kan legge frem programmering på måter som motiverer for læring',
+'TPACK2' = '75. Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå mer av naturfaglige begreper og fenomener',
+'TPACK3' = '76. Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå naturvitenskapelige praksiser og tenkemåter.',
+'TPACK4' = '77. Jeg vet hvordan jeg kan bruke programmering i naturfag til å lære elever hvordan teknologiske systemer fungerer.')
+
+# remove blank rows / NA
+# kilde: https://www.statology.org/remove-rows-in-r/
+# select columns:
+# kilde: https://www.statology.org/r-select-columns-by-index/
+data_trelis_q_ren        <- data_trelis_q_ren[ , 54:77]
+
+# select rows: (fjern to nederste rader med gjennomsnitt og standardavvik)
+data_trelis_q_ren        <- data_trelis_q_ren[ 1:48, ]
+
+# convert into numbers
+data_trelis_q_ren[,5:24] <- apply(data_trelis_q_ren[,5:24], 2,
+                    function(x) as.numeric(as.character(x)))
+```
+
+```{r merge_data_pilot_emne2}
+# begin with text-values: spring 2022
+data_prepost_emne2_ren <- bind_rows(data_pre_emne2_ren,
+                                    data_post_emne2_ren)
+data_pilot_emne2_ren   <- bind_rows(data_prepost_emne2_ren, data_pilot_ren)
+```
+
+```{r convert_likert_to_numbers}
+# data_list (enig = 1, osv.)
+data_pilot_emne2_ren[data_pilot_emne2_ren == 'Veldig uenig'] <- "1"
+data_pilot_emne2_ren[data_pilot_emne2_ren == 'Veldig enig']  <- "5"
+data_pilot_emne2_ren[data_pilot_emne2_ren == 'Hverken enig eller uenig'] <- "3"
+data_pilot_emne2_ren[data_pilot_emne2_ren == 'Uenig']        <- "2"
+data_pilot_emne2_ren[data_pilot_emne2_ren == 'Enig']         <- "4"
+
+# turn values into integer from character
+# kilde: https://statisticsglobe.com/convert-data-frame-column-to-numeric-in-r
+data_pilot_emne2_ren[,8:ncol(data_pilot_emne2_ren)] <- apply(
+                    data_pilot_emne2_ren[,8:ncol(data_pilot_emne2_ren)], 2,
+                    function(x) as.numeric(as.character(x)))
+```
+
+```{r merge_all_datasets}
+data_TPACK_pilot <- bind_rows(data_pilot_emne2_ren, data_trelis_q_ren)
+```
+
+```{r identify_dataset_in_dataframe}
+datasett <- c(replicate(21, "pre"),
+              replicate(20, "post"),
+              replicate(16, "pilot"),
+              replicate(48, "trelis_q"))
+
+data_TPACK_pilot <- cbind(datasett, data_TPACK_pilot)
+```
+
+##### Save reduced datasets
+
+Size of dataset
+
+* emne2_pre:   `r dim(data_pre_emne2_ren)`
+* emne2_post:  `r dim(data_post_emne2_ren)`
+* pilot:       `r dim(data_pilot_ren)`
+* trelis_q:    `r dim(data_trelis_q_ren)`
+
+```{r save_data}
+save(data_TPACK_pilot, file="../data/processed_data/tpack-pilot_2022.Rdata")
+```
 
@@
 
+From stage 3 there are three steps which will be discussed: step 7 (tests of dimensionality, i.e., CFA), step 8 (tests of reliability) and step 9 (tests of validity).
+
 ## Step 7: Tests of dimensionality (i.e., CFA)

### 2024-09-02 Rmd/Paper1.Rmd
Added loading of TRELIS-Q data

@@
 
-### Reducing the raw data ... (transforming from xlsx/ods to Rdata)
-
 The raw data were downloaded from nettskjema.no in the xlsx-format and transformed to the Rdata-format in Rstudio.
@@
 
-* Step 4a: convert from ods to Rdata; rename items using TPACK typology; treat missing
-
-#### pilot
-
 The aim is to create the following dataset:
@@
 
-df[1:21,] =   emne2_pre        (n=21)  
-df[22:41,] =  emne2_post       (n=20)  
-df[42:57,] =  pilot            (n=16)  
-df[58:105,] = trelis_q pilot   (n=48)
+* df[1:21,] =   emne2_pre        (n=21)  
+* df[22:41,] =  emne2_post       (n=20)  
+* df[42:57,] =  pilot            (n=16)  
+* df[58:105,] = pilot trelis_q   (n=48)
 
-#### Load datasets
+### Load datasets
 
@@
 ```{r load_pilot_trelis_q}
+folder_pilot_trelis_q <- "../data/rawdata/"
+pilot_trelis_q        <- "Datafil-utproving-trelis-q-institute4-institute1-aug2022.xlsx"
+
+path_pilot_trelis_q   <- paste(folder_pilot_trelis_q, pilot_trelis_q, sep='')
+
+data_pilot_trelis_q   <- read_excel(path_pilot_trelis_q, sheet=1)
+```
+
+```{r load_TRELIS_Q}
 folder_trelis_q       <- "../data/rawdata/"
-trelis_q        <- "Datafil-utproving-trelis-q-institute4-institute1-aug2022.xlsx"
+trelis_q              <- "data-trelis-q-tatt-ut-2feb2023.xlsx"
 
@@
 
-data_trelis_q   <- read_excel(path_trelis_q, sheet=1)
+data_trelis_q         <- read_excel(path_trelis_q) #, sheet=2)
 ```
 
-##### Data cleaning
+### Data cleaning
 
 Rename items using the term "coding" so they use "programming" instead to enable merging of the datasets.
+
+#### emne2
 
@@
 
+The order of the two first questions was switched between emne2 and the pilot. Correct this. Use pilot as correct order.
+
+* TK1: Jeg kan bruke algoritmisk tenkning når jeg lager kode.
+* TK2: Jeg kan lage algoritmer ved hjelp av penn og papir og flytdiagrammer.
+
+```{r emne2_switch_order_TK1_2}
+data_pre_emne2_ren  <- data_pre_emne2_ren  %>% relocate(TK2, .after = TK1)
+data_post_emne2_ren <- data_post_emne2_ren %>% relocate(TK2, .after = TK1)
+```
+
+#### pilot
+
 One of the respondents chose the wrong institution, because their institution lacked amongst the options available.
@@
 
-The order of the two first questions was switched between emne2 and the pilot. Correct this. Use pilot as correct order.
-
-* TK1: Jeg kan bruke algoritmisk tenkning når jeg lager kode.
-* TK2: Jeg kan lage algoritmer ved hjelp av penn og papir og flytdiagrammer.
-
-```{r emne2_switch_order_TK1_2}
-data_pre_emne2_ren  <- data_pre_emne2_ren  %>% relocate(TK2, .after = TK1)
-data_post_emne2_ren <- data_post_emne2_ren %>% relocate(TK2, .after = TK1)
-```
+#### pilot trelis_q
 
@@
 # keep some columns
-data_trelis_q <- rename(data_trelis_q,
+data_pilot_trelis_q <- rename(data_pilot_trelis_q,
                             "Trinn" =
                             "80. Går du på grunnskolelærerutdanning for trinn 1-7 eller trinn 5-10?")
-data_trelis_q <- rename(data_trelis_q,
+data_pilot_trelis_q <- rename(data_pilot_trelis_q,
                             "Gender" =
                             "82. Er du:")
-data_trelis_q <- rename(data_trelis_q,
+data_pilot_trelis_q <- rename(data_pilot_trelis_q,
                             "studiested" =
                             "79. Hvilket studiested studerer du ved?")
-data_trelis_q <- rename(data_trelis_q,
+data_pilot_trelis_q <- rename(data_pilot_trelis_q,
                             "semester" =
@@
 # change from year to semester
-data_trelis_q$semester <- rep(7, 52)  # 4 år =  7. semester (høst)
+data_pilot_trelis_q$semester <- rep(7, 52)  # 4 år =  7. semester (høst)
 
 # reorder some columns
-data_trelis_q <- data_trelis_q %>% relocate(Trinn,      .before = "58. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
-data_trelis_q <- data_trelis_q %>% relocate(Gender,     .before = "58. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
-data_trelis_q <- data_trelis_q %>% relocate(studiested, .before = "Trinn")
-data_trelis_q <- data_trelis_q %>% relocate(semester,   .before = "Trinn")
+data_pilot_trelis_q <- data_pilot_trelis_q %>% relocate(Trinn,      .before = "58.
+Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
+data_pilot_trelis_q <- data_pilot_trelis_q %>% relocate(Gender,     .before = "58.
+Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
+data_pilot_trelis_q <- data_pilot_trelis_q %>% relocate(studiested, .before = "Trinn")
+data_pilot_trelis_q <- data_pilot_trelis_q %>% relocate(semester,   .before = "Trinn")
+
+# remove blank rows / NA
+# kilde: https://www.statology.org/remove-rows-in-r/
+# select columns:
+# kilde: https://www.statology.org/r-select-columns-by-index/
+data_pilot_trelis_q        <- data_pilot_trelis_q[ , 54:77]
+
+# select rows: (fjern to nederste rader med gjennomsnitt og standardavvik)
+data_pilot_trelis_q        <- data_pilot_trelis_q[ 1:48, ]
+
+# convert into numbers
+data_pilot_trelis_q[,5:24] <- apply(data_pilot_trelis_q[,5:24], 2,
+                    function(x) as.numeric(as.character(x)))
 ```
@@
 # kilde: https://www.sharpsightlabs.com/blog/rename-columns-in-r/
-data_trelis_q_ren <- rename(data_trelis_q, 
+data_pilot_trelis_q_ren <- rename(data_pilot_trelis_q, 
 'CK123'  = '58. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.',
@@
 'TPACK4' = '77. Jeg vet hvordan jeg kan bruke programmering i naturfag til å lære elever hvordan teknologiske systemer fungerer.')
+```
 
-# remove blank rows / NA
-# kilde: https://www.statology.org/remove-rows-in-r/
-# select columns:
-# kilde: https://www.statology.org/r-select-columns-by-index/
-data_trelis_q_ren        <- data_trelis_q_ren[ , 54:77]
+#### TRELIS-Q
 
-# select rows: (fjern to nederste rader med gjennomsnitt og standardavvik)
-data_trelis_q_ren        <- data_trelis_q_ren[ 1:48, ]
+```{r TRELIS_Q_cleaning}
+# only keep relevant columns
+data_trelis_q <- data_trelis_q[,51:74]
+
+# keep some columns; covariates
+# 68. Hvilket studiested studerer du ved?	
+# 69. Går du på grunnskolelærerutdanning for trinn 1-7 eller trinn 5-10?	
+# 70. Hvilket studieår er du i?	
+# 71. Er du:
+
+# likert-spørsmål: 48 - 66
+# åpne spørsmål: 67
+
+data_trelis_q <- rename(data_trelis_q,
+                            "Trinn" =
+                            "69. Går du på grunnskolelærerutdanning for trinn 1-7 eller trinn 5-10?")
+data_trelis_q <- rename(data_trelis_q,
+                            "Gender" =
+                            "71. Er du:")
+data_trelis_q <- rename(data_trelis_q,
+                            "studiested" =
+                            "68. Hvilket studiested studerer du ved?")
+data_trelis_q <- rename(data_trelis_q,
+                            "semester" =
+                            "70. Hvilket studieår er du i?")
+
+# change from year to semester
+#data_trelis_q$semester <- rep(7, 52)  # 4 år =  7. semester (høst)
+
+# reorder some columns
+data_trelis_q <- data_trelis_q %>% relocate(Trinn, .before = "48. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
+data_trelis_q <- data_trelis_q %>% relocate(Gender, .before = "48. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
+data_trelis_q <- data_trelis_q %>% relocate(studiested, .before = "Trinn")
+data_trelis_q <- data_trelis_q %>% relocate(semester, .before = "Trinn")
+```
+
+```{r TRELIS_Q_rename}
+# rename columns: (bruker dplyr)
+# kilde: https://www.sharpsightlabs.com/blog/rename-columns-in-r/
+data_trelis_q_ren <- rename(data_trelis_q, 
+'CK123'  = '48. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.',
+'CK4'    = "49. Jeg forstår teknologiske prinsipper og virkemåter som er relevante for naturfag",
+'CK5'    = '50. Jeg har kunnskaper om naturvitenskapelige praksiser og tenkemåter',
+'PCK2'   = '51. Jeg vet hvordan jeg kan støtte elevene i å utvikle kritisk tenkning og argumentasjonsferdigheter i naturfag',
+'TK2'    = "52. Jeg kan sette opp en algoritme når jeg programmerer",
+'TK1a'   = "53. Jeg kan bryte ned problemer i mindre deler når jeg programmerer",
+'TK34'   = "54. Jeg har kunnskap om forskjellige programmeringsverktøy, f.eks. Makecode for Micro:bit, Scratch, Python, og lignende",
+'TK1b'   = "55. Jeg kan oppdage og rette feil, når jeg programmerer.",
+'TK567'  = "56. Jeg vet hvordan jeg lager kode som inneholder f.eks. variabler, vilkår (if, else) og løkker (for, while)",
+'TCK1'   = "58. Jeg kan bruke programmering til å utvikle enkle teknologiske systemer som består av deler som virker sammen.",
+'TCK4'   = '57. Jeg kan bruke programmering til å utforske naturfaglige fenomener',
+'PK3'    = '59. Jeg kan støtte elevene sosialt i deres læring i ulike fag, f.eks deres nysgjerrighet, kreativitet og samarbeidsevne',
+'PK1'    = '60. Jeg vet hvordan jeg kan støtte elevenes refleksjon over egen læring i ulike fag',
+'PCK5'   = '61. Jeg vet hvordan jeg kan vurdere elevenes læring på ulike måter i naturfag',
+'TPK1'   = '62. Jeg vet hvordan jeg kan bruke programmering på måter som beriker undervisningen i ulike fag',
+'TPK4'   = '63. Jeg tenker kritisk igjennom hvordan jeg kan bruke programmering i undervisningen i ulike fag',
+'TPACK2' = '64. Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå naturfaglige begreper og fenomener',
+'TPACK3' = '65. Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå naturvitenskapelige praksiser og tenkemåter.',
+'TPACK4' = '66. Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå hvordan teknologiske systemer fungerer')
+
+# change "5 Stemmer veldig godt" and "1 Stemmer veldig dårlig" to 5 and 1
+data_trelis_q_ren[data_trelis_q_ren == '1 Stemmer veldig dårlig'] <- "1"
+data_trelis_q_ren[data_trelis_q_ren == '5 Stemmer veldig godt']   <- "5"
+
+# hva gjør jeg med 9 / VET IKKE? Gjør om til NA?
+# gjør om til NA; OBS! Ikke gjør disse NA om til tall med impute, pmm senere
+data_trelis_q_ren[data_trelis_q_ren == 'VET IKKE'] <- NA # "9"
 
 # convert into numbers
-data_trelis_q_ren[,5:24] <- apply(data_trelis_q_ren[,5:24], 2,
+data_trelis_q_ren[,5:23] <-  apply(data_trelis_q_ren[,5:23], 2,
                     function(x) as.numeric(as.character(x)))
 ```
+
+```{r rename_dataset}
+data_TRELIS_Q <- data_trelis_q_ren
+```
+
+#### merge dataframes
 
@@
 ```{r merge_all_datasets}
-data_TPACK_pilot <- bind_rows(data_pilot_emne2_ren, data_trelis_q_ren)
+data_TPACK_pilot <- bind_rows(data_pilot_emne2_ren, data_pilot_trelis_q_ren)
 ```
@@
               replicate(16, "pilot"),
-              replicate(48, "trelis_q"))
+              replicate(48, "pilot_trelis_q"))
 
@@
 
-##### Save reduced datasets
+### Save reduced datasets
 
@@
 * pilot:             `r dim(data_pilot_ren)`
-* trelis_q:    `r dim(data_trelis_q_ren)`
+* pilot_trelis_q:    `r dim(data_pilot_trelis_q_ren)`
 
-```{r save_data}
+```{r save_data_pilot}
 save(data_TPACK_pilot, file="../data/processed_data/tpack-pilot_2022.Rdata")
+```
+
+```{r save_data_TRELIS_Q}
+save(data_TRELIS_Q, 
+     file="../data/processed_data/tpack-trelis_q_paper1_2023.Rdata")
 ```

### 2024-09-02 Rmd/Paper1.Rmd
Added checks of linearity

@@
 library(readxl)                             # read .xlsx-files. Kilde: https://readxl.tidyverse.org/
+
+library(pander)                             # print pretty tables
 library(dplyr)                              # rename columns
+library(tidyr)                              # pivot_longer
+
+library(ggplot2)
+
+# multicollinearity checks
+library("corrplot")                         # multicollinearity 
+library(car)                                # vif                      
+library("olsrr")                            # vif
 
@@
 * data_TPACK_pilot, which uses TK, CK, etc for column names
+* data_TRELIS_Q
+
+data_TPACK_pilot will be used for the EFA, while data_TRELIS_Q will be used for the CFA.
 
@@
 
+In the following we will only consider the pre-datasett, so we remove the post-data, which is to dissimilar from the pilot data.
+
+```{r select_data_for_screening}
+# emne2_anon (uten post)
+data_TPACK_pilot_minuspost <- data_TPACK_pilot[
+  data_TPACK_pilot$datasett != "post" & 
+  data_TPACK_pilot$datasett != "pilot_trelis_q",]
+```
+
 ## Step 5: Item reduction
@@
 Treating the data as interval, requires checking of missing data, linearity (checking for outliers and collinearity, i.e., very strong correlation between variables), and normally distributed residuals, checks which are not necessary to do if the data are treated as ordinal.
+In addition, checks suggested by @izquierdo_exploratory_2014 were done, such as check of eigenvalues, item-test correlation and reliabilities (i.e., Cronbach's alpha).
 
@@
 * Step 5e: check eigenvalues, item-test correlation, reliability [@izquierdo_exploratory_2014]
+
+A list of items is compiled for easier retrieval.
+
+```{r}
+TK_pos <- c("TK1","TK2","TK3","TK4","TK5","TK6","TK7","TK8")
+PK_pos <- c("PK1","PK2","PK3","PK4","PK5","PK6")
+CK_pos <- c("CK1","CK2","CK3","CK4","CK5","CK123","CK6")
+PCK_pos <- c("PCK1","PCK2","PCK3","PCK4","PCK5")
+TCK_pos <- c("TCK1","TCK2","TCK3","TCK4")
+TPK_pos <- c("TPK1","TPK2","TPK3","TPK4","TPK5")
+TPACK_pos <- c("TPACK1","TPACK2","TPACK3","TPACK4","TPACK5")
+
+#TPACK_complete <- c(TK_pos, PCK_full_pos, TPACK_full_pos) # minus XK
+
+TPACK_list <- list(TK=TK_pos,
+                   CK=CK_pos,
+                   PK=PK_pos,
+                   PCK=PCK_pos,
+                   TCK=TCK_pos,
+                   TPK=TPK_pos,
+                   TPACK=TPACK_pos)
+
+# CK6 and CK123 is only used for the TRELIS-Q dataset, not the pilot dataset
+TPACK_list$CK <- TPACK_list$CK[!TPACK_list$CK %in% c("CK6","CK123")]
+```
 
@@
 To support this choice further, we also presented the five alternatives as points on a linear scale in our questionnaire to help the respondents visualise the alternatives as being on an interval scale.
+
+### Linearity?
+
+Here I do checks for linearity, collinearity and multicollinearity.
+
+Through an iterative process, the following items were identified as contributing to high collinearity (r>0.9). They were therefore excluded from the dataset.
+
+```{r remove_collinear_items}
+data_TPACK_pilot_minuspost <- data_TPACK_pilot_minuspost %>%
+    # 25.5.2023: velg bort items som ikke er i TRELIS-Q
+    select(-c("TK8","PK2","PK4","PK5","PK6","PCK1","PCK3","PCK4","TCK2","TCK3",
+              "TPK2","TPK3","TPK5","TPACK1","TPACK5",
+              "TPACK3", # colinear in trelis-q; "TK1","TPK1",
+              "TK6","TK4", "PK1", "PK3"))
+```
+
+```{r linear_scatterplots}
+# scatterplot som korrigerer for overlap mellom datapunkter
+teller <- 0                          # hvor mange plot lager jeg?
+datasett_index <- 1
+start <- 1                          # siste item = datasett
+
+#for (dataset in names(data_list)) {
+  #datasett <- data_TPACK_pilot_minuspost
+  end <- ncol(data_TPACK_pilot_minuspost)
+  i <- 1                               # start posisjon
+  for (var1 in names(TPACK_list)) {           # TPACK-element
+    items <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% TPACK_list[[var1]]]
+    ifelse (length(items)>1, 
+      y <- rowMeans(data_TPACK_pilot_minuspost[, items]), # y-variabel
+      y <- data_TPACK_pilot_minuspost[, items])           # y består av bare ett item
+    i <- i + 1                         # for å unngå duplikater
+    for (var2 in items) {               # x-variabel
+      # scatterplot
+      # kilde: https://ggplot2-book.org/statistical-summaries.html
+      # kilde: https://stackoverflow.com/questions/32004292/r-ggplot-for-loop-aes-variable
+      x <- data_TPACK_pilot_minuspost[,var2]
+      xy <- data.frame(t(rbind(y,x)))
+      colnames(xy) <- c(var1,var2)
+
+      # check for collinearity, r>0.85 / 0.90
+      #print("collinearity: r > 0.85?")
+      r <- cor.test(x,y)
+      #print(r)   # marker verdier større enn 0.85
+
+      # 0.4 er moderat korrelasjon ift Christophersen (2018) kap. 3
+      if (abs(r$estimate)>0.4) { # only plot scatterplots which may be linear?
+        xy_plot <- ggplot(xy, aes_string(var2,var1)) +
+          geom_point() +                            # shape=1,alpha=1/5
+          geom_jitter() +
+          ggtitle(paste("scatterplot", data_TPACK_pilot_minuspost$datasett))
+
+        print(xy_plot)
+        teller <- teller + 1   # count number of scatterplots
+      }
+    }
+  }
+#  datasett_index <- datasett_index + 1
+#}
+print(teller) # antall plot
+```
+
+```{r linear_scatterplots_aggregated}
+teller <- 0                          # hvor mange plot lager jeg?
+datasett_index <- 1
+start <- 1                          # siste item = datasett
+
+#for (dataset in names(data_list)) {
+  #datasett <- data_TPACK_pilot_minuspost
+  end <- ncol(data_TPACK_pilot_minuspost)
+  i <- 1                               # start posisjon
+  for (var1 in names(TPACK_list)) {           # TPACK-element
+    items <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% TPACK_list[[var1]]]
+
+    if (length(items) == 0) {next} # skip empty items
+
+    ifelse (length(items)>1,  # calculate average value
+      y <- rowMeans(data_TPACK_pilot_minuspost[, items]), # y-variabel
+      y <- data_TPACK_pilot_minuspost[, items])
+
+    x <- data_TPACK_pilot_minuspost[, items]
+
+    xy <- data.frame(y,x)
+    colnames(xy) <- c(var1,items)
+
+    # pivot_longer
+    # Warning: Using an external vector in selections was deprecated in tidyselect 1.1.0.
+    # ℹ Please use `all_of()` or `any_of()` instead.
+    xy_long <- pivot_longer(xy,!c(all_of(var1)), names_to="items",values_to = "likert")
+
+    #Warning: `aes_string()` was deprecated in ggplot2 3.0.0.
+    #ℹ Please use tidy evaluation idioms with `aes()`.
+    #ℹ See also `vignette("ggplot2-in-packages")` for more information.
+
+       xy_plot <- ggplot(xy_long, aes_string("likert",var1)) +
+          geom_point() +                            # shape=1,alpha=1/5
+          geom_jitter() +
+          ggtitle(paste("scatterplot", data_TPACK_pilot_minuspost$datasett))
+
+        print(xy_plot)
+        teller <- teller + 1   # count number of scatterplots
+    }
+  #datasett_index <- datasett_index + 1
+  #}
+print(teller) # antall plot
+```
+
+```{r check_collinearity}
+
+### HELP FUNCTIONS
+
+lag_eigenvalue_tabell <- function(ev_list) {
+  no_ev <- length(ev_list) # antall eigenvalues
+  kum_var <- 0
+
+  # gjør om til tabell
+  ev_tabell <- data.frame(matrix(ncol=4,nrow=0))
+  for (i in 1:no_ev) {
+    ev <- ev_list[i]
+    kum_var <- kum_var + ev/no_ev
+    faktor <- cbind(as.integer(i), ev, ev/no_ev, kum_var)
+    #print(faktor)
+    ev_tabell <- rbind(ev_tabell, faktor)
+  }
+
+  colnames(ev_tabell) <- c("Factor","Eigenvalue","Variance","Cummulativ variance")
+
+  pander(format(ev_tabell,digits=2))
+}
+
+collinear <- function (dataset) {
+  ###
+  # Christophersen (2018): r > 0.9?
+  ###
+
+  teller <- 0                          # hvor mange sjekker gjør jeg?
+  start <- 1                           
+  end <- ncol(dataset)
+  i <- 1                               # start posisjon
+  for (var1 in colnames(dataset[start:(end-2)])) {
+    i <- i + 1                         # for å unngå duplikater
+    for (var2 in colnames(dataset[i:(end-1)])) {
+      # check for collinearity, r>0.90
+      r <- cor.test(dataset[,var1], dataset[,var2])
+
+      if (abs(r$estimate)>0.9) { 
+        print("collinearity: r > 0.90")
+        print(c(dataset, var1 ,var2, r$estimate))   # marker verdier større enn 0.90
+
+        teller <- teller + 1   # count number of scatterplots
+      }
+    }
+  }
+
+  if (teller == 0) {
+    print("Ingen kolinearitet oppdaget") # antall plot
+  }
+
+  ###
+  # Flora et al. (2012): eigenvalues <= 0? condition index < 30?
+  ###
+  ev <- eigen(cor(dataset,use="pairwise.complete.obs"), only.values=TRUE)
+  ev_table <- lag_eigenvalue_tabell (ev$values)
+  #ev
+  # beregn condition index: ci < 30
+  ci <- ev$values[1]**2 - ev$values[length(ev$values)]**2 
+  print(c("Condition index(<30?):", ci))
+  return(ev_table)
+}
+
+#datasett <- data_list[[1]]
+# eigenvalue_table
+ev_table <- collinear(data_TPACK_pilot_minuspost[,colnames(data_TPACK_pilot_minuspost) %in% unlist(TPACK_list)])
+ev_table
+```
+
+```{r check_multicollinearity}
+# multiple regression model
+
+# I browse through the items, and remove the current item by subtracting it at the end of the formula
+#for (dataset in names(data_list)) {
+  mc <- NULL      # list of items with multicollinearity
+
+      # kilde: https://www.codingprof.com/3-ways-to-test-for-multicollinearity-in-r-examples/
+      corrplot(cor(data_TPACK_pilot_minuspost[,colnames(data_TPACK_pilot_minuspost) %in% unlist(TPACK_list)]), method = "number", number.cex=0.3) # r bør være under 0.85
+
+      # forenkle navn på dataset for prettyprint
+      df <- data_TPACK_pilot_minuspost
+
+      #items <- unlist(TPACK_list)
+      items <- colnames(df[,colnames(df) %in% unlist(TPACK_list)]) # ikke ha med subset-kolonnen
+
+      # lag tekststreng for uavhengige variable
+      #antall_tpack <- length(grep("TPACK",items))
+      independent_var <- paste("'] ~ df [,'", items[1], sep="") #TK1"
+      for (item in items[2:(length(items))]) { #-antall_tpack)]) { # omit TPACK
+        independent_var <- paste(independent_var, "'] + df [,'", item, sep="")
+      }
+      independent_var <- paste(independent_var, "'] - df  [,'", sep="")
+
+      for (item in items) {
+        model <- lm( as.formula(
+        paste( "df [,'", item, independent_var, item, "']", sep="")))
+
+        print("############################")
+        print("######## NEW ITEM ##########")
+        print("############################")
+
+        print(item)
+
+        # check for multicollinearity: R > 0.9
+        print("Any R>0.9?")
+        mc_items <- model$coefficients[abs(model$coefficients)>0.9]
+        if (length(mc_items) > 0) {
+          print(c("multicollinearitet R > 0.9?", mc_items))
+          print(summary(model))
+          mc <- c(mc, item)
+        }
+
+        #if (vif == 1) {
+          # kilde: https://www.statology.org/multicollinearity-in-r/
+          print("car::vif")
+          print(vif(model))
+
+          print("olsrr::vif")
+          # kilde: https://www.codingprof.com/3-ways-to-test-for-multicollinearity-in-r-examples/
+          print("tolerance bør være over 0.1 og VIF under 10.")
+          print(ols_vif_tol(model)) # tolerance bør være over 0.1 og VIF under 10.
+          #print("CI bør både være under 10 og varians lav")
+          #print(ols_eigen_cindex(model)) # CI bør både være under 10 og varians lav
+        #}
+      }
+
+# print items with multicollinearity
+print(mc)
+#}
+```
+
+No multicollinearity. The items added to mc are due to the intercepts, not the slopes.
 

### 2024-09-02 Rmd/Paper1.Rmd
Cleaned up linearity code

@@
 teller <- 0                          # hvor mange plot lager jeg?
-datasett_index <- 1
-start <- 1                          # siste item = datasett
+#datasett_index <- 1
+#start <- 1                          # siste item = datasett
 
-#for (dataset in names(data_list)) {
-  #datasett <- data_TPACK_pilot_minuspost
 end <- ncol(data_TPACK_pilot_minuspost)
@@
 
-      # check for collinearity, r>0.85 / 0.90
-      #print("collinearity: r > 0.85?")
+    # check for collinearity, r>0.90
     r <- cor.test(x,y)
-      #print(r)   # marker verdier større enn 0.85
 
@@
 }
-#  datasett_index <- datasett_index + 1
-#}
 print(teller) # antall plot
@@
 }
-  #datasett_index <- datasett_index + 1
-  #}
+
 print(teller) # antall plot
@@
 # I browse through the items, and remove the current item by subtracting it at the end of the formula
-#for (dataset in names(data_list)) {
 mc <- NULL      # list of items with multicollinearity
@@
   #print(ols_eigen_cindex(model)) # CI bør både være under 10 og varians lav
-        #}
 }
@@
 print(mc)
-#}
 ```

### 2024-09-02 Rmd/Paper1.Rmd
Added checks for outliers and missing

@@
 library("olsrr")                            # vif
+
+library(psych)                              # for skew
+library(mice)                               # for missing
 
@@
 
+```{r}
+### HELP FUNCTION
+# describe data (mean, sd, skew, curtosis)
+# requires library(moments) for skew
+describe_data <- function (datalist, QUIET=1) {
+
+  im  <- mean(datalist, na.rm=TRUE)
+  isd <- sd(datalist, na.rm=TRUE)
+  isk <- psych::skew(datalist, na.rm=TRUE) # from psych
+  ik  <- kurtosi(datalist, na.rm=TRUE) # from psych
+  mis <- sum(is.na(datalist)) # missing
+  ir  <- max(datalist, na.rm=TRUE) - min(datalist, na.rm=TRUE) # range
+
+  # source: https://stackoverflow.com/questions/34351598/getting-name-of-variable-in-r
+  if (QUIET != 1) {
+    print(sprintf("%7s = %s",   "Item",    deparse(substitute(datalist))))
+    print(sprintf("%7s = %.3f", "mean",    im))
+    print(sprintf("%7s = %.3f", "std.dev", isd))
+    print(sprintf("%7s = %.3f", "skew",    isk))
+    print(sprintf("%7s = %.3f", "kurt",    ik))
+    print(sprintf("%7s = %.3f", "range",   ir))
+    print(sprintf("%7s = %i",   "missing", mis))
+
+    if (isk > 2 | ik > 7) {
+      print("                            Skew or curtosis TOO LARGE.")
+    }
+  }
+
+  # outliers
+  lowerq = quantile(datalist, na.rm=TRUE)[2]
+  upperq = quantile(datalist, na.rm=TRUE)[4]
+  iqr    = upperq - lowerq               #Or use IQR(df)
+  # gang med 3 for extreme outliers, 1.5 for mild outliers
+
+  mild.threshold.upper = (iqr * 3) + upperq
+  mild.threshold.lower = lowerq - (iqr * 3)
+
+  # print outliers
+  # kilde: https://stackoverflow.com/questions/19353116/how-do-i-print-values-in-a-list-that-are-greater-than-a-certain-number-along-wit#19353192
+  outliers_upper <- which(datalist>mild.threshold.upper)
+  outliersu      <- datalist[outliers_upper]
+  outliers_lower <- which(datalist<mild.threshold.lower)
+  outliersl      <- datalist[outliers_lower]
+
+  if (length(outliersu) > 0 | length(outliersl) > 0) {
+    print("outliers upper =")
+    print(outliersu)
+    print("outliers lower =")
+    print(outliersl)
+  }
+
+  #hist (datalist, main=i) # histogram
+  #readline(prompt = "Press enter")
+
+  return(c("mean"=round(im,3), "skew"=round(isk,3), "kurtosi"=round(ik,3)))
+           #round(outliersu,3), round(outliersl,3)))
+}
+
+# skip first column which is list of subsets
+#datasett_index <- 1
+#for (datasett in names(data_list)) {
+  for (i in colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% unlist(TPACK_list)]) {
+    print(paste(i))
+    describe_data(data_TPACK_pilot_minuspost[,i])
+  }
+#  datasett_index <- datasett_index + 1
+#}
+
+print(describe(data_TPACK_pilot_minuspost)) # independent variables, x
+#print(describe(scoreItems(TPACK_list,data_TPACK_pilot_minuspost)$scores)) # dependent variables
+```
+
 ### Normal distribution of residuals?
@@
 
+y = ax + b + restledd
+
+Residuals should have an average value of 0, equal variance (homoscedasticity) and no interdependence (no autocorrelation) (Christophersen, 2018, kap. 7).
+
+```{r residuals_singleitems}
+check_residuals <- function(dataset, items) {
+
+  table_residuals <- NULL
+
+  teller <- 0                          # hvor mange kombinasjoner sjekker jeg?
+  i      <- 1                          # start posisjon
+  end    <- length(items)
+  for (var1 in items[1:(end-1)]) {
+    ifelse (length(items)>1, 
+      y  <- rowMeans(dataset[, items]), # y-variabel
+      y  <- dataset[, items])           # y består av bare ett item
+
+    i    <- i + 1                      # for å unngå duplikater
+    for (var2 in items[i:end]) {       # x-variabel
+      x  <- dataset[,var2]
+      xy <- data.frame(t(rbind(y,x)))
+      colnames(xy) <- c(var1,var2)
+
+      model <- lm( as.formula(
+      paste( "xy [,'", var1, "'] ~ xy [,'", var2, "']", sep="")))
+
+      print(c(var1, "=", var2))
+
+      # check for outliers in residuals
+      descriptive <- (describe_data(model$residuals)) # returns ("mean"=im, "skew"=isk, "kurtosi"=ik, outliersu, outliersl)
+
+      # grense for betydningsfulle outliers / leverage; s. 78 i Christophersen, 2018
+      k          <- length(items) # antall uavhengige variables
+      n          <- nrow(xy)
+      ifelse (k<10, 
+        leverage <- 3*(k+1)/n,
+        leverage <- 2*(k+1)/n
+      )
+      #print(c("betydningsfull grense:", leverage))
+
+      # check for normalfordeling
+      # 1: normalfordeling av residuals?
+      hist(residuals(model), col = "steelblue")
+      # 2: jevn fordeling av varians, homoskedastisitet
+      print(plot(fitted(model), residuals(model)))
+      print(abline(h = 0, lty = 2))
+
+      teller <- teller + 1   # count number of scatterplots
+
+      # check for autocorrelation
+      dw     <- durbinWatsonTest(model)
+      print(dw)
+
+      table_residuals <- rbind(table_residuals, c("var1"=var1, "var2"=var2, 
+                                                  descriptive, 
+                                                  "dw=2"=round(dw$dw,2)))
+    } 
+  }
+  print(teller) # antall plot
+
+  # gjør om til tabell
+  table_residuals         <- data.frame(table_residuals)
+
+  # gjør om til tall
+  table_residuals$mean    <- as.numeric(table_residuals$mean)
+  table_residuals$skew    <- as.numeric(table_residuals$skew)
+  table_residuals$kurtosi <- as.numeric(table_residuals$kurtosi)
+  table_residuals$'dw.2'  <- as.numeric(table_residuals$'dw.2')
+
+  return (table_residuals)
+}
+
+items     <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% unlist(TPACK_list)]
+table_res <- check_residuals(data_TPACK_pilot_minuspost, items)
+
+# print items with large skew og kurtosi or autocollinearity
+pander(table_res[abs(table_res$skew) > 1. | abs(table_res$kurtosi) > 3. | table_res$dw.2 > 2.5 | table_res$dw.2 < 1.5,])
+```
+
+Are there correlating residuals?  
+Autocorrelation: Durbin-Watsons d = 2*(1-r)
+
+If d = 2, then OK  
+If d < 1 eller d > 3, then not OK
+
+If autocorrelation, use robust SE
+
+Autocorrelation can be due to structured data. Then use multilevel-analysis.
+
+```{r distribution_residuals_aggregated}
+teller         <- 0                             # hvor mange plot lager jeg?
+datasett_index <- 1
+#for (dataset in names(data_list)) {
+#  datasett <- data_list[[dataset]]      # access data object
+  for (var1 in names(TPACK_list)) {     # TPACK-element
+    items      <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% TPACK_list[[var1]]]
+    ifelse (length(items)>1,            # ikke gjør analyse hvis bare ett item
+      y        <- rowMeans(data_TPACK_pilot_minuspost[, items]), # y-variabel
+      y        <- 0)
+
+   print(c(var1, "=", items))
+
+   if (!length(y)==1) {                # sjekk for at vi har flere enn et item
+
+      x        <- data_TPACK_pilot_minuspost[, items]
+
+      xy       <- data.frame(y,x)
+      colnames(xy) <- c(var1,items)
+
+      # lag tekststreng til formel y= x1 + x2 + osv.
+      independent_var <- paste("'] ~ xy [,'", items[1], sep="") #TK1"
+      for (item in items[2:(length(items))]) { 
+        independent_var <- paste(independent_var, "'] + xy [,'", item, sep="")
+      }
+      independent_var <- paste(independent_var, "'] - xy  [,'", sep="")
+
+      #for (item in items) {
+        model <- lm( as.formula(
+        paste( "xy [,'", item, independent_var, item, "']", sep="")))
+
+        # check for outliers in residuals
+        print(describe_data(model$residuals))
+
+        # grense for betydningsfulle outliers / leverage; s. 78 i Christophersen, 2018
+        k <- length(items) # antall uavhengige variables
+        n <- nrow(xy)
+        ifelse (k<10, 
+                leverage <- 3*(k+1)/n,
+                leverage <- 2*(k+1)/n
+                )
+         print(c("betydningsfull grense:", leverage))
+
+        # check for normalfordeling
+        # 1: normalfordeling av residuals?
+        print(hist(residuals(model), col = "steelblue"))
+        # 2: jevn fordeling av varians, homoskedastisitet
+        print(plot(fitted(model), residuals(model)))
+        print(abline(h = 0, lty = 2))
+
+        teller <- teller + 1   # count number of scatterplots
+      #}
+    }
+  }
+  #datasett_index <- datasett_index + 1
+#} 
+print(teller) # antall plot
+
+```
+
 ### Missing?
 
-Missing data reduces the size of the dataset used for analysis, which leads to lower power (i.e., larger chances for Type II-error, that is not detecting significant effects, obtaining false negatives).
+Missing data reduces the size of the dataset used for analysis, which leads to lower power (i.e., larger chances for Type II-error, that is, not detecting significant effects, obtaining false negatives).
 There are different advice for how to handle missing data, depending on whether they are missing at random (MAR), completely at random (MCAR) or not at random (NMAR) [@rubin_inference_1975; @schafer_missing_2002].
@@
 
+```{r}
+# ref: https://gabriellajg.github.io/EPSY-579-R-Cookbook-for-SEM/lavaan-lab-12-sem-for-missing-data.html
+md.pattern(data_TPACK_pilot_minuspost)
+```
+
+n = 37  
+9 are missing andre_fag  
+16 are missing subject and gender
+
 ## Step 6: Factor extraction (i.e., EFA)

### 2024-09-02 Rmd/Paper1.Rmd
Final version for the day?

@@
 author: "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
-date: "2024-05-28"
+date: "2024-09-02"
 bibliography: "../data/meta_data/references.bib"
@@
 
+These are the libraries we use for special functions.
+
 ```{r}
@@
 
-Rename items using the term "coding" so they use "programming" instead to enable merging of the datasets.
+Clean the data in the different datasets, so they can be more easily compared.
+Differences in question formulations between datasets are adjusted so they are the same.
+Missing information (covariates) is added when it is known.
 
@@
 Rename items
+
+Rename items using the term "coding" so they use "programming" instead to enable merging of the datasets.
+
+Item labels are given based on the intended TPACK-structure when the items were created.
 

### 2024-09-03 Rmd/Paper1.Rmd
Added data cleaning for TRELIS-Q dataset

@@
 author: "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
-date: "2024-09-02"
+date: "2024-09-03"
 bibliography: "../data/meta_data/references.bib"
@@
 
+```{r}
+# remove open-ended question column
+data_TRELIS_Q <- data_TRELIS_Q[,1:23]
+```
+
+```{r}
+# make tibble into data.frame to treat it similarly to pilot dataset
+data_TRELIS_Q <- data.frame(data_TRELIS_Q)
+```
+
 #### merge dataframes
@@
   data_TPACK_pilot$datasett != "pilot_trelis_q",]
+
 ```
@@
 
-```{r}
+```{r items_pilot}
 TK_pos <- c("TK1","TK2","TK3","TK4","TK5","TK6","TK7","TK8")
 PK_pos <- c("PK1","PK2","PK3","PK4","PK5","PK6")
-CK_pos <- c("CK1","CK2","CK3","CK4","CK5","CK123","CK6")
+CK_pos <- c("CK1","CK2","CK3","CK4","CK5")
 PCK_pos <- c("PCK1","PCK2","PCK3","PCK4","PCK5")
@@
 
-TPACK_list <- list(TK=TK_pos,
+pilot_list <- list(TK=TK_pos,
                    CK=CK_pos,
@@
 
-# CK6 and CK123 is only used for the TRELIS-Q dataset, not the pilot dataset
-TPACK_list$CK <- TPACK_list$CK[!TPACK_list$CK %in% c("CK6","CK123")]
+```
+
+```{r items_TRELIS_Q}
+TK_pos_trelisq <- c("TK1a","TK1b","TK2","TK34","TK567")
+CK_pos_trelisq <- c("CK4","CK5","CK123")
+PCK_pos_trelisq <- c("PCK2","PCK5")
+TCK_pos_trelisq <- c("TCK1","TCK4")
+PK_pos_trelisq <- c("PK1","PK3")
+TPK_pos_trelisq <- c("TPK1","TPK4")
+TPACK_pos_trelisq <- c("TPACK2","TPACK3","TPACK4")
+
+TRELISQ_list <- list(TK=TK_pos_trelisq,
+                     CK=CK_pos_trelisq,
+                     PK=PK_pos_trelisq,
+                     PCK=PCK_pos_trelisq,
+                     TCK=TCK_pos_trelisq,
+                     TPK=TPK_pos_trelisq,
+                     TPACK=TPACK_pos_trelisq)
+
 ```
@@
 
+```{r help_functions_check_collinearity}
+### HELP FUNCTIONS
+
+lag_eigenvalue_tabell <- function(ev_list) {
+  no_ev <- length(ev_list) # antall eigenvalues
+  kum_var <- 0
+
+  # gjør om til tabell
+  ev_tabell <- data.frame(matrix(ncol=4,nrow=0))
+  for (i in 1:no_ev) {
+    ev <- ev_list[i]
+    kum_var <- kum_var + ev/no_ev
+    faktor <- cbind(as.integer(i), ev, ev/no_ev, kum_var)
+    #print(faktor)
+    ev_tabell <- rbind(ev_tabell, faktor)
+  }
+
+  colnames(ev_tabell) <- c("Factor","Eigenvalue","Variance","Cummulativ variance")
+
+  pander(format(ev_tabell,digits=2))
+}
+
+collinear <- function (dataset) {
+  ###
+  # Christophersen (2018): r > 0.9?
+  ###
+
+  teller <- 0                          # hvor mange sjekker gjør jeg?
+  start <- 1                           
+  end <- ncol(dataset)
+  i <- 1                               # start posisjon
+  for (var1 in colnames(dataset[start:(end-2)])) {
+    i <- i + 1                         # for å unngå duplikater
+    for (var2 in colnames(dataset[i:(end-1)])) {
+      # check for collinearity, r>0.90
+      r <- cor.test(dataset[,var1], dataset[,var2])
+
+      if (abs(r$estimate)>0.9) { 
+        print("collinearity: r > 0.90")
+        print(c(dataset, var1 ,var2, r$estimate))   # marker verdier større enn 0.90
+
+        teller <- teller + 1   # count number of scatterplots
+      }
+    }
+  }
+
+  if (teller == 0) {
+    print("Ingen kolinearitet oppdaget") # antall plot
+  }
+
+  ###
+  # Flora et al. (2012): eigenvalues <= 0? condition index < 30?
+  ###
+  ev <- eigen(cor(dataset,use="pairwise.complete.obs"), only.values=TRUE)
+  ev_table <- lag_eigenvalue_tabell (ev$values)
+  #ev
+  # beregn condition index: ci < 30
+  ci <- ev$values[1]**2 - ev$values[length(ev$values)]**2 
+  print(c("Condition index(<30?):", ci))
+  return(ev_table)
+}
+```
+
+#### pilot
+
 Through an iterative process, the following items were identified as contributing to high collinearity (r>0.9). They were therefore excluded from the dataset.
 
-```{r remove_collinear_items}
+```{r remove_collinear_items_pilot}
+### pilot
 data_TPACK_pilot_minuspost <- data_TPACK_pilot_minuspost %>%
@@
               "TK6","TK4", "PK1", "PK3"))
-```
-
-```{r linear_scatterplots}
+
+```
+
+##### Linearity
+
+First I plot scatterplots for indvidiual items against their intended construct (e.g., TK2 vs all TK-items).
+
+```{r linear_scatterplots_pilot}
 # scatterplot som korrigerer for overlap mellom datapunkter
 teller <- 0                          # hvor mange plot lager jeg?
-#datasett_index <- 1
-#start <- 1                          # siste item = datasett
 
@@
 i <- 1                               # start posisjon
-for (var1 in names(TPACK_list)) {           # TPACK-element
-  items <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% TPACK_list[[var1]]]
+for (var1 in names(pilot_list)) {           # TPACK-element
+  items <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% pilot_list[[var1]]]
   ifelse (length(items)>1, 
@@
 
-```{r linear_scatterplots_aggregated}
+Then I plot scatterplots for the average value of each intended construct against all the individual items for each intended construct (e.g., average of all TK-items against individual values of all TK-items).
+
+```{r linear_scatterplots_aggregated_pilot}
 teller <- 0                          # hvor mange plot lager jeg?
-#datasett_index <- 1
-#start <- 1                          # siste item = datasett
 
@@
 i <- 1                               # start posisjon
-for (var1 in names(TPACK_list)) {           # TPACK-element
-  items <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% TPACK_list[[var1]]]
+for (var1 in names(pilot_list)) {           # TPACK-element
+  items <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% pilot_list[[var1]]]
 
@@
 
-```{r check_collinearity}
-
-### HELP FUNCTIONS
-
-lag_eigenvalue_tabell <- function(ev_list) {
-  no_ev <- length(ev_list) # antall eigenvalues
-  kum_var <- 0
-
-  # gjør om til tabell
-  ev_tabell <- data.frame(matrix(ncol=4,nrow=0))
-  for (i in 1:no_ev) {
-    ev <- ev_list[i]
-    kum_var <- kum_var + ev/no_ev
-    faktor <- cbind(as.integer(i), ev, ev/no_ev, kum_var)
-    #print(faktor)
-    ev_tabell <- rbind(ev_tabell, faktor)
-  }
-
-  colnames(ev_tabell) <- c("Factor","Eigenvalue","Variance","Cummulativ variance")
-
-  pander(format(ev_tabell,digits=2))
-}
-
-collinear <- function (dataset) {
-  ###
-  # Christophersen (2018): r > 0.9?
-  ###
-
-  teller <- 0                          # hvor mange sjekker gjør jeg?
-  start <- 1                           
-  end <- ncol(dataset)
-  i <- 1                               # start posisjon
-  for (var1 in colnames(dataset[start:(end-2)])) {
-    i <- i + 1                         # for å unngå duplikater
-    for (var2 in colnames(dataset[i:(end-1)])) {
-      # check for collinearity, r>0.90
-      r <- cor.test(dataset[,var1], dataset[,var2])
-
-      if (abs(r$estimate)>0.9) { 
-        print("collinearity: r > 0.90")
-        print(c(dataset, var1 ,var2, r$estimate))   # marker verdier større enn 0.90
-
-        teller <- teller + 1   # count number of scatterplots
-      }
-    }
-  }
-
-  if (teller == 0) {
-    print("Ingen kolinearitet oppdaget") # antall plot
-  }
-
-  ###
-  # Flora et al. (2012): eigenvalues <= 0? condition index < 30?
-  ###
-  ev <- eigen(cor(dataset,use="pairwise.complete.obs"), only.values=TRUE)
-  ev_table <- lag_eigenvalue_tabell (ev$values)
-  #ev
-  # beregn condition index: ci < 30
-  ci <- ev$values[1]**2 - ev$values[length(ev$values)]**2 
-  print(c("Condition index(<30?):", ci))
-  return(ev_table)
-}
-
-#datasett <- data_list[[1]]
+##### Collinearity
+
+Eigenvalues are examined for possible collinearity. Eigenvalues should be significantly above zero.
+
+```{r collinearity_pilot}
 # eigenvalue_table
-ev_table <- collinear(data_TPACK_pilot_minuspost[,colnames(data_TPACK_pilot_minuspost) %in% unlist(TPACK_list)])
+ev_table <- collinear(data_TPACK_pilot_minuspost[,colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)])
 ev_table
@@
 
-```{r check_multicollinearity}
+##### Multicollinearity
+
+```{r check_multicollinearity_pilot}
 # multiple regression model
@@
 # kilde: https://www.codingprof.com/3-ways-to-test-for-multicollinearity-in-r-examples/
-corrplot(cor(data_TPACK_pilot_minuspost[,colnames(data_TPACK_pilot_minuspost) %in% unlist(TPACK_list)]), method = "number", number.cex=0.3) # r bør være under 0.85
+corrplot(cor(data_TPACK_pilot_minuspost[,colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)]), method = "number", number.cex=0.3) # r bør være under 0.85
 
@@
 
-#items <- unlist(TPACK_list)
-items <- colnames(df[,colnames(df) %in% unlist(TPACK_list)]) # ikke ha med subset-kolonnen
+#items <- unlist(pilot_list)
+items <- colnames(df[,colnames(df) %in% unlist(pilot_list)]) # ikke ha med subset-kolonnen
 
@@
 
+#### TRELIS_Q
+
+```{r remove_collinear_items_pilot_TRELISQ}
+### pilot trelis_q
+data_TPACK_trelis_q <- data_TRELIS_Q %>%
+    select(-c("TPACK3"))
+```
+
+##### Linearity
+
+```{r linear_scatterplots_TRELISQ}
+# scatterplot som korrigerer for overlap mellom datapunkter
+teller <- 0                          # hvor mange plot lager jeg?
+
+end <- ncol(data_TPACK_trelis_q)
+i <- 1                               # start posisjon
+for (var1 in names(TRELISQ_list)) {           # TPACK-element
+  items <- colnames(data_TPACK_trelis_q)[colnames(data_TPACK_trelis_q) %in% TRELISQ_list[[var1]]]
+  ifelse (length(items)>1, 
+    y <- rowMeans(data_TPACK_trelis_q[, items]), # y-variabel
+    y <- data_TPACK_trelis_q[, items])           # y består av bare ett item
+  i <- i + 1                         # for å unngå duplikater
+  for (var2 in items) {               # x-variabel
+    # scatterplot
+    # kilde: https://ggplot2-book.org/statistical-summaries.html
+    # kilde: https://stackoverflow.com/questions/32004292/r-ggplot-for-loop-aes-variable
+    x <- data_TPACK_trelis_q[,var2]
+    xy <- data.frame(t(rbind(y,x)))
+    colnames(xy) <- c(var1,var2)
+
+    # check for collinearity, r>0.90
+    r <- cor.test(x,y)
+
+    # 0.4 er moderat korrelasjon ift Christophersen (2018) kap. 3
+    if (abs(r$estimate)>0.4) { # only plot scatterplots which may be linear?
+      xy_plot <- ggplot(xy, aes_string(var2,var1)) +
+        geom_point() +                            # shape=1,alpha=1/5
+        geom_jitter() +
+        ggtitle(paste("scatterplot", data_TPACK_trelis_q$datasett))
+
+      print(xy_plot)
+    teller <- teller + 1   # count number of scatterplots
+    }
+  }
+}
+print(teller) # antall plot
+```
+
+```{r linear_scatterplots_aggregated_TRELISQ}
+teller <- 0                          # hvor mange plot lager jeg?
+#datasett_index <- 1
+#start <- 1                          # siste item = datasett
+
+end <- ncol(data_TPACK_trelis_q)
+i <- 1                               # start posisjon
+for (var1 in names(TRELISQ_list)) {           # TPACK-element
+  items <- colnames(data_TPACK_trelis_q)[colnames(data_TPACK_trelis_q) %in% TRELISQ_list[[var1]]]
+
+  if (length(items) == 0) {next} # skip empty items
+
+  ifelse (length(items)>1,  # calculate average value
+    y <- rowMeans(data_TPACK_trelis_q[, items]), # y-variabel
+    y <- data_TPACK_trelis_q[, items])
+
+  x <- data_TPACK_trelis_q[, items]
+
+  xy <- data.frame(y,x)
+  colnames(xy) <- c(var1,items)
+
+  # pivot_longer
+  # Warning: Using an external vector in selections was deprecated in tidyselect 1.1.0.
+  # ℹ Please use `all_of()` or `any_of()` instead.
+  xy_long <- pivot_longer(xy,!c(all_of(var1)), names_to="items",values_to = "likert")
+
+  #Warning: `aes_string()` was deprecated in ggplot2 3.0.0.
+  #ℹ Please use tidy evaluation idioms with `aes()`.
+  #ℹ See also `vignette("ggplot2-in-packages")` for more information.
+
+  xy_plot <- ggplot(xy_long, aes_string("likert",var1)) +
+    geom_point() +                            # shape=1,alpha=1/5
+    geom_jitter() +
+    ggtitle(paste("scatterplot", data_TPACK_trelis_q$datasett))
+
+  print(xy_plot)
+  teller <- teller + 1   # count number of scatterplots
+}
+
+print(teller) # antall plot
+```
+
+##### Collinearity
+
+```{r check_collinearity_TRELISQ}
+# eigenvalue_table
+ev_table <- collinear(data_TPACK_trelis_q[,colnames(data_TPACK_trelis_q) %in% unlist(TRELISQ_list)])
+ev_table
+```
+
+##### Multicollinearity
+
+```{r check_multicollinearity_TRELISQ}
+# multiple regression model
+
+# I browse through the items, and remove the current item by subtracting it at the end of the formula
+mc <- NULL      # list of items with multicollinearity
+
+# kilde: https://www.codingprof.com/3-ways-to-test-for-multicollinearity-in-r-examples/
+corrplot(cor(data_TPACK_trelis_q[,colnames(data_TPACK_trelis_q) %in% unlist(TRELISQ_list)]), method = "number", number.cex=0.3) # r bør være under 0.85
+
+# forenkle navn på dataset for prettyprint
+df <- data_TPACK_trelis_q
+
+#items <- unlist(TRELISQ_list)
+items <- colnames(df[,colnames(df) %in% unlist(TRELISQ_list)]) # ikke ha med subset-kolonnen
+
+# lag tekststreng for uavhengige variable
+#antall_tpack <- length(grep("TPACK",items))
+independent_var <- paste("'] ~ df [,'", items[1], sep="") #TK1"
+for (item in items[2:(length(items))]) { #-antall_tpack)]) { # omit TPACK
+  independent_var <- paste(independent_var, "'] + df [,'", item, sep="")
+}
+independent_var <- paste(independent_var, "'] - df  [,'", sep="")
+
+for (item in items) {
+  model <- lm( as.formula(
+  paste( "df [,'", item, independent_var, item, "']", sep="")))
+
+  print("############################")
+  print("######## NEW ITEM ##########")
+  print("############################")
+
+  print(item)
+
+  # check for multicollinearity: R > 0.9
+  print("Any R>0.9?")
+  mc_items <- model$coefficients[abs(model$coefficients)>0.9]
+  if (length(mc_items) > 0) {
+    print(c("multicollinearitet R > 0.9?", mc_items))
+    print(summary(model))
+    mc <- c(mc, item)
+  }
+
+  #if (vif == 1) {
+  # kilde: https://www.statology.org/multicollinearity-in-r/
+  print("car::vif")
+  print(vif(model))
+
+  print("olsrr::vif")
+  # kilde: https://www.codingprof.com/3-ways-to-test-for-multicollinearity-in-r-examples/
+  print("tolerance bør være over 0.1 og VIF under 10.")
+  print(ols_vif_tol(model)) # tolerance bør være over 0.1 og VIF under 10.
+  #print("CI bør både være under 10 og varians lav")
+  #print(ols_eigen_cindex(model)) # CI bør både være under 10 og varians lav
+}
+
+# print items with multicollinearity
+print(mc)
+```
+
 ### Outliers?
@@
 
-```{r}
+```{r help_function_outlier}
 ### HELP FUNCTION
@@
 
+```
+
+#### pilot
+
+```{r outlier_pilot}
 # skip first column which is list of subsets
@@
 #for (datasett in names(data_list)) {
-  for (i in colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% unlist(TPACK_list)]) {
+  for (i in colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)]) {
     print(paste(i))
@@
 print(describe(data_TPACK_pilot_minuspost)) # independent variables, x
-#print(describe(scoreItems(TPACK_list,data_TPACK_pilot_minuspost)$scores)) # dependent variables
+#print(describe(scoreItems(pilot_list,data_TPACK_pilot_minuspost)$scores)) # dependent variables
+```
+
+#### TRELIS_Q
+
+```{r outlier_TRELISQ}
+# skip first column which is list of subsets
+#datasett_index <- 1
+#for (datasett in names(data_list)) {
+  for (i in colnames(data_TPACK_trelis_q)[colnames(data_TPACK_trelis_q) %in% unlist(TRELISQ_list)]) {
+    print(paste(i))
+    describe_data(data_TPACK_trelis_q[,i])
+  }
+#  datasett_index <- datasett_index + 1
+#}
+
+print(describe(data_TPACK_trelis_q)) # independent variables, x
+#print(describe(scoreItems(TRELISQ_list,data_TPACK_trelis_q)$scores)) # dependent variables
 ```
@@
 
-```{r residuals_singleitems}
+```{r help_function_residuals_singleitems}
+### HELP FUNCTION
+
 check_residuals <- function(dataset, items) {
@@
 
-items     <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% unlist(TPACK_list)]
+```
+
+#### pilot
+
+```{r autocorrelation_pilot}
+items     <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)]
 table_res <- check_residuals(data_TPACK_pilot_minuspost, items)
@@
 
-```{r distribution_residuals_aggregated}
+```{r distribution_residuals_aggregated_pilot}
 teller         <- 0                             # hvor mange plot lager jeg?
@@
 #  datasett <- data_list[[dataset]]      # access data object
-  for (var1 in names(TPACK_list)) {     # TPACK-element
-    items      <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% TPACK_list[[var1]]]
+  for (var1 in names(pilot_list)) {     # TPACK-element
+    items      <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% pilot_list[[var1]]]
     ifelse (length(items)>1,            # ikke gjør analyse hvis bare ett item
@@
 
+#### TRELIS_Q
+
+```{r autocorrelation_TRELISQ}
+items     <- colnames(data_TPACK_trelis_q)[colnames(data_TPACK_trelis_q) %in% unlist(TRELISQ_list)]
+table_res <- check_residuals(data_TPACK_trelis_q, items)
+
+# print items with large skew og kurtosi or autocollinearity
+pander(table_res[abs(table_res$skew) > 1. | abs(table_res$kurtosi) > 3. | table_res$dw.2 > 2.5 | table_res$dw.2 < 1.5,])
+```
+
+Are there correlating residuals?  
+Autocorrelation: Durbin-Watsons d = 2*(1-r)
+
+If d = 2, then OK  
+If d < 1 eller d > 3, then not OK
+
+If autocorrelation, use robust SE
+
+Autocorrelation can be due to structured data. Then use multilevel-analysis.
+
+```{r distribution_residuals_aggregated_TRELISQ}
+teller         <- 0                             # hvor mange plot lager jeg?
+datasett_index <- 1
+#for (dataset in names(data_list)) {
+#  datasett <- data_list[[dataset]]      # access data object
+  for (var1 in names(TRELISQ_list)) {     # TPACK-element
+    items      <- colnames(data_TPACK_trelis_q)[colnames(data_TPACK_trelis_q) %in% TRELISQ_list[[var1]]]
+    ifelse (length(items)>1,            # ikke gjør analyse hvis bare ett item
+      y        <- rowMeans(data_TPACK_trelis_q[, items]), # y-variabel
+      y        <- 0)
+
+   print(c(var1, "=", items))
+
+   if (!length(y)==1) {                # sjekk for at vi har flere enn et item
+
+      x        <- data_TPACK_trelis_q[, items]
+
+      xy       <- data.frame(y,x)
+      colnames(xy) <- c(var1,items)
+
+      # lag tekststreng til formel y= x1 + x2 + osv.
+      independent_var <- paste("'] ~ xy [,'", items[1], sep="") #TK1"
+      for (item in items[2:(length(items))]) { 
+        independent_var <- paste(independent_var, "'] + xy [,'", item, sep="")
+      }
+      independent_var <- paste(independent_var, "'] - xy  [,'", sep="")
+
+      #for (item in items) {
+        model <- lm( as.formula(
+        paste( "xy [,'", item, independent_var, item, "']", sep="")))
+
+        # check for outliers in residuals
+        print(describe_data(model$residuals))
+
+        # grense for betydningsfulle outliers / leverage; s. 78 i Christophersen, 2018
+        k <- length(items) # antall uavhengige variables
+        n <- nrow(xy)
+        ifelse (k<10, 
+                leverage <- 3*(k+1)/n,
+                leverage <- 2*(k+1)/n
+                )
+         print(c("betydningsfull grense:", leverage))
+
+        # check for normalfordeling
+        # 1: normalfordeling av residuals?
+        print(hist(residuals(model), col = "steelblue"))
+        # 2: jevn fordeling av varians, homoskedastisitet
+        print(plot(fitted(model), residuals(model)))
+        print(abline(h = 0, lty = 2))
+
+        teller <- teller + 1   # count number of scatterplots
+      #}
+    }
+  }
+  #datasett_index <- datasett_index + 1
+#} 
+print(teller) # antall plot
+
+```
+
 ### Missing?
@@
 
-```{r}
+#### pilot
+
+```{r missing_pilot}
 # ref: https://gabriellajg.github.io/EPSY-579-R-Cookbook-for-SEM/lavaan-lab-12-sem-for-missing-data.html
@@
 
+#### TRELIS_Q
+
+```{r missing_TRELISQ}
+# ref: https://gabriellajg.github.io/EPSY-579-R-Cookbook-for-SEM/lavaan-lab-12-sem-for-missing-data.html
+md.pattern(data_TPACK_trelis_q)
+```
+
+At spørsmål mangler i spørreskjemaet kalles missing at random (MAR), siden missing skyldes at spørsmålene ikke var med i spørreskjemaet (jf.
+forklaring på MAR i Schafer og Graham, 2002, som gir nettopp dette som et eksempel på MAR).
+(random betyr ikke her at det er tilfeldig hvilke spørsmål som mangler, hvilket tilsvarer MCAR, men at det nettopp ikke er tilfeldig hvilke spørsmål som mangler, som kan være litt forvirrende begrepsbruk.)
+
+På TPACK-PST mangler også 5% av spørsmålene (109 items).
+Ser man på mønsteret her er det 4 respondenter som stort sett ikke har gitt svar, og står for mange av disse missing (6 respondenter: Rad 34, 44-46, 95 og 102 står for 90 missing = 83%).
+Disse er NMAR (fordi om de svarer har med spørsmålet å gjøre) eller MAR (fordi om de svarer har med spørreskjemaet å gjøre??).
+Det er altså 19 MCAR (altså at 2 eller færre missing per respondent).
+Rad 61 er kanskje MAR, siden det er de fire siste spørsmålene som ikke har blitt besvart.
+Kanskje respondenten ikke gadd mer?
+
+```{r}
+# kilde: https://www.statology.org/r-find-missing-values/
+which(is.na(data_TPACK_trelis_q))
+sum(is.na(data_TPACK_trelis_q))
+sum(is.na(data_TPACK_trelis_q[c(34,44,45,46,95,102),]))
+
+for (i in 1:110) {
+  if (sum(is.na(data_TPACK_trelis_q[i,])) > 0) {
+  print(c("rad",i))
+  print(c("kolonner",which(is.na(data_TPACK_trelis_q[i,]))))
+  print(c("antall", sum(is.na(data_TPACK_trelis_q[i,]))))
+  print("")
+}}
+```
+
+109 missing. (jf. data_screening-2023.Rmd)
+
+På TPACK-PST mangler også 5% av spørsmålene (109 items).
+Ser man på mønsteret her er det 4 respondenter som stort sett ikke har gitt svar, og står for mange av disse missing (6 respondenter: Rad 34, 44-46, 95 og 102 står for 90 missing = 83%).
+Disse er NMAR (fordi om de svarer har med spørsmålet å gjøre) eller MAR (fordi om de svarer har med spørreskjemaet å gjøre??).
+Det er altså 19 MCAR (altså at 2 eller færre missing per respondent).
+Rad 61 er kanskje MAR, siden det er de fire siste spørsmålene som ikke har blitt besvart.
+Kanskje respondenten ikke gadd mer?
+
+90 er NMAR.
+19 er MCAR eller MAR.
+
 ## Step 6: Factor extraction (i.e., EFA)
@@
 ## Step 7: Tests of dimensionality (i.e., CFA)
+
+```{r remove_NMAR_TRELISQ}
+# remove NMAR + NTNU (19)
+data_TPACK_PST_noNMAR <-
+  data_trelis_q_ren[-c(19,34,44,45,46,95,102),]
+
+```
 

### 2024-09-03 Rmd/Paper1.Rmd
Added EFA

@@
 
-## Step 5: Item reduction
+## Step 5: Item reduction (assumption checking)
+
+```{r}
+do_screening <- 0 # skip screening during testing
+```
 
@@
 # eigenvalue_table
+if (do_screening == 1) {
+
 ev_table <- collinear(data_TPACK_pilot_minuspost[,colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)])
-ev_table
+print(ev_table)
+}
 ```
@@
 # multiple regression model
+
+if (do_screening == 1) {
 
@@
 print(mc)
+}
 ```
@@
 # eigenvalue_table
+
+if (do_screening == 1) {
   ev_table <- collinear(data_TPACK_trelis_q[,colnames(data_TPACK_trelis_q) %in% unlist(TRELISQ_list)])
-ev_table
+  print(ev_table)
+}
 ```
@@
 print(mc)
+
+}
 ```
@@
 ```{r autocorrelation_pilot}
+if (do_screening == 1) {
+
 items     <- colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)]
@@
 pander(table_res[abs(table_res$skew) > 1. | abs(table_res$kurtosi) > 3. | table_res$dw.2 > 2.5 | table_res$dw.2 < 1.5,])
+
+}
 ```
@@
 ```{r autocorrelation_TRELISQ}
+if (do_screening == 1) {
+
 items     <- colnames(data_TPACK_trelis_q)[colnames(data_TPACK_trelis_q) %in% unlist(TRELISQ_list)]
@@
 
+### Describe variables
+
+To get an overview of the quality of the data and to see what data it is appropiate to use further in the analysis.
+
+First count distribution of answers on the likert-scale used.
+
+```{r likert_distribution}
+# kilde: https://solomonkurz.netlify.app/post/2021-05-11-yes-you-can-fit-an-exploratory-factor-analysis-with-lavaan/
+data_TPACK_pilot_minuspost[,colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)] %>% 
+  pivot_longer(everything(),values_to="value") %>% 
+  count(value) %>% 
+  mutate(percent = (100 * n / sum(n)) %>% round(digits = 1))
+```
+
+Plot histogram for each items, to see how they compare.
+
+```{r spring_22}
+# plot multi-panel histograms with facet_wrap
+# kilde: https://www.zevross.com/blog/2019/04/02/easy-multi-panel-plots-in-r-using-facet_wrap-and-facet_grid-from-ggplot2/
+
+data_long <- pivot_longer(data_TPACK_pilot_minuspost, 
+                          cols=colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)],
+                          names_to="item", 
+                          values_to="likert_verdi",
+                          values_drop_na = TRUE)
+
+ggplot(data_long, aes(factor(likert_verdi))) +
+  geom_histogram(stat = "count") + 
+  facet_wrap("item")
+```
+
+<!-- Poor items: y10 (CK2), y13 (CK5) (because of range=2). -->
+
+### Preliminary analysis
+
+"It is recommended to carry out a preliminary analysis of the metric
+quality of the items to subject the most adequate items to EFA. For
+this purpose, it is recommended to analyze and report the mean,
+standard deviation, and item-test correlation of each one of the
+items, as well as the Cronbach’s alpha of the scales of the test. The
+researcher should decide whether to eliminate certain items" (Izquierdo et al, 2014)
+
+Analyze and report the mean, standard deviation, and item-test correlation of each one of the items, as well as the Cronbach’s alpha of the scales of the test
+
+```{r help_function_correlation_matrix}
+### HELP FUNCTION
+
+antall_desimaler <- 2
+
+# print effect size with p-value
+print_starvalue <- function (pvalue, effectsize, bonferroni=1) {
+  ifelse (pvalue < 0.001/bonferroni, star <- "***",
+  ifelse (pvalue < 0.01/bonferroni, star <- "**",
+  ifelse (pvalue < 0.05/bonferroni, star <- "*",
+          star <- "")))
+  return (paste(format(effectsize,digits=antall_desimaler,nsmall=antall_desimaler), star, sep="")) # antall desimaler=2
+}
+
+create_matrix <- function (df, method="pearson") {
+  # create correlation matrix for df (and covariance matrix combined, cf. Weston, 2006, table 2, p. 736) (cov lower left, cor upper right)
+
+  # since ordered data, use spearman; option in function call?
+
+  # left side
+  # use casewise deletion if missing
+  cov_matrix <- cov(df, use="complete.obs", method=method)
+  # right side
+  cor_matrix <- data.frame(matrix(ncol = ncol(df), nrow = 0)) # correlation matrix
+
+  i <- 1                               # start posisjon
+  teller <- 0                          # hvor mange plot lager jeg?
+  for (var1 in df[1:ncol(df)-1]) {
+    i <- i + 1
+
+    # create right side: correlation 
+    right_side <- data.frame(matrix(ncol = 0, nrow = 1)) # row values
+    for (var2 in df[i:ncol(df)]) {
+      teller <- teller + 1   
+      cor_var1_var2 <- cor.test(var1,var2, method=method)
+      est_p <- print_starvalue(cor_var1_var2$p.value, cor_var1_var2$estimate)
+      right_side <- cbind(right_side,est_p) # add column
+    }
+
+    # add left-side of matrix: covariance
+    #left_side <- data.frame(matrix(ncol = i-1, nrow = 1))
+    left_side <- format(cov_matrix[i-1,1:i-1],digits=antall_desimaler)
+
+    temp_row <- c(left_side, right_side)  # merge left and right
+    names(temp_row) <- colnames(df)
+    cor_matrix <- rbind(cor_matrix, temp_row)   # add row
+  }
+
+  # add final row
+  #final_row <- data.frame(matrix(ncol = ncol(df), nrow = 1))
+  final_row <- format(cov_matrix[i,1:i],digits=antall_desimaler)
+  names(final_row) <- colnames(df)
+  cor_matrix <- rbind(cor_matrix, final_row)
+
+  # add row with mean and sd based on rows/persons
+  row_mean <- round(colMeans(df, na.rm=TRUE), digits=antall_desimaler)
+  row_sd <- format(psych::SD(df), digits=antall_desimaler)
+  cor_matrix <- rbind(cor_matrix,"mean"=row_mean)
+  cor_matrix <- rbind(cor_matrix,"sd"=row_sd)
+
+  # set names of columns and rows
+  row.names(cor_matrix)<- c(colnames(df),"mean","sd")
+  return(cor_matrix)
+}
+
+```
+
+```{r mean_sd_spring}
+# velg ut data
+df <- data_TPACK_pilot_minuspost[,colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)]
+
+# fjern items der over 1/2 av respondentene mangler NA?
+df_noNA <- df[ , colSums(is.na(df)) < nrow(df)/2] 
+```
+
+```{r correlation_matrix, eval=FALSE}
+pander(create_matrix(df_noNA)) # ordered=FALSE
+```
+
+Fjerne TK4, TK7, CK4, TCK2, PK4, PK5, PCK1,
+fordi disse korrelerer svakt med sine meditems (ikke signifikant).
+TK4 har også signifikant lavere gjennomsnitt enn de andre TK-itemsene. (noe kanskje også TK8 har?)
+
+#### eigenvalues
+
+```{r}
+# check for collinearity; values near or below 0
+ev <- eigen(cor(df_noNA[,colnames(df_noNA) %in%
+                                unlist(pilot_list)],
+                use="pairwise.complete.obs"), only.values=TRUE)
+lag_eigenvalue_tabell (ev$values)
+```
+
+#### item-test correlation
+
+Formålet er å sjekke om et item svarer godt til testresultatet sett over ett, f.eks. hvilke TK-spørsmål som best predikerer TK.
+
+Ref.: https://www.questionmark.com/resources/blog/psychometrics-101-item-total-correlation/ https://en.wikipedia.org/wiki/Item-total_correlation
+
+#### emne2-anon
+
+```{r item_test_spring, warning=FALSE}
+# get data for each test / construct
+TK <- df_noNA[,colnames(df_noNA) %in% TK_pos]
+PK <- df_noNA[,colnames(df_noNA) %in%PK_pos]
+CK <- df_noNA[,colnames(df_noNA) %in%CK_pos] # NA
+TCK <- df_noNA[,colnames(df_noNA) %in%TCK_pos]
+TPK <- df_noNA[,colnames(df_noNA) %in%TPK_pos]
+PCK <- df_noNA[,colnames(df_noNA) %in%PCK_pos]
+
+# check correlation of items to construct
+pander(create_matrix(data.frame(TK,"TK"=rowMeans(TK, na.rm=TRUE))))
+pander(create_matrix(data.frame(PK,"PK"=rowMeans(PK, na.rm=TRUE))))
+pander(create_matrix(data.frame(CK,"CK"=rowMeans(CK, na.rm=TRUE))))
+pander(create_matrix(data.frame(TCK,"TCK"=rowMeans(TCK, na.rm=TRUE))))
+pander(create_matrix(data.frame(TPK,"TPK"=rowMeans(TPK, na.rm=TRUE))))
+pander(create_matrix(data.frame(PCK,"PCK"=rowMeans(PCK, na.rm=TRUE))))
+
+ TPACK <- df_noNA[,colnames(df_noNA) %in% TPACK_pos]
+  pander(create_matrix(data.frame(TPACK,"TPACK"=rowMeans(TPACK, na.rm=TRUE))))
+
+```
+
+TCK2 har lav item-test correlation.
+
+#### Reliability: Cronbachs alpha
+
+```{r}
+# hvis grupperingen i prinsippet er mulig å skille fra hverandre, så kan Cronbachs alpha brukes til å vurdere likheten mellom gruppene (reliabilitet)
+# alpha beregnes per construct, altså TK, TCK, TPACK, osv.
+#psych::alpha(data.frame(pre,post,anon,trelis_q))
+```
+
+Assumptions: uni-dimensionality (one factor = factor analysis), tau-equality (equal covariance) (ref Revelle, 2019)
+
+##### TK
+
+###### uni-dimensionality
+
+```{r warning=FALSE}
+parallel <- fa.parallel(TK,fm='uls',fa='fa',SMC=TRUE) #, cor="poly", correct=0 ,use="complete.obs") #)
+parallel$fa.values
+```
+
+###### tau-equality
+
+```{r warning=FALSE}
+pander(create_matrix(TK))
+```
+
+Forutsetninger ikke oppfylt: ikke uni-dimensional (2 faktorer fra parallell analysis, og 1 fra eigenvalue), og ikke tau-equality (TK8 passer ikke på tau-equality).
+
+###### alpha
+
+```{r}
+psych::alpha(TK)
+```
+
+##### CK
+
+```{r warning=FALSE}
+parallel <- fa.parallel(CK,fm='uls',fa='fa',SMC=TRUE,  cor="poly", correct=0,use="complete.obs") #,cor="poly")
+parallel$fa.values
+pander(create_matrix(CK))
+psych::alpha(CK) 
+```
+
+Uni-dimensional, og nesten tau-equality. CK4 "ødelegger". CK4 er allerede tagget for fjerning.
+
+##### PK
+
+```{r warning=FALSE,eval=F}
+parallel <- fa.parallel(PK,fm='uls',fa='fa',SMC=TRUE,  cor="poly", correct=0,use="complete.obs") #,cor="poly")
+parallel$fa.values
+pander(create_matrix(PK))
+psych::alpha(PK)
+
+```
+
+Uni-dimensional. Bortimo tau-equality. PK4 og PK5 "ødelegger" litt. Begge disse er allerede tagget for fjerning. PK6 senker også reliabiliteten. (Denne er dog tatt med videre til TRELIS-q)
+
+##### TCK
+
+```{r warning=FALSE}
+parallel <- fa.parallel(TCK,fm='uls',fa='fa',SMC=TRUE,  cor="poly", correct=0,use="complete.obs") #,cor="poly")
+parallel$fa.values
+pander(create_matrix(TCK))
+psych::alpha(TCK)
+
+```
+
+Uni-dimensional. Ikke tau-equality. TCK1 ødelegger.
+
+##### TPK
+
+```{r warning=FALSE,eval=T}
+parallel <- fa.parallel(TPK,fm='uls',fa='fa',SMC=TRUE) #,  cor="poly", correct=0,use="complete.obs") #,cor="poly")
+parallel$fa.values
+pander(create_matrix(TPK))
+psych::alpha(TPK)
+
+```
+
+Uni-dimensional og omtrent tau-equality.
+
+##### PCK
+
+```{r warning=FALSE}
+parallel <- fa.parallel(PCK,fm='uls',fa='fa',SMC=TRUE,  cor="poly", correct=0,use="complete.obs") #,cor="poly")
+parallel$fa.values
+pander(create_matrix(PCK))
+psych::alpha(PCK)
+
+```
+
+Uni-dimensional og omtrent tau-equality. PCK1 og PCK3 kan fjernes. (men PCK1 er tatt med videre)
+
+##### TPACK
+
+```{r warning=FALSE, eval=T}
+  parallel <- fa.parallel(TPACK,fm='uls',fa='fa',SMC=TRUE,
+                          cor="poly", correct=0,use="complete.obs") #,cor="poly")
+  parallel$fa.values
+  pander(create_matrix(TPACK))
+  psych::alpha(TPACK)
+
+```
+
+Uni-dimensional og halvveis tau-equality. TPACK1 "ødelegger" litt.
+
+#### Results of preliminary analysis
+
+25.5.2023:
+
+TK7 skiller seg ut.
+CK4 skiller seg ut.
+
+Fjerner items basert på preTK4 har lav korrelasjon med de andre.
+
+```{r}
+# remove columns
+
+# 25.5.2023
+dTp_mod <- within(data_TPACK_pilot_minuspost, rm("TK7","CK4")) # "PK2", "TK3","TK8",
+
+#   data, since items removed above
+# velg ut data
+df <- dTp_mod[,colnames(dTp_mod) %in% unlist(pilot_list)]
+
+# fjern items der over 1/2 av respondentene mangler NA?
+df_noNA <- df[ , colSums(is.na(df)) < nrow(df)/2]
+
+TK <- df_noNA[,colnames(df_noNA) %in% TK_pos]
+PK <- df_noNA[,colnames(df_noNA) %in%PK_pos]
+CK <- df_noNA[,colnames(df_noNA) %in%CK_pos] # NA
+TCK <- df_noNA[,colnames(df_noNA) %in%TCK_pos]
+TPK <- df_noNA[,colnames(df_noNA) %in%TPK_pos]
+PCK <- df_noNA[,colnames(df_noNA) %in%PCK_pos]
+
+  TPACK <- df_noNA[,colnames(df_noNA) %in%TPACK_pos]
+
+```
+
+### EFA
+
+#### run 1
+
+Suitability of the data for doing an EFA.
+
+```{r}
+# KMO og Bartlett
+KMO_base <- KMO(df_noNA) # addition of PCK1 changed KMO from 0.5 to 0.36
+bartlett_base <- cortest.bartlett(df_noNA, n=5,diag=TRUE)
+```
+
+##### Number of factors
+
+```{r parallel_analysis_run1}
+# fm = mle, alpha, pa
+
+# number of factors: eigenvalue, scree or parallel (or theoretical)
+# parallel analysis (Worthington, 2006)
+# polychoric is used for ordered data with skewness, Watkins (2018)
+parallel_base <- fa.parallel(df_noNA,fm='mle',fa='fa',SMC=TRUE,  use="complete.obs") #, )correct=0,cor="poly",
+#parallel <- fa.parallel(df,fm='alpha',fa='fa',nfactors=7,use="complete.obs",cor='poly', error.bars=TRUE, SMC=TRUE) # complete.obs = delete NA; poly for fa recommended av Revelle & Condon, 2019
+
+# eigenvalues
+#parallel$fa.values
+lag_eigenvalue_tabell (parallel_base$fa.values)
+
+```
+
+```{r}
+print(format(list("KMO" = KMO_base$MSA, "Bartlett p =" = bartlett_base$p.value),digits=2))
+```
+
+##### 2-faktor
+
+```{r efa_ordinal2}
+# FA
+# ref.: https://www.rdocumentation.org/packages/psych/versions/2.2.5/topics/fa
+
+# print covariation matrix all items
+# bruk fm=uls, ikke pa (Flora et al, 2012)
+# bruk polykorisk korrelasjon (Flora et al, 2012)
+# use covariance matrix (correlation matrix fails because of missing)
+factor_model1 <- fa(df_noNA, nfactors = 2,rotate = "oblimin", fm="mle", scores="tenBerge") # missing=TRUE,impute="mean", ,oblique.scores=TRUE, cor="poly", correct=0 covar=TRUE,
+
+print.psych(fa.sort(factor_model1))
+```
+
+##### 3-faktor
+
+```{r efa_ordinal3}
+# FA
+# ref.: https://www.rdocumentation.org/packages/psych/versions/2.2.5/topics/fa
+
+# print covariation matrix all items
+# bruk fm=uls, ikke pa (Flora et al, 2012)
+# bruk polykorisk korrelasjon (Flora et al, 2012)
+# use covariance matrix (correlation matrix fails because of missing)
+factor_model1 <- fa(df_noNA, nfactors = 3,rotate = "oblimin", fm="mle", scores="tenBerge") # missing=TRUE,impute="mean", covar=TRUE,,oblique.scores=TRUE, cor="poly", correct=0
+
+print.psych(fa.sort(factor_model1))
+```
+
+##### 4-faktor
+
+```{r efa_ordinal4}
+# FA
+# ref.: https://www.rdocumentation.org/packages/psych/versions/2.2.5/topics/fa
+
+# print covariation matrix all items
+# bruk fm=uls, ikke pa (Flora et al, 2012)
+# bruk polykorisk korrelasjon (Flora et al, 2012)
+# use covariance matrix (correlation matrix fails because of missing)
+factor_model1 <- fa(df_noNA, nfactors = 4,rotate = "oblimin", fm="mle", scores="tenBerge") # missing=TRUE,impute="mean", covar=TRUE,,oblique.scores=TRUE, cor="poly", correct=0
+
+print.psych(fa.sort(factor_model1))
+```
+
+```{r}
+# eigenvalues
+print(factor_model1$values)
+```
+
+Eigenvalues suggests 6 factors
+
+```{r}
+# cutoff verdi gitt av formel i Watkins, 2018 (avsnitt: Interpretation of results)
+n <- factor_model1$n.obs
+c <- 4.084/sqrt(n-2)   # n=100 gir 3.92 p<5% (z-dist); n=30 gir 4.084 for p<5% (t-dist)
+print(c)
+# bare print loadings med verdier over 0.32
+print(factor_model1$loadings,cutoff = c)
+
+print.psych(factor_model1)
+```
+
+```{r communality}
+# h2 = communality; bare print items med communality under 0.4 siden disse kan fjernes (Worthington, 2006); eller 0.6 for små utvalg? (ref)
+# kilde: https://stackoverflow.com/questions/30467834/r-print-cutoff-values-under-a-certain-value
+print(factor_model1$communality[factor_model1$communality<0.6])
+```
+
+```{r}
+# fjern items med lav communality; remove columns
+#df_mod <- within(df_noNA, rm("PK5"))
+
+# 26.5.2023, remove cross-loadings
+df_mod <- within(df_noNA, rm("TK5"))
+```
+
+Kan vurdere å beholde CK1 med h2 = 0.37.
+
+#### run 2
+
+Suitability of the data for doing an EFA.
+
+```{r}
+# KMO og Bartlett
+KMO_base <- KMO(df_mod) # addition of PCK1 changed KMO from 0.5 to 0.36
+bartlett_base <- cortest.bartlett(df_mod, n=4,diag=TRUE)
+```
+
+##### Number of factors
+
+```{r parallel_analysis_run2}
+# fm = mle, alpha, pa
+
+#if (do_parallel == 1) {
+
+# number of factors: eigenvalue, scree or parallel (or theoretical)
+# parallel analysis (Worthington, 2006)
+# polychoric is used for ordered data with skewness, Watkins (2018)
+  # "pa (parallel analysis) most accurately estimated the number of factors" (Li et al., 2020)
+parallel_base <- fa.parallel(df_mod,fm='mle',fa='fa',SMC=TRUE) #, ),  correct=0,cor="poly",use="complete.obs"
+#parallel <- fa.parallel(df,fm='alpha',fa='fa',nfactors=7,use="complete.obs",cor='poly', error.bars=TRUE, SMC=TRUE) # complete.obs = delete NA; poly for fa recommended av Revelle & Condon, 2019
+
+# eigenvalues
+#parallel$fa.values
+lag_eigenvalue_tabell (parallel_base$fa.values)
+#}
+```
+
+```{r}
+print(format(list("KMO" = KMO_base$MSA, "Bartlett p =" = bartlett_base$p.value),digits=2))
+```
+
+##### 2 faktorer
+
+```{r}
+factor_model2 <- fa(df_mod, nfactors = 2, rotate = "oblimin", fm="mle", scores="tenBerge") # ,oblique.scores=TRUE, cor="poly", correct=0
+
+print.psych(fa.sort(factor_model2))
+```
+
+##### 3 faktorer
+
+```{r}
+# equamax for complex structure, ref. Schmitt & Sass, 2011
+# CI with bootstrapping, n.iter=1000
+# p = .8 (Cohen, 1990)
+factor_model2 <- fa(df_mod, nfactors = 3, rotate = "oblimin", fm="mle", scores="tenBerge") #, missing=TRUE, impute="mean") # n.iter=1000, p=.8,  #,oblique.scores=TRUE, cor="poly", correct=0)
+
+print.psych(fa.sort(factor_model2))
+
+#factor_model2
+```
+
+##### 4 faktorer
+
+```{r}
+factor_model2 <- fa(df_mod, nfactors = 4, rotate = "equamax", fm="uls", scores="tenBerge") #,oblique.scores=TRUE, cor="poly", correct=0)
+
+print.psych(fa.sort(factor_model2))
+```
+
+```{r}
+# korrelation matrix factors
+#format(round(factor_model$r.scores,1), nsmall=1) 
+order2 <- factor_model2$Phi # 2nd order factors
+print(order2)
+
+# eigenvalues
+print(factor_model2$values)
+
+# loadings
+print.psych(factor_model2) #, cut=c)
+```
+
+Remove items "with factor loadings less than .32 or cross-loadings less than .15 difference from an item’s highest factor loading.
+In addition, they should also delete items that contain absolute loadings higher than a certain value (e.g., .32) on two or more factors." (Worthington and Whittaker, 2006).
+
+#### Vurdering av EFA-resultater
+
+Felles varians: hvor mye felles varians forklarer faktorene? Faktorene bør tilsammen forklare minst 50% av variansen (Christophersen, 2018).
+
+```{r}
+p <- factor_model2$Vaccounted
+print(p)                           # se på cumulative var
+```
+
+Kommunaliteten (h2) beskriver hvor stor del av variansen til indikatoren som faktorene "forklarer".
+Residualvariansen (u2) er den uforklarte variansen til indikatoren (item/spørsmålet).
+Det er i praksis vanskelig å få særlig høyere kommunalitet enn 65% i følge Christophersen (2018).
+
+```{r communality7}
+# Mundfrom et al., 2005: Råd: høy kommunalitet (>0.6) 
+# dette er unstandardized loadings / covariance matrix ??
+print(factor_model2$communality[factor_model2$communality>0.6])
+# dette er standardized loadings ??
+#print(efa7_ml_pro$communalities[efa7_ml_pro$communalities>0.6])
+
+# avoid below 0.4
+print(factor_model2$communality[factor_model2$communality<0.4])
+```
+
+Vurder modellen: Hvor sterkt korrelerer indikatorene i modellen?
+Indikatorer med r>0.80 er for sterkt korrelert (Christophersen, 2018).
+Indikatorer som korrelerer svært svakt kan tyde på at et vesentlig begrepsaspekt er utelatt av modellen.
+
+Når det er korrelerte faktorer vi bruker, så vil strukturmatrisen gi korrelasjonene, mens pattern matrisen gir loadings.
+
+```{r}
+# for sterke korrelasjoner
+print(factor_model2$Structure, cutoff=0.8)
+
+factor_model2$Structure
+
+# meget svake korrelasjoner
+#efa7_ml_pro$Structure[efa7_ml_pro$Structure<0.1]
+```
+
+restledd skal helst ikke være større enn +- 0.05
+
+```{r}
+round(factor_model2$residual, digits=2)
+```
+
+```{r}
+# check reliability
+#cronbach.alpha(df_mod, na.rm=TRUE)
+
+# tegn factor_model
+stimodel <- fa.diagram(factor_model2)
+#print(stimodel)
+```
+
+```{r}
+# compare with intervall pearson (appx normal distributed)
+pander(create_matrix(df_mod))
+```
+
+### McDonalds omega
+
+EFA som forberedelse til omega (informal test)
+
+Fra help-file til omega: "A recommendation that should be heeded, regardless of the method chosen to estimate \omega_hω, is to always examine the pattern of the estimated general factor loadings prior to estimating \omega_hω.
+Such an examination constitutes an informal test of the assumption that there is a latent variable common to all of the scale's indicators that can be conducted even in the context of EFA.
+If the loadings were salient for only a relatively small subset of the indicators, this would suggest that there is no true general factor underlying the covariance matrix.
+Just such an informal assumption test would have afforded a great deal of protection against the possibility of misinterpreting the misleading \omega_hω estimates occasionally produced in the simulations reported here." (Zinbarg et al., 2006, p 137)"
+
+```{r eval=T}
+alpha(df_mod[,c("TK1", "TK2", "TK3")])
+alpha(df_mod[,c("CK1", "PCK2", "PCK5", "CK5", "CK3", "CK2")])
+alpha(df_mod[,c("TPK1","TPK4","TCK1","TCK4", "TPACK2","TPACK4")]) # , "TPK4"
+```
+
+```{r}
+omega(df_mod, fm="mle") #, rotate="Promax")
+```
+
 # Stage 3

### 2024-09-03 Rmd/Paper1.Rmd
Added CFA

@@
 
+### read files
 library(readODS)                            # read .ods-files. Kilde: https://github.com/ropensci/readODS/
@@
 
+### presentation
 library(pander)                             # print pretty tables
@@
 library(tidyr)                              # pivot_longer
-
 library(ggplot2)
 
-# multicollinearity checks
-library("corrplot")                         # multicollinearity 
-library(car)                                # vif       
+### screening
+library("corrplot")                         # multicollinearity library(car)                                # vif       
 library("olsrr")                            # vif
-
-library(psych)                              # for skew
 library(mice)                               # for missing
+
+### EFA
+library(psych)                              # for skew and EFA
+
+### CFA
+library(lavaan)
+#library(lavaanPlot)
+library(semTools)                           # to find omega
 
@@
 
+Choices:
+
+* treat the data as interval or ordinal? (Watkins, 2018)
+* missing data: use imputation (McNeish, 2017; Schafer and Graham, 2002)
+* linearity-check: scatterplot, collinearity (Flora et al., 2012)
+* normal or nonnormal distribution?
+Normality is important for test statistic.
+For nonnormal use Satorra-Bentler scaled chi-square and robust SE (Flora et al., 2012; McNeish, 2017).
+Nonnormal means skew>2 and kurtosis>9 (ref?).
+* aim for approximate fit, not exact fit (Flora et al., 2012; Weston and Gore, 2006)
+* item selection: communalities (h2) - high (0.6 < h2 > 0.8) or wide (0.2 < h2 > 0.8)
+
+"The swain and FAiR packages in R came perform the Swain correction and the FAiR package can also compute the Bartlett correction"
+
+"Swain correction tends to provide the smallest correction but provides researchers with the most power to detect whether their model fits poorly"
+
+yuan.bentler forutsetter adf-estimering.
+(McNeish, 2017) som krever at utvalgsstørrelser er minst p(p+1)/2, der p er antall items i modellen.
+Med p=12 blir det 78 respondenter, som er mer enn jeg har (p=106).
+
+test="satorra.bentler"  # for robust test når ikke-normalfordelte data
+test="yuan.bentler(.mplus)" (McNiesh, 2017: small samples? NB: adf)  
+estimator="ULS"?  
+se="robust.sem" # standard error  
+information="observed" # ref. Schafer & Graham, 2002, p. 162
+
 ```{r remove_NMAR_TRELISQ}
 # remove NMAR + NTNU (19)
-data_TPACK_PST_noNMAR <-
-  data_trelis_q_ren[-c(19,34,44,45,46,95,102),]
+CFA_TPACK_PST <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),]
+
+```
+
+### Spesifisering av modellene
+
+Hensikt: bekrefte measurement model, altså indikatorenes kobling til de latente faktorene. (ser ikke på korrelasjoner her?)
+
+```{r specify_models}
+
+pseudo_1faktor <- '
+PCK =~ CK123 + PCK2 + PCK5 + CK5 
+TK =~ TK1b + TK2 + TK34 
+TPACK =~ TPACK2 + TPACK4 + TCK4 + TPK4 + TCK1 + TPK1
+'
+
+tinker2 <- '
+TPACK =~  TK1b + TK2 + TK34 + TPACK2 + TPACK4 + TCK4 + TPK4 + TCK1 + TPK1
+PCK =~ CK123 + PCK2 + PCK5 + CK5
+'
+
+```
+
+### Test av modellene
+
+```{r help_functions_cfa}
+
+### HELP FUNCTIONS
+do_cfa <- function (cfaspec, cfadata) {
+  # 7 faktor modell basert på EFA
+# estimator og missing: Rhemtulla et al. (2012)
+# ref.: https://statistics.ohlsen-web.de/multiple-imputation-with-mice/
+# Predictive mean matching (McNeish, 2017)
+TPACK_PST_0 <- cfa(cfaspec, data= cfadata, std.lv=TRUE,
+                   #fixed.x=TRUE, 
+                   auto.fix.single=TRUE, 
+                   #auto.var=TRUE, auto.cov.lv.x=TRUE,
+                  #estimator="MLR", 
+                  missing="fiml", #robust.two.stage",
+             # DWLS URL: https://lavaan.ugent.be/tutorial/est.html
+             estimator="ML", #missing="pairwise", 
+             #estimator="WLS",# missing="ml",
+             #estimator="ULS", 
+             #ordered=TRUE,
+             #test="satorra.bentler", 
+             test="yuan.bentler",
+             information="observed",
+             se="robust.huber.white") # robust.sem
+             #se="boot", bootstrap=1000, # 10000 parallel=
+             #parallel = "snow", ncpus = 4) # ref: https://github.com/yrosseel/lavaan/issues/23
+
+return (TPACK_PST_0)
+}
+
+print_model <- function (cfamodel) {
+# The argument standardized = TRUE augments the output with standardized parameter values.
+Two extra columns of standardized parameter values are printed.
+In the first column (labeled Std.lv), only the latent variables are standardized.
+In the second column (labeled Std.all), both latent and observed variables are standardized.
+The latter is often called the ‘completely standardized solution’.
+(ref: https://lavaan.ugent.be/tutorial/sem.html)
+  print("summary")
+  print(summary(cfamodel, standardized=TRUE, fit.measures=TRUE)) 
+  print("estimates")
+  # CI suggested by Cheung et al. (2023) for discriminant validity < 1 (one-tailed)
+  print(parameterEstimates(cfamodel, rsquare = TRUE, level=0.90))
+  print("modindices")
+  print(modindices(cfamodel, sort=TRUE, maximum.number = 5))            # mi should be below 3.84
+
+  # ref Rönkkö and Cho (2022)
+#discriminantValidity(TPACK_PST_1, merge=TRUE)
+# reliability, ref.: https://groups.google.com/g/lavaan/c/tPEPXQqQR48
+#semTools::reliability(TPACK_PST_1)                         # deprecated
+  print("reliability")
+  rel <- compRelSEM(cfamodel, dropSingle = FALSE)
+  print(rel)
+  return (rel)   # for later comparison of reliability between models
+}
+
+```
+
+#### tinker2
+
+```{r estimate_9}
+# egenfaktor
+model9 <- do_cfa(tinker2, CFA_TPACK_PST)
+rel9 <- print_model(model9)
+
+```
+
+#### pseudo_1faktor
+
+```{r estimate_10}
+# egenfaktor
+model10 <- do_cfa(pseudo_1faktor, CFA_TPACK_PST)
+rel10 <- print_model(model10)
+
+```
+
+### Oppsummering av fit
+
+```{r CFA_results_comparison}
+kolonnenavn <- c("AIC", "chi-square", "df", "p-value","RMSEA",
+                 "RMSEA (90%) CI lower","RMSEA (90%) CI upper", 
+                 "CFI","TLI", "SRMR") #, "chi2/df", 
+                 #"chi-square scaled Swain", "p-value Swain", 
+                 #"swain chi2/df")
+anbefalte_verdier <- c("x", "x", "x", ">0.05", "<0.06(0.05-0.10)", "x", "x", ">0.95(0.9-0.95)", ">0.95", "<0.08(0.08-0.15)") #, "<2", "x", ">0.05", "<2") # Weston and Gore (2006); Gallagher and Brown (2013)
+
+# skal jeg bruke scaled eller robust?
+# bruk tre siffer for å kunne skille mellom CFI på 0.002 (discriminatory validity, ref. Shaffer et al. (2016))
+
+#                  fitMeasures(TPACK_PST_0,"chisq.scaled")/
+#                    fitMeasures(TPACK_PST_0,"df.scaled"),
+#                80.35, 0.002, 80.35/fitMeasures(TPACK_PST_0,"df.scaled")),3)
+
+fit9 <- round(fitMeasures(model9,
+                  c("aic","chisq.scaled", "df.scaled",
+                    "pvalue.scaled","rmsea.robust",
+                    "rmsea.ci.lower.robust","rmsea.ci.upper.robust",
+                    "cfi.robust","tli.robust","srmr")),3)
+fit10 <- round(fitMeasures(model10,
+                  c("aic","chisq.scaled", "df.scaled",
+                    "pvalue.scaled","rmsea.robust",
+                    "rmsea.ci.lower.robust","rmsea.ci.upper.robust",
+                    "cfi.robust","tli.robust","srmr")),3)
+
+radnavn <- c("anbefalt", "tinker2", "pseudo_1faktor")
+
+fit_table <- rbind(anbefalte_verdier,fit9)
+fit_table <- rbind(fit_table, fit10)
+
+colnames(fit_table) <- kolonnenavn
+row.names(fit_table) <- radnavn
+
+pander(fit_table)
+```
+
+### Results
+
+```{r covariance_matrix}
+
+#CFA_TPACK_PST <- CFA_TPACK_PST - 3 # flytt nullpunkt
+# covariance matrix / "TK1a","TK567"
+pander(create_matrix(CFA_TPACK_PST[,c("CK123", "CK5", "PCK2" , "PCK5" , "TK1b", "TK2", "TK34", "TCK1","TCK4", "TPK1","TPK4","TPACK2","TPACK4")])) # ,"TPK4"
+```
+
+To calculate the students TPACK-score we would only calculate the average of all the items, i.e. an integrative TPACK-model, where all the elements contributes to the students TPACK.
+
+```{r scores}
+
+TK <- c("TK1b", "TK2", "TK34") # "TK1a","TK567"
+PCK <- c("CK123", "PCK2" , "PCK5" , "CK5")
+TPACK <- c("TPK1","TPK4","TPACK2","TCK1","TPACK4", "TCK4") # , "TPK4"
+
+TPACK_PST_items <- list("TK" = TK,
+                        "PCK" = PCK,
+                        "TPACK" = TPACK) # everyone contributes til TPACK; c(TK,"TCK4",TPK, "CK123","PCK2","PCK5"))
+df <- CFA_TPACK_PST[,unlist(TPACK_PST_items[1:3])]
+describe(df)      # with missing (see n)? (na.rm=TRUE)
+score_TPACK <- scoreItems(TPACK_PST_items,df)
+#print(score_TPACK,option=FALSE)
+describe(score_TPACK$scores)
 
@@
 
+```{r omega_mle4}
+omega0 <- omegaFromSem(model10)
+omega0
+#print(c(omega1$alpha, omega1$omega.tot)
+
+# get correlation matrix of observed variables
+cor_mat <- lavInspect(model10, what="cor.ov")
+omega(cor_mat, nfactors=3)
+
+omega(CFA_TPACK_PST[,c("CK123", "CK5", "PCK2" , "PCK5" , "TCK1","TCK4", "TPK1","TPK4","TPACK2","TPACK4", "TK1b", "TK2", "TK34")]) # ,"TPK4"
+```
+
 ## Step 9: Tests of validity
+
+I test for discriminatory and convergent validity following @ronkko_updated_2022
+
+```{r}
+# ref.: https://link.springer.com/article/10.1007/s10490-023-09871-y
+source("../../scripts/measureQ.R")
+```
+
+```{r eval=T}
+#measureQ(model9, CFA_TPACK_PST)
+
+measureQ(model10, CFA_TPACK_PST)
+```
 

### 2024-09-03 Rmd/Paper1.Rmd
First complete version of Rmd for paper 1.

--- 2024-09-03 Rmd/Paper1.Rmd
@@
 library(tidyr)                              # pivot_longer
-
 library(ggplot2)
 
-# multicollinearity checks
+### screening
 library("corrplot")                         # multicollinearity
@@
 library("olsrr")                            # vif
-
-library(psych)                              # for skew
 library(mice)                               # for missing
+
+### EFA
+library(psych)                              # for skew and EFA
+
+### CFA
+library(lavaan)
+library(semTools)                           # to find omega
 
@@
 
+```{r}
+# clean environment to reset for next step
+# ref.: https://www.statology.org/clear-environment-in-r/ [3.9.2024]
+rm(list=ls())
+```
+
+## Step 5: Item reduction (assumption checking)
+
+```{r}
+do_screening <- 1 # skip screening during testing
+```
+
+```{r}
+# load data
+load(file="../data/processed_data/tpack-pilot_2022.Rdata") 
+load(file="../data/processed_data/tpack-trelis_q_paper1_2023.Rdata")
+```
+
 In the following we will only consider the pre-datasett, so we remove the post-data, which is to dissimilar from the pilot data.
 
-```{r select_data_for_screening}
+```{r select_data_for_screening_step5}
 # emne2_anon (uten post)
@@
 ```
-
-## Step 5: Item reduction
 
@@
 
+```{r}
+# clean environment to reset for next step
+# ref.: https://www.statology.org/clear-environment-in-r/ [3.9.2024]
+rm(list=ls())
+```
+
 ## Step 6: Factor extraction (i.e., EFA)
 
+Here I do an exploratory factor analysis (EFA) to examine how the items correlate and can be structured. The aim is to reduce the items to a limited set of factors that represent the data.
+
+```{r}
+# load data
+load(file="../data/processed_data/tpack-pilot_2022.Rdata") 
+```
+
+```{r items_EFA}
+TK_pos <- c("TK1","TK2","TK3","TK4","TK5","TK6","TK7","TK8")
+PK_pos <- c("PK1","PK2","PK3","PK4","PK5","PK6")
+CK_pos <- c("CK1","CK2","CK3","CK4","CK5")
+PCK_pos <- c("PCK1","PCK2","PCK3","PCK4","PCK5")
+TCK_pos <- c("TCK1","TCK2","TCK3","TCK4")
+TPK_pos <- c("TPK1","TPK2","TPK3","TPK4","TPK5")
+TPACK_pos <- c("TPACK1","TPACK2","TPACK3","TPACK4","TPACK5")
+
+#TPACK_complete <- c(TK_pos, PCK_full_pos, TPACK_full_pos) # minus XK
+
+pilot_list <- list(TK=TK_pos,
+                   CK=CK_pos,
+                   PK=PK_pos,
+                   PCK=PCK_pos,
+                   TCK=TCK_pos,
+                   TPK=TPK_pos,
+                   TPACK=TPACK_pos)
+
+```
+
+In the following we will only consider the pre-datasett, so we remove the post-data, which is to dissimilar from the pilot data.
+
+```{r select_data_for_screening_step6}
+# emne2_anon (uten post)
+data_TPACK_pilot_minuspost <- data_TPACK_pilot[
+  data_TPACK_pilot$datasett != "post" & 
+  data_TPACK_pilot$datasett != "pilot_trelis_q",]
+```
+
+```{r remove_collinear_items_EFA}
+### pilot
+data_TPACK_pilot_minuspost <- data_TPACK_pilot_minuspost %>%
+    # 25.5.2023: velg bort items som ikke er i TRELIS-Q
+    select(-c("TK8","PK2","PK4","PK5","PK6","PCK1","PCK3","PCK4","TCK2","TCK3",
+              "TPK2","TPK3","TPK5","TPACK1","TPACK5",
+              "TPACK3", # colinear in trelis-q; "TK1","TPK1",
+              "TK6","TK4", "PK1", "PK3"))
+
+```
+
+### Describe variables
+
+To get an overview of the quality of the data and to see what data it is appropiate to use further in the analysis.
+
+First count distribution of answers on the likert-scale used.
+
+```{r likert_distribution}
+# kilde: https://solomonkurz.netlify.app/post/2021-05-11-yes-you-can-fit-an-exploratory-factor-analysis-with-lavaan/
+data_TPACK_pilot_minuspost[,colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)] %>% 
+  pivot_longer(everything(),values_to="value") %>% 
+  count(value) %>% 
+  mutate(percent = (100 * n / sum(n)) %>% round(digits = 1))
+```
+
+Plot histogram for each items, to see how they compare.
+
+```{r histogram}
+# plot multi-panel histograms with facet_wrap
+# kilde: https://www.zevross.com/blog/2019/04/02/easy-multi-panel-plots-in-r-using-facet_wrap-and-facet_grid-from-ggplot2/
+
+data_long <- pivot_longer(data_TPACK_pilot_minuspost, 
+                          cols=colnames(data_TPACK_pilot_minuspost)[colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)],
+                          names_to="item", 
+                          values_to="likert_verdi",
+                          values_drop_na = TRUE)
+
+ggplot(data_long, aes(factor(likert_verdi))) +
+  geom_histogram(stat = "count") + 
+  facet_wrap("item")
+```
+
+<!-- Poor items: y10 (CK2), y13 (CK5) (because of range=2). -->
+
+### Preliminary analysis
+
+"It is recommended to carry out a preliminary analysis of the metric quality of the items to subject the most adequate items to EFA.
+For this purpose, it is recommended to analyze and report the mean, standard deviation, and item-test correlation of each one of the items, as well as the Cronbach’s alpha of the scales of the test.
+The researcher should decide whether to eliminate certain items" (Izquierdo et al, 2014, p.
+??)
+
+```{r help_function_correlation_matrix}
+### HELP FUNCTION
+
+antall_desimaler <- 2
+
+# print effect size with p-value
+print_starvalue <- function (pvalue, effectsize, bonferroni=1) {
+  ifelse (pvalue < 0.001/bonferroni, star <- "***",
+  ifelse (pvalue < 0.01/bonferroni, star <- "**",
+  ifelse (pvalue < 0.05/bonferroni, star <- "*",
+          star <- "")))
+  return (paste(format(effectsize,
+                       digits=antall_desimaler,
+                       nsmall=antall_desimaler),
+                star, sep="")) # antall desimaler=2
+}
+
+create_matrix <- function (df, method="pearson") {
+  # create correlation matrix for df (and covariance matrix combined, cf. Weston, 2006, table 2, p. 736) (cov lower left, cor upper right)
+
+  # since ordered data, use spearman; option in function call?
+
+  # left side
+  # use casewise deletion if missing
+  cov_matrix     <- cov(df, use="complete.obs", method=method)
+  # right side
+  cor_matrix     <- data.frame(matrix(ncol = ncol(df), nrow = 0)) # correlation matrix
+
+  i              <- 1          # start posisjon
+  teller         <- 0          # hvor mange plot lager jeg?
+  for (var1 in df[1:ncol(df)-1]) {
+    i            <- i + 1
+
+    # create right side: correlation 
+    right_side   <- data.frame(matrix(ncol = 0, nrow = 1)) # row values
+    for (var2 in df[i:ncol(df)]) {
+      teller     <- teller + 1   
+      cor_var1_var2 <- cor.test(var1,var2, method=method)
+      est_p      <- print_starvalue(cor_var1_var2$p.value,
+                                    cor_var1_var2$estimate)
+      right_side <- cbind(right_side,est_p) # add column
+    }
+
+    # add left-side of matrix: covariance
+    left_side    <- format(cov_matrix[i-1,1:i-1],
+                           digits=antall_desimaler)
+
+    temp_row     <- c(left_side, right_side)  # merge left and right
+    names(temp_row) <- colnames(df)
+    cor_matrix   <- rbind(cor_matrix, temp_row)   # add row
+  }
+
+  # add final row
+  final_row      <- format(cov_matrix[i,1:i],digits=antall_desimaler)
+  names(final_row) <- colnames(df)
+  cor_matrix     <- rbind(cor_matrix, final_row)
+
+  # add row with mean and sd based on rows/persons
+  row_mean       <- round(colMeans(df, na.rm=TRUE),
+                          digits=antall_desimaler)
+  row_sd         <- format(psych::SD(df),
+                          digits=antall_desimaler)
+  cor_matrix     <- rbind(cor_matrix,"mean"=row_mean)
+  cor_matrix     <- rbind(cor_matrix,"sd"=row_sd)
+
+  # set names of columns and rows
+  row.names(cor_matrix) <- c(colnames(df),"mean","sd")
+  return(cor_matrix)
+}
+
+lag_eigenvalue_tabell <- function(ev_list) {
+  no_ev <- length(ev_list) # antall eigenvalues
+  kum_var <- 0
+
+  # gjør om til tabell
+  ev_tabell <- data.frame(matrix(ncol=4,nrow=0))
+  for (i in 1:no_ev) {
+    ev <- ev_list[i]
+    kum_var <- kum_var + ev/no_ev
+    faktor <- cbind(as.integer(i), ev, ev/no_ev, kum_var)
+    #print(faktor)
+    ev_tabell <- rbind(ev_tabell, faktor)
+  }
+
+  colnames(ev_tabell) <- c("Factor","Eigenvalue","Variance","Cummulativ variance")
+
+  pander(format(ev_tabell,digits=2))
+}
+
+```
+
+```{r calculate_mean_sd}
+# select Likert columns
+data_pilot <- data_TPACK_pilot_minuspost[,colnames(data_TPACK_pilot_minuspost) %in% unlist(pilot_list)]
+
+# remove items with over 50 % NA
+df_noNA <- data_pilot[ , colSums(is.na(data_pilot)) < nrow(data_pilot)/2] 
+```
+
+#### eigenvalues
+
+```{r}
+# check for collinearity; values near or below 0
+ev <- eigen(cor(df_noNA[,colnames(df_noNA) %in%
+                                unlist(pilot_list)],
+                use="pairwise.complete.obs"), only.values=TRUE)
+
+lag_eigenvalue_tabell (ev$values)
+```
+
+#### item-test correlation
+
+Check to see if an item corresponds to the average test-result for the intended factor (e.g., if the TK-items correspond to the other TK-items).
+
+Ref.: https://www.questionmark.com/resources/blog/psychometrics-101-item-total-correlation/ https://en.wikipedia.org/wiki/Item-total_correlation
+
+```{r item_test}
+# get data for each test / construct
+TK    <- df_noNA[,colnames(df_noNA) %in% TK_pos]
+PK    <- df_noNA[,colnames(df_noNA) %in%PK_pos]
+CK    <- df_noNA[,colnames(df_noNA) %in%CK_pos] # NA
+TCK   <- df_noNA[,colnames(df_noNA) %in%TCK_pos]
+TPK   <- df_noNA[,colnames(df_noNA) %in%TPK_pos]
+PCK   <- df_noNA[,colnames(df_noNA) %in%PCK_pos]
+TPACK <- df_noNA[,colnames(df_noNA) %in% TPACK_pos]
+
+# check correlation of items to construct
+pander(create_matrix(data.frame(TK,    "TK"=rowMeans(TK, na.rm=TRUE))))
+pander(create_matrix(data.frame(PK,    "PK"=rowMeans(PK, na.rm=TRUE))))
+pander(create_matrix(data.frame(CK,    "CK"=rowMeans(CK, na.rm=TRUE))))
+pander(create_matrix(data.frame(TCK,   "TCK"=rowMeans(TCK, na.rm=TRUE))))
+pander(create_matrix(data.frame(TPK,   "TPK"=rowMeans(TPK, na.rm=TRUE))))
+pander(create_matrix(data.frame(PCK,   "PCK"=rowMeans(PCK, na.rm=TRUE))))
+pander(create_matrix(data.frame(TPACK, "TPACK"=rowMeans(TPACK, na.rm=TRUE))))
+
+```
+
+TCK2 has low item-test correlation.
+
+#### Reliability: Cronbach's alpha
+
+Assumptions for calculating Cronbach's alpha: uni-dimensionality (one factor = factor analysis) and tau-equality (equal covariance) [@revelle_reliability_2019].
+
+##### TK
+
+```{r uni-dimensionality}
+parallel <- fa.parallel(TK,fm='uls',fa='fa',SMC=TRUE) #, cor="poly", correct=0 ,use="complete.obs") #)
+parallel$fa.values
+```
+
+```{r tau-equality, warning=T}
+pander(create_matrix(TK))
+```
+
+Forutsetninger ikke oppfylt: ikke uni-dimensional (2 faktorer fra parallell analysis, og 1 fra eigenvalue), og ikke tau-equality (TK8 passer ikke på tau-equality).
+
+```{r alpha}
+psych::alpha(TK)
+```
+
+##### CK
+
+```{r warning=FALSE}
+parallel <- fa.parallel(CK,fm='uls',fa='fa',SMC=TRUE,  cor="poly", correct=0,use="complete.obs") #,cor="poly")
+parallel$fa.values
+pander(create_matrix(CK))
+psych::alpha(CK) 
+```
+
+Uni-dimensional, og nesten tau-equality. CK4 "ødelegger". CK4 er allerede tagget for fjerning.
+
+##### PK
+
+```{r warning=FALSE,eval=F}
+parallel <- fa.parallel(PK,fm='uls',fa='fa',SMC=TRUE,  cor="poly", correct=0,use="complete.obs") #,cor="poly")
+parallel$fa.values
+pander(create_matrix(PK))
+psych::alpha(PK)
+
+```
+
+Uni-dimensional. Bortimo tau-equality. PK4 og PK5 "ødelegger" litt. Begge disse er allerede tagget for fjerning. PK6 senker også reliabiliteten. (Denne er dog tatt med videre til TRELIS-q)
+
+##### TCK
+
+```{r warning=FALSE}
+parallel <- fa.parallel(TCK,fm='uls',fa='fa',SMC=TRUE,  cor="poly", correct=0,use="complete.obs") #,cor="poly")
+parallel$fa.values
+pander(create_matrix(TCK))
+psych::alpha(TCK)
+
+```
+
+Uni-dimensional. Ikke tau-equality. TCK1 ødelegger.
+
+##### TPK
+
+```{r warning=FALSE,eval=T}
+parallel <- fa.parallel(TPK,fm='uls',fa='fa',SMC=TRUE) #,  cor="poly", correct=0,use="complete.obs") #,cor="poly")
+parallel$fa.values
+pander(create_matrix(TPK))
+psych::alpha(TPK)
+
+```
+
+Uni-dimensional og omtrent tau-equality.
+
+##### PCK
+
+```{r warning=FALSE}
+parallel <- fa.parallel(PCK,fm='uls',fa='fa',SMC=TRUE,  cor="poly", correct=0,use="complete.obs") #,cor="poly")
+parallel$fa.values
+pander(create_matrix(PCK))
+psych::alpha(PCK)
+
+```
+
+Uni-dimensional og omtrent tau-equality. PCK1 og PCK3 kan fjernes. (men PCK1 er tatt med videre)
+
+##### TPACK
+
+```{r warning=FALSE, eval=T}
+  parallel <- fa.parallel(TPACK,fm='uls',fa='fa',SMC=TRUE,
+                          cor="poly", correct=0,use="complete.obs") #,cor="poly")
+  parallel$fa.values
+  pander(create_matrix(TPACK))
+  psych::alpha(TPACK)
+
+```
+
+Uni-dimensional og halvveis tau-equality. TPACK1 "ødelegger" litt.
+
+#### Results of preliminary analysis
+
+TK7 and CK4 stand apart as poor items.
+
+Reset data for EFA
+
+```{r}
+# remove columns
+data_modified <- within(data_TPACK_pilot_minuspost, rm("TK7","CK4"))
+
+# pick Likert columns
+df <- data_modified[,colnames(data_modified) %in% unlist(pilot_list)]
+
+# remove respondents with more than 50% NA
+df_run1 <- df[ , colSums(is.na(df)) < nrow(df)/2]
+
+```
+
+### EFA
+
+The aim of the exploratory factor analysis is to find the lowest number of items that is needed to represent the dataset.
+
+Estimation is done using maximum-likelihood estimation. Rotation is oblique (oblimin). tenBerge is used to estimate scores (ref. Psych-package help-file).
+
+#### run 1
+
+The suitability of the data for doing an EFA is tested by calculating KMO og Bartlett.
+
+```{r}
+# KMO og Bartlett
+KMO_base      <- KMO(df_run1) # addition of PCK1 changed KMO from 0.5 to 0.36
+bartlett_base <- cortest.bartlett(df_run1, n=5,diag=TRUE)
+```
+
+The number of factors is estimated using parallel analysis [@worthington_scale_2006].
+
+```{r parallel_analysis_run1}
+# number of factors: eigenvalue, scree or parallel (or theoretical)
+# polychoric is used for ordered data with skewness, Watkins (2018)
+
+parallel_base <- fa.parallel(df_run1,fm='mle',fa='fa',SMC=TRUE,  use="complete.obs") #, )correct=0,cor="poly",
+
+#parallel <- fa.parallel(df,fm='alpha',fa='fa',nfactors=7,use="complete.obs",cor='poly', error.bars=TRUE, SMC=TRUE) # complete.obs = delete NA; poly for fa recommended av Revelle & Condon, 2019
+```
+
+```{r}
+# eigenvalues
+lag_eigenvalue_tabell (parallel_base$fa.values)
+
+print(format(list("KMO" = KMO_base$MSA, "Bartlett p =" = bartlett_base$p.value),digits=2))
+```
+
+Due to the ambiguity of the parallel analysis, several factor solutions are tested (i.e., 2-factor, 3-factor and 4-factor).
+
+##### 2-factor
+
+```{r efa_ordinal2}
+# FA
+# ref.: https://www.rdocumentation.org/packages/psych/versions/2.2.5/topics/fa
+
+# print covariation matrix all items
+# bruk fm=uls, ikke pa (Flora et al, 2012)
+# bruk polykorisk korrelasjon (Flora et al, 2012)
+# use covariance matrix (correlation matrix fails because of missing)
+
+model_2factor_run1 <- fa(df_run1, nfactors = 2,rotate = "oblimin", fm="mle", scores="tenBerge") # missing=TRUE,impute="mean", ,oblique.scores=TRUE, cor="poly", correct=0 covar=TRUE,
+
+print.psych(fa.sort(model_2factor_run1))
+```
+
+##### 3-factor
+
+```{r efa_ordinal3}
+# FA
+# ref.: https://www.rdocumentation.org/packages/psych/versions/2.2.5/topics/fa
+
+# print covariation matrix all items
+# bruk fm=uls, ikke pa (Flora et al, 2012)
+# bruk polykorisk korrelasjon (Flora et al, 2012)
+# use covariance matrix (correlation matrix fails because of missing)
+
+model_3factor_run1 <- fa(df_run1, nfactors = 3,rotate = "oblimin", fm="mle", scores="tenBerge") # missing=TRUE,impute="mean", covar=TRUE,,oblique.scores=TRUE, cor="poly", correct=0
+
+print.psych(fa.sort(model_3factor_run1))
+```
+
+##### 4-factor
+
+```{r efa_ordinal4}
+# FA
+# ref.: https://www.rdocumentation.org/packages/psych/versions/2.2.5/topics/fa
+
+# print covariation matrix all items
+# bruk fm=uls, ikke pa (Flora et al, 2012)
+# bruk polykorisk korrelasjon (Flora et al, 2012)
+# use covariance matrix (correlation matrix fails because of missing)
+
+model_4factor_run1 <- fa(df_run1, nfactors = 4,rotate = "oblimin", fm="mle", scores="tenBerge") # missing=TRUE,impute="mean", covar=TRUE,,oblique.scores=TRUE, cor="poly", correct=0
+
+print.psych(fa.sort(model_4factor_run1))
+```
+
+```{r}
+# eigenvalues
+print(model_3factor_run1$values)
+```
+
+Eigenvalues suggests 3 factors
+
+The cutoff for significant loadings ...
+
+```{r}
+# cutoff verdi gitt av formel i Watkins, 2018 (avsnitt: Interpretation of results)
+n <- model_3factor_run1$n.obs
+c <- 4.084/sqrt(n-2)   # n=100 gir 3.92 p<5% (z-dist); n=30 gir 4.084 for p<5% (t-dist)
+print(c)
+# bare print loadings med verdier over 0.32
+print(model_3factor_run1$loadings,cutoff = c)
+
+print.psych(model_3factor_run1)
+```
+
+Weak communality ...
+
+```{r communality}
+# h2 = communality; bare print items med communality under 0.4 siden disse kan fjernes (Worthington, 2006); eller 0.6 for små utvalg? (ref)
+# kilde: https://stackoverflow.com/questions/30467834/r-print-cutoff-values-under-a-certain-value
+print(model_3factor_run1$communality[model_3factor_run1$communality<0.6])
+```
+
+Remove poor items from run 1 due to cross-loadings ...
+
+```{r}
+# 26.5.2023, remove cross-loadings
+df_run2 <- within(df_run1, rm("TK5"))
+```
+
+Kan vurdere å beholde CK1 med h2 = 0.37.
+
+#### run 2
+
+The EFA is repeated, since an item was removed.
+
+Suitability of the data for doing an EFA.
+
+```{r}
+# KMO og Bartlett
+KMO_base      <- KMO(df_run2) # addition of PCK1 changed KMO from 0.5 to 0.36
+bartlett_base <- cortest.bartlett(df_run2, n=4,diag=TRUE)
+```
+
+```{r parallel_analysis_run2}
+# fm = mle, alpha, pa
+
+#if (do_parallel == 1) {
+
+# number of factors: eigenvalue, scree or parallel (or theoretical)
+# parallel analysis (Worthington, 2006)
+# polychoric is used for ordered data with skewness, Watkins (2018)
+  # "pa (parallel analysis) most accurately estimated the number of factors" (Li et al., 2020)
+
+parallel_base <- fa.parallel(df_run2,fm='mle',fa='fa',SMC=TRUE) #, ),  correct=0,cor="poly",use="complete.obs"
+#parallel <- fa.parallel(df,fm='alpha',fa='fa',nfactors=7,use="complete.obs",cor='poly', error.bars=TRUE, SMC=TRUE) # complete.obs = delete NA; poly for fa recommended av Revelle & Condon, 2019
+```
+
+```{r}
+# eigenvalues
+#parallel$fa.values
+lag_eigenvalue_tabell (parallel_base$fa.values)
+
+print(format(list("KMO" = KMO_base$MSA, "Bartlett p =" = bartlett_base$p.value),digits=2))
+```
+
+##### 2 factor
+
+```{r}
+model_2factor_run2 <- fa(df_run2, nfactors = 2, rotate = "oblimin", fm="mle", scores="tenBerge") # ,oblique.scores=TRUE, cor="poly", correct=0
+
+print.psych(fa.sort(model_2factor_run2))
+```
+
+##### 3 factor
+
+```{r}
+# equamax for complex structure, ref. Schmitt & Sass, 2011
+# CI with bootstrapping, n.iter=1000
+# p = .8 (Cohen, 1990)
+model_3factor_run2 <- fa(df_run2, nfactors = 3, rotate = "oblimin", fm="mle", scores="tenBerge") #, missing=TRUE, impute="mean") # n.iter=1000, p=.8,  #,oblique.scores=TRUE, cor="poly", correct=0)
+
+print.psych(fa.sort(model_3factor_run2))
+
+#factor_model2
+```
+
+##### 4 factor
+
+```{r}
+model_4factor_run2 <- fa(df_run2, nfactors = 4, rotate = "oblimin", fm="mle", scores="tenBerge") #,oblique.scores=TRUE, cor="poly", correct=0)
+
+print.psych(fa.sort(model_4factor_run2))
+```
+
+Check 2nd order solution.
+
+```{r}
+# korrelation matrix factors
+#format(round(factor_model$r.scores,1), nsmall=1) 
+order2 <- model_3factor_run2$Phi # 2nd order factors
+print(order2)
+```
+
+```{r}
+# eigenvalues
+print(model_3factor_run2$values)
+
+# loadings
+print.psych(model_3factor_run2) #, cut=c)
+```
+
+Remove items "with factor loadings less than .32 or cross-loadings less than .15 difference from an item’s highest factor loading.
+In addition, they should also delete items that contain absolute loadings higher than a certain value (e.g., .32) on two or more factors." (Worthington and Whittaker, 2006).
+
+#### Assessment
+
+Felles varians: hvor mye felles varians forklarer faktorene? Faktorene bør tilsammen forklare minst 50% av variansen (Christophersen, 2018).
+
+```{r}
+p <- model_3factor_run2$Vaccounted
+print(p)                           # se på cumulative var
+```
+
+Kommunaliteten (h2) beskriver hvor stor del av variansen til indikatoren som faktorene "forklarer".
+Residualvariansen (u2) er den uforklarte variansen til indikatoren (item/spørsmålet).
+Det er i praksis vanskelig å få særlig høyere kommunalitet enn 65% i følge Christophersen (2018).
+
+```{r communality7}
+# Mundfrom et al., 2005: Råd: høy kommunalitet (>0.6) 
+# dette er unstandardized loadings / covariance matrix ??
+print(model_3factor_run2$communality[model_3factor_run2$communality>0.6])
+# dette er standardized loadings ??
+#print(efa7_ml_pro$communalities[efa7_ml_pro$communalities>0.6])
+
+# avoid below 0.4
+print(model_3factor_run2$communality[model_3factor_run2$communality<0.4])
+```
+
+Vurder modellen: Hvor sterkt korrelerer indikatorene i modellen?
+Indikatorer med r>0.80 er for sterkt korrelert (Christophersen, 2018).
+Indikatorer som korrelerer svært svakt kan tyde på at et vesentlig begrepsaspekt er utelatt av modellen.
+
+Når det er korrelerte faktorer vi bruker, så vil strukturmatrisen gi korrelasjonene, mens pattern matrisen gir loadings.
+
+```{r}
+# for sterke korrelasjoner
+print(model_3factor_run2$Structure, cutoff=0.8)
+
+model_3factor_run2$Structure
+
+# meget svake korrelasjoner
+#efa7_ml_pro$Structure[efa7_ml_pro$Structure<0.1]
+```
+
+restledd skal helst ikke være større enn +- 0.05
+
+```{r}
+round(model_3factor_run2$residual, digits=2)
+```
+
+```{r}
+# check reliability
+#cronbach.alpha(df_run2, na.rm=TRUE)
+
+# tegn factor_model
+stimodel <- fa.diagram(model_3factor_run2)
+#print(stimodel)
+```
+
+```{r}
+# compare with intervall pearson (appx normal distributed)
+pander(create_matrix(df_run2))
+```
+
+### McDonald's omega
+
+Fra help-file til omega: "A recommendation that should be heeded, regardless of the method chosen to estimate \omega_hω, is to always examine the pattern of the estimated general factor loadings prior to estimating \omega_hω.
+Such an examination constitutes an informal test of the assumption that there is a latent variable common to all of the scale's indicators that can be conducted even in the context of EFA.
+If the loadings were salient for only a relatively small subset of the indicators, this would suggest that there is no true general factor underlying the covariance matrix.
+Just such an informal assumption test would have afforded a great deal of protection against the possibility of misinterpreting the misleading \omega_hω estimates occasionally produced in the simulations reported here." (Zinbarg et al., 2006, p 137)"
+
+```{r eval=T}
+alpha(df_run2[,c("TK1", "TK2", "TK3")])
+alpha(df_run2[,c("CK1", "PCK2", "PCK5", "CK5", "CK3", "CK2")])
+alpha(df_run2[,c("TPK1","TPK4","TCK1","TCK4", "TPACK2","TPACK4")]) # , "TPK4"
+```
+
+```{r}
+omega(df_run2, fm="mle") #, rotate="Promax")
+```
+
+```{r}
+# clean environment to reset for next step
+# ref.: https://www.statology.org/clear-environment-in-r/ [3.9.2024]
+rm(list=ls())
+```
+
 # Stage 3
@@
 ## Step 7: Tests of dimensionality (i.e., CFA)
+
+The following choices were made when determining how to test the data:
+
+* treat the data as interval data.
+* missing data: use maximum-likelihood [@mcneish_challenging_2017; @schafer_missing_2002].
+* Normal or nonnormal distribution of residuals?
+Normality is an important assumption for doing test statistics.
+Nonnormal means skew>2 and kurtosis>9 (ref?).
+Some of the residuals showed high skewness (i.e., non-normality) (e.g., TPACK2).
+In such cases, WLS can be an option to OLS [@christophersen_introduksjon_2018].
+Some items also showed evidence of being homoscedastic (i.e., non-equal variance in the residuals): CK123, PCK5, TCK4.
+In case of nonnormal data the Satorra-Bentler scaled chi-square test statistic and robust SE can be used [@flora_old_2012; @mcneish_challenging_2017].
+* I aim for approximate fit, not exact fit [@flora_old_2012; weston_brief_2006].
+
+yuan.bentler forutsetter adf-estimering.
+(McNeish, 2017) som krever at utvalgsstørrelser er minst p(p+1)/2, der p er antall items i modellen.
+Med p=13 blir det 91 respondenter.
+Mitt datautvalg er større enn dette (n=103).
+
+test="yuan.bentler(.mplus)" (McNiesh, 2017: small samples? NB: adf)  
+se="robust.sem" # standard error  
+information="observed" # ref. Schafer & Graham, 2002, p. 162
+
+### Specification of models to be tested
+
+The purpose is to confirm the measurement model, that is, the indicators loading to the latent factors.
+
+The two models are based on the EFA and are nested (e.g., the two_factor model is a constrained version of the three_factor model, i.e., TK and TPACK are merged).
+
+```{r specify_models}
+three_factor <- '
+PCK   =~ CK123 + PCK2 + PCK5 + CK5 
+TK    =~ TK1b + TK2 + TK34 
+TPACK =~ TPACK2 + TPACK4 + TCK4 + TPK4 + TCK1 + TPK1
+'
+
+two_factor   <- '
+PCK   =~ CK123 + PCK2 + PCK5 + CK5
+TPACK =~ TK1b + TK2 + TK34 + TPACK2 + TPACK4 + TCK4 + TPK4 + TCK1 + TPK1
+'
+```
+
+### Test of models
+
+Two models are compared. A two factor and a three factor solution derived from the EFA.
+
+```{r}
+# load data
+load(file="../data/processed_data/tpack-trelis_q_paper1_2023.Rdata")
+```
+
+NMAR are removed from the dataset.
 
@@
 # remove NMAR + NTNU (19)
-data_TPACK_PST_noNMAR <-
-  data_trelis_q_ren[-c(19,34,44,45,46,95,102),]
-
+CFA_TPACK_PST <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),] 
+```
+
+Some help functions are defined.
+
+```{r help_functions_cfa}
+
+### HELP FUNCTIONS
+do_cfa        <- function (cfaspec, cfadata) {
+  # estimator og missing: Rhemtulla et al. (2012)
+  # ref.: https://statistics.ohlsen-web.de/multiple-imputation-with-mice/
+  TPACK_PST_0 <- cfa(cfaspec, data= cfadata, std.lv=TRUE,
+                     missing="fiml", #robust.two.stage",
+                     estimator="ML", #missing="pairwise", 
+                     #test="satorra.bentler", 
+                     test="yuan.bentler",
+                     information="observed",
+                     se="robust.huber.white") # robust.sem
+
+  return (TPACK_PST_0)
+}
+
+print_model <- function (cfamodel) {
+  # print results of CFA
+
+  print("summary")
+  print(summary(cfamodel, standardized=TRUE, fit.measures=TRUE))
+
+  print("estimates")
+  # CI suggested by Cheung et al. (2023) for discriminant validity < 1 (one-tailed)
+  print(parameterEstimates(cfamodel, rsquare = TRUE, level=0.90))
+
+  print("modindices")
+  print(modindices(cfamodel, sort=TRUE, maximum.number = 5))            # mi should be below 3.84
+
+  print("reliability")
+  rel <- compRelSEM(cfamodel, dropSingle = FALSE)
+  print(rel)
+  return (rel)   # for later comparison of reliability between models
+}
+
+```
+
+#### two_factor
+
+```{r estimate_2factor}
+# egenfaktor
+model_two_factor <- do_cfa(two_factor, CFA_TPACK_PST)
+rel_two_factor   <- print_model(model_two_factor)
+```
+
+#### three_factor
+
+```{r estimate_3factor}
+# egenfaktor
+model_three_factor <- do_cfa(three_factor, CFA_TPACK_PST)
+rel_three_factor   <- print_model(model_three_factor)
+```
+
+### Summary of fit
+
+```{r CFA_results_comparison}
+# fit indices
+columnnames  <- c("AIC", "chi-square", "df", "p-value",
+                  "RMSEA", "RMSEA (90%) CI lower","RMSEA (90%) CI upper",
+                  "CFI","TLI", "SRMR") #, "chi2/df", 
+                 #"chi-square scaled Swain", "p-value Swain", 
+                 #"swain chi2/df")
+
+# fit criteria
+fit_criteria <- c("x", "x", "x", ">0.05", 
+                  "<0.06(0.05-0.10)", "x", "x",
+                  ">0.95(0.9-0.95)", ">0.95", "<0.08(0.08-0.15)") #, "<2", "x", ">0.05", "<2") # Weston and Gore (2006); Gallagher and Brown (2013)
+
+fit_2factor  <- round(fitMeasures(model_two_factor,
+                  c("aic","chisq.scaled", "df.scaled",
+                    "pvalue.scaled",
+                    "rmsea.robust", "rmsea.ci.lower.robust", "rmsea.ci.upper.robust",
+                    "cfi.robust","tli.robust","srmr")),3)
+
+fit_3factor  <- round(fitMeasures(model_three_factor,
+                  c("aic","chisq.scaled", "df.scaled",
+                    "pvalue.scaled",
+                    "rmsea.robust", "rmsea.ci.lower.robust", "rmsea.ci.upper.robust",
+                    "cfi.robust","tli.robust","srmr")),3)
+
+rownames     <- c("lower bounds", "two_factor", "three_factor")
+
+fit_table    <- rbind(fit_criteria, fit_2factor)
+fit_table    <- rbind(fit_table, fit_3factor)
+
+colnames(fit_table)  <- columnnames
+row.names(fit_table) <- rownames
+
+pander(fit_table)
+```
+
+### Results
+
+The correlation matrix helps assess the data.
+
+```{r covariance_matrix}
+# covariance matrix / "TK1a","TK567"
+pander(create_matrix(CFA_TPACK_PST[,c("CK123", "CK5", "PCK2" , "PCK5" , "TK1b", "TK2", "TK34", "TCK1","TCK4", "TPK1","TPK4","TPACK2","TPACK4")])) # ,"TPK4"
+```
+
+To calculate the students TPACK-score we would only calculate the average of all the items, i.e. an integrative TPACK-model, where all the elements contributes to the students TPACK.
+
+```{r scores}
+TK    <- c("TK1b", "TK2", "TK34") # "TK1a","TK567"
+PCK   <- c("CK123", "PCK2" , "PCK5" , "CK5")
+TPACK <- c("TPK1","TPK4","TPACK2","TCK1","TPACK4", "TCK4") # , "TPK4"
+
+TPACK_PST_items <- list("TK" = TK,
+                        "PCK" = PCK,
+                        "TPACK" = TPACK) # everyone contributes til TPACK; c(TK,"TCK4",TPK, "CK123","PCK2","PCK5"))
+
+df    <- CFA_TPACK_PST[,unlist(TPACK_PST_items[1:3])]
+describe(df)      # with missing (see n)? (na.rm=TRUE)
+
+score_TPACK <- scoreItems(TPACK_PST_items,df)
+describe(score_TPACK$scores)
 ```
@@
 
+Three models are tested for reliability. The CFA-model (the estimated correlation matrix), the correlation matrix from the observed variables, and the observed variables.
+
+```{r omega_mle4}
+omega0  <- omegaFromSem(model_three_factor)
+omega0
+
+# get correlation matrix of observed variables
+cor_mat <- lavInspect(model_three_factor, what="cor.ov")
+omega(cor_mat, nfactors=3)
+
+omega(CFA_TPACK_PST[,c("CK123", "CK5", "PCK2" , "PCK5" , "TCK1","TCK4", "TPK1","TPK4","TPACK2","TPACK4", "TK1b", "TK2", "TK34")]) # ,"TPK4"
+```
+
 ## Step 9: Tests of validity
 
+I test for discriminatory and convergent validity following @ronkko_updated_2022
+
+```{r}
+# ref.: https://link.springer.com/article/10.1007/s10490-023-09871-y
+source("../../scripts/measureQ.R")
+```
+
+```{r measureQ, eval=T}
+measureQ(model_three_factor, CFA_TPACK_PST)
+```
+
 # Discussion and conclusion
+
+The results are discussed in @karlsen_assessing_2024.
 

### 2024-09-04 Rmd/Paper1.Rmd
Increased readability of document

@@
 author:       "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
-date:         "2024-09-03"
+date:         "2024-09-04"
 bibliography: "../data/meta_data/references.bib"
@@
     toc: true
-    toc_depth: 3
+    toc_depth: 4
     toc_float: true
@@
 The document is inspired by the idea of a computational essay [@odden_using_2023].
+
+This document is large.
+This is mainly due to the analyses done in step 5: item reduction.
+Especially the scatterplots to check for linearity, checks for autocorrelation and multicollinearity, and checks of the normal distribution of residuals, produce a lot of output.
+They are provided here for transparency.
+Using the menu on the left, you can skip this part if you like.
 

### 2024-09-04 Rmd/Paper1.Rmd
Changed filename of dataset to correspond with paper2

@@
 ```{r}
+# readxls stores as tibble, not dataframe?
 # make tibble into data.frame to treat it similarly to pilot dataset
@@
 save(data_TRELIS_Q, 
-     file="../data/processed_data/tpack-trelis_q_paper1_2023.Rdata")
+     file="../data/processed_data/tpack-trelis_q_partial_2023-paper1.Rdata")
 ```
@@
 load(file="../data/processed_data/tpack-pilot_2022.Rdata") 
-load(file="../data/processed_data/tpack-trelis_q_paper1_2023.Rdata")
+load(file="../data/processed_data/tpack-trelis_q_partial_2023-paper1.Rdata")
 ```

### 2024-09-04 Rmd/Paper2.Rmd
Several corrections

--- 2024-04-02 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
@@
 ---
-title: "Multiple regression: TRELIS-Q WP5 data"
+title:        "Paper 2: survey analysis"
 author:       "Niklas Karlsen"
-date: "2024-04-02"
+date:         "2024-09-04"
+bibliography: "../data/meta_data/references.bib"
+csl:          "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl" # (source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html)
 output: 
@@
 
-knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/bookchapter.TRELIS-Q-WP5.html')) })
+knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/paper2.html')) })
 ---
 
-# Initialize
+# Introduction
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE) #,results="asis")
-
-Sys.setenv(LANG = "no_NB.UTF-8")            # for norwegian æ,ø og å. Kilde:
-Sys.setlocale("LC_ALL","Norwegian")
-# https://stackoverflow.com/questions/16347731/how-to-change-the-locale-of-r
-
-plot_to_file = 1
-```
+This document presents the quantitative analyses done in paper 2. Descriptive and inferential analyses are done. The inferences are based on multiple regression analyses.
 
@@
 
-```{r load_libraries, echo=FALSE, warning=FALSE, message=FALSE}
+```{r load_libraries}
+
+### prepare raw data
+library(readxl)                             # read .xlsx-files. Kilde: https://readxl.tidyverse.org/
+
+### presentation of results
 #library(pander)                             # prettyprint for tables
+
+# NB! Collision between plyr and dplyr!
 library(plyr)
+library(dplyr)                              # rename columns
+
 library(ggplot2)
@@
 
+### analyses
 library(mice)                               # for md.pattern (missing)
@@
 #library(lmtest)
+library(nlme)                               # for ICC - variance HLM
+```
+
+## Define constructs
+
+The constructs are based on @karlsen_assessing_2024.
+
+```{r load_constructs}
+## TPACK-PST for TRELIS-Q
+TK_pos_trelisq    <- c("TK1b","TK2","TK34") 
+PCK_pos_trelisq   <- c("CK123","CK5","PCK2","PCK5")
+TPACK_pos_trelisq <- c("TPACK2","TPACK4","TCK1","TCK4","TPK1","TPK4")
+
+TPACK_trelis_q    <- list(TK=TK_pos_trelisq,
+                     PCK=PCK_pos_trelisq,
+                     TPACK=TPACK_pos_trelisq)
+```
+
+# Method
+
+The following approach is used in this document.
+First, the raw data is loaded and prepared for analyses.
+Second, descriptive statistics are created.
+Third, inferential statistics (i.e., regression analyses) are created.
+Finally, some comments are provided on the results.
+
+## Data management
+
+For the survey, the complete TRELIS-Q dataset was used.
+
+### Load data
+
+```{r}
+folder_trelis_q <- "../data/rawdata/"
+trelis_q <- "trelis-q-datafil-endelig-juni2023.xlsx"
+
+path_trelis_q <- paste(folder_trelis_q, trelis_q, sep='')
+
+data_trelis_q <- read_excel(path_trelis_q, sheet=4)
+```
+
+### Data cleaning
+
+```{r}
+# only keep relevant columns
+data_trelis_q <- data_trelis_q[,51:74]
+```
+
+```{r}
+# keep some columns; covariates
+# 68. Hvilket studiested studerer du ved?	
+# 69. Går du på grunnskolelærerutdanning for trinn 1-7 eller trinn 5-10?	
+# 70. Hvilket studieår er du i?	
+# 71. Er du:
+
+# likert-spørsmål: 48 - 66
+# åpne spørsmål: 67
+
+data_trelis_q <- rename(data_trelis_q,
+                            "Trinn" =
+                            "69. Går du på grunnskolelærerutdanning for trinn 1-7 eller trinn 5-10?")
+data_trelis_q <- rename(data_trelis_q,
+                            "Gender" =
+                            "71. Er du:")
+data_trelis_q <- rename(data_trelis_q,
+                            "studiested" =
+                            "68. Hvilket studiested studerer du ved?")
+data_trelis_q <- rename(data_trelis_q,
+                            "semester" =
+                            "70. Hvilket studieår er du i?")
+
+# change from year to semester
+#data_trelis_q$semester <- rep(7, 52)  # 4 år =  7. semester (høst)
+
+# reorder some columns
+data_trelis_q <- data_trelis_q %>% relocate(Trinn, .before = "48. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
+data_trelis_q <- data_trelis_q %>% relocate(Gender, .before = "48. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
+data_trelis_q <- data_trelis_q %>% relocate(studiested, .before = "Trinn")
+data_trelis_q <- data_trelis_q %>% relocate(semester, .before = "Trinn")
+```
+
+Rename items to TPACK-lingo. (introduce bias in interpretation? Should it be named something neutral instead?)
+
+```{r}
+# rename columns: (bruker dplyr)
+# kilde: https://www.sharpsightlabs.com/blog/rename-columns-in-r/
+data_trelis_q_ren <- rename(data_trelis_q, 
+'CK123' = '48. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.',
+'CK4' = "49. Jeg forstår teknologiske prinsipper og virkemåter som er relevante for naturfag",
+'CK5' = '50. Jeg har kunnskaper om naturvitenskapelige praksiser og tenkemåter',
+'PCK2' = '51. Jeg vet hvordan jeg kan støtte elevene i å utvikle kritisk tenkning og argumentasjonsferdigheter i naturfag',
+'TK2' = "52. Jeg kan sette opp en algoritme når jeg programmerer",
+'TK1a' = "53. Jeg kan bryte ned problemer i mindre deler når jeg programmerer",
+'TK34' = "54. Jeg har kunnskap om forskjellige programmeringsverktøy, f.eks. Makecode for Micro:bit, Scratch, Python, og lignende",
+'TK1b' = "55. Jeg kan oppdage og rette feil, når jeg programmerer.",
+'TK567' = "56. Jeg vet hvordan jeg lager kode som inneholder f.eks. variabler, vilkår (if, else) og løkker (for, while)",
+'TCK1' = "58. Jeg kan bruke programmering til å utvikle enkle teknologiske systemer som består av deler som virker sammen.",
+'TCK4' = '57. Jeg kan bruke programmering til å utforske naturfaglige fenomener',
+'PK3' = '59. Jeg kan støtte elevene sosialt i deres læring i ulike fag, f.eks deres nysgjerrighet, kreativitet og samarbeidsevne',
+'PK1' = '60. Jeg vet hvordan jeg kan støtte elevenes refleksjon over egen læring i ulike fag',
+'PCK5' = '61. Jeg vet hvordan jeg kan vurdere elevenes læring på ulike måter i naturfag',
+'TPK1' = '62. Jeg vet hvordan jeg kan bruke programmering på måter som beriker undervisningen i ulike fag',
+'TPK4' = '63. Jeg tenker kritisk igjennom hvordan jeg kan bruke programmering i undervisningen i ulike fag',
+'TPACK2' = '64. Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå naturfaglige begreper og fenomener',
+'TPACK3' = '65. Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå naturvitenskapelige praksiser og tenkemåter.',
+'TPACK4' = '66. Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå hvordan teknologiske systemer fungerer')
+
+# change "5 Stemmer veldig godt" and "1 Stemmer veldig dårlig" to 5 and 1
+data_trelis_q_ren[data_trelis_q_ren == '1 Stemmer veldig dårlig'] <- "1"
+data_trelis_q_ren[data_trelis_q_ren == '5 Stemmer veldig godt'] <- "5"
+```
+
+```{r}
+# hva gjør jeg med 9 / VET IKKE? Gjør om til NA?
+# tell antall VET IKKE
+antall_VETIKKE <- length(which(data_trelis_q_ren == 'VET IKKE'))
+print(antall_VETIKKE)
+# gjør om til NA; OBS! Ikke gjør disse NA om til tall med impute, pmm senere
+data_trelis_q_ren[data_trelis_q_ren == 'VET IKKE'] <- NA # "9"
+
+# convert into numbers
+data_trelis_q_ren[,colnames(data_trelis_q_ren) %in% unlist(TPACK_trelis_q)] <-
+  apply(
+    data_trelis_q_ren[,colnames(data_trelis_q_ren) %in% unlist(TPACK_trelis_q)], 2,
+                    function(x) as.numeric(as.character(x)))
+```
+
+Hvor mange respondenter er det?
+
+```{r}
+nrow(data_trelis_q)
+nrow(data_trelis_q_ren)
+```
+
+```{r rename_dataset}
+data_TRELIS_Q <- data_trelis_q_ren
+```
+
+```{r}
+# remove open-ended question column
+data_TRELIS_Q <- data_TRELIS_Q[,1:23]
+```
+
+```{r}
+# readxls stores as tibble, not dataframe?
+# make tibble into data.frame to treat it similarly to pilot dataset
+data_TRELIS_Q <- data.frame(data_TRELIS_Q)
+```
+
+### save data
+
+```{r}
+save(data_TRELIS_Q,  
+     file="../data/processed_data/tpack-TRELIS-Q_complete_2023-paper2.Rdata")
+```
+
+```{r}
+# clean environment to reset for next step
+# ref.: https://www.statology.org/clear-environment-in-r/ [3.9.2024]
+rm(list=ls())
+```
+
+## Prepare data
+
+```{r load_data_cfa}
+# Rdata-file contains reduced / prepared datasets
+load(file="../data/processed_data/tpack-TRELIS-Q_complete_2023-paper2.Rdata")
+```
+
+```{r reload_constructs}
+## TPACK-PST for TRELIS-Q
+TK_pos_trelisq    <- c("TK1b","TK2","TK34") 
+PCK_pos_trelisq   <- c("CK123","CK5","PCK2","PCK5")
+TPACK_pos_trelisq <- c("TPACK2","TPACK4","TCK1","TCK4","TPK1","TPK4")
+
+TPACK_trelis_q    <- list(TK=TK_pos_trelisq,
+                     PCK=PCK_pos_trelisq,
+                     TPACK=TPACK_pos_trelisq)
 ```
@@
 
-## Define constructs
-
-```{r load_constructs}
-## TPACK-PST for TRELIS-Q
-TK_pos_trelisq    <- c("TK1b","TK2","TK34") 
-PCK_pos_trelisq   <- c("CK123","CK5","PCK2","PCK5")
-TPACK_pos_trelisq <- c("TPACK2","TPACK4","TCK1","TCK4","TPK1","TPK4")
-
-TPACK_PST_list    <- list(TK=TK_pos_trelisq,
-                     PCK=PCK_pos_trelisq,
-                     TPACK=TPACK_pos_trelisq)
-```
-
-# Data management
-
-## Load data
-
-```{r load_data_cfa}
-# Rdata-file contains reduced / prepared datasets
-load(file="../data/processed_data/tpack-TRELIS-Q_2023.Rdata")
-```
-
-## Prepare data
-
 Flytt origo
@@
 ```{r}
-data_trelis_q_ren[,unlist(TPACK_PST_list)] <- data_trelis_q_ren[,unlist(TPACK_PST_list)] - 3 
+data_TRELIS_Q[,unlist(TPACK_trelis_q)] <- data_TRELIS_Q[,unlist(TPACK_trelis_q)] - 3 
 ```
@@
 # identify NA as studiested 5(?) 
-data_trelis_q_ren[data_trelis_q_ren$studiested == "Nord universitet","Trinn"] <- "Trinn 5-10"
+data_TRELIS_Q[data_TRELIS_Q$studiested == "Institute3","Trinn"] <- "Trinn 5-10"
 ```
@@
 
+The complete TRELIS_Q dataset is almost identical to the trelis_q dataset used in paper 1, except that TRELIS_Q includes all answers from institute2.
+These were not all collected when paper 1 was written.
+The addition of these final answers did not add any new NMAR to the dataset.
+
 ```{r}
 # covariates
-md.pattern(data_trelis_q_ren[,c("studiested","Gender", "Trinn", "semester")])
+md.pattern(data_TRELIS_Q[,c("studiested","Gender", "Trinn", "semester")])
 
 # items / likert data
-md.pattern(data_trelis_q_ren[, unlist(TPACK_PST_list)])
+md.pattern(data_TRELIS_Q[, unlist(TPACK_trelis_q)])
 ```
@@
 ```{r fix_missing, eval=T}
-# remove NMAR + NTNU (19)
-data_clean <- data_trelis_q_ren[-c(19,34,44,45,46,95,102),]
+# remove NMAR + Institute9 (19)
+data_clean <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),]
 
@@
 ```{r}
-scores_v23 <- scoreItems(TPACK_PST_list, data_clean)
+scores_v23 <- scoreItems(TPACK_trelis_q, data_clean)
 
@@
 data_clean$studiested2 <- factor(data_clean$studiested, 
-                                 level=c("HVL","UiT","Nord universitet","Høgskolen i Innlandet","OsloMet"),
+                                 level=c("Institute1","Institute5","Institute3","Institute2","Institute4"),
                                  label=c("A", "A", "B", "C", "C"))
@@
 
-```{r}
-# save plot
-if (plot_to_file) {
-  ggsave("../output/TPACK_stolpediagram.png") #,height=8,width=15)
-  }
+# Multiple regression analysis
 
+The data is multilevel, as the PSTs (level 1) are grouped at the institutional level (level 2).
+
+## Check assumptions for regression
+
+### ICC
+
+To check whether to treat the data as multilevel, ICC is calculated [@christophersen_introduksjon_2018] (p. 109-111).
+
+```{r ICC}
+# variance decomposition
+null.mod <- lme(TPACK ~ 1, random =~ 1|studiested, data = data_clean)
+vcn <- VarCorr(null.mod)
+vcn
+```
+
+The variance of indviduals is given by the residual.
+The variance of institutions is givene by the intercept.
+If the variance of the intercept is larger than 0.05 of the total variance (intercept + residual), then the institutional variance contributes to explain the outcome.
+
+### Normality
+
+```{r normality}
+# covariates; 80/20 distribution
+# ref http://stackoverflow.com/questions/21639392/ddg#21639445
+barplot(prop.table(table(data_clean$Gender)))
+barplot(prop.table(table(data_clean$Trinn)))
+barplot(prop.table(table(data_clean$semester)))
+#hist(Institute4_r)
+#hist(Institute2_r)
+#hist(Institute1_r)
+#hist(Institute5_r)
+
+# variable
+hist(data_clean$TK)
+hist(data_clean$PCK)
+hist(data_clean$TPACK)
 ```
@@
 ```{r}
-ggplot(dc, aes(x=TPACK, y=var, fill=studiested2)) + 
-  geom_bar(stat="identity", position="dodge") +
-  #scale_fill_manual(values=viridis(3)) + 
-  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
-                  linewidth=.3,    # Thinner lines
-                  width=.2,
-                  position=position_dodge(.9)) +
-  ggtitle("PCK, TK og TPACK fordelt på studiested") +
-  guides(fill="none") +
-  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
-#  coord_cartesian(ylim = c(1,5)) +
-  ylab("verdi") +
-  xlab("") +
-  theme_bw() +
-  facet_wrap(~studiested2)
+describe_data(data_clean$TK)
+describe_data(data_clean$PCK)
+describe_data(data_clean$TPACK)
 ```
 
+### Linearity
+
+```{r scatterplot_linearity}
+# ref: https://stackoverflow.com/questions/29425892/how-do-i-loop-through-column-names-and-make-a-ggplot-scatteplot-for-each-one
+for (var1 in names(data_clean)[names(data_clean) %in% c("Gender", "Trinn", "semester", "TK", "PCK")]) {
+  xy_plot <- ggplot(data_clean, aes(.data[[var1]],TPACK)) +
+        geom_point() +                            # shape=1,alpha=1/5
+        geom_jitter() +
+        ggtitle(paste("scatterplot",var1,"TPACK"))
+
+  print(xy_plot)
+}
+
+#plot(data_clean$Trinn,         data_clean$TPACK)
+#plot(data_clean$semester,      data_clean$TPACK)
+#plot(data_clean$TK,            data_clean$TPACK)
+#plot(data_clean$PCK,           data_clean$TPACK)
+```
+
+```{r multicollinearity}
+#cor(data_clean$PCK,   data_clean$semester)
+#cor(data_clean$TPACK, data_clean$Gender)
+
+#cor(data_clean$TK,    data_clean$Gender)
+#cor(data_clean$TK,    data_clean$Trinn)
+
+cor(data_clean$TK,    data_clean$PCK)
+cor(data_clean$TK,    data_clean$TPACK)
+cor(data_clean$PCK,   data_clean$TPACK)
+```
+
+TK korrelererer veldig sterkt (0.84) med TPACK, noe som indikerer kollinearitet. (Jeong and Jung, 2016)
+
 ## Check for interactions
+
+Should I include interactions in the regression analysis?
 
@@
 
-# Regression analysis
-
-## Check assumptions for regression
-
-```{r normality}
-# covariates; 80/20 distribution
-# ref http://stackoverflow.com/questions/21639392/ddg#21639445
-barplot(prop.table(table(data_clean$Gender)))
-barplot(prop.table(table(data_clean$Trinn)))
-barplot(prop.table(table(data_clean$semester)))
-#hist(oslomet_r)
-#hist(innlandet_r)
-#hist(HVL_r)
-#hist(UIT_r)
-
-# variable
-hist(data_clean$TK)
-hist(data_clean$PCK)
-hist(data_clean$TPACK)
-```
-
-```{r}
-describe_data(data_clean$TK)
-describe_data(data_clean$PCK)
-describe_data(data_clean$TPACK)
-```
-
-```{r scatterplot_linearity, eval=F}
-plot(data_clean$Gender,        data_clean$TPACK)
-plot(data_clean$Trinn,         data_clean$TPACK)
-plot(data_clean$semester,      data_clean$TPACK)
-#plot(data_clean$oslomet_r,         data_clean$TPACK_v23)
-#plot(data_clean$innlandet_r,       data_clean$TPACK_v23)
-#plot(data_clean$HVL_r,             data_clean$TPACK_v23)
-#plot(data_clean$UIT_r,             data_clean$TPACK_v23)
-
-plot(data_clean$TK,            data_clean$TPACK)
-#plot(data_clean$TK[oslomet_r],data_clean$TPACK[oslomet_r]) # gr.A
-#plot(data_clean$TK[innlandet_r],data_clean$TPACK[innlandet_r]) # gr.B
-#plot(data_clean$TK[HVL_r],data_clean$TPACK[HVL_r]) # gr.A
-#plot(data_clean$TK[UIT_r],data_clean$TPACK[UIT_r]) # gr.B
-
-#scatterplot(TPACK_regdata, c("TK", "TPACK"), factor(studiested))
-#scatterplot(TPACK_regdata, c("TK", "TPACK"), factor(semester))
-
-plot(data_clean$PCK,           data_clean$TPACK)
-```
-
-```{r multicollinearity, eval=F}
-#cor(TK_v23, studiested_r)
-#cor(PCK_v23, studiested_r)
-cor(data_clean$PCK,   data_clean$semester)
-#cor(TPACK_v23, studiested_r)
-cor(data_clean$TPACK, data_clean$Gender)
-
-cor(data_clean$TK,    data_clean$Gender)
-cor(data_clean$TK,    data_clean$Trinn)
-
-cor(data_clean$TK,    data_clean$PCK)
-cor(data_clean$TK,    data_clean$TPACK)
-cor(data_clean$PCK,   data_clean$TPACK)
-```
-
-TK korrelererer veldig sterkt (0.84) med TPACK, noe som indikerer kollinearitet. (Jeong and Jung, 2016)
-
 ## Hierarchical RA for TK
@@
 ```{r TKlevel2_main}
-# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
 m3 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
@@
 ```{r TKlevel2_interaction}
-# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
 m4 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
@@
 ```{r PCKlevel2_main}
-# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
 m3 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
@@
 ```{r PCKlevel2_interaction}
-# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
 m4 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
@@
 ```{r level2_main}
-# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
 m3 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
@@
 ```{r level2_interaction}
-# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
 m4 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
@@
 ```
+
+# Discussion and conclusion
+
+The results are discussed in @karlsen_assessing_2024.
+
+# References
+
+<!-- citation(package) -->
+
+<div id="refs"></div>
+
+<!-- source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html
+
+# (APPENDIX) Appendix {-} -->
+
+# Session info
+<!-- # ref.: https://intro2r.com/proj_doc.html -->
+<!-- #sessionInfo() -->
+
+```{r}
+xfun::session_info()
+```

### 2024-09-04 Rmd/Paper2.Rmd
Added correct calculations to reproduce paper 2 results

@@
 
-This document presents the quantitative analyses done in paper 2. Descriptive and inferential analyses are done. The inferences are based on multiple regression analyses.
+This document presents the quantitative analyses done in paper 2.
+Descriptive and inferential analyses are done.
+The inferences are based on multiple regression analyses.
+The document is inspired by the idea of a computational essay [@odden_using_2023].
+
+The purpose is foremost to reproduce the results presented in paper 2.
+But I will also do a reanalysis excluding some interaction effects, which were included in the paper 2 analysis.
+In addition, I will compare the multiple regression analysis with multilevel modeling, to see if the results are similar, as well as when including cross-level interaction effects due to year of study (semester), since there could be differences in different years either due to changes in the student population or at the teacher education institutes.
 
@@
 ### presentation of results
-#library(pander)                             # prettyprint for tables
+library(pander)                             # prettyprint for tables
 
@@
 
+Covariates:
+
+* target grade level (Trinn): level 1-7 or level 5-10
+* institute: 5 institutes
+* gender: male or female
+* year of study (semester): 2., 3. or 5. year
+
 ```{r}
@@
 
-The data is multilevel, as the PSTs (level 1) are grouped at the institutional level (level 2).
+The purpose of the multiple regression analysis is to explain the variation in the data.
+I therefore use stepwise regression and add the variables step-by-step to examine changes in R2 (explained variance).
+
+lm_robust is used to calculate cluster corrected standard errors. (or robust SE? Actually, CR2 SE = cluster robust standard errors)
+
+The data is however multilevel, as the PSTs (level 1) are grouped at the institutional level (level 2).
+I therefore also have to test whether I need to use hierarchical level modeling (HLM) by examining the intraclass correlation coefficient (ICC) of the data.
 
 ## Check assumptions for regression
+
+ICC is calculated to check whether HLM must be done.
+
+Multiple regression assumes independent observations and linearity between dependent and independent variables.
+In addition, the residuals should be normally distributed, to calculate test statistics for doing inferences.
 
@@
 If the variance of the intercept is larger than 0.05 of the total variance (intercept + residual), then the institutional variance contributes to explain the outcome.
+
+The ICC indicates that the data are nested. But since there is only 5 institutes, I used cluster corrected standard errors, instead of HLM, as suggested by @christophersen_introduksjon_2018.
 
@@
 Should I include interactions in the regression analysis?
+Interaction plots are drawn to see for effects of interaction.
+Any crossing lines indicate interaction effects.
+Parallel lines indicate no interaction.
 
@@
                  ylab = "TK",
-                 xlab = "Year of study",
+                 xlab = "Target grade level",
                  col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
@@
                  ylab = "PCK",
-                 xlab = "Year of study",
+                 xlab = "Target grade level",
                  col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
@@
                  ylab = "TPACK",
-                 xlab = "Year of study",
+                 xlab = "Target grade level",
                  col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
@@
 
-## Hierarchical RA for TK
+### Summary of interaction effects
+
+TK level 1 (PST):
+
+* possible interaction between gender and target grade level (Trinn)
+* possible interaction between gender and year of study (semester)
+
+TK level 2 (TEI):
+
+* possible interaction between institute and year of study (semester)
+* possible weak interaction between institute and target grade level (Trinn)
+* possible interaction between gender and institute
+
+PCK level 1:
+
+* possible weak interaction between gender and target grade level
+* possible interaction between gender and year of study (semester)
+* possible weak interaction between target grade level and year of study
+
+PCK level 2:
+
+* possible interaction between institute and year of study
+* possible weak interaction between institute and target grade level 
+* possible interaction between institute and gender
+
+TPACK level 1:
+
+* possible interaction between gender and year of study
+
+TPACK level 2:
+
+* possible interaction between institute and year of study
+* possible weak interaction between institute and gender
+
+The significance of these interaction effects have to be tested for when doing the regression analysis.
+
+## Replication of paper 2 results
+
+### Hierarchical RA for TK
 
@@
 
-### Level 1
+Interaction effects are examined by calculating F-statistics to see if models with and without interaction are significantly different.
+
+#### Level 1
+
+First, all covariates at the individual PST level 1 (i.e., gender, target grade level and year of study) are introduced as a block.
+
+Without interaction
+
+```{r TKlevel1_main_paper2}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+model_TK_covariates_l1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
+                clusters = factor(studiested), data=data_clean)
+summary(model_TK_covariates_l1)
+```
+
+With interaction
+
+when calculating interaction effects, TK needs to be standardized [@christophersen_introduksjon_2018].
+
+```{r TKlevel1_interaction_paper2}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+m2 <- lm_robust(scale(TK, center = T, scale = T) ~ factor(Gender) * (factor(Trinn) + factor(semester)),
+  clusters = factor(studiested), data = data_clean
+)
+summary(m2)
+```
+
+Calculate F-stat by comparing R2 to see if they are significantly different
+
+```{r}
+# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
+# source: p. 257 i Howitt & Cramer, 2014
+fstat <- m2$adj.r.squared / model_TK_covariates_l1$adj.r.squared
+df1 <- length(data_clean$TK) - 1
+df2 <- df1
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+The F-statistic indicate that the two models are not significantly different. This means that interaction effects at TK level 1 can be excluded from the rest of the analyses.
+
+#### Level 2
+
+Then I add level 2-variables: studiested
+
+without interaction
+
+```{r TKlevel2_main_paper2}
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
+model_TK_covariates_l2 <- lm_robust(
+  scale(TK) ~
+    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+    factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_TK_covariates_l2)
+```
+
+```{r}
+# main
+fstat <- model_TK_covariates_l2$adj.r.squared / model_TK_covariates_l1$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+
+# interaction
+fstat <- model_TK_covariates_l2$adj.r.squared / m2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+with interaction
+
+Trinn is omitted from interaction since it indicated the weakest interaction.
+
+```{r TKlevel2_interaction_paper2}
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
+m4 <- lm_robust(
+  scale(TK) ~
+    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+    factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
+  clusters = factor(studiested), data = data_clean
+)
+summary(m4)
+```
+
+Some missing data leads to NA.
+
+```{r}
+# interaction
+fstat <- m4$adj.r.squared / model_TK_covariates_l2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+The f-statistic indicate no significant differences with or without interaction, so interaction effects at level 2 are excluded from further analysis.
+
+#### Summary
+
+All main variables are included in the further analyses, but no interaction effects.
+
+### Hierarchical RA for PCK
+
+Level 1: students  
+Level 2: institution
+
+We are using cluster corrected standard errors to correct for clustering on institution.
+
+Add student covariates
+
+#### Level 1
+
+without interaction, only main effects
+
+```{r PCKlevel1_main_paper2}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+model_PCK_covariates_l1 <- lm_robust(PCK ~ factor(Gender) + factor(Trinn) + factor(semester),
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_PCK_covariates_l1)
+```
+
+with interaction
+
+```{r PCKlevel1_interaction_paper2}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+m2 <- lm_robust(scale(PCK, center = T, scale = T) ~ 
+                  factor(semester) * (factor(Gender) + factor(Trinn)),
+  clusters = factor(studiested), data = data_clean
+)
+summary(m2)
+```
+
+Calculate F-stat to see if they are significantly different
+
+```{r}
+# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
+# source: p. 257 i Howitt & Cramer, 2014
+fstat <- m2$adj.r.squared / model_PCK_covariates_l1$adj.r.squared
+df1 <- length(data_clean$PCK) - 1
+df2 <- df1
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There is no significant difference when including interaction effects. They are therefore excluded from further analysis.
+
+#### Level 2
+
+only main effects
+
+```{r PCKlevel2_main_paper2}
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
+model_PCK_covariates_l2 <- lm_robust(
+  scale(PCK) ~ 
+    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+    factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_PCK_covariates_l2)
+```
+
+```{r}
+# main
+fstat <- model_PCK_covariates_l2$adj.r.squared / model_PCK_covariates_l1$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+
+# interaction
+fstat <- model_PCK_covariates_l2$adj.r.squared / m2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There is no significant difference between level 1 and level 2, indicating no difference between TEIs.
+
+interaction effects
+
+target grade level is excluded from the interaction effects
+
+```{r PCKlevel2_interaction_paper2}
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
+m4 <- lm_robust(scale(PCK) ~ 
+                  factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+                  factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
+                clusters = factor(studiested), data=data_clean)
+summary(m4)
+```
+
+```{r}
+fstat <- m4$adj.r.squared / model_PCK_covariates_l2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There is no significant difference when including interaction effects.
+
+#### Summary
+
+Only level 1 variables contribute to explain variance in PCK.
+
+### Hierarchical RA for TPACK
+
+Level 1: students  
+Level 2: institution
+
+We are using cluster corrected standard errors to correct for clustering on institution.
+
+Add student covariates
+
+#### Level 1
+
+Covariates are first added as a block.
+
+Main effects
+
+```{r level1_main_paper2}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+model_TPACK_covariates_l1 <- lm_robust(
+  TPACK ~
+    factor(Gender) + factor(Trinn) + factor(semester),
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_TPACK_covariates_l1)
+```
+
+Interaction effects
+
+```{r level1_interaction_paper2}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+m2 <- lm_robust(
+  scale(TPACK, center = T, scale = T) ~
+    factor(Gender) * (factor(semester) + factor(Trinn)),
+  clusters = factor(studiested), data = data_clean
+)
+summary(m2)
+```
+
+Calculate F-stat to see if they are significantly different
+
+```{r}
+# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
+# source: p. 257 i Howitt & Cramer, 2014
+fstat <- m2$adj.r.squared / model_TPACK_covariates_l1$adj.r.squared
+df1 <- length(data_clean$TPACK) - 1
+df2 <- df1
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There is no significant difference when including interaction effects, so they are omitted from further analysis.
+
+#### Level 2
+
+Main effects
+
+```{r level2_main_paper2}
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
+model_TPACK_covariates_l2 <- lm_robust(
+  scale(TPACK) ~
+    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+    factor(studiested) + (factor(semester) + factor(Gender)),    # level 2
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_TPACK_covariates_l2)
+```
+
+```{r}
+#main
+fstat <- model_TPACK_covariates_l2$adj.r.squared / model_TPACK_covariates_l1$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+
+# interaction
+fstat <- model_TPACK_covariates_l2$adj.r.squared / m2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+Interaction effects
+
+```{r level2_interaction_paper2}
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
+m4 <- lm_robust(
+  scale(TPACK) ~
+    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+    factor(studiested) * (factor(semester) + factor(Gender)),    # level 2
+  clusters = factor(studiested), data = data_clean
+)
+summary(m4)
+```
+
+```{r}
+fstat <- m4$adj.r.squared / model_TPACK_covariates_l2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There are no significant interaction effects, so they are excluded from further analysis.
+
+#### PCK
+
+After the covariates are added as a block, PCK is added.
+
+```{r PCK_paper2}
+# PCK
+model_TPACK_PCK <- lm_robust(
+  scale(TPACK) ~
+    factor(Gender) * (factor(Trinn) + factor(semester)) +         # level 1
+    factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
+    scale(PCK),
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_TPACK_PCK)
+
+# main
+fstat <- model_TPACK_PCK$adj.r.squared / model_TPACK_covariates_l2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+
+# interaction
+fstat <- model_TPACK_PCK$adj.r.squared / m4$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There is a significant difference when adding PCK as a main effect.
+
+#### TK
+
+Finally, TK is added.
+
+```{r TK_paper2}
+# TK
+model_TPACK_TK <- lm_robust(
+  scale(TPACK) ~
+    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+    factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
+    scale(PCK) + scale(TK),
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_TPACK_TK)
+
+fstat <- model_TPACK_TK$adj.r.squared / model_TPACK_PCK$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There is a significant difference, when adding TK.
+
+### Results summarised
+
+The results differ somewhat from paper 2 in that some interaction effects are excluded, which was included in the analysis done for paper 2. The main trends are however similar.
+
+```{r}
+# table
+
+# TK
+
+mtkcl1rs  <- model_TK_covariates_l1$r.squared
+mtkcl1ars <- model_TK_covariates_l1$adj.r.squared
+
+mtkcl2rs  <- model_TK_covariates_l2$r.squared
+mtkcl2ars <- model_TK_covariates_l2$adj.r.squared
+
+TK_R2     <- data.frame(
+  "R2"    = c(mtkcl1rs, mtkcl2rs),
+  "R2adj" = c(mtkcl1ars, mtkcl2ars),
+  "diff"  = c(mtkcl1ars, mtkcl2ars - mtkcl1ars)
+)
+
+rownames(TK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute")
+
+pander(TK_R2)
+```
+
+```{r}
+# PCK
+
+mpcl1rs   <- model_PCK_covariates_l1$r.squared
+mpcl1ars  <- model_PCK_covariates_l1$adj.r.squared
+
+mpcl2rs   <- model_PCK_covariates_l2$r.squared
+mpcl2ars  <- model_PCK_covariates_l2$adj.r.squared
+
+PCK_R2    <- data.frame(
+  "R2"    = c(mpcl1rs, mpcl2rs),
+  "R2adj" = c(mpcl1ars, mpcl2ars),
+  "diff"  = c(mpcl1ars, mpcl2ars - mpcl1ars)
+)
+
+rownames(PCK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute")
+
+pander(PCK_R2)
+```
+
+```{r}
+# TPACK
+
+mtcl1rs  <- model_TPACK_covariates_l1$r.squared
+mtcl1ars <- model_TPACK_covariates_l1$adj.r.squared
+
+mtcl2rs  <- model_TPACK_covariates_l2$r.squared
+mtcl2ars <- model_TPACK_covariates_l2$adj.r.squared
+
+mtprs    <- model_TPACK_PCK$r.squared
+mtpars   <- model_TPACK_PCK$adj.r.squared
+
+mttrs    <- model_TPACK_TK$r.squared
+mttars   <- model_TPACK_TK$adj.r.squared
+
+TPACK_R2 <- data.frame(
+  "R2" = c(mtcl1rs, mtcl2rs, mtprs, mttrs),
+  "R2adj" = c(mtcl1ars, mtcl2ars, mtpars, mttars),
+  "diff" = c(mtcl1ars, mtcl2ars - mtcl1ars, 
+             mtpars - mtcl2ars, mttars - mtpars)
+)
+
+rownames(TPACK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute", "PCK", "TK")
+
+pander(TPACK_R2)
+```
+
+## Reanalysis with interaction effects omitted
+
+### Hierarchical RA for TK
+
+Level 1: students  
+Level 2: institution
+
+We are using cluster corrected standard errors to correct for clustering on institution.
+
+Add student covariates
+
+Interaction effects are examined by calculating F-statistics to see if models with and without interaction are significantly different.
+
+#### Level 1
+
+First, all covariates at the individual PST level 1 (i.e., gender, target grade level and year of study) are introduced as a block.
+
+Without interaction
 
@@
 # sjekk: er det riktig å skalere?
-m1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
+model_TK_covariates_l1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                 clusters = factor(studiested), data=data_clean)
-summary(m1)
-```
+summary(model_TK_covariates_l1)
+```
+
+With interaction
+
+when calculating interaction effects, TK needs to be standardized [@christophersen_introduksjon_2018].
 
@@
 # sjekk: er det riktig å skalere?
-m2 <- lm_robust(scale(TK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
-                clusters = factor(studiested), data=data_clean)
+m2 <- lm_robust(scale(TK, center = T, scale = T) ~ factor(Gender) * (factor(Trinn) + factor(semester)),
+  clusters = factor(studiested), data = data_clean
+)
 summary(m2)
@@
 
-Calculate F-stat to see if they are significantly different
+Calculate F-stat by comparing R2 to see if they are significantly different
 
@@
 # source: p. 257 i Howitt & Cramer, 2014
-fstat <- m2$adj.r.squared / m1$adj.r.squared
+fstat <- m2$adj.r.squared / model_TK_covariates_l1$adj.r.squared
 df1 <- length(data_clean$TK) - 1
@@
 
-### Level 2
+The F-statistic indicate that the two models are not significantly different. This means that interaction effects at TK level 1 can be excluded from the rest of the analyses.
+
+#### Level 2
+
+Then I add level 2-variables: studiested
+
+without interaction
 
@@
 # studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-m3 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
-                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
-                clusters = factor(studiested), data=data_clean)
-summary(m3)
+model_TK_covariates_l2 <- lm_robust(
+  scale(TK) ~
+    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
+    factor(studiested),                                 # level 2
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_TK_covariates_l2)
 ```
@@
 # main
-fstat <- m3$adj.r.squared / m1$adj.r.squared
+fstat <- model_TK_covariates_l2$adj.r.squared / model_TK_covariates_l1$adj.r.squared
 pf(fstat, df1, df2, lower.tail = FALSE)
@@
 # interaction
-fstat <- m3$adj.r.squared / m2$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
+fstat <- model_TK_covariates_l2$adj.r.squared / m2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+with interaction
+
+Trinn is omitted from interaction since it indicated the weakest interaction.
 
@@
 # studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-m4 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
-                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
-                clusters = factor(studiested), data=data_clean)
+m4 <- lm_robust(
+  scale(TK) ~
+    factor(Gender) + factor(Trinn) + factor(semester) +       # level 1
+    factor(studiested) * (factor(semester) + factor(Gender)), # level 2
+  clusters = factor(studiested), data = data_clean
+)
 summary(m4)
@@
 
+Some missing data leads to NA.
+
 ```{r}
 # interaction
-fstat <- m4$adj.r.squared / m3$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-## Hierarchical RA for PCK
+fstat <- m4$adj.r.squared / model_TK_covariates_l2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+The f-statistic indicate no significant differences with or without interaction, so interaction effects at level 2 are excluded from further analysis.
+
+#### Summary
+
+All main variables are included in the further analyses, but no interaction effects.
+
+### Hierarchical RA for PCK
 
@@
 
-### Level 1
+#### Level 1
+
+without interaction, only main effects
 
@@
 # sjekk: er det riktig å skalere?
-m1 <- lm_robust(PCK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
-                clusters = factor(studiested), data=data_clean)
-summary(m1)
-```
+model_PCK_covariates_l1 <- lm_robust(PCK ~ factor(Gender) + factor(Trinn) + factor(semester),
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_PCK_covariates_l1)
+```
+
+with interaction
 
@@
 # sjekk: er det riktig å skalere?
-m2 <- lm_robust(scale(PCK, center=T, scale=T) ~ factor(semester) * (factor(Gender) + factor(Trinn)), 
-                clusters = factor(studiested), data=data_clean)
+m2 <- lm_robust(scale(PCK, center = T, scale = T) ~ 
+                  factor(Trinn) * (factor(Gender) + factor(semester)) +
+                  factor(semester) * factor(Gender),
+  clusters = factor(studiested), data = data_clean
+)
 summary(m2)
@@
 # source: p. 257 i Howitt & Cramer, 2014
-fstat <- m2$adj.r.squared / m1$adj.r.squared
+fstat <- m2$adj.r.squared / model_PCK_covariates_l1$adj.r.squared
 df1 <- length(data_clean$PCK) - 1
@@
 
-### Level 2
+There is no significant difference when including interaction effects. They are therefore excluded from further analysis.
+
+#### Level 2
+
+only main effects
 
@@
 # studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-m3 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
-                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
-                clusters = factor(studiested), data=data_clean)
-summary(m3)
+model_PCK_covariates_l2 <- lm_robust(
+  scale(PCK) ~ 
+    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
+    factor(studiested),                                 # level 2
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_PCK_covariates_l2)
 ```
@@
 # main
-fstat <- m3$adj.r.squared / m1$adj.r.squared
+fstat <- model_PCK_covariates_l2$adj.r.squared / model_PCK_covariates_l1$adj.r.squared
 pf(fstat, df1, df2, lower.tail = FALSE)
@@
 # interaction
-fstat <- m3$adj.r.squared / m2$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
+fstat <- model_PCK_covariates_l2$adj.r.squared / m2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There is no significant difference between level 1 and level 2, indicating no difference between TEIs.
+
+interaction effects
+
+target grade level is excluded from the interaction effects
 
@@
 # studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-m4 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
-                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
+m4 <- lm_robust(scale(PCK) ~ 
+                  factor(Gender) + factor(Trinn) + factor(semester) +        # level 1
+                  factor(studiested) * (factor(semester) + factor(Gender)),  # level 2
                 clusters = factor(studiested), data=data_clean)
@@
 ```{r}
-fstat <- m4$adj.r.squared / m3$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-## Hierarchical RA for TPACK
+fstat <- m4$adj.r.squared / model_PCK_covariates_l2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There is no significant difference when including interaction effects.
+
+#### Summary
+
+Only level 1 variables contribute to explain variance in PCK.
+
+### Hierarchical RA for TPACK
 
@@
 
-### Level 1
+#### Level 1
+
+Covariates are first added as a block.
+
+Main effects
 
@@
 # sjekk: er det riktig å skalere?
-m1 <- lm_robust(TPACK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
-                clusters = factor(studiested), data=data_clean)
-summary(m1)
-```
+model_TPACK_covariates_l1 <- lm_robust(
+  TPACK ~
+    factor(Gender) + factor(Trinn) + factor(semester),
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_TPACK_covariates_l1)
+```
+
+Interaction effects
 
@@
 # sjekk: er det riktig å skalere?
-m2 <- lm_robust(scale(TPACK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
-                clusters = factor(studiested), data=data_clean)
+m2 <- lm_robust(
+  scale(TPACK, center = T, scale = T) ~
+    factor(Gender) * factor(semester) + factor(Trinn),
+  clusters = factor(studiested), data = data_clean
+)
 summary(m2)
@@
 # source: p. 257 i Howitt & Cramer, 2014
-fstat <- m2$adj.r.squared / m1$adj.r.squared
+fstat <- m2$adj.r.squared / model_TPACK_covariates_l1$adj.r.squared
 df1 <- length(data_clean$TPACK) - 1
@@
 
-### Level 2
+There is no significant difference when including interaction effects, so they are omitted from further analysis.
+
+#### Level 2
+
+Main effects
 
@@
 # studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-m3 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
-                               factor(studiested) + (factor(semester) + factor(Gender)),    # level 2
-                clusters = factor(studiested), data=data_clean)
-summary(m3)
-
+model_TPACK_covariates_l2 <- lm_robust(
+  scale(TPACK) ~
+    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
+    factor(studiested),                                 # level 2
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_TPACK_covariates_l2)
+```
+
+```{r}
 #main
-fstat <- m3$adj.r.squared / m1$adj.r.squared
+fstat <- model_TPACK_covariates_l2$adj.r.squared / model_TPACK_covariates_l1$adj.r.squared
 pf(fstat, df1, df2, lower.tail = FALSE)
@@
 # interaction
-fstat <- m3$adj.r.squared / m2$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
+fstat <- model_TPACK_covariates_l2$adj.r.squared / m2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+Interaction effects
 
@@
 # studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-m4 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
+m4 <- lm_robust(
+  scale(TPACK) ~
+    factor(Gender) + factor(Trinn) + factor(semester) +       # level 1
     factor(studiested) * (factor(semester) + factor(Gender)), # level 2
-                clusters = factor(studiested), data=data_clean)
+  clusters = factor(studiested), data = data_clean
+)
 summary(m4)
-
-fstat <- m4$adj.r.squared / m3$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-### PCK
+```
+
+```{r}
+fstat <- m4$adj.r.squared / model_TPACK_covariates_l2$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There are no significant interaction effects, so they are excluded from further analysis.
+
+#### PCK
+
+After the covariates are added as a block, PCK is added.
 
@@
 # PCK
-m5 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +         # level 1
-                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
+model_TPACK_PCK <- lm_robust(
+  scale(TPACK) ~
+    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
+    factor(studiested) +                                # level 2
     scale(PCK),
-                clusters = factor(studiested), data=data_clean)
-summary(m5)
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_TPACK_PCK)
 
 # main
-fstat <- m5$adj.r.squared / m3$adj.r.squared
+fstat <- model_TPACK_PCK$adj.r.squared / model_TPACK_covariates_l2$adj.r.squared
 pf(fstat, df1, df2, lower.tail = FALSE)
@@
 # interaction
-fstat <- m5$adj.r.squared / m4$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-### TK
+fstat <- model_TPACK_PCK$adj.r.squared / m4$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There is a significant difference when adding PCK as a main effect.
+
+#### TK
+
+Finally, TK is added.
 
@@
 # TK
-m6 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
-                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
+model_TPACK_TK <- lm_robust(
+  scale(TPACK) ~
+    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
+    factor(studiested) +                                # level 2
     scale(PCK) + scale(TK),
-                clusters = factor(studiested), data=data_clean)
-summary(m6)
-
-fstat <- m6$adj.r.squared / m5$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
+  clusters = factor(studiested), data = data_clean
+)
+summary(model_TPACK_TK)
+
+fstat <- model_TPACK_TK$adj.r.squared / model_TPACK_PCK$adj.r.squared
+pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+There is a significant difference, when adding TK.
+
+### Results summarised
+
+The results differ somewhat from paper 2 in that some interaction effects are excluded, which was included in the analysis done for paper 2. The main trends are however similar.
+
+```{r}
+# table
+
+# TK
+
+mtkcl1rs  <- model_TK_covariates_l1$r.squared
+mtkcl1ars <- model_TK_covariates_l1$adj.r.squared
+
+mtkcl2rs  <- model_TK_covariates_l2$r.squared
+mtkcl2ars <- model_TK_covariates_l2$adj.r.squared
+
+TK_R2     <- data.frame(
+  "R2"    = c(mtkcl1rs, mtkcl2rs),
+  "R2adj" = c(mtkcl1ars, mtkcl2ars),
+  "diff"  = c(mtkcl1ars, mtkcl2ars - mtkcl1ars)
+)
+
+rownames(TK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute")
+
+pander(TK_R2)
+```
+
+```{r}
+# PCK
+
+mpcl1rs   <- model_PCK_covariates_l1$r.squared
+mpcl1ars  <- model_PCK_covariates_l1$adj.r.squared
+
+mpcl2rs   <- model_PCK_covariates_l2$r.squared
+mpcl2ars  <- model_PCK_covariates_l2$adj.r.squared
+
+PCK_R2    <- data.frame(
+  "R2"    = c(mpcl1rs, mpcl2rs),
+  "R2adj" = c(mpcl1ars, mpcl2ars),
+  "diff"  = c(mpcl1ars, mpcl2ars - mpcl1ars)
+)
+
+rownames(PCK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute")
+
+pander(PCK_R2)
+```
+
+```{r}
+# TPACK
+
+mtcl1rs  <- model_TPACK_covariates_l1$r.squared
+mtcl1ars <- model_TPACK_covariates_l1$adj.r.squared
+
+mtcl2rs  <- model_TPACK_covariates_l2$r.squared
+mtcl2ars <- model_TPACK_covariates_l2$adj.r.squared
+
+mtprs    <- model_TPACK_PCK$r.squared
+mtpars   <- model_TPACK_PCK$adj.r.squared
+
+mttrs    <- model_TPACK_TK$r.squared
+mttars   <- model_TPACK_TK$adj.r.squared
+
+TPACK_R2 <- data.frame(
+  "R2" = c(mtcl1rs, mtcl2rs, mtprs, mttrs),
+  "R2adj" = c(mtcl1ars, mtcl2ars, mtpars, mttars),
+  "diff" = c(mtcl1ars, mtcl2ars - mtcl1ars, 
+             mtpars - mtcl2ars, mttars - mtpars)
+)
+
+rownames(TPACK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute", "PCK", "TK")
+
+pander(TPACK_R2)
+```
+
+Level 1: gender, target grade level  
+Level 2: year of study, institute
+
+Burde semester vært en level 2 variabel, siden det er mer en egenskap ved instituttet enn studenten?
+
+Spørsmålet er om forskjeller på semester skyldes endringer på instituttet (level 2) eller endringer i studentmassen (level 1)? Både-og? Samspill på tvers av nivåer!
+
+## Multi-level modeling with cross-level interaction effects
 

### 2024-09-04 Rmd/Paper2.Rmd
Added MLM

@@
 
+TODO
+
+* 4.9.2024: 
+** sjekk om det var riktig å gjøre om "Annet" til NA, da det påvirker analyseresultater (R2) for de grupperte kovariatene.
+** burde semester (year of study) være en nivå 1 eller nivå 2- variabel? Både-og? Samspill?
+
 # Introduction
@@
 library(dplyr)                              # rename columns
+#library(tidyr)
 
@@
 # jeg skjuler Annet, siden det bare er en person; alternativt NA?
-data_clean$Trinn[data_clean$Trinn == "Annet"] <- NA # "Trinn 5-10"
+#data_clean$Trinn[data_clean$Trinn == "Annet"] <- NA # "Trinn 5-10"
 #trinn_r[is.na(trinn_r)] <- 1
@@
 # skjuler som Kvinne, siden det er flest av de
-data_clean$Gender[data_clean$Gender == "Annet/ønsker ikke å oppgi"] <- NA # "Kvinne"
+#data_clean$Gender[data_clean$Gender == "Annet/ønsker ikke å oppgi"] <- NA # "Kvinne"
 #gender_r[is.na(gender_r)] <- 1
@@
 
+To compare fit, the number of observations need to be similar. Missing is therefore replaced using predictive mean matching.
+
+First, the baseline model is calculated as a reference and to find the ICC.
+
+```{r}
+# baseline model / nullmodell
+
+# variables 
+#  scale(TPACK) ~
+#    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
+#    factor(studiested) +                                # level 2
+#    scale(PCK) + scale(TK),
+#  clusters = factor(studiested), data = data_imputed
+
+# REML is used for parameter estimation
+model_baseline <- lme(TPACK~1, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
+
+# ML is used for model fit
+model_baseline_fit <- lme(TPACK~1, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
+
+summary(model_baseline)
+summary(model_baseline_fit)
+```
+
+```{r}
+# ICC
+nu_00 <- var(ranef(model_baseline)) # intercept (i.e., TEIs)
+nu_e  <- var(resid(model_baseline)) # residuals (i.e., PSTs)
+ICC   <- nu_00 / (nu_00 + nu_e)
+ICC
+```
+
+30 % av variansen i TPACK ligger på nivå 2.
+Dette er større enn den anbefalte grensen på 5 %.
+Vurder om OLS-regresjon med klyngekorrigert standardfeil egner seg bedre enn flernivåanalyse [@christophersen_introduksjon_2018] (p.
+109).
+
+```{r}
+# effektiv utvalgsstørrelse
+n_e <- 147 / (1 + ((147 / 5) - 1) * ICC)
+n_e
+```
+
+Utvalgsstørrelsen korrigert for avhengighet er lik 15.
+
+### random intercept (level 1)
+
+#### Level 1 variables
+
+```{r}
+# REML is used for parameter estimation
+model_level1 <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
+
+# ML is used for model fit
+model_level1_fit <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
+
+summary(model_level1)
+summary(model_level1_fit)
+```
+
+Er level 1-modellen bedre enn nullmodellen?
+
+```{r eval=T}
+anova(model_baseline_fit, model_level1_fit)
+```
+
+#### Level 2 variables
+
+La kovariatene variere mellom studiesteder
+
+```{r eval=F}
+# REML is used for parameter estimation
+model_level2 <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="REML", na.action="na.omit", random= ~Gender + semester + Trinn|studiested)
+
+# ML is used for model fit
+model_level2_fit <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="ML", na.action="na.omit", random= ~Gender + semester + Trinn|studiested)
+
+summary(model_level2)
+summary(model_level2_fit)
+```
+
+Er level 2-modellen bedre enn level 1?
+
+```{r eval=F}
+anova(model_level2_fit, model_level1_fit)
+```
+
+#### PCK
+
+```{r}
+# REML is used for parameter estimation
+model_PCK <- lme(TPACK~Gender + semester + Trinn + PCK, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
+
+# ML is used for model fit
+model_PCK_fit <- lme(TPACK~Gender + semester + Trinn + PCK, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
+
+summary(model_PCK)
+summary(model_PCK_fit)
+```
+
+```{r eval=T}
+anova(model_PCK_fit, model_level1_fit)
+```
+
+#### TK
+
+```{r}
+# REML is used for parameter estimation
+model_TK <- lme(TPACK~Gender + semester + Trinn + PCK + TK, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
+
+# ML is used for model fit
+model_TK_fit <- lme(TPACK~Gender + semester + Trinn + PCK + TK, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
+
+summary(model_TK)
+summary(model_TK_fit)
+```
+
+```{r eval=T}
+anova(model_PCK_fit, model_TK_fit)
+```
+
 # Discussion and conclusion

### 2024-09-05 Rmd/Paper2.Rmd
Fixed and tidied the presentation of the results.

--- 2024-09-04 Rmd/Paper2.Rmd
@@
 author:       "Niklas Karlsen"
-date:         "2024-09-04"
+date:         "2024-09-05"
 bibliography: "../data/meta_data/references.bib"
@@
 
+Since the data is structured at two levels (i.e., PSTs and TEIs), the data is not independent as is assumed when doing OLS regression.
+Since only 5 TEIs are in the dataset, which is below the recommended miminum of 10, OLS with cluster corrected standard errors may be a feasible option to MLM.
+
+The ICC is checked to see how much of the variance is accounted for by the level 2 units, to assess whether it is necessary to account for the level 2-dependence.
+ICC should be above 5 % for MLM to be considered.
+
+The OLS regression is hierarchical.
+The covariates are introduced block-wise.
+Gender and target grade is related to the individual PST, while year of study can reflect both changes at level 1 and level 2.
+
+For the OLS the TEIs are introduced as dummy variables, as are the other covariates with more than two levels.
+
+For the multi-level modeling (MLM) the PSTs are considered level 1 units and the TEIs are level 2.
+The covariates gender and target grade are related to the level 1 units, while year of study can vary at both levels.
+
+In addition, TK and PCK are considered independent variables, while TPACK is the dependent variable.
+
+Both the OLS and the MLM are done by introducing the variables sequentially.
+For the OLS done in paper 2, the level 1 variables (i.e., gender, trinn and semester) are introduced first, then the level 2 variables (i.e., TEIs), and finally, PCK and TK.
+For the MLM, a baseline model (null-model) is created first, where the TEIs are set as random (i.e., the intercept is allowed to vary among TEIs).
+Then the level 1 variables are introduced to create the level 1-model.
+Then PCK and TK are added to the model.
+Since the order of introducing TEI was altered between the OLS in paper 2 and the MLM, a reanalysis of the OLS was done with the TEIs introduced before the covariates.
+This could affect how much of the variance was explained by the covariates.
+
+The explained variance (i.e., R2) is calculated for both methods, and can be used to compare the results of the two methods.
+
+The following models are tested:
+
+Model | Paper 2 OLS | Reanalysis OLS | New analysis MLM 
+------------- | ------------- | ------------- | ------------- 
+Options | 3xAnnet = NA | N/A | N/A
+Nullmodell | N/A | TPACK ~ TEIs | TPACK ~ 1
+Level 1 | TPACK ~ covariates + TEIs | TPACK ~ TEIs + covariates | TPACK ~ covariates
+Level 2 | N/A | N/A | + year of study | TEI
+PCK | TPACK ~ covariates + TEIs + PCK | TPACK ~ TEIs + covariates + PCK | TPACK ~ covariates + PCK_sentrert
+TK | TPACK ~ covariates + TEIs + PCK + TK | TPACK ~ TEIs + covariates + PCK + TK | TPACK ~ covariates + PCK_s + TK_sentrert
+
+The analysis in paper 2 changed the answers of three respondents from "Other" to NA.
+This made it impossible to compare the fit for the MLMs since the number of respondents were dissimilar for the models compared when including these covariates.
+For the reanalyses I therefore kept these values unchanged.
+
+In the MLM the TEIs are added as random variables (i.e., the intercepts are allowed to differ between TEIs), while the TEIs are added as dummy variables in the OLS. This should give similar results.
+
+In the MLM the covariates are added as fixed variables (i.e., the slopes are the same for all PSTs independent of TEI).
+
+The order of the models was also changed in the reanalysis, as the baseline model in the MLM corresponds to having the TEIs as dummy variables.
+In paper 2 this model was tested after having controlled for the covariates.
+In the reanalysis this order was switched to correspond to the MLM.
+
+Only the MLM has the option to test for level 2 covariates. I allowed slope of year of study vary between TEIs. The level2 model is therefore not tested. It did not converge.
+
+For the MLM the independendent variables were centred using the grand mean, following @christophersen_introduksjon_2018.
+
+The different methods can be compared by comparing R-squared, and diff R-squared.
+
 ## Data management
 
-For the survey, the complete TRELIS-Q dataset was used.
+For the survey, the complete TRELIS-Q dataset was used. Here I load the raw data and prepare it for analysis.
 
@@
 
-```{r load_data_cfa}
+The environment is reset and the data reloaded.
+
+```{r load_data}
 # Rdata-file contains reduced / prepared datasets
@@
 
-OBS! Anonymiser data i datafil på studiested før det legges ut åpent på nett
-
 ```{r}
@@
 data_clean <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),] 
-
+```
+
+To reproduce paper 2 values, I set the answers "Other" to NA.
+
+```{r}
 # jeg skjuler Annet, siden det bare er en person; alternativt NA?
@@
 The purpose of the multiple regression analysis is to explain the variation in the data.
-I therefore use stepwise regression and add the variables step-by-step to examine changes in R2 (explained variance).
-
-lm_robust is used to calculate cluster corrected standard errors. (or robust SE? Actually, CR2 SE = cluster robust standard errors)
+I therefore use hierarchical regression [@howitt_introduction_2014] and add the variables step-by-step to examine changes in R2 (explained variance).
+
+## Check assumptions for regression
 
@@
 I therefore also have to test whether I need to use hierarchical level modeling (HLM) by examining the intraclass correlation coefficient (ICC) of the data.
-
-## Check assumptions for regression
 
@@
 
+lm_robust [@blair_estimatr_2022] is used to calculate cluster corrected standard errors. (or robust SE? Actually, CR2 SE = cluster robust standard errors)
+
 ### Normality
+
+The residuals need to be normally distributed for the statistical tests. For dichotomous variables, the distribution should be less than 80/20 skewed.
 
@@
 
+The data should be linearly related.
+
 ```{r scatterplot_linearity}
@@
 
-TK korrelererer veldig sterkt (0.84) med TPACK, noe som indikerer kollinearitet. (Jeong and Jung, 2016)
+TK korrelererer veldig sterkt (0.84) med TPACK, noe som indikerer kollinearitet (Jeong and Jung, 2016). This issue is discussed in @karlsen_assessing_2024 and the synopsis.
 
@@
 
+Here I reproduce the results reported in paper 2. Interaction effects were also tested to see if they needed to be included in the analysis.
+
 ### Hierarchical RA for TK
@@
 
-The results differ somewhat from paper 2 in that some interaction effects are excluded, which was included in the analysis done for paper 2. The main trends are however similar.
+Here I reproduce the tables presented in paper 2. The significance values can be found in the f-tests done above.
 
@@
 
-## Reanalysis with interaction effects omitted
+## Reanalysis of OLS
+
+A reanalysis of the OLS is done here, due to some methodical reconsiderations after a closer look at the MLM.
+The dataset is slightly closer to the rawdata (3 NAs were reset to other, as in the original dataset).
+The order of the variables are also slightly different (TEIs are introduced before covariates).
+Finally, interaction effects are not included, as they were in paper 2.
+The statistical tests suggested that the interaction effects were non-significant.
+Otherwise, the analysis is similar to paper 2.
+
+### Reload data
+
+```{r}
+# ref: https://www.statology.org/clear-environment-in-r/ [5.9.2024]
+#clear all data frames from environment
+rm(list=ls(all=TRUE)[sapply(mget(ls(all=TRUE)), class) == "data.frame"])
+```
+
+Then I reload the dataset from file.
+
+```{r load_data_reanalysis}
+# Rdata-file contains reduced / prepared datasets
+load(file="../data/processed_data/tpack-TRELIS-Q_complete_2023-paper2.Rdata")
+```
+
+Then I recenter the data.
+
+```{r}
+data_TRELIS_Q[,unlist(TPACK_trelis_q)] <- data_TRELIS_Q[,unlist(TPACK_trelis_q)] - 3 
+```
+
+And add some missing information.
+
+```{r}
+# identify NA as studiested 5(?) 
+data_TRELIS_Q[data_TRELIS_Q$studiested == "Institute3","Trinn"] <- "Trinn 5-10"
+```
+
+NMAR are removed.
+
+```{r fix_missing_2, eval=T}
+# remove NMAR + Institute9 (19)
+data_clean <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),] 
+```
+
+```{r}
+scores_v23 <- scoreItems(TPACK_trelis_q, data_clean)
+
+describe(scores_v23$scores)
+
+# constructs for multiple regression
+TK_v23    <- scores_v23$scores[,"TK"]
+PCK_v23   <- scores_v23$scores[,"PCK"]
+TPACK_v23 <- scores_v23$scores[,"TPACK"]
+
+# add constructs to data.frame
+data_clean$TK    <- TK_v23
+data_clean$PCK   <- PCK_v23
+data_clean$TPACK <- TPACK_v23
+```
 
@@
 Interaction effects are examined by calculating F-statistics to see if models with and without interaction are significantly different.
+
+#### baseline model (nullmodell)
+
+```{r TK_baseline}
+# kovariater; we use scale to standardize when using interaction effects
+# sjekk: er det riktig å skalere?
+model_TK_baseline <- lm_robust(TK ~ factor(studiested), 
+                clusters = factor(studiested), data=data_clean)
+summary(model_TK_baseline)
+```
 
@@
 # sjekk: er det riktig å skalere?
-model_TK_covariates_l1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
+model_TK_level1 <- lm_robust(TK ~ factor(studiested) + factor(Gender) + (factor(Trinn) + factor(semester)), 
                 clusters = factor(studiested), data=data_clean)
-summary(model_TK_covariates_l1)
-```
-
-With interaction
-
-when calculating interaction effects, TK needs to be standardized [@christophersen_introduksjon_2018].
-
-```{r TKlevel1_interaction}
-# kovariater; we use scale to standardize when using interaction effects
-# sjekk: er det riktig å skalere?
-m2 <- lm_robust(scale(TK, center = T, scale = T) ~ factor(Gender) * (factor(Trinn) + factor(semester)),
-  clusters = factor(studiested), data = data_clean
-)
-summary(m2)
+summary(model_TK_level1)
 ```
@@
 # source: p. 257 i Howitt & Cramer, 2014
-fstat <- m2$adj.r.squared / model_TK_covariates_l1$adj.r.squared
+fstat <- model_TK_level1$adj.r.squared / model_TK_baseline$adj.r.squared
 df1 <- length(data_clean$TK) - 1
@@
 
-The F-statistic indicate that the two models are not significantly different. This means that interaction effects at TK level 1 can be excluded from the rest of the analyses.
-
-#### Level 2
-
-Then I add level 2-variables: studiested
-
-without interaction
-
-```{r TKlevel2_main}
+The F-statistic indicate that the two models are not significantly different. This means that level1 effects do not contribute significantly to the variance.
+
+### Hierarchical RA for PCK
+
+Level 1: students  
+Level 2: institution
+
+We are using cluster corrected standard errors to correct for clustering on institution.
+
+Add student covariates
+
+#### Baseline
+
+```{r PCKlevel2_main}
 # studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-model_TK_covariates_l2 <- lm_robust(
-  scale(TK) ~
-    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
+model_PCK_baseline <- lm_robust(
+  PCK ~ 
     factor(studiested),                                 # level 2
@@
 )
-summary(model_TK_covariates_l2)
-```
-
-```{r}
-# main
-fstat <- model_TK_covariates_l2$adj.r.squared / model_TK_covariates_l1$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-
-# interaction
-fstat <- model_TK_covariates_l2$adj.r.squared / m2$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-with interaction
-
-Trinn is omitted from interaction since it indicated the weakest interaction.
-
-```{r TKlevel2_interaction}
-# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-m4 <- lm_robust(
-  scale(TK) ~
-    factor(Gender) + factor(Trinn) + factor(semester) +       # level 1
-    factor(studiested) * (factor(semester) + factor(Gender)), # level 2
-  clusters = factor(studiested), data = data_clean
-)
-summary(m4)
-```
-
-Some missing data leads to NA.
-
-```{r}
-# interaction
-fstat <- m4$adj.r.squared / model_TK_covariates_l2$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-The f-statistic indicate no significant differences with or without interaction, so interaction effects at level 2 are excluded from further analysis.
-
-#### Summary
-
-All main variables are included in the further analyses, but no interaction effects.
-
-### Hierarchical RA for PCK
-
-Level 1: students  
-Level 2: institution
-
-We are using cluster corrected standard errors to correct for clustering on institution.
-
-Add student covariates
+summary(model_PCK_baseline)
+```
 
@@
 # sjekk: er det riktig å skalere?
-model_PCK_covariates_l1 <- lm_robust(PCK ~ factor(Gender) + factor(Trinn) + factor(semester),
+model_PCK_level1 <- lm_robust(PCK ~ factor(studiested) + factor(Gender) + factor(Trinn) + factor(semester),
   clusters = factor(studiested), data = data_clean
 )
-summary(model_PCK_covariates_l1)
-```
-
-with interaction
-
-```{r PCKlevel1_interaction}
-# kovariater; we use scale to standardize when using interaction effects
-# sjekk: er det riktig å skalere?
-m2 <- lm_robust(scale(PCK, center = T, scale = T) ~ 
-                  factor(Trinn) * (factor(Gender) + factor(semester)) +
-                  factor(semester) * factor(Gender),
-  clusters = factor(studiested), data = data_clean
-)
-summary(m2)
+summary(model_PCK_level1)
 ```
@@
 # source: p. 257 i Howitt & Cramer, 2014
-fstat <- m2$adj.r.squared / model_PCK_covariates_l1$adj.r.squared
+fstat <- model_PCK_level1$adj.r.squared / model_PCK_baseline$adj.r.squared
 df1 <- length(data_clean$PCK) - 1
@@
 
-There is no significant difference when including interaction effects. They are therefore excluded from further analysis.
-
-#### Level 2
-
-only main effects
-
-```{r PCKlevel2_main}
+There were significant differences when including covariates.
+
+### Hierarchical RA for TPACK
+
+Level 1: students  
+Level 2: institution
+
+We are using cluster corrected standard errors to correct for clustering on institution.
+
+Add student covariates
+
+#### Baseline
+
+```{r level2_main}
 # studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-model_PCK_covariates_l2 <- lm_robust(
-  scale(PCK) ~ 
-    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
+model_TPACK_baseline <- lm_robust(
+  TPACK ~
     factor(studiested),                                 # level 2
@@
 )
-summary(model_PCK_covariates_l2)
-```
-
-```{r}
-# main
-fstat <- model_PCK_covariates_l2$adj.r.squared / model_PCK_covariates_l1$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-
-# interaction
-fstat <- model_PCK_covariates_l2$adj.r.squared / m2$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-There is no significant difference between level 1 and level 2, indicating no difference between TEIs.
-
-interaction effects
-
-target grade level is excluded from the interaction effects
-
-```{r PCKlevel2_interaction}
-# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-m4 <- lm_robust(scale(PCK) ~ 
-                  factor(Gender) + factor(Trinn) + factor(semester) +        # level 1
-                  factor(studiested) * (factor(semester) + factor(Gender)),  # level 2
-                clusters = factor(studiested), data=data_clean)
-summary(m4)
-```
-
-```{r}
-fstat <- m4$adj.r.squared / model_PCK_covariates_l2$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-There is no significant difference when including interaction effects.
-
-#### Summary
-
-Only level 1 variables contribute to explain variance in PCK.
-
-### Hierarchical RA for TPACK
-
-Level 1: students  
-Level 2: institution
-
-We are using cluster corrected standard errors to correct for clustering on institution.
-
-Add student covariates
+summary(model_TPACK_baseline)
+```
 
@@
 # sjekk: er det riktig å skalere?
-model_TPACK_covariates_l1 <- lm_robust(
+model_TPACK_level1 <- lm_robust(
   TPACK ~
-    factor(Gender) + factor(Trinn) + factor(semester),
+    factor(studiested) + factor(Gender) + factor(Trinn) + factor(semester),
   clusters = factor(studiested), data = data_clean
 )
-summary(model_TPACK_covariates_l1)
-```
-
-Interaction effects
-
-```{r level1_interaction}
-# kovariater; we use scale to standardize when using interaction effects
-# sjekk: er det riktig å skalere?
-m2 <- lm_robust(
-  scale(TPACK, center = T, scale = T) ~
-    factor(Gender) * factor(semester) + factor(Trinn),
-  clusters = factor(studiested), data = data_clean
-)
-summary(m2)
+summary(model_TPACK_level1)
 ```
@@
 # source: p. 257 i Howitt & Cramer, 2014
-fstat <- m2$adj.r.squared / model_TPACK_covariates_l1$adj.r.squared
+fstat <- model_TPACK_level1$adj.r.squared / model_TPACK_baseline$adj.r.squared
 df1 <- length(data_clean$TPACK) - 1
@@
 
-There is no significant difference when including interaction effects, so they are omitted from further analysis.
-
-#### Level 2
-
-Main effects
-
-```{r level2_main}
-# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-model_TPACK_covariates_l2 <- lm_robust(
-  scale(TPACK) ~
-    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
-    factor(studiested),                                 # level 2
-  clusters = factor(studiested), data = data_clean
-)
-summary(model_TPACK_covariates_l2)
-```
-
-```{r}
-#main
-fstat <- model_TPACK_covariates_l2$adj.r.squared / model_TPACK_covariates_l1$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-
-# interaction
-fstat <- model_TPACK_covariates_l2$adj.r.squared / m2$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-Interaction effects
-
-```{r level2_interaction}
-# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-m4 <- lm_robust(
-  scale(TPACK) ~
-    factor(Gender) + factor(Trinn) + factor(semester) +       # level 1
-    factor(studiested) * (factor(semester) + factor(Gender)), # level 2
-  clusters = factor(studiested), data = data_clean
-)
-summary(m4)
-```
-
-```{r}
-fstat <- m4$adj.r.squared / model_TPACK_covariates_l2$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-There are no significant interaction effects, so they are excluded from further analysis.
+There is no significant difference when including covariates.
 
@@
 model_TPACK_PCK <- lm_robust(
-  scale(TPACK) ~
+  TPACK ~
     factor(Gender) + factor(Trinn) + factor(semester) + # level 1
     factor(studiested) +                                # level 2
-    scale(PCK),
+    PCK,
   clusters = factor(studiested), data = data_clean
@@
 # main
-fstat <- model_TPACK_PCK$adj.r.squared / model_TPACK_covariates_l2$adj.r.squared
+fstat <- model_TPACK_PCK$adj.r.squared / model_TPACK_level1$adj.r.squared
 pf(fstat, df1, df2, lower.tail = FALSE)
 
-# interaction
-fstat <- model_TPACK_PCK$adj.r.squared / m4$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
 ```
@@
 model_TPACK_TK <- lm_robust(
-  scale(TPACK) ~
+  TPACK ~
     factor(Gender) + factor(Trinn) + factor(semester) + # level 1
     factor(studiested) +                                # level 2
-    scale(PCK) + scale(TK),
+    PCK + TK,
   clusters = factor(studiested), data = data_clean
@@
 
-mtkcl1rs  <- model_TK_covariates_l1$r.squared
-mtkcl1ars <- model_TK_covariates_l1$adj.r.squared
-
-mtkcl2rs  <- model_TK_covariates_l2$r.squared
-mtkcl2ars <- model_TK_covariates_l2$adj.r.squared
+mtkcl1rs  <- model_TK_baseline$r.squared
+mtkcl1ars <- model_TK_baseline$adj.r.squared
+
+mtkcl2rs  <- model_TK_level1$r.squared
+mtkcl2ars <- model_TK_level1$adj.r.squared
 
@@
 
-rownames(TK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute")
+rownames(TK_R2) <- c("Teacher education institute", "Gender, year of study, target grade level")
 
@@
 
-mpcl1rs   <- model_PCK_covariates_l1$r.squared
-mpcl1ars  <- model_PCK_covariates_l1$adj.r.squared
-
-mpcl2rs   <- model_PCK_covariates_l2$r.squared
-mpcl2ars  <- model_PCK_covariates_l2$adj.r.squared
+mpcl1rs   <- model_PCK_baseline$r.squared
+mpcl1ars  <- model_PCK_baseline$adj.r.squared
+
+mpcl2rs   <- model_PCK_level1$r.squared
+mpcl2ars  <- model_PCK_level1$adj.r.squared
 
@@
 
-rownames(PCK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute")
+rownames(PCK_R2) <- c("Teacher education institute", "Gender, year of study, target grade level")
 
@@
 
-mtcl1rs  <- model_TPACK_covariates_l1$r.squared
-mtcl1ars <- model_TPACK_covariates_l1$adj.r.squared
-
-mtcl2rs  <- model_TPACK_covariates_l2$r.squared
-mtcl2ars <- model_TPACK_covariates_l2$adj.r.squared
+mtcl1rs  <- model_TPACK_baseline$r.squared
+mtcl1ars <- model_TPACK_baseline$adj.r.squared
+
+mtcl2rs  <- model_TPACK_level1$r.squared
+mtcl2ars <- model_TPACK_level1$adj.r.squared
 
@@
 
-rownames(TPACK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute", "PCK", "TK")
+rownames(TPACK_R2) <- c("Teacher education institute", "Gender, year of study, target grade level", "PCK", "TK")
 
@@
 
+When the data is nested, multi-level modeling can be used.
+However, when the number of level 2-units is below 10, as it is in our dataset (we have only 5 TEIs), OLS with cluster corrected standard errors can be used.
+I, however, do a MLM here to compare with the OLS results.
+MLM is only done for TPACK as dependent variable to illustrate the method.
+
+The r package NLME [@pinheiro_nlme_2023] is used for the analysis.
+
+### Baseline model
+
+First, the baseline model is calculated as a reference and to find the ICC.
+
+```{r}
+# baseline model / nullmodell
+
+# REML is used for parameter estimation
+model_MLM_baseline <- lme(TPACK~1, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
+
+# ML is used for model fit
+model_MLM_baseline_fit <- lme(TPACK~1, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
+
+summary(model_MLM_baseline)
+summary(model_MLM_baseline_fit)
+```
+
+```{r}
+# ICC
+nu_00 <- var(ranef(model_MLM_baseline)) # intercept (i.e., TEIs)
+nu_e  <- var(resid(model_MLM_baseline)) # residuals (i.e., PSTs)
+ICC   <- nu_00 / (nu_00 + nu_e)
+ICC
+```
+
+30 % av variansen i TPACK ligger på nivå 2.
+Dette er større enn den anbefalte grensen på 5 %.
+Vurder om OLS-regresjon med klyngekorrigert standardfeil egner seg bedre enn flernivåanalyse [@christophersen_introduksjon_2018] (p.
+109).
+
+```{r}
+# effektiv utvalgsstørrelse
+n_e <- 147 / (1 + ((147 / 5) - 1) * ICC)
+n_e
+```
+
+Utvalgsstørrelsen korrigert for avhengighet er lik 15.
+
+### Level 1 model (random intercept)
+
+The level 1 model extends the baseline model by including the covariates.
+
+#### Level 1 variables
+
+```{r}
+# REML is used for parameter estimation
+model_MLM_level1 <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
+
+# ML is used for model fit
+model_MLM_level1_fit <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
+
+summary(model_MLM_level1)
+summary(model_MLM_level1_fit)
+```
+
+Er level 1-modellen bedre enn nullmodellen?
+
+```{r eval=T}
+anova(model_MLM_baseline_fit, model_MLM_level1_fit)
+```
+
+#### Level 2 variables
+
+Since Level 2 variables are unique to MLM, and the solution does not converge, level 2 analysis is not done here.
+
+La kovariatene variere mellom studiesteder
+
+```{r eval=F}
+# REML is used for parameter estimation
+model_level2 <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="REML", na.action="na.omit", random= ~ semester|studiested)
+
+# ML is used for model fit
+model_level2_fit <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="ML", na.action="na.omit", random= ~ semester|studiested)
+
+summary(model_level2)
+summary(model_level2_fit)
+```
+
+Er level 2-modellen bedre enn level 1?
+
+```{r eval=F}
+anova(model_level2_fit, model_MLM_level1_fit)
+```
+
+### PCK
+
+```{r}
+# grand mean centering: scale(TK, center = T, scale = F)
+
+data_clean$PCK_cent <- scale(data_clean$PCK, center = T, scale = F)
+data_clean$TK_cent  <- scale(data_clean$TK, center = T, scale = F)
+
+# REML is used for parameter estimation
+model_MLM_PCK <- lme(TPACK~Gender + semester + Trinn + PCK_cent, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
+
+# ML is used for model fit
+model_MLM_PCK_fit <- lme(TPACK~Gender + semester + Trinn + PCK_cent, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
+
+summary(model_MLM_PCK)
+summary(model_MLM_PCK_fit)
+```
+
+```{r eval=T}
+anova(model_MLM_PCK_fit, model_MLM_level1_fit)
+```
+
+### TK
+
+```{r}
+# REML is used for parameter estimation
+model_MLM_TK <- lme(TPACK~Gender + semester + Trinn + PCK_cent + TK_cent, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
+
+# ML is used for model fit
+model_MLM_TK_fit <- lme(TPACK~Gender + semester + Trinn + PCK_cent + TK_cent, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
+
+summary(model_MLM_TK)
+summary(model_MLM_TK_fit)
+```
+
+```{r eval=T}
+anova(model_MLM_PCK_fit, model_MLM_TK_fit)
+```
+
+## Comparison of OLS and MLM
+
+### Comparison of variance
+
+I follow @christophersen_introduksjon_2018 (p. 115) and use the variance estimates as an indicator of R2.
+
+```{r}
+# MLM
+# variance baseline
+nu_00_baseline <- var(ranef(model_MLM_baseline)) # intercept (i.e., TEIs)
+nu_e_baseline  <- var(resid(model_MLM_baseline)) # residuals (i.e., PSTs)
+
+# variance level 1
+nu_00_level1 <- var(ranef(model_MLM_level1)) # intercept (i.e., TEIs)
+nu_e_level1  <- var(resid(model_MLM_level1)) # residuals (i.e., PSTs)
+
+# variance PCK
+nu_00_PCK <- var(ranef(model_MLM_PCK)) # intercept (i.e., TEIs)
+nu_e_PCK  <- var(resid(model_MLM_PCK)) # residuals (i.e., PSTs)
+
+# variance TK
+nu_00_TK <- var(ranef(model_MLM_TK)) # intercept (i.e., TEIs)
+nu_e_TK  <- var(resid(model_MLM_TK)) # residuals (i.e., PSTs)
+
+# OLS
+mtcl1ars <- model_TPACK_baseline$adj.r.squared
+mtcl2ars <- model_TPACK_level1$adj.r.squared
+mtpars   <- model_TPACK_PCK$adj.r.squared
+mttars   <- model_TPACK_TK$adj.r.squared
+
+# table of difference in variance
+
+table_variance <- data.frame(
+  c(nu_e_baseline, (nu_e_baseline - nu_e_level1) / nu_e_baseline, (nu_e_level1 - nu_e_PCK) / nu_e_level1, (nu_e_PCK - nu_e_TK) / nu_e_PCK),
+  c(nu_00_baseline, (nu_00_baseline - nu_00_level1) / nu_00_baseline, (nu_00_baseline - nu_00_PCK) / nu_00_baseline, (nu_00_baseline - nu_00_TK) / nu_00_baseline),
+  c(mtcl1ars, mtcl2ars - mtcl1ars, mtpars - mtcl2ars, mttars - mtpars)
+  )
+
+rownames(table_variance) <- c("TEI", "covariates", "PCK", "TK")
+colnames(table_variance) <- c("MLM Level 1 % var", "MLM Level 2 % var", "OLS reanalysis diff R2")
+
+pander(table_variance)
+```
+
+The variance differ somewhat from the standard deviation (when squared) reported in the summary of the MLMs. Why? Is the OLS a kind of average of the MLM, which is separated into level 1 and level 2?
+
 # Discussion and conclusion
 
-The results are discussed in @karlsen_assessing_2024.
+I was able to replicate the results discussed in @karlsen_assessing_2024.
+
+The reanalysis using OLS gave similar results as for paper 2.
+Only TEIs explained variance in TK, not the covariates.
+The covariates did however significantly explain the variance in PCK, but the TEIs did not (or to a small degree, only 3%).
+The variance in TPACK was significantly explained by TEI, PCK and TK, not the covariates.
+
+The results in paper 2 indicated that the covariates contributed to explain the variance in TPACK.
+The reanalysis done here, suggest that this may be due to the order in which the variables were added to the model being tested.
+If I follow the order used in the MLM, the reanalysis suggests that the covariates do not contribute to TPACK.
+
+The MLM gave supporting results.
+Including the covariates did not improve the model, whereas including PCK and TK did.
+The variance at level 2 (between TEIs) was due to TK, not PCK or covariates, that is, there were significant differences in TK between TEIs.
+At level 1 (between PSTs) 6% of the variance were due to the covariates, 17 % due to PCK and 55 % due to TK.
+
+The OLS reanalysis suggested that 30 % of the variance in TPACK was due to TEIs, 11 % due to PCK and 32 % due to TPACK.
+The covariates did not contribute to the variance in TPACK.
+Even though the percentages differ between the OLS and the MLM, they show the same trends.
+The different percentages may be due to how R2 (or difference in variance) is calculated differently in OLS and MLM.
 

### 2024-09-05 Rmd/Paper2.Rmd
Restructured some code to make it more stylish

@@
 
-knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/paper2.html')) })
+knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), "../output/paper2.html") )})
 ---
@@
 ## TPACK-PST for TRELIS-Q
-TK_pos_trelisq    <- c("TK1b","TK2","TK34") 
-PCK_pos_trelisq   <- c("CK123","CK5","PCK2","PCK5")
-TPACK_pos_trelisq <- c("TPACK2","TPACK4","TCK1","TCK4","TPK1","TPK4")
-
-TPACK_trelis_q    <- list(TK=TK_pos_trelisq,
-                     PCK=PCK_pos_trelisq,
-                     TPACK=TPACK_pos_trelisq)
+TK_pos_trelisq    <- c("TK1b", "TK2", "TK34")
+PCK_pos_trelisq   <- c("CK123", "CK5", "PCK2", "PCK5")
+TPACK_pos_trelisq <- c("TPACK2", "TPACK4", "TCK1", "TCK4", "TPK1", "TPK4")
+
+TPACK_trelis_q <- list(
+  TK = TK_pos_trelisq,
+  PCK = PCK_pos_trelisq,
+  TPACK = TPACK_pos_trelisq
+)
 ```
@@
 # convert into numbers
-data_trelis_q_ren[,colnames(data_trelis_q_ren) %in% unlist(TPACK_trelis_q)] <-
+data_trelis_q_ren[, colnames(data_trelis_q_ren) %in% unlist(TPACK_trelis_q)] <-
   apply(
-    data_trelis_q_ren[,colnames(data_trelis_q_ren) %in% unlist(TPACK_trelis_q)], 2,
-                    function(x) as.numeric(as.character(x)))
+    data_trelis_q_ren[, colnames(data_trelis_q_ren) %in% unlist(TPACK_trelis_q)], 2,
+    function(x) as.numeric(as.character(x))
+  )
 ```
@@
 
-TPACK_trelis_q    <- list(TK=TK_pos_trelisq,
-                     PCK=PCK_pos_trelisq,
-                     TPACK=TPACK_pos_trelisq)
+TPACK_trelis_q    <- list(
+  TK = TK_pos_trelisq,
+  PCK = PCK_pos_trelisq,
+  TPACK = TPACK_pos_trelisq
+)
 ```
@@
 # describe data (mean, sd, skew, curtosis)
-describe_data <- function (datalist, QUIET=1) {
-  im <- mean(datalist, na.rm=TRUE)
-  isd <- sd(datalist, na.rm=TRUE)
-  isk <- psych::skew(datalist, na.rm=TRUE) # from psych
-  ik <- kurtosi(datalist, na.rm=TRUE) # from psych
+describe_data <- function(datalist, QUIET = 1) {
+  im  <- mean(datalist, na.rm = TRUE)
+  isd <- sd(datalist, na.rm = TRUE)
+  isk <- psych::skew(datalist, na.rm = TRUE) # from psych
+  ik  <- kurtosi(datalist, na.rm = TRUE) # from psych
   mis <- sum(is.na(datalist)) # missing
-  ir <- max(datalist, na.rm=TRUE) -          
-    min(datalist, na.rm=TRUE) # range
+  ir  <- max(datalist, na.rm = TRUE) -
+    min(datalist, na.rm = TRUE) # range
 
@@
   # outliers
-  lowerq = quantile(datalist, na.rm=TRUE)[2]
-  upperq = quantile(datalist, na.rm=TRUE)[4]
-  iqr = upperq - lowerq #Or use IQR(df)
+  lowerq <- quantile(datalist, na.rm = TRUE)[2]
+  upperq <- quantile(datalist, na.rm = TRUE)[4]
+  iqr    <- upperq - lowerq # Or use IQR(df)
   # gang med 3 for extreme outliers, 1.5 for mild outliers
 
-  mild.threshold.upper = (iqr * 3) + upperq
-  mild.threshold.lower = lowerq - (iqr * 3)
+  mild.threshold.upper <- (iqr * 3) + upperq
+  mild.threshold.lower <- lowerq - (iqr * 3)
 
@@
   # kilde: https://stackoverflow.com/questions/19353116/how-do-i-print-values-in-a-list-that-are-greater-than-a-certain-number-along-wit#19353192
-  outliers_upper <- which(datalist>mild.threshold.upper)
+  outliers_upper <- which(datalist > mild.threshold.upper)
   outliersu      <- datalist[outliers_upper]
-  outliers_lower <- which(datalist<mild.threshold.lower)
+  outliers_lower <- which(datalist < mild.threshold.lower)
   outliersl      <- datalist[outliers_lower]
@@
 
-  #hist (datalist, main=i) # histogram
-  #readline(prompt = "Press enter")
-
-  return(c("mean"=round(im,3), "skew"=round(isk,3), "kurtosi"=round(ik,3)))
-           #round(outliersu,3), round(outliersl,3)))
+  # hist (datalist, main=i) # histogram
+  # readline(prompt = "Press enter")
+
+  return(c("mean" = round(im, 3), "skew" = round(isk, 3), "kurtosi" = round(ik, 3)))
+  # round(outliersu,3), round(outliersl,3)))
 }
@@
 ```{r}
-stolpediagram <- ggplot(dc, aes(x=studiested2, y=var, fill=studiested2)) + 
-  geom_bar(stat="identity", position="dodge") +
-  #scale_fill_manual(values=viridis(3)) + 
-  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
-                  linewidth=.3,    # Thinner lines
-                  width=.2,
-                  position=position_dodge(.9)) +
+stolpediagram <- ggplot(dc, aes(x = studiested2, y = var, fill = studiested2)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  # scale_fill_manual(values=viridis(3)) +
+  geom_errorbar(aes(ymin = var - sd, ymax = var + sd), # se
+    linewidth = .3, # Thinner lines
+    width = .2,
+    position = position_dodge(.9)
+  ) +
   ggtitle("PCK, TK og TPACK fordelt på studiested") +
-  guides(fill="none") +
+  guides(fill = "none") +
   # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
@@
 for (var1 in names(data_clean)[names(data_clean) %in% c("Gender", "Trinn", "semester", "TK", "PCK")]) {
-  xy_plot <- ggplot(data_clean, aes(.data[[var1]],TPACK)) +
+  xy_plot <- ggplot(data_clean, aes(.data[[var1]], TPACK)) +
     geom_point() + # shape=1,alpha=1/5
     geom_jitter() +
-        ggtitle(paste("scatterplot",var1,"TPACK"))
+    ggtitle(paste("scatterplot", var1, "TPACK"))
 
@@
 ```{r TPACK_gender_institute}
-interaction.plot(x.factor = factor(data_clean$Gender),
-                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
+interaction.plot(
+  x.factor = factor(data_clean$Gender),
+  trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
   response = data_clean$TPACK,
@@
   xlab = "Gender",
-                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
-                 trace.label = "Institute")
+  col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
+  trace.label = "Institute"
+)
 ```
@@
 model_TK_covariates_l1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)),
-                clusters = factor(studiested), data=data_clean)
+  clusters = factor(studiested), data = data_clean
+)
 summary(model_TK_covariates_l1)
@@
 # studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-m4 <- lm_robust(scale(PCK) ~ 
+m4 <- lm_robust(
+  scale(PCK) ~
     factor(Gender) * (factor(Trinn) + factor(semester)) + # level 1
     factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)), # level 2
-                clusters = factor(studiested), data=data_clean)
+  clusters = factor(studiested), data = data_clean
+)
 summary(m4)
@@
 table_variance <- data.frame(
-  c(nu_e_baseline, (nu_e_baseline - nu_e_level1) / nu_e_baseline, (nu_e_level1 - nu_e_PCK) / nu_e_level1, (nu_e_PCK - nu_e_TK) / nu_e_PCK),
-  c(nu_00_baseline, (nu_00_baseline - nu_00_level1) / nu_00_baseline, (nu_00_baseline - nu_00_PCK) / nu_00_baseline, (nu_00_baseline - nu_00_TK) / nu_00_baseline),
+  c(
+    nu_e_baseline,
+    (nu_e_baseline - nu_e_level1) / nu_e_baseline,
+    ((nu_e_baseline - nu_e_PCK) / nu_e_baseline) - ((nu_e_baseline - nu_e_level1) / nu_e_baseline),
+    ((nu_e_baseline - nu_e_TK) / nu_e_baseline) - ((nu_e_baseline - nu_e_PCK) / nu_e_baseline) - ((nu_e_baseline - nu_e_level1) / nu_e_baseline)
+  ),
+  c(
+    nu_00_baseline,
+    (nu_00_baseline - nu_00_level1) / nu_00_baseline,
+    ((nu_00_baseline - nu_00_PCK) / nu_00_baseline) - ((nu_00_baseline - nu_00_level1) / nu_00_baseline),
+    ((nu_00_baseline - nu_00_TK) / nu_00_baseline) - ((nu_00_baseline - nu_00_PCK) / nu_00_baseline) + ((nu_00_baseline - nu_00_level1) / nu_00_baseline)
+  ),
   c(mtcl1ars, mtcl2ars - mtcl1ars, mtpars - mtcl2ars, mttars - mtpars)

### 2024-09-06 Rmd/Paper2.Rmd
Endret farger på figur.

@@
   geom_bar(stat = "identity", position = "dodge") +
-  # scale_fill_manual(values=viridis(3)) +
+  #scale_fill_manual(values=viridis(3)) +
+  scale_fill_manual(values=c("#812244", "#3fabb5", "#bee2e9")) +
   geom_errorbar(aes(ymin = var - sd, ymax = var + sd), # se
@@
 print(stolpediagram) # vis plot
+
+#ggsave("../output/TPACK_stolpediagram_new.png")
 ```

### 2024-09-07 Rmd/Paper2.Rmd
Added points for evaluation / discussion

--- 2024-09-05 Rmd/Paper2.Rmd
@@
 ---
+
+TODO
+
+Check baseline model. Hvordan skal jeg tolke modellen jeg tester? Hva representerer variablene jeg putter inn i modellen, og hva representerer y-verdien?
+
+MLM: TPACK~factor(studiested) er lik OLS: TPACK~factor(studiested)
+
+MLM: TPACK~1 er ikke lik OLS: TPACK~factor(studiested)
+
+Hva er rett måte å sette opp en baseline model på i OLS? (med lm_robust og clusters)
+
+Skal studiested være med som variabel i likningen (fixed effect), eller bare som en cluster (random effect)?
+
+Beregner jeg individuell TPACK eller studiested TPACK? TPACK for den enkelte PST, eller TPACK for hvert studiested?
+
+Hvis studiested er med som variable, beregner jeg TPACK per studiested, ikke student? Hvis studiested ikke er med som variabel, regner jeg TPACK for den enkelte student?
 
@@
 The purpose of the multiple regression analysis is to explain the variation in the data.
-I therefore use hierarchical regression [@howitt_introduction_2014] and add the variables step-by-step to examine changes in R2 (explained variance).
+I therefore use hierarchical regression [@howitt_introduction_2014; @kelley_multiple_2013] and add the variables step-by-step to examine changes in R2 (explained variance).
 
@@
 # sjekk: er det riktig å skalere?
-model_TK_baseline <- lm_robust(TK ~ factor(studiested), 
+model_TK_baseline <- lm_robust(TK ~ 1, #factor(studiested), 
                 clusters = factor(studiested), data=data_clean)
@@
 
+```{r}
+hierarchical_fstat <- function(R_qr, R_q, df_num, df_den) {
+  # calculate fstat
+  # p. 92 in Kelley & Bolin (2013)
+  # numerator df = r; r = diff antall variable
+  # denominator df = N - q - r - 1; N = sample size, q = antall variable i enkleste modell
+
+  fstat <- ((R_qr - R_q) / df_num) / ((1 - R_qr) / df_den)
+  p     <- pf(fstat, df_num, df_den, lower.tail=F)
+
+  return(p)
+}
+
+df_num <- 4 + length(model_TK_level1$term) - length(model_TK_baseline$term)
+df_den <- 4 + length(data_clean$TK) - length(model_TK_level1$term) - 1
+R_qr   <- model_TK_level1$adj.r.squared
+R_q    <- model_TK_baseline$adj.r.squared
+
+hierarchical_fstat(R_qr, R_q, df_num, df_den)
+```
+
 The F-statistic indicate that the two models are not significantly different. This means that level1 effects do not contribute significantly to the variance.
@@
 
+```{r}
+df_num <- 4 + length(model_PCK_level1$term) - length(model_PCK_baseline$term)
+df_den <- 4 + length(data_clean$PCK) - length(model_PCK_level1$term) - 1
+R_qr   <- model_PCK_level1$adj.r.squared
+R_q    <- model_PCK_baseline$adj.r.squared
+
+hierarchical_fstat(R_qr, R_q, df_num, df_den)
+```
+
 There were significant differences when including covariates.
@@
   TPACK ~ 
-    factor(studiested),                                 # level 2
+    1, #factor(studiested),                       # level 2
   clusters = factor(studiested), data = data_clean
@@
 summary(model_TPACK_baseline)
+```
+
+test: baseline OLS er ikke lik baseline MLM. Hvorfor ikke?
+
+```{r level2_main_test, eval=F}
+# ref: https://www.r-bloggers.com/2021/05/clustered-standard-errors-with-r/ [5.9.2024]
+
+# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
+model_TPACK_baseline_test <- lm_robust(
+  TPACK ~ 1,                       # level 2
+  clusters = factor(studiested), 
+  fixed_effects = ~factor(studiested),
+  data = data_clean
+)
+summary(model_TPACK_baseline_test)
 ```
@@
 pf(fstat, df1, df2, lower.tail = FALSE)
+```
+
+```{r}
+df_num <- 4 + length(model_TPACK_level1$term) - length(model_TPACK_baseline$term)
+df_den <- 4 + length(data_clean$TPACK) - length(model_TPACK_level1$term) - 1
+R_qr   <- model_TPACK_level1$adj.r.squared
+R_q    <- model_TPACK_baseline$adj.r.squared
+
+hierarchical_fstat(R_qr, R_q, df_num, df_den)
 ```
@@
 
+```{r}
+df_num <- 4 + length(model_TPACK_PCK$term) - length(model_TPACK_level1$term)
+df_den <- 4 + length(data_clean$TPACK) - length(model_TPACK_PCK$term) - 1
+R_qr   <- model_TPACK_PCK$adj.r.squared
+R_q    <- model_TPACK_level1$adj.r.squared
+
+hierarchical_fstat(R_qr, R_q, df_num, df_den)
+```
+
 There is a significant difference when adding PCK as a main effect.
@@
 
+```{r}
+df_num <- 4 + length(model_TPACK_TK$term) - length(model_TPACK_PCK$term)
+df_den <- 4 + length(data_clean$TPACK) - length(model_TPACK_TK$term) - 1
+R_qr   <- model_TPACK_TK$adj.r.squared
+R_q    <- model_TPACK_PCK$adj.r.squared
+
+hierarchical_fstat(R_qr, R_q, df_num, df_den)
+```
+
 There is a significant difference, when adding TK.
@@
 
+"all level 1 predictors were centered at the group mean, and all level 2 predictors were centered at the grand mean" (p. 192) (Osborne and Neupert in Teo, 2013)
+
+level 1 predictors: properties related to the PST
+level 2 predictors: properties related to the TEI
+
 ### Baseline model
@@
 # baseline model / nullmodell
+
+# MLM: TPACK~factor(studiested) er lik OLS: TPACK~factor(studiested)
+# MLM: TPACK~1 er ikke det
 

### 2024-10-08 Rmd/Paper1.Rmd
Corrections

@@
 author:       "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
-date:         "2024-09-04"
+date:         "2024-10-08"
 bibliography: "../data/meta_data/references.bib"
@@
 data_pilot$semester[data_pilot$studiested == "Institute4"] <- 4 
-# Institute7/8, ?. semester 2/3?
+# Institute7/8, ?. år 2/3
+data_pilot$semester[data_pilot$studiested == "Institute7"] <- 6 
 data_pilot$semester[data_pilot$studiested == "Institute5"] <- 8
@@
 # load data
-load(file="../data/processed_data/tpack-trelis_q_paper1_2023.Rdata")
+load(file="../data/processed_data/tpack-trelis_q_partial_2023-paper1.Rdata")
 ```
@@
 ```{r remove_NMAR_TRELISQ}
-# remove NMAR + NTNU (19)
+# remove NMAR (19)
 CFA_TPACK_PST <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),] 

### 2024-10-08 Rmd/Paper2.Rmd
Removed reanalysis done for synopsis

--- 2024-09-06 Rmd/Paper2.Rmd
@@
 title:        "Paper 2: survey analysis"
-author:       "Niklas Karlsen"
-date:         "2024-09-05"
+author:       "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
+date:         "2024-09-23"
 bibliography: "../data/meta_data/references.bib"
@@
 
+TODO
+
+Check baseline model. Hvordan skal jeg tolke modellen jeg tester? Hva representerer variablene jeg putter inn i modellen, og hva representerer y-verdien?
+
+MLM: TPACK~factor(studiested) er lik OLS: TPACK~factor(studiested)
+
+MLM: TPACK~1 er ikke lik OLS: TPACK~factor(studiested)
+
+Hva er rett måte å sette opp en baseline model på i OLS? (med lm_robust og clusters)
+
+Skal studiested være med som variabel i likningen (fixed effect), eller bare som en cluster (random effect)?
+
+Beregner jeg individuell TPACK eller studiested TPACK? TPACK for den enkelte PST, eller TPACK for hvert studiested?
+
+Hvis studiested er med som variable, beregner jeg TPACK per studiested, ikke student? Hvis studiested ikke er med som variabel, regner jeg TPACK for den enkelte student?
+
 # Introduction
@@
 library(nlme)                               # for ICC - variance HLM
+
+library(texreg)
 ```
@@
 
-## Reanalysis of OLS
-
-A reanalysis of the OLS is done here, due to some methodical reconsiderations after a closer look at the MLM.
-The dataset is slightly closer to the rawdata (3 NAs were reset to other, as in the original dataset).
-The order of the variables are also slightly different (TEIs are introduced before covariates).
-Finally, interaction effects are not included, as they were in paper 2.
-The statistical tests suggested that the interaction effects were non-significant.
-Otherwise, the analysis is similar to paper 2.
-
-### Reload data
-
-```{r}
-# ref: https://www.statology.org/clear-environment-in-r/ [5.9.2024]
-#clear all data frames from environment
-rm(list=ls(all=TRUE)[sapply(mget(ls(all=TRUE)), class) == "data.frame"])
-```
-
-Then I reload the dataset from file.
-
-```{r load_data_reanalysis}
-# Rdata-file contains reduced / prepared datasets
-load(file="../data/processed_data/tpack-TRELIS-Q_complete_2023-paper2.Rdata")
-```
-
-Then I recenter the data.
-
-```{r}
-data_TRELIS_Q[,unlist(TPACK_trelis_q)] <- data_TRELIS_Q[,unlist(TPACK_trelis_q)] - 3 
-```
-
-And add some missing information.
-
-```{r}
-# identify NA as studiested 5(?) 
-data_TRELIS_Q[data_TRELIS_Q$studiested == "Institute3","Trinn"] <- "Trinn 5-10"
-```
-
-NMAR are removed.
-
-```{r fix_missing_2, eval=T}
-# remove NMAR + Institute9 (19)
-data_clean <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),] 
-```
-
-```{r}
-scores_v23 <- scoreItems(TPACK_trelis_q, data_clean)
-
-describe(scores_v23$scores)
-
-# constructs for multiple regression
-TK_v23    <- scores_v23$scores[,"TK"]
-PCK_v23   <- scores_v23$scores[,"PCK"]
-TPACK_v23 <- scores_v23$scores[,"TPACK"]
-
-# add constructs to data.frame
-data_clean$TK    <- TK_v23
-data_clean$PCK   <- PCK_v23
-data_clean$TPACK <- TPACK_v23
-```
-
-### Hierarchical RA for TK
-
-Level 1: students  
-Level 2: institution
-
-We are using cluster corrected standard errors to correct for clustering on institution.
-
-Add student covariates
-
-Interaction effects are examined by calculating F-statistics to see if models with and without interaction are significantly different.
-
-#### baseline model (nullmodell)
-
-```{r TK_baseline}
-# kovariater; we use scale to standardize when using interaction effects
-# sjekk: er det riktig å skalere?
-model_TK_baseline <- lm_robust(TK ~ factor(studiested), 
-                clusters = factor(studiested), data=data_clean)
-summary(model_TK_baseline)
-```
-
-#### Level 1
-
-First, all covariates at the individual PST level 1 (i.e., gender, target grade level and year of study) are introduced as a block.
-
-Without interaction
-
-```{r TKlevel1_main}
-# kovariater; we use scale to standardize when using interaction effects
-# sjekk: er det riktig å skalere?
-model_TK_level1 <- lm_robust(TK ~ factor(studiested) + factor(Gender) + (factor(Trinn) + factor(semester)), 
-                clusters = factor(studiested), data=data_clean)
-summary(model_TK_level1)
-```
-
-Calculate F-stat by comparing R2 to see if they are significantly different
-
-```{r}
-# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
-# source: p. 257 i Howitt & Cramer, 2014
-fstat <- model_TK_level1$adj.r.squared / model_TK_baseline$adj.r.squared
-df1   <- length(data_clean$TK) - 1
-df2   <- df1
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-The F-statistic indicate that the two models are not significantly different. This means that level1 effects do not contribute significantly to the variance.
-
-### Hierarchical RA for PCK
-
-Level 1: students  
-Level 2: institution
-
-We are using cluster corrected standard errors to correct for clustering on institution.
-
-Add student covariates
-
-#### Baseline
-
-```{r PCKlevel2_main}
-# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-model_PCK_baseline <- lm_robust(
-  PCK ~ 
-    factor(studiested),                                 # level 2
-  clusters = factor(studiested), data = data_clean
-)
-summary(model_PCK_baseline)
-```
-
-#### Level 1
-
-without interaction, only main effects
-
-```{r PCKlevel1_main}
-# kovariater; we use scale to standardize when using interaction effects
-# sjekk: er det riktig å skalere?
-model_PCK_level1 <- lm_robust(PCK ~ factor(studiested) + factor(Gender) + factor(Trinn) + factor(semester),
-  clusters = factor(studiested), data = data_clean
-)
-summary(model_PCK_level1)
-```
-
-Calculate F-stat to see if they are significantly different
-
-```{r}
-# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
-# source: p. 257 i Howitt & Cramer, 2014
-fstat <- model_PCK_level1$adj.r.squared / model_PCK_baseline$adj.r.squared
-df1   <- length(data_clean$PCK) - 1
-df2   <- df1
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-There were significant differences when including covariates.
-
-### Hierarchical RA for TPACK
-
-Level 1: students  
-Level 2: institution
-
-We are using cluster corrected standard errors to correct for clustering on institution.
-
-Add student covariates
-
-#### Baseline
-
-```{r level2_main}
-# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
-model_TPACK_baseline <- lm_robust(
-  TPACK ~
-    factor(studiested),                                 # level 2
-  clusters = factor(studiested), data = data_clean
-)
-summary(model_TPACK_baseline)
-```
-
-#### Level 1
-
-Covariates are first added as a block.
-
-Main effects
-
-```{r level1_main}
-# kovariater; we use scale to standardize when using interaction effects
-# sjekk: er det riktig å skalere?
-model_TPACK_level1 <- lm_robust(
-  TPACK ~
-    factor(studiested) + factor(Gender) + factor(Trinn) + factor(semester),
-  clusters = factor(studiested), data = data_clean
-)
-summary(model_TPACK_level1)
-```
-
-Calculate F-stat to see if they are significantly different
-
-```{r}
-# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
-# source: p. 257 i Howitt & Cramer, 2014
-fstat <- model_TPACK_level1$adj.r.squared / model_TPACK_baseline$adj.r.squared
-df1 <- length(data_clean$TPACK) - 1
-df2 <- df1
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-There is no significant difference when including covariates.
-
-#### PCK
-
-After the covariates are added as a block, PCK is added.
-
-```{r PCK}
-# PCK
-model_TPACK_PCK <- lm_robust(
-  TPACK ~
-    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
-    factor(studiested) +                                # level 2
-    PCK,
-  clusters = factor(studiested), data = data_clean
-)
-summary(model_TPACK_PCK)
-
-# main
-fstat <- model_TPACK_PCK$adj.r.squared / model_TPACK_level1$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-
-```
-
-There is a significant difference when adding PCK as a main effect.
-
-#### TK
-
-Finally, TK is added.
-
-```{r TK}
-# TK
-model_TPACK_TK <- lm_robust(
-  TPACK ~
-    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
-    factor(studiested) +                                # level 2
-    PCK + TK,
-  clusters = factor(studiested), data = data_clean
-)
-summary(model_TPACK_TK)
-
-fstat <- model_TPACK_TK$adj.r.squared / model_TPACK_PCK$adj.r.squared
-pf(fstat, df1, df2, lower.tail = FALSE)
-```
-
-There is a significant difference, when adding TK.
-
-### Results summarised
-
-The results differ somewhat from paper 2 in that some interaction effects are excluded, which was included in the analysis done for paper 2. The main trends are however similar.
-
-```{r}
-# table
-
-# TK
-
-mtkcl1rs  <- model_TK_baseline$r.squared
-mtkcl1ars <- model_TK_baseline$adj.r.squared
-
-mtkcl2rs  <- model_TK_level1$r.squared
-mtkcl2ars <- model_TK_level1$adj.r.squared
-
-TK_R2     <- data.frame(
-  "R2"    = c(mtkcl1rs, mtkcl2rs),
-  "R2adj" = c(mtkcl1ars, mtkcl2ars),
-  "diff"  = c(mtkcl1ars, mtkcl2ars - mtkcl1ars)
-)
-
-rownames(TK_R2) <- c("Teacher education institute", "Gender, year of study, target grade level")
-
-pander(TK_R2)
-```
-
-```{r}
-# PCK
-
-mpcl1rs   <- model_PCK_baseline$r.squared
-mpcl1ars  <- model_PCK_baseline$adj.r.squared
-
-mpcl2rs   <- model_PCK_level1$r.squared
-mpcl2ars  <- model_PCK_level1$adj.r.squared
-
-PCK_R2    <- data.frame(
-  "R2"    = c(mpcl1rs, mpcl2rs),
-  "R2adj" = c(mpcl1ars, mpcl2ars),
-  "diff"  = c(mpcl1ars, mpcl2ars - mpcl1ars)
-)
-
-rownames(PCK_R2) <- c("Teacher education institute", "Gender, year of study, target grade level")
-
-pander(PCK_R2)
-```
-
-```{r}
-# TPACK
-
-mtcl1rs  <- model_TPACK_baseline$r.squared
-mtcl1ars <- model_TPACK_baseline$adj.r.squared
-
-mtcl2rs  <- model_TPACK_level1$r.squared
-mtcl2ars <- model_TPACK_level1$adj.r.squared
-
-mtprs    <- model_TPACK_PCK$r.squared
-mtpars   <- model_TPACK_PCK$adj.r.squared
-
-mttrs    <- model_TPACK_TK$r.squared
-mttars   <- model_TPACK_TK$adj.r.squared
-
-TPACK_R2 <- data.frame(
-  "R2" = c(mtcl1rs, mtcl2rs, mtprs, mttrs),
-  "R2adj" = c(mtcl1ars, mtcl2ars, mtpars, mttars),
-  "diff" = c(mtcl1ars, mtcl2ars - mtcl1ars, 
-             mtpars - mtcl2ars, mttars - mtpars)
-)
-
-rownames(TPACK_R2) <- c("Teacher education institute", "Gender, year of study, target grade level", "PCK", "TK")
-
-pander(TPACK_R2)
-```
-
-Level 1: gender, target grade level  
-Level 2: year of study, institute
-
-Burde semester vært en level 2 variabel, siden det er mer en egenskap ved instituttet enn studenten?
-
-Spørsmålet er om forskjeller på semester skyldes endringer på instituttet (level 2) eller endringer i studentmassen (level 1)? Både-og? Samspill på tvers av nivåer!
-
-## Multi-level modeling with cross-level interaction effects
-
-When the data is nested, multi-level modeling can be used.
-However, when the number of level 2-units is below 10, as it is in our dataset (we have only 5 TEIs), OLS with cluster corrected standard errors can be used.
-I, however, do a MLM here to compare with the OLS results.
-MLM is only done for TPACK as dependent variable to illustrate the method.
-
-The r package NLME [@pinheiro_nlme_2023] is used for the analysis.
-
-### Baseline model
-
-First, the baseline model is calculated as a reference and to find the ICC.
-
-```{r}
-# baseline model / nullmodell
-
-# REML is used for parameter estimation
-model_MLM_baseline <- lme(TPACK~1, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
-
-# ML is used for model fit
-model_MLM_baseline_fit <- lme(TPACK~1, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
-
-summary(model_MLM_baseline)
-summary(model_MLM_baseline_fit)
-```
-
-```{r}
-# ICC
-nu_00 <- var(ranef(model_MLM_baseline)) # intercept (i.e., TEIs)
-nu_e  <- var(resid(model_MLM_baseline)) # residuals (i.e., PSTs)
-ICC   <- nu_00 / (nu_00 + nu_e)
-ICC
-```
-
-30 % av variansen i TPACK ligger på nivå 2.
-Dette er større enn den anbefalte grensen på 5 %.
-Vurder om OLS-regresjon med klyngekorrigert standardfeil egner seg bedre enn flernivåanalyse [@christophersen_introduksjon_2018] (p.
-109).
-
-```{r}
-# effektiv utvalgsstørrelse
-n_e <- 147 / (1 + ((147 / 5) - 1) * ICC)
-n_e
-```
-
-Utvalgsstørrelsen korrigert for avhengighet er lik 15.
-
-### Level 1 model (random intercept)
-
-The level 1 model extends the baseline model by including the covariates.
-
-#### Level 1 variables
-
-```{r}
-# REML is used for parameter estimation
-model_MLM_level1 <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
-
-# ML is used for model fit
-model_MLM_level1_fit <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
-
-summary(model_MLM_level1)
-summary(model_MLM_level1_fit)
-```
-
-Er level 1-modellen bedre enn nullmodellen?
-
-```{r eval=T}
-anova(model_MLM_baseline_fit, model_MLM_level1_fit)
-```
-
-#### Level 2 variables
-
-Since Level 2 variables are unique to MLM, and the solution does not converge, level 2 analysis is not done here.
-
-La kovariatene variere mellom studiesteder
-
-```{r eval=F}
-# REML is used for parameter estimation
-model_level2 <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="REML", na.action="na.omit", random= ~ semester|studiested)
-
-# ML is used for model fit
-model_level2_fit <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="ML", na.action="na.omit", random= ~ semester|studiested)
-
-summary(model_level2)
-summary(model_level2_fit)
-```
-
-Er level 2-modellen bedre enn level 1?
-
-```{r eval=F}
-anova(model_level2_fit, model_MLM_level1_fit)
-```
-
-### PCK
-
-```{r}
-# grand mean centering: scale(TK, center = T, scale = F)
-
-data_clean$PCK_cent <- scale(data_clean$PCK, center = T, scale = F)
-data_clean$TK_cent  <- scale(data_clean$TK, center = T, scale = F)
-
-# REML is used for parameter estimation
-model_MLM_PCK <- lme(TPACK~Gender + semester + Trinn + PCK_cent, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
-
-# ML is used for model fit
-model_MLM_PCK_fit <- lme(TPACK~Gender + semester + Trinn + PCK_cent, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
-
-summary(model_MLM_PCK)
-summary(model_MLM_PCK_fit)
-```
-
-```{r eval=T}
-anova(model_MLM_PCK_fit, model_MLM_level1_fit)
-```
-
-### TK
-
-```{r}
-# REML is used for parameter estimation
-model_MLM_TK <- lme(TPACK~Gender + semester + Trinn + PCK_cent + TK_cent, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)
-
-# ML is used for model fit
-model_MLM_TK_fit <- lme(TPACK~Gender + semester + Trinn + PCK_cent + TK_cent, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)
-
-summary(model_MLM_TK)
-summary(model_MLM_TK_fit)
-```
-
-```{r eval=T}
-anova(model_MLM_PCK_fit, model_MLM_TK_fit)
-```
-
-## Comparison of OLS and MLM
-
-### Comparison of variance
-
-I follow @christophersen_introduksjon_2018 (p. 115) and use the variance estimates as an indicator of R2.
-
-```{r}
-# MLM
-# variance baseline
-nu_00_baseline <- var(ranef(model_MLM_baseline)) # intercept (i.e., TEIs)
-nu_e_baseline  <- var(resid(model_MLM_baseline)) # residuals (i.e., PSTs)
-
-# variance level 1
-nu_00_level1   <- var(ranef(model_MLM_level1)) # intercept (i.e., TEIs)
-nu_e_level1    <- var(resid(model_MLM_level1)) # residuals (i.e., PSTs)
-
-# variance PCK
-nu_00_PCK      <- var(ranef(model_MLM_PCK)) # intercept (i.e., TEIs)
-nu_e_PCK       <- var(resid(model_MLM_PCK)) # residuals (i.e., PSTs)
-
-# variance TK
-nu_00_TK       <- var(ranef(model_MLM_TK)) # intercept (i.e., TEIs)
-nu_e_TK        <- var(resid(model_MLM_TK)) # residuals (i.e., PSTs)
-
-# OLS
-mtcl1ars       <- model_TPACK_baseline$adj.r.squared
-mtcl2ars       <- model_TPACK_level1$adj.r.squared
-mtpars         <- model_TPACK_PCK$adj.r.squared
-mttars         <- model_TPACK_TK$adj.r.squared
-
-# table of difference in variance
-
-table_variance <- data.frame(
-  c(
-    nu_e_baseline,
-    (nu_e_baseline - nu_e_level1) / nu_e_baseline,
-    ((nu_e_baseline - nu_e_PCK) / nu_e_baseline) - ((nu_e_baseline - nu_e_level1) / nu_e_baseline),
-    ((nu_e_baseline - nu_e_TK) / nu_e_baseline) - ((nu_e_baseline - nu_e_PCK) / nu_e_baseline) - ((nu_e_baseline - nu_e_level1) / nu_e_baseline)
-  ),
-  c(
-    nu_00_baseline,
-    (nu_00_baseline - nu_00_level1) / nu_00_baseline,
-    ((nu_00_baseline - nu_00_PCK) / nu_00_baseline) - ((nu_00_baseline - nu_00_level1) / nu_00_baseline),
-    ((nu_00_baseline - nu_00_TK) / nu_00_baseline) - ((nu_00_baseline - nu_00_PCK) / nu_00_baseline) + ((nu_00_baseline - nu_00_level1) / nu_00_baseline)
-  ),
-  c(mtcl1ars, mtcl2ars - mtcl1ars, mtpars - mtcl2ars, mttars - mtpars)
-)
-
-rownames(table_variance) <- c("TEI", "covariates", "PCK", "TK")
-colnames(table_variance) <- c("MLM Level 1 % var", "MLM Level 2 % var", "OLS reanalysis diff R2")
-
-pander(table_variance)
-```
-
-The variance differ somewhat from the standard deviation (when squared) reported in the summary of the MLMs. Why? Is the OLS a kind of average of the MLM, which is separated into level 1 and level 2?
-
 # Discussion and conclusion
@@
 I was able to replicate the results discussed in @karlsen_assessing_2024.
-
-The reanalysis using OLS gave similar results as for paper 2.
-Only TEIs explained variance in TK, not the covariates.
-The covariates did however significantly explain the variance in PCK, but the TEIs did not (or to a small degree, only 3%).
-The variance in TPACK was significantly explained by TEI, PCK and TK, not the covariates.
-
-The results in paper 2 indicated that the covariates contributed to explain the variance in TPACK.
-The reanalysis done here, suggest that this may be due to the order in which the variables were added to the model being tested.
-If I follow the order used in the MLM, the reanalysis suggests that the covariates do not contribute to TPACK.
-
-The MLM gave supporting results.
-Including the covariates did not improve the model, whereas including PCK and TK did.
-The variance at level 2 (between TEIs) was due to TK, not PCK or covariates, that is, there were significant differences in TK between TEIs.
-At level 1 (between PSTs) 6% of the variance were due to the covariates, 17 % due to PCK and 55 % due to TK.
-
-The OLS reanalysis suggested that 30 % of the variance in TPACK was due to TEIs, 11 % due to PCK and 32 % due to TPACK.
-The covariates did not contribute to the variance in TPACK.
-Even though the percentages differ between the OLS and the MLM, they show the same trends.
-The different percentages may be due to how R2 (or difference in variance) is calculated differently in OLS and MLM.
 

### 2024-10-08 Rmd/Paper2.Rmd
Removed comments on MLM and OLS reanalysis

@@
 author:       "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
-date:         "2024-09-23"
+date:         "2024-10-08"
 bibliography: "../data/meta_data/references.bib"
@@
     toc_float: true
-
-knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), "../output/paper2.html") )})
+    number_sections: true
+
+knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), "../output/Paper2.html") )})
 ---
-
-TODO
-
-Check baseline model. Hvordan skal jeg tolke modellen jeg tester? Hva representerer variablene jeg putter inn i modellen, og hva representerer y-verdien?
-
-MLM: TPACK~factor(studiested) er lik OLS: TPACK~factor(studiested)
-
-MLM: TPACK~1 er ikke lik OLS: TPACK~factor(studiested)
-
-Hva er rett måte å sette opp en baseline model på i OLS? (med lm_robust og clusters)
-
-Skal studiested være med som variabel i likningen (fixed effect), eller bare som en cluster (random effect)?
-
-Beregner jeg individuell TPACK eller studiested TPACK? TPACK for den enkelte PST, eller TPACK for hvert studiested?
-
-Hvis studiested er med som variable, beregner jeg TPACK per studiested, ikke student? Hvis studiested ikke er med som variabel, regner jeg TPACK for den enkelte student?
 
@@
 
-For the multi-level modeling (MLM) the PSTs are considered level 1 units and the TEIs are level 2.
-The covariates gender and target grade are related to the level 1 units, while year of study can vary at both levels.
-
 In addition, TK and PCK are considered independent variables, while TPACK is the dependent variable.
-
-Both the OLS and the MLM are done by introducing the variables sequentially.
-For the OLS done in paper 2, the level 1 variables (i.e., gender, trinn and semester) are introduced first, then the level 2 variables (i.e., TEIs), and finally, PCK and TK.
-For the MLM, a baseline model (null-model) is created first, where the TEIs are set as random (i.e., the intercept is allowed to vary among TEIs).
-Then the level 1 variables are introduced to create the level 1-model.
-Then PCK and TK are added to the model.
-Since the order of introducing TEI was altered between the OLS in paper 2 and the MLM, a reanalysis of the OLS was done with the TEIs introduced before the covariates.
-This could affect how much of the variance was explained by the covariates.
-
-The explained variance (i.e., R2) is calculated for both methods, and can be used to compare the results of the two methods.
-
-The following models are tested:
-
-Model | Paper 2 OLS | Reanalysis OLS | New analysis MLM 
-------------- | ------------- | ------------- | ------------- 
-Options | 3xAnnet = NA | N/A | N/A
-Nullmodell | N/A | TPACK ~ TEIs | TPACK ~ 1
-Level 1 | TPACK ~ covariates + TEIs | TPACK ~ TEIs + covariates | TPACK ~ covariates
-Level 2 | N/A | N/A | + year of study | TEI
-PCK | TPACK ~ covariates + TEIs + PCK | TPACK ~ TEIs + covariates + PCK | TPACK ~ covariates + PCK_sentrert
-TK | TPACK ~ covariates + TEIs + PCK + TK | TPACK ~ TEIs + covariates + PCK + TK | TPACK ~ covariates + PCK_s + TK_sentrert
-
-The analysis in paper 2 changed the answers of three respondents from "Other" to NA.
-This made it impossible to compare the fit for the MLMs since the number of respondents were dissimilar for the models compared when including these covariates.
-For the reanalyses I therefore kept these values unchanged.
-
-In the MLM the TEIs are added as random variables (i.e., the intercepts are allowed to differ between TEIs), while the TEIs are added as dummy variables in the OLS. This should give similar results.
-
-In the MLM the covariates are added as fixed variables (i.e., the slopes are the same for all PSTs independent of TEI).
-
-The order of the models was also changed in the reanalysis, as the baseline model in the MLM corresponds to having the TEIs as dummy variables.
-In paper 2 this model was tested after having controlled for the covariates.
-In the reanalysis this order was switched to correspond to the MLM.
-
-Only the MLM has the option to test for level 2 covariates. I allowed slope of year of study vary between TEIs. The level2 model is therefore not tested. It did not converge.
-
-For the MLM the independendent variables were centred using the grand mean, following @christophersen_introduksjon_2018.
-
-The different methods can be compared by comparing R-squared, and diff R-squared.
 

### 2024-10-08 Rmd/Paper4.Rmd
Prepared for publication

@@
 ---
-title: "Paper 4"
-author: "Niklas Karlsen"
-date: "2024-08-22"
+title: "Paper 4: quasi-experiment"
+author: "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
+date: "2024-10-08"
+bibliography: "../data/meta_data/references.bib"
+csl:          "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl" # (source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html)
 output: 
@@
     toc: true
-    toc_depth: 3
+    toc_depth: 4
     toc_float: true
-
-knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/paper4.html')) })
+    number_sections: true
+
+knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/Paper4.html')) })
 ---
 
-This is a computational essay?
-
-Changelog
-
-* 22.8.2024: added comparison of two hybrid models
-
-TODO
-
-# Preamble / technical details
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE) #,results="asis")
-
-Sys.setenv(LANG = "no_NB.UTF-8")            # for æ,ø og å. Kilde:
-Sys.setlocale("LC_ALL","Norwegian")
-# https://stackoverflow.com/questions/16347731/how-to-change-the-locale-of-r
-
-# variable
-plot_to_file = 0
-```
-
-```{r last_bibliotek, echo=FALSE, warning=FALSE, message=FALSE}
-## Last bibliotek: rekkefølge - de siste overstyrer de første
-
+# Introduction
+
+The purpose of this script is to present the results of the quasi-experiment in paper 4, by presenting the calculations presented in paper 4 for prototype 1 and 2. Three effect sizes are calculated:
+
+* Difference (gain) scores / Hedges' g
+* Fixed-effects analysis
+* Within-between (hybrid) model
+
+## Load libraries
+
+```{r last_bibliotek}
 ## bibliotek for tabell-behandling
-# last plyr før dplyr
-#library(plyr)                               # for plot of mean and sd
 library(dplyr)
@@
 library(pander)                             # prettyprint for tables
-# kilde: https://stackoverflow.com/questions/31626419/in-knitr-no-output-from-pander-in-for-loop
-#panderOptions('knitr.auto.asis', FALSE)     # to pander in loop
-#library(tidyverse)
+library(texreg)                               # screenreg (print compare results regresion)
 
 ## bibliotek for plotting
-#library(plotrix)                            # SE(m) (plot mean+CI)
 library(ggplot2)
-#library(GGally)                             # for plot av ggpairs
-# kilde: https://ggplot2-book.org/arranging-plots.html#arranging-plots
-#library(patchwork)                          # for inset_element
+library(viridis)
 
 ## bibliotek for statistikk
-#library(bruceR,quietly=TRUE)                           # for manova
-#library(DescTools)                          # for lambda, categorical data
 library(psych)                              # for ICC, SD
 library(effsize)                            # for cohen's d
-#library(corrplot)
-#library(rcompanion)                          # for Cramers V
-#library(Rfit)                               # ranked regression # https://journal.r-project.org/archive/2013/RJ-2013-027/RJ-2013-027.pdf
 
@@
 library(nlme)                                 # for lme
-#library(rmcorr)
 library(plm)                                  # paneldata analysis
@@
 library(lmtest) # logLikelihood tests, AIC, BIC ,...
-#library(MuMIn) # AICc
 library(aod) # wald.test
-
-library(texreg)                               # screenreg (print compare results regresion)
-library(viridis)
-
-#library(ppcor) # partiell correlation
-#library(mice) # missing
 library(MuMIn) # AICc
@@
 
-## Definer noen hjelpefunksjoner.
-
-```{r last_hjelpefunksjoner}
-source("../../scripts/help-functions.R")
-```
-
 ## Definer constructs for lettere access i kolonnene
 
+These are the items that were used. The TK-items differ somewhat from the items used in @karlsen_assessing_2024.
+
 ```{r load_constructs}
-source("../../scripts/constructs.R")
-```
-
-# Introduction
-
-The purpose of this script is to do the calculations presented in paper 4 for prototype 1 and 2.
-
-* Difference (gain) scores / Hedges' g
-* Fixed-effects analysis
-* Within-between (hybrid) model
+#source("../../scripts/constructs.R")
+
+# TPACK-PST 22
+TK_PST22_pos <- c("TK2","TK5","TK6") # ,"TK5","TK6","TCK1"
+PCK_PST22_pos <- c("CK1","CK2","CK3","CK5","PCK2","PCK5") # "CK2", "CK5",
+TPACK_PST22_pos <- c("TCK1","TCK4","TPK1","TPK4","TPACK2", "TPACK4")
+
+TPACK_PST_emne2_v22 <- list("TK"=TK_PST22_pos,
+                            "PCK"=PCK_PST22_pos,
+                            "TPACK"=TPACK_PST22_pos)
+
+# TPACK-PST 22 adjusted to 23
+TK_PST23_pos <- c("TK2","TK5","TK6") # "TK4","TK5","TK6"
+PCK_PST23_pos <- c("CK123", "PCK2")
+TPACK_PST23_pos <- c("TPACK2","TCK4","TPK1","TPK4","TPACK4") 
+TPACK_PST_emne2_v23 <- list("TK"=TK_PST23_pos,
+                            "PCK"=PCK_PST23_pos, 
+                            "TPACK"=TPACK_PST23_pos) #, 
+```
 
@@
 
-Properties of samples
+### Prepare dataset
 
@@
 
-add group information
-
-```{r add_groups_v22}
-# add column with group numbers
-data_emne2_v22$group <- c(1:nrow(data_emne2_v22))
-data_emne2_v22[data_emne2_v22$Subject == 'Gabriel' |
-                 data_emne2_v22$Subject == 'Hulda' |
-                 data_emne2_v22$Subject == 'Isak' |
-                 data_emne2_v22$Subject == 'Maria' |
-                 data_emne2_v22$Subject == 'Nicoline',"group"] <-
-                 "armageddon" #1
-
-data_emne2_v22[data_emne2_v22$Subject == 'Conny' |
-                 data_emne2_v22$Subject == 'Fride' |
-                 data_emne2_v22$Subject == 'Jonas' |
-                 data_emne2_v22$Subject == 'Julia' |
-                 data_emne2_v22$Subject == 'Lea',"group"] <-
-                 "fasetilstander" # 2
-
-data_emne2_v22[data_emne2_v22$Subject == 'Berta' |
-                 data_emne2_v22$Subject == 'Dina' |
-                 data_emne2_v22$Subject == 'Gloria' |
-                 data_emne2_v22$Subject == 'Klaudia',"group"] <- 
-                 "atomeksitasjon" # 3
-
-data_emne2_v22[data_emne2_v22$Subject == 'Benjamin' |
-                 data_emne2_v22$Subject == 'Ester' |
-                 data_emne2_v22$Subject == 'Harald' |
-                 data_emne2_v22$Subject == 'Inga' |
-                 data_emne2_v22$Subject == 'Olava',"group"] <- 
-                 "smittespredning" # 4
-
-data_emne2_v22[data_emne2_v22$Subject == 'David' |
-                 data_emne2_v22$Subject == 'Elias' |
-                 data_emne2_v22$Subject == 'Filip',"group"] <-
-                 "trykkbeholder" # 5
-
-data_emne2_v22 <- data_emne2_v22 %>% relocate(group, .after = Subject)           # bruker dplyr
-```
-
-```{r add_groups_v23}
-# add column with group numbers
-data_emne2_v23$group <- rep(NA,nrow(data_emne2_v23))
-data_emne2_v23[data_emne2_v23$Subject == 'Berta' |
-                 data_emne2_v23$Subject == 'Fride' |
-                 data_emne2_v23$Subject == 'Gloria' |
-                 data_emne2_v23$Subject == 'Nora',"group"] <-
-                 "kokeplate" #1/A
-
-data_emne2_v23[data_emne2_v23$Subject == 'Conny' |
-                 data_emne2_v23$Subject == 'Idun' |
-                 data_emne2_v23$Subject == 'Magrit' |
-                 data_emne2_v23$Subject == 'Vaarin',"group"] <-
-                 "arv øyefarge" # 2/B
-
-data_emne2_v23[data_emne2_v23$Subject == 'Julia' |
-                 data_emne2_v23$Subject == 'Dina' |
-                 data_emne2_v23$Subject == 'Ruben',"group"] <- 
-                 "partikkelmodellen gass" # 3/C
-
-data_emne2_v23[data_emne2_v23$Subject == 'Emmanuel' |
-                 data_emne2_v23$Subject == 'Klaudia' |
-                 data_emne2_v23$Subject == 'Noah' | 
-                 data_emne2_v23$Subject == "Christian","group"] <- 
-                 "partikkelmodellen vannglass" # 4/E
-
-data_emne2_v23[data_emne2_v23$Subject == 'Gabriel' |
-                 data_emne2_v23$Subject == 'Harald' |
-                 data_emne2_v23$Subject == 'Lukas' |
-                 data_emne2_v23$Subject == 'Mikael',"group"] <-
-                 "friksjon" # 5/F
-
-data_emne2_v23[data_emne2_v23$Subject == 'Olava' |
-               data_emne2_v23$Subject == "Yosita" | 
-               data_emne2_v23$Subject == "Nicoline" ,"group"] <-
-                 "faseoverganger grønn rute" # 6/D
-
-data_emne2_v23 <- data_emne2_v23 %>% relocate(group, .after = Subject)           # bruker dplyr
-```
-
 ## Instruments (data collection)
 
-Data were collected using the measure developed by Karlsen et al. (2024) / paper 1.
+Data were collected using the measure developed by @karlsen_assessing_2024 (paper 1).
 
@@
 ## Analyses
+
+### Choose subsample of PSTs
 
@@
 
-To decide which statistical tests we should use to compare pre-post scores, the Levene’s (Mean) test for equal variances and the Kolmogorov-Smirnov test for normality were used as preliminary tests (Pearce and Derrick, 2019).
+To decide which statistical tests we should use to compare pre-post scores, the Levene’s (Mean) test for equal variances and the Kolmogorov-Smirnov test for normality were used as preliminary tests [@pearce_preliminary_2019].
 Due to the small sample sizes, the data are probably non-normally distributed.
-If the variance of the pre-post differences are unequal, this suggests the use of the non-parametric Yuen d-test with trimmed means (Mair & Wilcox, 2020), which show good power for low samples (Wanishsakpong et al., 2023) and can be appropriate when one is not interested in generalizing to a population (Ng  & Cribbie, 2019).
+If the variance of the pre-post differences are unequal, this suggests the use of the non-parametric Yuen d-test with trimmed means [@mair_robust_2020], which show good power for low samples [@wanishsakpong_comparison_2023] and can be appropriate when one is not interested in generalizing to a population [@ng_gamma_2019].
 If, on the other hand, the pre-post data have equal variance, this suggests the use of the Mann-Whitney (i.e., Wilcox) test with arithmetic mean.
@@
 
-beregn SID (Estrada et al., 2019).
+beregn SID [@estrada_statistics_2019].
 
@@
 
-Ref.: Marsden og Torgerson (2012)
+Ref.: @marsden_single_2012
 
@@
 
+Fixed effects were used, comparing only pre-post.
+
 ### Within-between models
@@
 
-Hybrid with variable coefficients was chosen since there was only one significant dependent variable in the fixed-effects model (Bell et al., 2019).
+Hybrid with variable coefficients was chosen since there was only one significant dependent variable in the fixed-effects model [@bell_fixed_2019].
 
@@
 
-hybrid_TK <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
+hybrid_TPACK_n19 <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                    random=~1|Subject, method="REML", data=data_emne2_v22[alle_studenter_v22,])
 
-summary(hybrid_TK)
+summary(hybrid_TPACK_n19)
 ```
@@
 
-hybrid_TK <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
+hybrid_TPACK_varcof <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                    random=~TK_S|Subject, method="REML", data=data_emne2_v22[alle_studenter_v22,])
 
-summary(hybrid_TK)
+summary(hybrid_TPACK_varcof)
 ```
@@
 
-hybrid_TK <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
+hybrid_TPACK <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                    random=~TK_S|Subject, method="REML", data=data_emne2_v22)
 
-summary(hybrid_TK)
+summary(hybrid_TPACK)
 ```
@@
 
-Use FIML (ML) for model comparison, REML for parameter estimates (Christophersen, 2018).
+Use FIML (ML) for model comparison, REML for parameter estimates [@christophersen_introduksjon_2018].
 
@@
 
-summary of details for best fit models
+#### Summary of details for best fit models
 
@@
 # Results
+
+Here I reproduce the results presented in paper 4.
+
+## Difference
+
+```{r}
+# v22
+pre_scores_v22 <- describe(scores_v22_pre$scores)
+post_scores_v22 <- describe(scores_v22_post$scores)
+
+diff_TK_v22 <- yuend(scores_v22_post$scores[,"TK"],
+                  scores_v22_pre$scores[,"TK"]
+                  )
+diff_PCK_v22 <- yuend(scores_v22_post$scores[,"PCK"],
+                  scores_v22_pre$scores[,"PCK"]
+                  )
+diff_TPACK_v22 <- yuend(scores_v22_post$scores[,"TPACK"],
+                  scores_v22_pre$scores[,"TPACK"]
+                  )
+
+# v23
+pre_scores_v23 <- describe(scores_v23_pre$scores)
+post_scores_v23 <- describe(scores_v23_post$scores)
+
+diff_TK_v23 <- post_scores_v23["TK","mean"] - pre_scores_v23["TK","mean"]
+diff_p_TK_v23 <- wilcox.test(scores_v23_post$scores[,"TK"], scores_v23_pre$scores[,"TK"], paired=T)$p.value
+
+diff_PCK_v23 <- post_scores_v23["PCK","mean"] - pre_scores_v23["PCK","mean"]
+diff_p_PCK_v23 <- wilcox.test(scores_v23_post$scores[,"PCK"], scores_v23_pre$scores[,"PCK"], paired=T)$p.value
+
+diff_TPACK_v23 <- post_scores_v23["TPACK","mean"] - pre_scores_v23["TPACK","mean"]
+diff_p_TPACK_v23 <- wilcox.test(scores_v23_post$scores[,"TPACK"], scores_v23_pre$scores[,"TPACK"], paired=T)$p.value
+
+# create table
+rad_TK_v22 <- list("TK",
+  pre_scores_v22["TK","mean"],
+            pre_scores_v22["TK","sd"],
+            post_scores_v22["TK","mean"],
+            post_scores_v22["TK","sd"],
+            paste0(format(diff_TK_v22$diff,digit=2),
+                   " (p = ", 
+                   format(diff_TK_v22$p.value, digits=2), 
+                   ")"))
+
+rad_PCK_v22 <- list("PCK",
+  pre_scores_v22["PCK","mean"],
+            pre_scores_v22["PCK","sd"],
+            post_scores_v22["PCK","mean"],
+            post_scores_v22["PCK","sd"],
+            paste0(format(diff_PCK_v22$diff,digit=2),
+                   " (p = ", 
+                   format(diff_PCK_v22$p.value, digits=2), 
+                   ")"))
+
+rad_TPACK_v22 <- list("TPACK",
+  pre_scores_v22["TPACK","mean"],
+            pre_scores_v22["TPACK","sd"],
+            post_scores_v22["TPACK","mean"],
+            post_scores_v22["TPACK","sd"],
+            paste0(format(diff_TPACK_v22$diff,digit=2),
+                   " (p = ", 
+                   format(diff_TPACK_v22$p.value, digits=2), 
+                   ")"))
+
+rad_TK_v23 <- list("TK",
+            pre_scores_v23["TK","mean"],
+            pre_scores_v23["TK","sd"],
+            post_scores_v23["TK","mean"],
+            post_scores_v23["TK","sd"],
+            paste0(format(diff_TK_v23, digit=2),
+                   " (p = ",
+                   format(diff_p_TK_v23, digit=2),
+                   ")")
+            )
+
+rad_PCK_v23 <- list("PCK",
+            pre_scores_v23["PCK","mean"],
+            pre_scores_v23["PCK","sd"],
+            post_scores_v23["PCK","mean"],
+            post_scores_v23["PCK","sd"],
+            paste0(format(diff_PCK_v23, digit=2),
+                   " (p = ",
+                   format(diff_p_PCK_v23, digit=2),
+                   ")")
+            )
+
+rad_TPACK_v23 <- list("TPACK",
+            pre_scores_v23["TPACK","mean"],
+            pre_scores_v23["TPACK","sd"],
+            post_scores_v23["TPACK","mean"],
+            post_scores_v23["TPACK","sd"],
+            paste0(format(diff_TPACK_v23, digit=2, nsmall=2),
+                   " (p = ",
+                   format(diff_p_TPACK_v23, digit=2, nsmall=2),
+                   ")")
+            )
+
+row1_p1 <- c("", "Prototype 1", "", "", "", "")
+row1_p2 <- c("", "Prototype 2", "", "", "", "")
+row2 <- c("", "Pre", "", "Post", "", "")
+row3 <- c("", "Mean", "SD", "Mean", "SD", "diff")
+
+prepost_table_v22 <- t(data.frame(
+  row1_p1, row2, row3,
+  format(rad_TK_v22,digit=2,nsmall=2), # only 2 digits
+  format(rad_PCK_v22,digit=2,nsmall=2), # only 2 digits
+  format(rad_TPACK_v22,digit=2,nsmall=2) # only 2 digits
+))
+
+prepost_table_v23 <- t(data.frame(
+  row1_p2, row2, row3,
+  format(rad_TK_v23,digit=2,nsmall=2), # only 2 digits
+  format(rad_PCK_v23,digit=2,nsmall=2), # only 2 digits
+  format(rad_TPACK_v23,digit=2,nsmall=2) # only 2 digits
+))
+
+rownames(prepost_table_v22) <- NULL # remove row names
+rownames(prepost_table_v23) <- NULL # remove row names
+
+pander(prepost_table_v22)
+pander(prepost_table_v23)
+```
+
+## Fixed effects
+
+```{r}
+screenreg(list("within22" = within22,
+               "within23pp" = within23_prepost))
+```
+
+## Within-between
+
+```{r}
+summary(hybrid_TPACK)
+```
+
+TK_S and PCK_S are the within-effects, while TK_mean and PCK_mean are the between-effects.
+
+This model is however not directly comparably to the fixed effects model, since the hybrid model uses all the students in the analysis (n=21+20), while the fixed effects only uses the students that are in both pre-post datasets (n=19+19) (cf.
+model hybrid_TPACK_n19 above).
+
+```{r}
+# print summary of models
+screenreg(list("Hybrid (all students)" = hybrid_TPACK, 
+               "Hybrid (pre-post students)" = hybrid_TPACK_n19))
+```
+
+# Discussion
+
+The hybrid model reported in paper 4, should perhaps be switched to the n=19 model, for more direct comparison with the fixed effects model.
+
+# References
+
+<!-- citation(package) -->
+
+<div id="refs"></div>
+
+<!-- source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html
+
+# (APPENDIX) Appendix {-} -->
+
+# Session info
+<!-- # ref.: https://intro2r.com/proj_doc.html -->
+<!-- #sessionInfo() -->
+
+```{r}
+xfun::session_info()
+```

### 2024-10-08 Rmd/Paper4.Rmd
Added some comments and section headings

@@
 
+The file tpack-pilot_2022.Rdata was created by the Paper1.Rmd-script.
+
+The file tpack-emne2_2023.Rdata was created by the datareduksjon-TPACK-PST-2023.Rmd-script, which is not published, but available by request. It is similar to the 2022-script.
+
 ```{r load_data}
@@
 
+### Recalculate values
+
 ```{r}
@@
 diff_p_TPACK_v23 <- wilcox.test(scores_v23_post$scores[,"TPACK"], scores_v23_pre$scores[,"TPACK"], paired=T)$p.value
-
+```
+
+### Create table
+
+```{r}
 # create table
@@
 rownames(prepost_table_v23) <- NULL # remove row names
-
+```
+
+### Present results in table
+
+This is the table that was presented in paper 4
+
+```{r}
 pander(prepost_table_v22)

### 2024-10-08 Rmd/Paper1.Rmd
Added separate presentation of results in paper for easy access

@@
 ```{r eval=T}
-alpha(df_run2[,c("TK1", "TK2", "TK3")])
-alpha(df_run2[,c("CK1", "PCK2", "PCK5", "CK5", "CK3", "CK2")])
-alpha(df_run2[,c("TPK1","TPK4","TCK1","TCK4", "TPACK2","TPACK4")]) # , "TPK4"
+alpha_TK <- alpha(df_run2[,c("TK1", "TK2", "TK3")])
+alpha_PCK <- alpha(df_run2[,c("CK1", "PCK2", "PCK5", "CK5", "CK3", "CK2")])
+alpha_TPACK <- alpha(df_run2[,c("TPK1","TPK4","TCK1","TCK4", "TPACK2","TPACK4")]) # , "TPK4"
+
+alpha_TK
+alpha_PCK
+alpha_TPACK
 ```
@@
 omega(df_run2, fm="mle") #, rotate="Promax")
+```
+
+# Results of EFA
+
+Here I reproduce the results presented in paper 1.
+
+table 1
+
+```{r}
+print.psych(fa.sort(model_3factor_run2))
+```
+
+```{r}
+alpha_TK$total$std.alpha
+alpha_PCK$total$std.alpha
+alpha_TPACK$total$std.alpha
 ```
@@
 
-df    <- CFA_TPACK_PST[,unlist(TPACK_PST_items[1:3])]
-describe(df)      # with missing (see n)? (na.rm=TRUE)
-
-score_TPACK <- scoreItems(TPACK_PST_items,df)
+df_CFA    <- CFA_TPACK_PST[,unlist(TPACK_PST_items[1:3])]
+describe(df_CFA)      # with missing (see n)? (na.rm=TRUE)
+
+score_TPACK <- scoreItems(TPACK_PST_items,df_CFA)
 describe(score_TPACK$scores)
@@
 cor_mat <- lavInspect(model_three_factor, what="cor.ov")
-omega(cor_mat, nfactors=3)
+observed_omega <- omega(cor_mat, nfactors=3)
+observed_omega
 
@@
 ```{r measureQ, eval=T}
-measureQ(model_three_factor, CFA_TPACK_PST)
+table.2 <- measureQ(model_three_factor, CFA_TPACK_PST)
+```
+
+# Results of CFA
+
+Table 2
+
+```{r}
+pander(fit_table)
+```
+
+Figure 3
+
+std.all (column to the far right) shows the standardized results used in figure 3.
+
+```{r}
+parameterEstimates(model_three_factor, standardized=TRUE)
+```
+
+Table 3
+
+```{r}
+describe(df_CFA)
+```
+
+F1 equals TPACK, F2 equals PCK and F3 equals TK. Omega_t is the first column, while omega_h is the second column.
+
+```{r}
+observed_omega$omega.group
+#observed_omega$omega.group$general
+#observed_omega$omega.group$total
+```
+
+Table 4
+
+```{r}
+
+so.lv <- NULL # temp variable second order level
+
+cat("\n", "  Table 2. Descriptive Statistics (Observed Mean, Latent s.d., AVE, Construct Reliability, Latent Correlation)", rep("\n", 2))
+rownames(table.2) <- rep("    ", nrow(table.2))
+print(table.2, quote=FALSE, right=TRUE)
+cat("\n", "  Note: AVE = Average Variance Extracted; * = AVE significantly lower than 0.5 (p < .05)","\n")
+if (length(so.lv) == 0) {
+  cat("         diagonal elements in brackets = Construct Reliability","\n")
+} else {
+  cat(paste0("         diagonal elements in brackets = Construct Reliability for first-order factor and ", omega, " for second-order factor"), "\n")
+}
+cat("         A = Construct Reliability significantly lower than 0.7; B = Construct Reliability significantly lower than 0.8 (p < .05)", "\n")
+cat("         Correlation coefficient: a = significantly larger than 0.85; b = significantly larger than 0.8; c = significantly larger than 0.7 (p < .05)", "\n")
+if (length(so.lv) == 0) {
+  cat("         # = AVE is significantly less than squared-correlation (p < .05)", rep("\n", 3))
+} else {
+  cat("         # = AVE is significantly less than squared-correlation (p < .05)", rep("\n"))
+  cat("         Observed mean of second-order factor is based on all items ignoring the second-order structure", rep("\n", 3))
+}
 ```
